---
title: "HR Predictive Analytics Project – Turnover Intention"
author: "Kevin Allen Cauley"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    theme: flatly
    highlight: tango
    df_print: paged
---

```{r 00-setup, include=FALSE}
# Set working directory to the project root
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

# Create output folders if they don't exist
if (!dir.exists("outputs")) dir.create("outputs", recursive = TRUE)
if (!dir.exists("outputs/figures")) dir.create("outputs/figures", recursive = TRUE)
if (!dir.exists("data/processed")) dir.create("data/processed", recursive = TRUE)

# Global chunk options
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.path = "outputs/figures/",
  fig.align = "center",
  cache = TRUE
)

# Load previous session
if (file.exists("outputs/my_environment.RData")) {
  load("outputs/my_environment.RData")
}
```

```{r 00-setup-save, include=FALSE}
# Save the current R session to restore later
save.image(file = "outputs/my_environment.RData")
```

```{r 00-setup-packages, include=FALSE}
# Load packages
library(car)
library(cowplot)
library(caret)
library(corrplot)
library(discrim)
library(dplyr)
library(e1071)
library(fastDummies)
library(forcats)
library(gbm)
library(GGally)
library(ggcorrplot)
library(ggplot2)
library(ggthemes)
library(glmnet)
library(grid)
library(gridExtra)
library(iml)
library(kableExtra)
library(keras3)
library(keras)
library(knitr)
library(mgcv)
library(naivebayes)
library(naniar)
library(patchwork)
library(pheatmap)
library(plotly)
library(psych)
library(randomForest)
library(ranger)
library(RColorBrewer)
library(readxl)
library(reshape2)
library(skimr)
library(tensorflow)
library(themis)
library(tidymodels)
library(tidyverse)
library(vip)
library(viridis)
library(xgboost)
```
---
## Predictive Modeling Strategy

Following the exploratory analysis, predictive models were developed to identify factors associated with employee turnover intention. The modeling process was designed to evaluate multiple algorithms and compare their predictive performance.

Several modeling approaches were implemented to capture both linear and non-linear relationships within the workforce dataset.

The modeling workflow consisted of three primary stages.

### Data Preparation

Before modeling, the dataset was prepared to ensure compatibility with machine learning algorithms. Numeric predictors were standardized where appropriate to ensure comparable feature scales. Categorical predictors were encoded to allow models to properly interpret categorical information. The primary target variable for this analysis is Turnover Intention.

### Model Development

Multiple modeling approaches were evaluated in order to compare predictive performance across both linear and non-linear algorithms. The models evaluated in this analysis include:

The models evaluated include:

- Logistic Regression
- Regularized Logistic Regression (LASSO / Ridge)
- Random Forest
- Gradient Boosted Trees
- Support Vector Machines

### Model Evaluation

Models were evaluated using:

- Accuracy
- Precision
- Recall
- F1 Score
- Area Under the ROC Curve (AUC)

Comparing models across these metrics allows identification of the approach that best captures the underlying patterns associated with employee turnover intention.

## Predictive Modeling on Turnover Intention

The goal of this modeling section is to predict **Turnover Intention** using a set of behavioral, demographic, and engagement-related predictors.

We will evaluate a range of classification models using multiple **resampling strategies** (No Sampling, Up-Sampling, Down-Sampling, and SMOTE). Models will be evaluated using six classification metrics:

-   Accuracy
-   Precision
-   Recall
-   F1 Score
-   Specificity
-   ROC AUC

This approach ensures fairness, robustness, and clarity in identifying the most effective model and configuration.

### Shared Model Setup - Turnover Intention

```{r 04-logit-shared-model-setup}
# Ensure Turnover Intention is a Factor
hr_df <- hr_df %>%
  mutate(Turnover_Intention = as.factor(Turnover_Intention))

# Set Seed
set.seed(42)

# 80/20 train/test split with stratification
turnover_split <- initial_split(hr_df, prop = 0.8, strata = Turnover_Intention)
turnover_train_data <- training(turnover_split)
turnover_test_data  <- testing(turnover_split)

# 10-Fold Cross Validation
turnover_folds <- vfold_cv(turnover_train_data, v = 10, strata = Turnover_Intention)

# Shared Pre-processing (no resampling)
base_recipe <- recipe(Turnover_Intention ~ ., data = turnover_train_data) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_numeric_predictors())

# Basic Logistic Model
logit_model <- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")
```

```{r 04-logit-turnover-shared-metrics-setup}
# Shared metric set for all models
all_metrics <- metric_set(accuracy, precision, recall, f_meas, yardstick::spec, roc_auc)
```

## Turnover Intention with Logistic Regression

### Logistic Regression - No Resampling

```{r 04-logit-none-workflow, message=FALSE}
# Workflow with Base Recipe
workflow_logit_none <- workflow() %>%
  add_model(logit_model) %>%
  add_recipe(base_recipe)

# Fit to training data
fit_logit_none <- fit(workflow_logit_none, data = turnover_train_data)

# Predict on test data
logit_preds_none <- predict(fit_logit_none, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(fit_logit_none, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(Turnover_Intention))
```

### Logistic Regression - No Resampling - Performance Metrics

```{r 04-logit-none-metrics, echo=TRUE}
# Metrics
logit_none_metrics <- all_metrics(logit_preds_none, truth = Turnover_Intention, estimate = .pred_class, .pred_Yes)

# Plot
logit_none_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#4B9CD3", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – LR (No Resampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

```{r 04-logit-none-confmat, echo=TRUE}
library(ggplot2)
library(yardstick)
library(forcats)

# Confusion matrix to dataframe
cm_logit_none <- conf_mat(logit_preds_none, truth = Turnover_Intention, estimate = .pred_class)
cm_df <- as.data.frame(cm_logit_none$table)

# Ensure proper order for plotting
cm_df$Truth <- factor(cm_df$Truth, levels = c("Yes", "No"))         
cm_df$Prediction <- factor(cm_df$Prediction, levels = c("No", "Yes")) 

# Plot
ggplot(cm_df, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#e6f2ff", high = "#1A5276", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – LR (No Resampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )
```

```{r 04-logit-none-roc-curve, echo=TRUE}
# Generate ROC curve
roc_curve(logit_preds_none, truth = Turnover_Intention, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – LR (No Resampling)",
    subtitle = "Model performance on predicting Turnover Intention",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )
```

```{r 04-logit-none-vip, echo=TRUE}
# Extract full importance and then limit to top 15
vip_logit_none <- fit_logit_none %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance))

# Plot
ggplot(vip_logit_none, aes(x = Importance, y = Variable)) +
  geom_col(fill = "#1A5276", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – LR (No Resampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

### Logistic Regression – SMOTE

```{r 04-logit-smote-workflow, message=FALSE}
# SMOTE recipe
recipe_logit_smote <- base_recipe %>%
  step_smote(Turnover_Intention)

workflow_logit_smote <- workflow() %>%
  add_model(logit_model) %>%
  add_recipe(recipe_logit_smote)

# 10-Fold CV
set.seed(42)
logit_cv_smote <- fit_resamples(
  workflow_logit_smote,
  resamples = turnover_folds,
  metrics = metric_set(accuracy, precision, recall, f_meas, yardstick::spec, roc_auc)
)

# Final model fit and prediction on test data for SMOTE
fit_logit_smote <- fit(workflow_logit_smote, data = turnover_train_data)

logit_preds_smote <- predict(fit_logit_smote, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(fit_logit_smote, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(Turnover_Intention))
```

### Logistic Regression – SMOTE - Performance Metrics

```{r 04-logit-smote-finalfit, message=FALSE}
# Metrics
logit_smote_metrics <- all_metrics(logit_preds_smote, truth = Turnover_Intention, estimate = .pred_class, .pred_Yes)

# Plot
logit_smote_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#4B9CD3", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – LR (SMOTE)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

```{r 04-logit-smote-confmat, echo=TRUE}
# Create confusion matrix
cm_logit_smote <- conf_mat(logit_preds_smote, truth = Turnover_Intention, estimate = .pred_class)
cm_df_smote <- as.data.frame(cm_logit_smote$table)

# Order axes
cm_df_smote$Truth <- factor(cm_df_smote$Truth, levels = c("Yes", "No"))
cm_df_smote$Prediction <- factor(cm_df_smote$Prediction, levels = c("No", "Yes"))

# Plot heatmap
ggplot(cm_df_smote, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#e6f2ff", high = "#1A5276", name = "Frequency") +
  scale_y_discrete(limits = rev) +  # YES on top
  labs(
    title = "Confusion Matrix – LR (SMOTE)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )
```

```{r 04-logit-smote-roc, echo=TRUE}
# ROC Curve
roc_curve(logit_preds_smote, truth = Turnover_Intention, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – LR (SMOTE)",
    subtitle = "Model performance on predicting Turnover Intention",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )
```

```{r 04-logit-smote-vip, echo=TRUE}
# Top Features only
fit_logit_smote %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#1A5276", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – LR (SMOTE)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

### Logistic Regression - Upsampling

```{r 04-logit-upsample-workflow, message=FALSE}
# Upsampling Recipe
recipe_logit_upsample <- base_recipe %>%
  step_upsample(Turnover_Intention)

# Workflow 
workflow_logit_upsample <- workflow() %>%
  add_model(logit_model) %>%
  add_recipe(recipe_logit_upsample)

# Fit
fit_logit_upsample <- fit(workflow_logit_upsample, data = turnover_train_data)

# Predict
logit_preds_upsample <- predict(fit_logit_upsample, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(fit_logit_upsample, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(Turnover_Intention))
```

### LR - Upsampling - Performance Metrics

```{r 04-logit-upsample-metrics, echo=TRUE}
# Metrics
logit_upsample_metrics <- all_metrics(logit_preds_upsample, truth = Turnover_Intention, estimate = .pred_class, .pred_Yes)

# Plot
logit_upsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#4B9CD3", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – LR (Upsampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

```{r 04-logit-upsample-confmat, echo=TRUE}
# Confusion matrix data prep
cm_logit_upsample <- conf_mat(logit_preds_upsample, truth = Turnover_Intention, estimate = .pred_class)
cm_df_upsample <- as.data.frame(cm_logit_upsample$table)

# Set factor levels to match layout: Yes on top, No bottom; No left, Yes right
cm_df_upsample$Truth <- factor(cm_df_upsample$Truth, levels = c("Yes", "No"))
cm_df_upsample$Prediction <- factor(cm_df_upsample$Prediction, levels = c("No", "Yes"))

# Plot
ggplot(cm_df_upsample, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#e6f2ff", high = "#1A5276", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – LR (Upsampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )
```

```{r 04-logit-upsample-roc-curve, echo=TRUE}
# Roc
roc_curve(logit_preds_upsample, truth = Turnover_Intention, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – LR (Upsampling)",
    subtitle = "Model performance on predicting Turnover Intention",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )
```

```{r 04-logit-upsample-vip, echo=TRUE}
# Styled VIP for Top Features
fit_logit_upsample %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#1A5276", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – LR (Upsampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

### Logistic Regression - Downsampling

```{r 04-logit-downsample-workflow, message=FALSE}
# Downsampling Recipe
recipe_logit_downsample <- base_recipe %>%
  step_downsample(Turnover_Intention)

# Workflow 
workflow_logit_downsample <- workflow() %>%
  add_model(logit_model) %>%
  add_recipe(recipe_logit_downsample)

# Fit on training data
fit_logit_downsample <- fit(workflow_logit_downsample, data = turnover_train_data)

# Predict on test data
logit_preds_downsample <- predict(fit_logit_downsample, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(fit_logit_downsample, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(Turnover_Intention))
```

### Logistic Regression - Downsampling - Performance Metrics

```{r 04-logit-downsample-metrics, echo=TRUE}
# Metrics
logit_downsample_metrics <- all_metrics(logit_preds_downsample, truth = Turnover_Intention, estimate = .pred_class, .pred_Yes)

# Plot
logit_downsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#4B9CD3", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – LR (Downsampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

```{r 04-logit-downsample-confmat, echo=TRUE}
cm_logit_down <- conf_mat(logit_preds_downsample, truth = Turnover_Intention, estimate = .pred_class)
cm_df_down <- as.data.frame(cm_logit_down$table)

cm_df_down$Truth <- factor(cm_df_down$Truth, levels = c("Yes", "No"))
cm_df_down$Prediction <- factor(cm_df_down$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_down, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#e6f2ff", high = "#1A5276", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – LR (Downsampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )
```

```{r 04-logit-downsample-roc-curve, echo=TRUE}
roc_curve(logit_preds_downsample, truth = Turnover_Intention, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – LR (Downsampling)",
    subtitle = "Model performance on predicting Turnover Intention",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )
```

```{r 04-logit-downsample-vip, echo=TRUE}
fit_logit_downsample %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#1A5276", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – LR (Downsampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

### Results Table for Logistic Regression (Turnover Intention)

```{r 04-logit-results-table, results='asis', echo=FALSE}
logit_results_long <- bind_rows(
  logit_none_metrics %>% mutate(Resample = "No Sampling"),
  logit_smote_metrics %>% mutate(Resample = "SMOTE"),
  logit_upsample_metrics %>% mutate(Resample = "Up Sampling"),
  logit_downsample_metrics %>% mutate(Resample = "Down Sampling")
) %>%
  mutate(
    Metric = factor(.metric,
                    levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                    labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC")),
    Score = .estimate,
    Resample = factor(Resample, levels = c("No Sampling", "SMOTE", "Up Sampling", "Down Sampling"))
  ) %>%
  filter(!is.na(Score))

highlight_logit <- logit_results_long %>%
  group_by(Metric) %>%
  filter(Score == max(Score, na.rm = TRUE)) %>%
  mutate(is_best = TRUE)

logit_results_highlight <- left_join(logit_results_long, highlight_logit %>% select(Resample, Metric, is_best),
                                     by = c("Resample", "Metric")) %>%
  mutate(is_best = ifelse(is.na(is_best), FALSE, is_best))

ggplot(logit_results_highlight, aes(x = Resample, y = Score, fill = Metric)) +
  geom_col(width = 0.7) +
  geom_col(data = subset(logit_results_highlight, is_best),
           fill = NA, color = "black", linewidth = 1.2, width = 0.7) +
  facet_wrap(~ Metric, nrow = 2, scales = "free_y") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_manual(values = rep("#4B9CD3", 6)) +
  labs(
    title = "LR Performance by Metric and Resampling Method",
    subtitle = "Black outline shows the best result for each metric",
    x = "Resampling Strategy", y = "Score"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 30, hjust = 1),
    strip.text = element_text(face = "bold"),
    legend.position = "none"
  )
```

### Executive Summary for Logistic Regression (Turnover Intention)

Logistic Regression provided a clear and interpretable baseline across all resampling methods. While it didn’t lead in any individual metric, it consistently delivered **balanced, moderate performance** — ideal for transparent reporting and stakeholder understanding.

- **SMOTE** offered the most balanced results overall, delivering the highest **recall** (**0.478**) and **F1 score** (**0.484**), making it suitable when catching more at-risk employees is a priority.

- **Upsampling** gave the best **precision** (**0.463**) and **specificity** (**0.596**), helping to reduce false positives in turnover predictions.

- **Downsampling** maintained stable performance across all metrics, with a shared peak in **recall**, **precision**, and **F1 score** (**0.478**) and modest **ROC AUC** (**0.437**).

- **No Resampling** produced strong **specificity** (**0.670**) and decent **accuracy** (**0.527**), though it struggled more with **recall** (**0.359**) due to class imbalance.
---

## Turnover Intention with Random Forest

### Random Forest - No Resampling

```{r 04-rf-no-resampling-workflow, message=FALSE}
# Define Random Forest model
rf_model <- rand_forest(mtry = tune(), trees = 500, min_n = tune()) %>%
  set_engine("ranger", importance = "impurity") %>%
  set_mode("classification")

# Workflow using base recipe
recipe_rf_none <- base_recipe

# Combine into Workflow
workflow_rf_none <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(recipe_rf_none)

# Tune the RF model with cross-validation
set.seed(42)
rf_tuned <- tune_grid(
  workflow_rf_none,
  resamples = turnover_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_rf <- select_best(rf_tuned, metric = "roc_auc")

# Final workflow with best RF model
final_rf_workflow <- finalize_workflow(workflow_rf_none, best_rf)

# Fit to training data
rf_fit_none <- fit(final_rf_workflow, data = turnover_train_data)

# Predict on test data
rf_preds_none <- predict(rf_fit_none, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(rf_fit_none, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(Turnover_Intention))
```

### Random Forest - No Resampling - Performance Metrics

```{r 04-rf-no-resampling-metrics, echo=TRUE}
# Metrics
rf_none_metrics <- all_metrics(rf_preds_none, truth = Turnover_Intention, estimate = .pred_class, .pred_Yes)

# Plot
rf_none_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#3FA34D", width = 0.7) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – RF (No Resampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

```{r 04-rf-no-resampling-confmat, echo=TRUE}
rf_cm_none <- conf_mat(rf_preds_none, truth = Turnover_Intention, estimate = .pred_class)
rf_cm_df_none <- as.data.frame(rf_cm_none$table)

rf_cm_df_none$Truth <- factor(rf_cm_df_none$Truth, levels = c("Yes", "No"))
rf_cm_df_none$Prediction <- factor(rf_cm_df_none$Prediction, levels = c("No", "Yes"))

ggplot(rf_cm_df_none, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#D5F5E3", high = "#145A32", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – RF (No Resampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )
```

```{r 04-rf-no-resampling-roc-curve, echo=TRUE}
# ROC Curve
roc_curve(rf_preds_none, truth = Turnover_Intention, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray70") +
  labs(
    title = "ROC Curve – RF (No Resampling)",
    subtitle = "Model performance on predicting Turnover Intention",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

```{r 04-rf-no-resampling-vip, echo=TRUE}
# Variable Importance
rf_fit_none %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#3FA34D", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – RF (No Resampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

### Random Forest - SMOTE

```{r 04-rf-smote-workflow, message=FALSE}
# Create Recipe from Base
recipe_rf_smote <- base_recipe %>%
  step_smote(Turnover_Intention)

# Workflow with Random Forest
workflow_rf_smote <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(recipe_rf_smote)

# Tune the RF model with cross-validation
set.seed(42)
rf_tuned_smote <- tune_grid(
  workflow_rf_smote,
  resamples = turnover_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_rf_smote <- select_best(rf_tuned_smote, metric = "roc_auc")

# Final workflow with best RF model
final_rf_workflow_smote <- finalize_workflow(workflow_rf_smote, best_rf_smote)

# Fit to training data
rf_fit_smote <- fit(final_rf_workflow_smote, data = turnover_train_data)

# Predict on test data
rf_preds_smote <- predict(rf_fit_smote, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(rf_fit_smote, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(Turnover_Intention))
```

### Random Forest - SMOTE - Performance Metrics

```{r 04-rf-smote-metrics, echo=TRUE}
# Metrics
rf_smote_metrics <- all_metrics(rf_preds_smote, truth = Turnover_Intention, estimate = .pred_class, .pred_Yes)

# Plot
rf_smote_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#3FA34D", width = 0.7) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – RF (SMOTE)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-rf-smote-confmat, echo=TRUE}
# Confusion Matrix
rf_cm_smote <- conf_mat(rf_preds_smote, truth = Turnover_Intention, estimate = .pred_class)
rf_cm_df_smote <- as.data.frame(rf_cm_smote$table)

rf_cm_df_smote$Truth <- factor(rf_cm_df_smote$Truth, levels = c("Yes", "No"))
rf_cm_df_smote$Prediction <- factor(rf_cm_df_smote$Prediction, levels = c("No", "Yes"))

ggplot(rf_cm_df_smote, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#D5F5E3", high = "#145A32", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – RF (SMOTE)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-rf-smote-roc-curve, echo=TRUE}
# ROC Curve
roc_curve(rf_preds_smote, truth = Turnover_Intention, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray70") +
  labs(
    title = "ROC Curve: RF on Turnover Intention (SMOTE)",
    subtitle = "Model performance on predicting Turnover Intention",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

```{r 04-rf-smote-vip, echo=TRUE}
# Variable Importance
rf_fit_smote %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#3FA34D", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – RF (SMOTE)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

### Random Forest - Upsampling

```{r 04-rf-upsample-workflow, message=FALSE}
# Create Recipe from Base
recipe_rf_upsample <- base_recipe %>%
  step_upsample(Turnover_Intention)

# Workflow with Random Forest
workflow_rf_upsample <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(recipe_rf_upsample)

# Tune the RF model with cross-validation
set.seed(42)
rf_tuned_upsample <- tune_grid(
  workflow_rf_upsample,
  resamples = turnover_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_rf_upsample <- select_best(rf_tuned_upsample, metric = "roc_auc")

# Final workflow with best RF model
final_rf_workflow_upsample <- finalize_workflow(workflow_rf_upsample, best_rf_upsample)

# Fit to training data
rf_fit_upsample <- fit(final_rf_workflow_upsample, data = turnover_train_data)

# Predict on test data
rf_preds_upsample <- predict(rf_fit_upsample, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(rf_fit_upsample, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(Turnover_Intention))
```

### Random Forest - Upsampling - Performance Metrics

```{r 04-rf-upsample-metrics, echo=TRUE}
# Metrics
rf_upsample_metrics <- all_metrics(rf_preds_upsample, truth = Turnover_Intention, estimate = .pred_class, .pred_Yes)

# Plot
rf_upsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#3FA34D", width = 0.7) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – RF (Upsampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-rf-upsample-confmat, echo=TRUE}
# Confusion Matrix
rf_cm_up <- conf_mat(rf_preds_upsample, truth = Turnover_Intention, estimate = .pred_class)
rf_cm_df_up <- as.data.frame(rf_cm_up$table)

rf_cm_df_up$Truth <- factor(rf_cm_df_up$Truth, levels = c("Yes", "No"))
rf_cm_df_up$Prediction <- factor(rf_cm_df_up$Prediction, levels = c("No", "Yes"))

ggplot(rf_cm_df_up, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#D5F5E3", high = "#145A32", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – RF (Upsampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-rf-upsample-roc-curve, echo=TRUE}
# ROC Curve
roc_curve(rf_preds_upsample, truth = Turnover_Intention, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray70") +
  labs(
    title = "ROC Curve – RF (Upsampling)",
    subtitle = "Model performance on predicting Turnover Intention",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

```{r 04-rf-upsample-var-imp, echo=TRUE}
# Variable Importance
rf_fit_upsample %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#3FA34D", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – RF (Upsampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

### Random Forest - Downsampling

```{r 04-rf-downsample-workflow, message=FALSE}
# Create Recipe from Base
recipe_rf_downsample <- base_recipe %>%
  step_downsample(Turnover_Intention)

# Workflow with Random Forest
workflow_rf_downsample <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(recipe_rf_downsample)

# Tune the RF model with cross-validation
set.seed(42)
rf_tuned_downsample <- tune_grid(
  workflow_rf_downsample,
  resamples = turnover_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_rf_downsample <- select_best(rf_tuned_downsample, metric = "roc_auc")

# Final workflow with best RF model
final_rf_workflow_downsample <- finalize_workflow(workflow_rf_downsample, best_rf_downsample)

# Fit to training data
rf_fit_downsample <- fit(final_rf_workflow_downsample, data = turnover_train_data)

# Predict on test data
rf_preds_downsample <- predict(rf_fit_downsample, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(rf_fit_downsample, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(Turnover_Intention))
```

### Random Forest - Downsampling - Performance Metrics

```{r 04-rf-downsample-metrics, echo=TRUE}
# Metrics
rf_downsample_metrics <- all_metrics(rf_preds_downsample, truth = Turnover_Intention, estimate = .pred_class, .pred_Yes)

# Plot
rf_downsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#3FA34D", width = 0.7) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – RF (Downsampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-rf-downsample-confmat, echo=TRUE}
# Confusion Matrix
rf_cm_down <- conf_mat(rf_preds_downsample, truth = Turnover_Intention, estimate = .pred_class)
rf_cm_df_down <- as.data.frame(rf_cm_down$table)

rf_cm_df_down$Truth <- factor(rf_cm_df_down$Truth, levels = c("Yes", "No"))
rf_cm_df_down$Prediction <- factor(rf_cm_df_down$Prediction, levels = c("No", "Yes"))

ggplot(rf_cm_df_down, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#D5F5E3", high = "#145A32", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – RF (Downsampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-rf-downsample-roc-curve, echo=TRUE}
# ROC Curve
roc_curve(rf_preds_downsample, truth = Turnover_Intention, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray70") +
  labs(
    title = "ROC Curve – RF (Downsampling)",
    subtitle = "Model performance on predicting Turnover Intention",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

```{r 04-rf-downsample-vip, echo=TRUE}
# Variable Importance
rf_fit_downsample %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#3FA34D", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – RF (Downsampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

### Results Table for Random Forest (Turnover Intention)

```{r 04-rf-results-table, results='asis', echo=TRUE}
rf_results_long <- rf_results_wide %>%
  pivot_longer(cols = -c(Model, Resample), names_to = "Metric", values_to = "Score") %>%
  mutate(
    Metric = factor(Metric,
                    levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                    labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC")),
    Resample = factor(Resample, levels = c("No Sampling", "SMOTE", "Up Sampling", "Down Sampling"))
  ) %>%
  filter(!is.na(Score))

highlight_rf <- rf_results_long %>%
  group_by(Metric) %>%
  filter(Score == max(Score, na.rm = TRUE)) %>%
  mutate(is_best = TRUE)

rf_results_highlight <- left_join(rf_results_long, highlight_rf %>% select(Resample, Metric, is_best),
                                  by = c("Resample", "Metric")) %>%
  mutate(is_best = ifelse(is.na(is_best), FALSE, is_best))

ggplot(rf_results_highlight, aes(x = Resample, y = Score, fill = Metric)) +
  geom_col(width = 0.7) +
  geom_col(data = subset(rf_results_highlight, is_best),
           fill = NA, color = "black", linewidth = 1.2, width = 0.7) +
  facet_wrap(~ Metric, nrow = 2, scales = "free_y") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_manual(values = rep("#3FA34D", 6)) +
  labs(
    title = "RF Performance by Metric and Resampling Method",
    subtitle = "Black outline shows the best result for each metric",
    x = "Resampling Strategy", y = "Score"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 30, hjust = 1),
    strip.text = element_text(face = "bold"),
    legend.position = "none"
  )
```

### Executive Summary for Random Forest (Turnover Intention)

Random Forest showed strong performance on several key metrics, particularly when handling imbalanced data. While not always the top F1 performer, it delivered impressive results for **precision**, **specificity**, and **ROC AUC**, confirming its robustness and predictive power across resampling strategies.

- **Upsampling** yielded the best **precision** (**0.556**), **specificity** (**0.927**), and **accuracy** (**0.552**), making it well-suited for minimizing false positives and identifying confident cases of turnover risk.

- **Downsampling** balanced metrics across the board with solid **recall** (**0.478**), **F1 score** (**0.463**), and **ROC AUC** (**0.495**), offering a steady and dependable configuration.

- **SMOTE** improved **recall** (**0.380**) and **F1 score** (**0.417**) over the baseline, helpful when capturing at-risk employees is critical.

- **No Resampling** still performed reasonably well, especially in **specificity** (**0.651**) and **ROC AUC** (**0.499**), showing Random Forest’s natural ability to handle imbalance without adjustments.

## Turnover Intention with XGBoost

### XGBoost - No Resampling

```{r 04-xgb-no-resampling-workflow, message=FALSE}
# Define XGBoost model
xgb_model <- boost_tree(
  trees = 1000,
  tree_depth = tune(),
  min_n = tune(),
  loss_reduction = tune(),        
  sample_size = tune(),           
  mtry = tune(),                  
  learn_rate = tune()             
) %>%
  set_engine("xgboost") %>%
  set_mode("classification")

# Recipe with no resampling
recipe_xgb_none <- base_recipe

# Combine into Workflow
workflow_xgb_none <- workflow() %>%
  add_model(xgb_model) %>%
  add_recipe(recipe_xgb_none)

# Tune the XGBoost model with cross-validation
set.seed(42)
xgb_tuned <- tune_grid(
  workflow_xgb_none,
  resamples = turnover_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_xgb <- select_best(xgb_tuned, metric = "roc_auc")

# Final workflow with best XGBoost model
final_xgb_workflow <- finalize_workflow(workflow_xgb_none, best_xgb)

# Fit to training data
xgb_fit_none <- fit(final_xgb_workflow, data = turnover_train_data)

# Predict on test data
xgb_preds_none <- predict(xgb_fit_none, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(xgb_fit_none, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(Turnover_Intention))
```

### XGBoost - No Resampling - Performance Metrics

```{r 04-xgb-no-resampling-metrics, echo=TRUE}
# Metrics
xgb_none_metrics <- all_metrics(xgb_preds_none, truth = Turnover_Intention, estimate = .pred_class, .pred_Yes)

# Plot
xgb_none_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#E87461", width = 0.7) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – XGB (No Resampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

```{r 04-xgb-no-resampling-confmat, echo=TRUE}
# Confusion Matrix
xgb_cm_none <- conf_mat(xgb_preds_none, truth = Turnover_Intention, estimate = .pred_class)
xgb_cm_df_none <- as.data.frame(xgb_cm_none$table)

xgb_cm_df_none$Truth <- factor(xgb_cm_df_none$Truth, levels = c("Yes", "No"))
xgb_cm_df_none$Prediction <- factor(xgb_cm_df_none$Prediction, levels = c("No", "Yes"))

ggplot(xgb_cm_df_none, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#FDEDEC", high = "#922B21", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – XGB (No Resampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-xgb-no-resampling-roc-curve, echo=TRUE}
# ROC Curve
roc_curve(xgb_preds_none, truth = Turnover_Intention, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray70") +
  labs(
    title = "ROC Curve – XGB (No Resampling)",
    subtitle = "Model performance on predicting Turnover Intention",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

```{r 04-xgb-no-resampling-vip, echo=TRUE}
# Variable Importance
xgb_fit_none %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#E87461", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – XGB (No Resampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

### XGBoost - SMOTE

```{r 04-xgb-smote-workflow, message=FALSE}
# Create SMOTE Recipe
recipe_xgb_smote <- base_recipe %>%
  step_smote(Turnover_Intention)

# Build XGBoost Workflow with SMOTE
workflow_xgb_smote <- workflow() %>%
  add_model(xgb_model) %>%
  add_recipe(recipe_xgb_smote)

# Tune Hyperparameters with Cross-Validation
set.seed(42)
xgb_tuned_smote <- tune_grid(
  workflow_xgb_smote,
  resamples = turnover_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
)

# Select Best Parameters Based on ROC AUC
best_xgb_smote <- select_best(xgb_tuned_smote, metric = "roc_auc")

# Finalize Workflow with Best Parameters
final_xgb_workflow_smote <- finalize_workflow(workflow_xgb_smote, best_xgb_smote)

# Fit Final Model to Training Data
xgb_fit_smote <- fit(final_xgb_workflow_smote, data = turnover_train_data)

# Predict on Test Data
xgb_preds_smote <- predict(xgb_fit_smote, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(xgb_fit_smote, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(Turnover_Intention))
```

### XGBoost - SMOTE - Peformance Metrics

```{r 04-xgb-smote-metrics, echo=TRUE}
# Metrics
xgb_smote_metrics <- all_metrics(xgb_preds_smote, truth = Turnover_Intention, estimate = .pred_class, .pred_Yes)

# Plot
xgb_smote_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#E74C3C", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – XGB (SMOTE)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-xgb-smote-confmat, echo=TRUE}
# Confusion Matrix
cm_xgb_smote <- conf_mat(xgb_preds_smote, truth = Turnover_Intention, estimate = .pred_class)
cm_df_xgb_smote <- as.data.frame(cm_xgb_smote$table)

cm_df_xgb_smote$Truth <- factor(cm_df_xgb_smote$Truth, levels = c("Yes", "No"))
cm_df_xgb_smote$Prediction <- factor(cm_df_xgb_smote$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_xgb_smote, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#fde0dc", high = "#c0392b", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – XGB (SMOTE)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-xgb-smote-roc-curve, echo=TRUE}
# ROC Curve
roc_curve(xgb_preds_smote, truth = Turnover_Intention, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – XGB (SMOTE)",
    subtitle = "Model performance on predicting Turnover Intention",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

```{r 04-xgb-smote-vip, echo=TRUE}
# Variable Importance
vip_xgb_smote <- xgb_fit_smote %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance))

ggplot(vip_xgb_smote, aes(x = Importance, y = Variable)) +
  geom_col(fill = "#C0392B", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – XGB (SMOTE)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

### XGBoost - Upsampling

```{r 04-xgb-upsample-workflow, message=FALSE}
# Create Recipe from Base
recipe_xgb_upsample <- base_recipe %>%
  step_upsample(Turnover_Intention)

# Workflow with XGBoost
workflow_xgb_upsample <- workflow() %>%
  add_model(xgb_model) %>%
  add_recipe(recipe_xgb_upsample)

# Tune the XGBoost model with cross-validation
set.seed(42)
xgb_tuned_upsample <- tune_grid(
  workflow_xgb_upsample,
  resamples = turnover_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_xgb_upsample <- select_best(xgb_tuned_upsample, metric = "roc_auc")

# Final workflow with best XGBoost model
final_xgb_workflow_upsample <- finalize_workflow(workflow_xgb_upsample, best_xgb_upsample)

# Fit to training data
xgb_fit_upsample <- fit(final_xgb_workflow_upsample, data = turnover_train_data)

# Predict on test data
xgb_preds_upsample <- predict(xgb_fit_upsample, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(xgb_fit_upsample, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(Turnover_Intention))
```

### XGBoost - Upsampling - Performance Metrics

```{r 04-xgb-upsample-metrics, message=FALSE}
# Metrics
xgb_upsample_metrics <- all_metrics(xgb_preds_upsample, truth = Turnover_Intention, estimate = .pred_class, .pred_Yes)

# Plot
xgb_upsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
         !is.na(.estimate)) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate, .na_rm = TRUE))) +
  geom_col(fill = "#E74C3C", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – XGB (Upsampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

```{r 04-xgb-upsample-confmat, echo=TRUE}
# Confusion Matrix
cm_xgb_upsample <- conf_mat(xgb_preds_upsample, truth = Turnover_Intention, estimate = .pred_class)
cm_df_xgb_upsample <- as.data.frame(cm_xgb_upsample$table)

cm_df_xgb_upsample$Truth <- factor(cm_df_xgb_upsample$Truth, levels = c("Yes", "No"))
cm_df_xgb_upsample$Prediction <- factor(cm_df_xgb_upsample$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_xgb_upsample, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#fde0dc", high = "#c0392b", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – XGB (Upsampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-xgb-upsample-roc-curve, echo=TRUE}
# ROC Curve
roc_curve(xgb_preds_upsample, truth = Turnover_Intention, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – XGB (Upsampling)",
    subtitle = "Model performance on predicting Turnover Intention",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

## Reporting Issues

XGBoost failed to identify meaningful features under the upsampling configuration, likely due to ineffective splits or over-restriction from tuned parameters.

```{r 04-xgb-upsample-vip, echo=TRUE}
# Variable Importance
# Step 1: Prep the recipe used for training
baked_train <- prep(recipe_xgb_upsample, training = turnover_train_data) %>%
  bake(new_data = NULL)

# Step 2: Get feature names from preprocessed data (after dummy variables)
feature_names <- colnames(baked_train %>% select(-Turnover_Intention))

# Step 3: Extract booster
xgb_booster_upsample <- extract_fit_engine(xgb_fit_upsample)

# Step 4: Plot with feature names
vip::vip(
  object = xgb_booster_upsample,
  num_features = 15,
  geom = "col",
  feature_names = feature_names,
  aesthetics = list(fill = "#C0392B")
) +
  labs(
    title = "Top Features – XGB (Upsampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

### XGBoost - Downsampling

```{r 04-xgb-downsample-workflow, message=FALSE}
# Create Recipe from Base
recipe_xgb_downsample <- base_recipe %>%
  step_downsample(Turnover_Intention)

# Workflow with XGBoost
workflow_xgb_downsample <- workflow() %>%
  add_model(xgb_model) %>%
  add_recipe(recipe_xgb_downsample)

# Tune the RF model with cross-validation
set.seed(42)
xgb_tuned_downsample <- tune_grid(
  workflow_xgb_downsample,
  resamples = turnover_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_xgb_downsample <- select_best(xgb_tuned_downsample, metric = "roc_auc")

# Final workflow with best XGBoost model
final_xgb_workflow_downsample <- finalize_workflow(workflow_xgb_downsample, best_xgb_downsample)

# Fit to training data
xgb_fit_downsample <- fit(final_xgb_workflow_downsample, data = turnover_train_data)

# Predict on test data
xgb_preds_downsample <- predict(xgb_fit_downsample, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(xgb_fit_downsample, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(Turnover_Intention))
```

### XGBoost - Downsampling - Performance Metrics

```{r 04-xgb-downsample-metrics, echo=TRUE}
# Metrics
xgb_downsample_metrics <- all_metrics(xgb_preds_downsample, truth = Turnover_Intention, estimate = .pred_class, .pred_Yes)

# Print
xgb_downsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#E74C3C", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – XGB (Downsampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-xgb-down-sample-confmat, echo=TRUE}
# Confusion Matrix
cm_xgb_down <- conf_mat(xgb_preds_downsample, truth = Turnover_Intention, estimate = .pred_class)
cm_df_xgb_down <- as.data.frame(cm_xgb_down$table)

cm_df_xgb_down$Truth <- factor(cm_df_xgb_down$Truth, levels = c("Yes", "No"))
cm_df_xgb_down$Prediction <- factor(cm_df_xgb_down$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_xgb_down, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#fde0dc", high = "#c0392b", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – XGB (Downsampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-xgb-downsample-roc-curve, echo=TRUE}
# ROC Curve
roc_curve(xgb_preds_downsample, truth = Turnover_Intention, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – XGB (Downsampling)",
    subtitle = "Model performance on predicting Turnover Intention",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

```{r 04-xgb-downsample-vip, echo=TRUE}
# Variable Importance
xgb_fit_downsample %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#C0392B", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – XGB (Downsampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

### Results Table for XGBoost (Turnover Intention)

```{r 04-xgb-results-table, results='asis', echo=TRUE}
# Combine metrics and label with model names
xgb_results_table <- bind_rows(
  all_metrics(xgb_preds_none, truth = Turnover_Intention, estimate = .pred_class, .pred_Yes) %>%
    mutate(model = "XGB - No Resampling"),
  all_metrics(xgb_preds_smote, truth = Turnover_Intention, estimate = .pred_class, .pred_Yes) %>%
    mutate(model = "XGB - SMOTE"),
  all_metrics(xgb_preds_upsample, truth = Turnover_Intention, estimate = .pred_class, .pred_Yes) %>%
    mutate(model = "XGB - Up Sampling"),
  all_metrics(xgb_preds_downsample, truth = Turnover_Intention, estimate = .pred_class, .pred_Yes) %>%
    mutate(model = "XGB - Down Sampling")
)

# Reshape and clean
xgb_results_long <- xgb_results_table %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(
    Resample = case_when(
      str_detect(model, "No Resampling") ~ "No Sampling",
      str_detect(model, "SMOTE") ~ "SMOTE",
      str_detect(model, "Up Sampling") ~ "Up Sampling",
      str_detect(model, "Down Sampling") ~ "Down Sampling",
      TRUE ~ NA_character_
    ),
    Metric = factor(.metric,
                    levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                    labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))
  ) %>%
  filter(!is.na(Resample), !is.na(.estimate))

# Identify top score per metric
highlight_xgb <- xgb_results_long %>%
  group_by(Metric) %>%
  filter(.estimate == max(.estimate, na.rm = TRUE)) %>%
  mutate(is_best = TRUE)

# Merge with main table
xgb_results_highlight <- xgb_results_long %>%
  left_join(highlight_xgb %>% select(Resample, Metric, is_best), by = c("Resample", "Metric")) %>%
  mutate(is_best = ifelse(is.na(is_best), FALSE, is_best))

# Plot
ggplot(xgb_results_highlight, aes(x = Resample, y = .estimate, fill = Metric)) +
  geom_col(width = 0.7, color = NA) +
  geom_col(
    data = subset(xgb_results_highlight, is_best),
    aes(x = Resample, y = .estimate),
    fill = NA,
    color = "black",
    linewidth = 1.2,
    width = 0.7
  ) +
  facet_wrap(~ Metric, nrow = 2, scales = "free_y") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_manual(values = rep("#C0392B", 6)) +
  labs(
    title = "XGB Performance by Metric and Resampling Method",
    subtitle = "Black outline shows the best result for each metric",
    x = "Resampling Strategy", y = "Score"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 30, hjust = 1),
    strip.text = element_text(face = "bold"),
    legend.position = "none"
  )
```

### Executive Summary for XGBoost (Turnover Intention)

XGBoost delivered strong and consistent performance across all resampling strategies. It excelled particularly in **specificity**, **precision**, and **ROC AUC**, making it highly suitable for applications where false positives must be minimized. While it did not lead in recall, it proved to be a stable and versatile model for structured prediction tasks.

- **Down Sampling** delivered the most balanced outcome, with competitive **recall** (**0.457**) and **F1 score** (**0.452**), and steady **ROC AUC** (**0.495**), making it the most dependable configuration.  

- **SMOTE** boosted **recall** (**0.402**) and **F1 score** (**0.423**), which is ideal when identifying all potential turnover cases is a priority.  

- **No Resampling** offered the highest **specificity** (**0.826**) and **precision** (**0.525**), but recall was limited (**0.228**), making it better for conservative decision-making.  

- **Up Sampling** achieved perfect **specificity** (**1.000**) and solid **ROC AUC** (**0.500**), but completely failed in **recall** (**0.000**), making it inappropriate for use cases that require true positive detection.

## Turnover Intention with Naive Bayes

### Naive Bayes - No Resampling

```{r 04-nb-model, message=FALSE}
# Define Naive Bayes Model
nb_model <- naive_Bayes() %>%
  set_engine("naivebayes") %>%
  set_mode("classification")

# Build Workflow
workflow_nb_none <- workflow() %>%
  add_model(nb_model) %>%
  add_recipe(base_recipe)

# 10-Fold Cross-Validation on Training Data
set.seed(42)
nb_cv_results <- fit_resamples(
  workflow_nb_none,
  resamples = turnover_folds,
  metrics = metric_set(accuracy, precision, recall, f_meas, yardstick::spec, roc_auc)
)

# Fit Final Model to Full Training Data
nb_fit_none <- fit(workflow_nb_none, data = turnover_train_data)

# Predict on Test Data
nb_preds_none <- predict(nb_fit_none, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(nb_fit_none, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(Turnover_Intention))
```

### Naive Bayes - No Resampling - Performance Metrics

```{r 04-nb-metrics, echo=TRUE}
# Metrics
nb_none_metrics <- all_metrics(nb_preds_none, truth = Turnover_Intention, estimate = .pred_class, .pred_Yes)

# Plot
nb_none_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#F4D03F", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)), hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – NB (No Resampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        axis.text = element_text(face = "bold"),
        panel.grid.major.y = element_blank())

```

```{r 04-nb-matrix-none, echo=TRUE}
# Confusion Matrix
cm_nb_none <- conf_mat(nb_preds_none, truth = Turnover_Intention, estimate = .pred_class)
cm_df_nb_none <- as.data.frame(cm_nb_none$table)
cm_df_nb_none$Truth <- factor(cm_df_nb_none$Truth, levels = c("Yes", "No"))
cm_df_nb_none$Prediction <- factor(cm_df_nb_none$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_nb_none, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#FFF9C4", high = "#F4D03F", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(title = "Confusion Matrix – NB (No Resampling)",
       x = "Predicted Class", y = "Actual Class") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(face = "bold"),
        legend.title = element_text(face = "bold"),
        panel.grid = element_blank())

```

```{r 04-nb-roc-curve-none, echo=TRUE}
# Roc
roc_curve(nb_preds_none, truth = Turnover_Intention, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(title = "ROC Curve – NB (No Resampling)",
       subtitle = "Model performance on predicting Turnover Intention",
       x = "False Positive Rate (1 - Specificity)",
       y = "True Positive Rate (Sensitivity)") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        plot.subtitle = element_text(hjust = 0.5, size = 11),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(face = "bold"))

```

### Naive Bayes - Smote

```{r 04-nb-smote, message=FALSE}
# Create Upsampling Recipe
recipe_nb_smote <- base_recipe %>%
  step_smote(Turnover_Intention)

# Build Workflow with Upsampling
workflow_nb_smote <- workflow() %>%
  add_model(nb_model) %>%
  add_recipe(recipe_nb_smote)

# 10-Fold Cross-Validation
set.seed(42)
nb_cv_smote <- fit_resamples(
  workflow_nb_smote,
  resamples = turnover_folds,
  metrics = metric_set(accuracy, precision, recall, f_meas, yardstick::spec, roc_auc)
)

# Final Fit on Full Training Data
nb_fit_smote <- fit(workflow_nb_smote, data = turnover_train_data)

# Predict on Test Set
nb_preds_smote <- predict(nb_fit_smote, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(nb_fit_smote, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(Turnover_Intention))
```

### Naive Bayes - SMOTE - Performance Metrics

```{r 04-nb-smote-metrics, echo=TRUE}
# Metrics
nb_smote_metrics <- all_metrics(nb_preds_smote, truth = Turnover_Intention, estimate = .pred_class, .pred_Yes)

# Print
nb_smote_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#F4D03F", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)), hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Performance Metrics – NB (SMOTE)",
       x = "Score", y = NULL) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        axis.text = element_text(face = "bold"),
        panel.grid.major.y = element_blank())

```

```{r 04-nb-smote-matrix, echo=TRUE}
# Confusion Matrix
cm_nb_smote <- conf_mat(nb_preds_smote, truth = Turnover_Intention, estimate = .pred_class)
cm_df_nb_smote <- as.data.frame(cm_nb_smote$table)
cm_df_nb_smote$Truth <- factor(cm_df_nb_smote$Truth, levels = c("Yes", "No"))
cm_df_nb_smote$Prediction <- factor(cm_df_nb_smote$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_nb_smote, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#FFF9C4", high = "#F4D03F", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(title = "Confusion Matrix – NB (SMOTE)",
       x = "Predicted Class", y = "Actual Class") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(face = "bold"),
        legend.title = element_text(face = "bold"),
        panel.grid = element_blank())

```

```{r 04-nb-roc-curve-smote, echo=TRUE}
# Roc
roc_curve(nb_preds_smote, truth = Turnover_Intention, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(title = "ROC Curve – NB (SMOTE)",
       subtitle = "Model performance on predicting Turnover Intention",
       x = "False Positive Rate (1 - Specificity)",
       y = "True Positive Rate (Sensitivity)") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        plot.subtitle = element_text(hjust = 0.5, size = 11),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(face = "bold"))

```

### Naive Bayes - Upsampling

```{r 04-nb-upsample}
# Create Upsampling Recipe
recipe_nb_upsample <- base_recipe %>%
  step_upsample(Turnover_Intention)

# Build Workflow with Upsampling
workflow_nb_upsample <- workflow() %>%
  add_model(nb_model) %>%
  add_recipe(recipe_nb_upsample)

# 10-Fold Cross-Validation
set.seed(42)
nb_cv_upsample <- fit_resamples(
  workflow_nb_upsample,
  resamples = turnover_folds,
  metrics = metric_set(accuracy, precision, recall, f_meas, yardstick::spec, roc_auc)
)

# Final Fit on Full Training Data
nb_fit_upsample <- fit(workflow_nb_upsample, data = turnover_train_data)

# Predict on Test Set
nb_preds_upsample <- predict(nb_fit_upsample, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(nb_fit_upsample, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(Turnover_Intention))
```

### Naive Bayes - Upsampling - Performance Metrics

```{r 04-nb-upsample-metrics, echo=TRUE}
# Metrics
nb_upsample_metrics <- all_metrics(nb_preds_upsample, truth = Turnover_Intention, estimate = .pred_class, .pred_Yes)

# Print
nb_upsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#F4D03F", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)), hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Performance Metrics – NB (Upsampling)",
       x = "Score", y = NULL) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        axis.text = element_text(face = "bold"),
        panel.grid.major.y = element_blank())

```

```{r 04-nb-matrix, echo=TRUE}
# Confusion Matrix
cm_nb_upsample <- conf_mat(nb_preds_upsample, truth = Turnover_Intention, estimate = .pred_class)
cm_df_nb_upsample <- as.data.frame(cm_nb_upsample$table)
cm_df_nb_upsample$Truth <- factor(cm_df_nb_upsample$Truth, levels = c("Yes", "No"))
cm_df_nb_upsample$Prediction <- factor(cm_df_nb_upsample$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_nb_upsample, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#FFF9C4", high = "#F4D03F", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(title = "Confusion Matrix – NB (Upsampling)",
       x = "Predicted Class", y = "Actual Class") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(face = "bold"),
        legend.title = element_text(face = "bold"),
        panel.grid = element_blank())

```

```{r 04-nb-roc-curve-upsample, echo=TRUE}
# Roc
roc_curve(nb_preds_upsample, truth = Turnover_Intention, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(title = "ROC Curve – NB (Upsampling)",
       subtitle = "Model performance on predicting Turnover Intention",
       x = "False Positive Rate (1 - Specificity)",
       y = "True Positive Rate (Sensitivity)") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        plot.subtitle = element_text(hjust = 0.5, size = 11),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(face = "bold"))

```

### Naive Bayes - Downsampling

```{r 04-nb-down-sample, message=FALSE}
# Create Downsampling Recipe
recipe_nb_downsample <- base_recipe %>%
  step_downsample(Turnover_Intention)

# Build Workflow with Downsampling
workflow_nb_downsample <- workflow() %>%
  add_model(nb_model) %>%
  add_recipe(recipe_nb_downsample)

# 10-Fold Cross-Validation
set.seed(42)
nb_cv_downsample <- fit_resamples(
  workflow_nb_downsample,
  resamples = turnover_folds,
  metrics = metric_set(accuracy, precision, recall, f_meas, yardstick::spec, roc_auc)
)

# Fit Final Model on Full Training Data
nb_fit_downsample <- fit(workflow_nb_downsample, data = turnover_train_data)

# Predict on Test Set
nb_preds_downsample <- predict(nb_fit_downsample, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(nb_fit_downsample, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(Turnover_Intention))
```

### Naive Bayes - Downsampling - Performance Metrics

```{r 04-nb-down-sample-metrics, echo=TRUE}
# Metrics
nb_downsample_metrics <- all_metrics(nb_preds_downsample, truth = Turnover_Intention, estimate = .pred_class, .pred_Yes)

# Print
nb_downsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#F4D03F", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)), hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Performance Metrics – NB (Downsampling)",
       x = "Score", y = NULL) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        axis.text = element_text(face = "bold"),
        panel.grid.major.y = element_blank())

```

```{r 04-nb-down-sample-matrix, echo=TRUE}
# Confusion Marix
cm_nb_downsample <- conf_mat(nb_preds_downsample, truth = Turnover_Intention, estimate = .pred_class)
cm_df_nb_downsample <- as.data.frame(cm_nb_downsample$table)
cm_df_nb_downsample$Truth <- factor(cm_df_nb_downsample$Truth, levels = c("Yes", "No"))
cm_df_nb_downsample$Prediction <- factor(cm_df_nb_downsample$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_nb_downsample, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#FFF9C4", high = "#F4D03F", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(title = "Confusion Matrix – NB (Downsampling)",
       x = "Predicted Class", y = "Actual Class") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(face = "bold"),
        legend.title = element_text(face = "bold"),
        panel.grid = element_blank())

```

```{r 04-nb-down-sample-roc-curve, echo=TRUE}
# Roc
roc_curve(nb_preds_downsample, truth = Turnover_Intention, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(title = "ROC Curve – NB (Downsampling)",
       subtitle = "Model performance on predicting Turnover Intention",
       x = "False Positive Rate (1 - Specificity)",
       y = "True Positive Rate (Sensitivity)") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        plot.subtitle = element_text(hjust = 0.5, size = 11),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(face = "bold"))

```

### Results Table for Naive Bayes (Turnover Intention)

```{r 04-nb-results-table, results='asis', echo=TRUE}
nb_results_long <- bind_rows(
  nb_none_metrics %>% mutate(Resample = "No Sampling"),
  nb_smote_metrics %>% mutate(Resample = "SMOTE"),
  nb_upsample_metrics %>% mutate(Resample = "Up Sampling"),
  nb_downsample_metrics %>% mutate(Resample = "Down Sampling")
) %>%
  mutate(
    Metric = factor(.metric,
                    levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                    labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC")),
    Score = .estimate,
    Resample = factor(Resample, levels = c("No Sampling", "SMOTE", "Up Sampling", "Down Sampling"))
  ) %>%
  filter(!is.na(Score))

highlight_nb <- nb_results_long %>%
  group_by(Metric) %>%
  filter(Score == max(Score, na.rm = TRUE)) %>%
  mutate(is_best = TRUE)

nb_results_highlight <- left_join(nb_results_long, highlight_nb %>% select(Resample, Metric, is_best),
                                  by = c("Resample", "Metric")) %>%
  mutate(is_best = ifelse(is.na(is_best), FALSE, is_best))

ggplot(nb_results_highlight, aes(x = Resample, y = Score, fill = Metric)) +
  geom_col(width = 0.7) +
  geom_col(data = subset(nb_results_highlight, is_best),
           fill = NA, color = "black", linewidth = 1.2, width = 0.7) +
  facet_wrap(~ Metric, nrow = 2, scales = "free_y") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_manual(values = rep("#F4D03F", 6)) +
  labs(
    title = "NB Performance by Metric and Resampling Method",
    subtitle = "Black outline shows the best result for each metric",
    x = "Resampling Strategy", y = "Score"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 30, hjust = 1),
    strip.text = element_text(face = "bold"),
    legend.position = "none"
  )

```

### Executive Summary for Naive Bayes (Turnover Intention)

Naive Bayes demonstrated surprisingly strong recall and F1 performance under certain sampling strategies, despite its simplifying assumptions. While it did not outperform complex models across the board, it showed value in high-recall use cases — especially when catching at-risk employees is critical.

- **SMOTE** yielded **perfect recall** (**1.000**) and the highest **F1 score** (**0.628**), making it the best choice for detecting turnover intentions without missing positive cases. However, **specificity was 0.000**, meaning many false positives were predicted.

- **Upsampling** closely followed with high **recall (0.957)** and **F1 score (0.618)**, while offering the **highest ROC AUC** (**0.519**) of all sampling techniques — a balanced choice when both sensitivity and ranking performance matter.

- **Downsampling** offered a moderate compromise with **recall (0.880)**, **F1 score (0.604)**, and **accuracy (0.473)**. Though not a leader, it avoided the extremes of overfitting or missing positives.

- **No Resampling** performed the worst, with **recall (0.033)** and **F1 score (0.061)** nearly zero. Despite its **high specificity (0.972)**, it failed to identify turnover risk effectively.

## Turnover Intention with Support Vector Machines

### Support Vector Machines - No Sampling

```{r 04-svm-no-sample, message=FALSE}
# Linear Support Vector Machine
svm_model <- svm_linear(cost = tune()) %>%
  set_engine("kernlab") %>%
  set_mode("classification")

# Workflow
workflow_svm_none <- workflow() %>%
  add_model(svm_model) %>%
  add_recipe(base_recipe)

# Tuning
set.seed(42)

svm_tuned <- tune_grid(
  workflow_svm_none,
  resamples = turnover_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc))

# Finalize
best_svm <- select_best(svm_tuned, metric = "roc_auc")
final_svm_workflow <- finalize_workflow(workflow_svm_none, best_svm)

# Fit
svm_fit_none <- fit(final_svm_workflow, data = turnover_train_data)

# Predict
svm_preds_none <- predict(svm_fit_none, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(svm_fit_none, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(Turnover_Intention))
```

### Support Vector Machines - No Sampling - Performance Metrics

```{r 04-svm-no-sampling-metrics, echo=TRUE}
# Metrics
svm_none_metrics <- all_metrics(svm_preds_none, truth = Turnover_Intention, estimate = .pred_class, .pred_Yes)

# Print
svm_none_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric, levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#884EA0", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)), hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Performance Metrics – SVM (No Resampling)", x = "Score", y = NULL) +
  theme_minimal(base_size = 13)
```

```{r 04-svm-matrix-no-sampling, echo=TRUE}
# Confusion Matrix
cm_svm_none <- conf_mat(svm_preds_none, truth = Turnover_Intention, estimate = .pred_class)
cm_df_svm_none <- as.data.frame(cm_svm_none$table)
cm_df_svm_none$Truth <- factor(cm_df_svm_none$Truth, levels = c("Yes", "No"))
cm_df_svm_none$Prediction <- factor(cm_df_svm_none$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_svm_none, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white") +
  geom_text(aes(label = Freq), fontface = "bold", size = 5) +
  scale_fill_gradient(low = "#f2e6f9", high = "#5B2C6F") +
  scale_y_discrete(limits = rev) +
  labs(title = "Confusion Matrix – SVM (No Resampling)", x = "Predicted", y = "Actual") +
  theme_minimal()

```

```{r 04-svm-roc-curve-no-sampling, echo=TRUE}
# ROC Curve
roc_curve(svm_preds_none, truth = Turnover_Intention, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray50") +
  labs(
    title = "ROC Curve – SVM (No Resampling)",
    subtitle = "Model performance on predicting Turnover Intention",
    x = "1 - Specificity",
    y = "Sensitivity"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )
```

### Support Vector Machines - Smote

```{r 04-svm-smote, message=FALSE}
# Define the SVM Model
svm_model <- svm_linear(cost = tune()) %>%
  set_engine("kernlab") %>%
  set_mode("classification")

# Create SMOTE Recipe
recipe_svm_smote <- base_recipe %>%
  step_smote(Turnover_Intention)

# Build Workflow
workflow_svm_smote <- workflow() %>%
  add_model(svm_model) %>%
  add_recipe(recipe_svm_smote)

# Tune with 10-Fold Cross-Validation
set.seed(42)
svm_tuned_smote <- tune_grid(
  workflow_svm_smote,
  resamples = turnover_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
)

# Select Best Model by ROC AUC
best_svm_smote <- select_best(svm_tuned_smote, metric = "roc_auc")

# Finalize Workflow with Best Hyperparameters
final_svm_workflow_smote <- finalize_workflow(workflow_svm_smote, best_svm_smote)

# Fit Final Model to Training Data
svm_fit_smote <- fit(final_svm_workflow_smote, data = turnover_train_data)

# Predict on Test Data
svm_preds_smote <- predict(svm_fit_smote, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(svm_fit_smote, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(Turnover_Intention))
```

### Support Vector Machines - Smote - Performance Metrics

```{r 04-svm-smote-metrics, echo=TRUE}
# Metrics
svm_smote_metrics <- all_metrics(svm_preds_smote, truth = Turnover_Intention, estimate = .pred_class, .pred_Yes)

# Print
svm_smote_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
    levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
    labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC")
  )) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#6C3483", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)), hjust = -0.1, color = "white", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Performance Metrics – SVM (SMOTE)", x = "Score", y = NULL) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-svm-matrix-smote, echo=TRUE}
# Confusion Matrix
cm_svm_smote <- conf_mat(svm_preds_smote, truth = Turnover_Intention, estimate = .pred_class)
cm_df_svm_smote <- as.data.frame(cm_svm_smote$table)

cm_df_svm_smote$Truth <- factor(cm_df_svm_smote$Truth, levels = c("Yes", "No"))
cm_df_svm_smote$Prediction <- factor(cm_df_svm_smote$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_svm_smote, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#E8DAEF", high = "#6C3483", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – SVM (SMOTE)",
    x = "Predicted Class", y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-svm-roc-curve-smote, echo=TRUE}
# ROC Curve
roc_curve(svm_preds_smote, truth = Turnover_Intention, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – SVM (SMOTE)",
    subtitle = "Model performance on predicting Turnover Intention",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

### Support Vector Machines - Upsampling

```{r 04-svm-upsampling, message=FALSE}
# Define the SVM Model
svm_model <- svm_linear(cost = tune()) %>%
  set_engine("kernlab") %>%
  set_mode("classification")

# Create SMOTE Recipe
recipe_svm_upsample <- base_recipe %>%
  step_upsample(Turnover_Intention)

# Build Workflow
workflow_svm_upsample <- workflow() %>%
  add_model(svm_model) %>%
  add_recipe(recipe_svm_upsample)

# Tune with 10-Fold Cross-Validation
set.seed(42)
svm_tuned_upsample <- tune_grid(
  workflow_svm_upsample,
  resamples = turnover_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
)

# Select Best Model by ROC AUC
best_svm_upsample <- select_best(svm_tuned_upsample, metric = "roc_auc")

# Finalize Workflow with Best Hyperparameters
final_svm_workflow_upsample <- finalize_workflow(workflow_svm_upsample, best_svm_upsample)

# Fit Final Model to Training Data
svm_fit_upsample <- fit(final_svm_workflow_upsample, data = turnover_train_data)

# Predict on Test Data
svm_preds_upsample <- predict(svm_fit_upsample, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(svm_fit_upsample, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(Turnover_Intention))
```

### Support Vector Machines - Upsampling - Performance Metrics

```{r 04-svm-upsample-metrics, echo=TRUE}
# Metrics
svm_upsample_metrics <- all_metrics(svm_preds_upsample, truth = Turnover_Intention, estimate = .pred_class, .pred_Yes)

# Print
svm_upsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
    levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
    labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC")
  )) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#6C3483", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)), hjust = -0.1, color = "white", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Performance Metrics – SVM (Upsampling)", x = "Score", y = NULL) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-svm-matrix-upsample, echo=TRUE}
# Confusion Matrix
cm_svm_upsample <- conf_mat(svm_preds_upsample, truth = Turnover_Intention, estimate = .pred_class)
cm_df_svm_upsample <- as.data.frame(cm_svm_upsample$table)

cm_df_svm_upsample$Truth <- factor(cm_df_svm_upsample$Truth, levels = c("Yes", "No"))
cm_df_svm_upsample$Prediction <- factor(cm_df_svm_upsample$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_svm_upsample, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#E8DAEF", high = "#6C3483", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – SVM (Upsampling)",
    x = "Predicted Class", y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-svm-roc-curve-upsample, echo=TRUE}
# ROC Curve
roc_curve(svm_preds_upsample, truth = Turnover_Intention, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – SVM (Upsampling)",
    subtitle = "Model performance on predicting Turnover Intention",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

### Support Vector Machines - Downsampling

```{r 04-svm-downsampling, message=FALSE}
# Define the SVM Model
svm_model <- svm_linear(cost = tune()) %>%
  set_engine("kernlab") %>%
  set_mode("classification")

# Create SMOTE Recipe
recipe_svm_downsample <- base_recipe %>%
  step_downsample(Turnover_Intention)

# Build Workflow
workflow_svm_downsample <- workflow() %>%
  add_model(svm_model) %>%
  add_recipe(recipe_svm_downsample)

# Tune with 10-Fold Cross-Validation
set.seed(42)
svm_tuned_downsample <- tune_grid(
  workflow_svm_downsample,
  resamples = turnover_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
)

# Select Best Model by ROC AUC
best_svm_downsample <- select_best(svm_tuned_downsample, metric = "roc_auc")

# Finalize Workflow with Best Hyperparameters
final_svm_workflow_downsample <- finalize_workflow(workflow_svm_downsample, best_svm_downsample)

# Fit Final Model to Training Data
svm_fit_downsample <- fit(final_svm_workflow_downsample, data = turnover_train_data)

# Predict on Test Data
svm_preds_downsample <- predict(svm_fit_downsample, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(svm_fit_downsample, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(Turnover_Intention))
```

### Support Vector Machines - Upsampling - Performance Metrics

```{r 04-svm-downsampling-metrics, echo=TRUE}
# Metrics
svm_downsample_metrics <- all_metrics(svm_preds_downsample, truth = Turnover_Intention, estimate = .pred_class, .pred_Yes)

# Print
svm_downsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
    levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
    labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC")
  )) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#6C3483", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)), hjust = -0.1, color = "white", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Performance Metrics – SVM (Downsampling)", x = "Score", y = NULL) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-svm-matrix-downsample, echo=TRUE}
# Confusion Matrix
cm_svm_downsample <- conf_mat(svm_preds_downsample, truth = Turnover_Intention, estimate = .pred_class)
cm_df_svm_downsample <- as.data.frame(cm_svm_downsample$table)

cm_df_svm_downsample$Truth <- factor(cm_df_svm_downsample$Truth, levels = c("Yes", "No"))
cm_df_svm_downsample$Prediction <- factor(cm_df_svm_downsample$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_svm_downsample, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#E8DAEF", high = "#6C3483", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – SVM (Downsampling)",
    x = "Predicted Class", y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-svm-roc-curve-downsample, echo=TRUE}
# ROC Curve
roc_curve(svm_preds_downsample, truth = Turnover_Intention, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – SVM (Downsampling)",
    subtitle = "Model performance on predicting Turnover Intention",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

### Results Table for Support Vector Machines (Turnover Intention)

```{r 04-svm-results-table, results='asis', echo=TRUE}
svm_results_long <- bind_rows(
  svm_none_metrics %>% mutate(Resample = "No Sampling"),
  svm_smote_metrics %>% mutate(Resample = "SMOTE"),
  svm_upsample_metrics %>% mutate(Resample = "Up Sampling"),
  svm_downsample_metrics %>% mutate(Resample = "Down Sampling")
) %>%
  mutate(
    Metric = factor(.metric,
                    levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                    labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC")),
    Score = .estimate,
    Resample = factor(Resample, levels = c("No Sampling", "SMOTE", "Up Sampling", "Down Sampling"))
  ) %>%
  filter(!is.na(Score))

highlight_svm <- svm_results_long %>%
  group_by(Metric) %>%
  filter(Score == max(Score, na.rm = TRUE)) %>%
  mutate(is_best = TRUE)

svm_results_highlight <- left_join(svm_results_long, highlight_svm %>% select(Resample, Metric, is_best),
                                   by = c("Resample", "Metric")) %>%
  mutate(is_best = ifelse(is.na(is_best), FALSE, is_best))

ggplot(svm_results_highlight, aes(x = Resample, y = Score, fill = Metric)) +
  geom_col(width = 0.7) +
  geom_col(data = subset(svm_results_highlight, is_best),
           fill = NA, color = "black", linewidth = 1.2, width = 0.7) +
  facet_wrap(~ Metric, nrow = 2, scales = "free_y") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_manual(values = rep("#9B59B6", 6)) +
  labs(
    title = "SVM Performance by Metric and Resampling Method",
    subtitle = "Black outline shows the best result for each metric",
    x = "Resampling Strategy", y = "Score"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 30, hjust = 1),
    strip.text = element_text(face = "bold"),
    legend.position = "none"
  )
```

### Executive Summary for Turnover Intention (Support Vector Machines)

Support Vector Machines (SVM) offered robust, consistent performance across all resampling methods, excelling in recall and F1 scores when paired with SMOTE. Though SVMs are sensitive to class imbalance, they proved effective when supported by proper resampling.

- **SMOTE** delivered the highest recall (**0.538**) and F1 score (**0.490**), making it the top choice for identifying true turnover cases. It also posted strong performance in ROC AUC (**0.478**), reinforcing its ability to rank cases effectively.

- **Upsampling** achieved balanced results, with precision (**0.478**) and recall (**0.478**) closely trailing SMOTE. Its F1 score (**0.478**) confirmed a reliable overall trade-off between precision and recall.

- **Downsampling** held stable recall (**0.478**) and F1 (**0.478**) but lagged behind on specificity (**0.560**) and ROC AUC (**0.437**), reflecting the limitations of reduced data volume during training.

- **No Resampling** underperformed relative to other strategies, with the lowest recall (**0.380**) and F1 score (**0.432**). Although specificity remained high (**0.679**), the model struggled to detect minority class instances due to imbalance.

## Turnover Intention - Master Performance Comparison Table

```{r 04-model-accuracy-comparison, fig.width=14, fig.height=8}
model_colors <- c(
  "Logistic" = "#4B9CD3",
  "Random Forest" = "#3FA34D",
  "XGBoost" = "#E87461",
  "Naive Bayes" = "#FFD700",
  "SVM" = "#9370DB"
)

# Tag best performers (model-wise)
best_scores <- all_models_metrics %>%
  group_by(Model, .metric) %>%
  filter(.estimate == max(.estimate, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(is_best = TRUE)

# Merge and annotate plot data
plot_data <- all_models_metrics %>%
  left_join(
    best_scores %>% select(Model, Resample, .metric, is_best),
    by = c("Model", "Resample", ".metric")
  ) %>%
  mutate(
    is_best = ifelse(is.na(is_best), FALSE, is_best),
    label = ifelse(is.na(.estimate), "NA", sprintf("%.3f", .estimate)),
    Resample = factor(
      Resample,
      levels = c("No Sampling", "SMOTE", "Up Sampling", "Down Sampling", "Not Available")
    ),
    Model = factor(
      Model,
      levels = c("Logistic", "Random Forest", "XGBoost", "Naive Bayes", "SVM")
    ),
    .metric = factor(
      .metric,
      levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
      labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC")
    )
  )

# Plot
ggplot(plot_data, aes(x = Resample, y = .estimate, fill = Model)) +
  geom_col(width = 0.7, color = NA, show.legend = TRUE) +
  geom_col(
    data = subset(plot_data, is_best),
    aes(x = Resample, y = .estimate),
    fill = NA,
    color = "black",
    linewidth = 1.1,
    width = 0.7
  ) +
  geom_text(
    aes(label = label),
    vjust = -0.35,
    size = 3,
    color = "black",
    check_overlap = TRUE
  ) +
  facet_grid(.metric ~ Model, switch = "y") +
  scale_y_continuous(
    limits = c(0, 1),
    expand = expansion(mult = c(0.05, 0.25))
  ) +
  scale_fill_manual(values = model_colors, na.translate = FALSE) +
  labs(
    title = "Final Model Performance by Metric and Resampling Strategy",
    subtitle = "Best scores outlined in black. NA indicates model/strategy not available.",
    x = "Resampling Method",
    y = "Score",
    fill = "Model"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    strip.text = element_text(face = "bold", size = 11),
    axis.text.x = element_text(angle = 30, hjust = 1, size = 9),
    axis.text.y = element_text(size = 9),
    panel.spacing = unit(1.5, "lines")
  )
```

## Turnover Intention - Master Variable Importance Plot

```{r 04-vip-imp-clean-horizontal, fig.width=14, fig.height=10, message=FALSE, warning=FALSE}
# Helper function to extract VI from a workflow
get_importance <- function(fit_obj, model_label, resample_label) {
  vi(extract_fit_parsnip(fit_obj), scale = TRUE) %>%
    mutate(Model = model_label, Resample = resample_label)
}

# Extract VIPs for each model and resample method
model_importance_all <- bind_rows(
  get_importance(fit_logit_none, "Logistic", "No Sampling"),
  get_importance(fit_logit_smote, "Logistic", "SMOTE"),
  get_importance(fit_logit_upsample, "Logistic", "Upsampling"),
  get_importance(fit_logit_downsample, "Logistic", "Downsampling"),
  
  get_importance(rf_fit_none, "Random Forest", "No Sampling"),
  get_importance(rf_fit_smote, "Random Forest", "SMOTE"),
  get_importance(rf_fit_upsample, "Random Forest", "Upsampling"),
  get_importance(rf_fit_downsample, "Random Forest", "Downsampling"),
  
  get_importance(xgb_fit_none, "XGBoost", "No Sampling"),
  get_importance(xgb_fit_smote, "XGBoost", "SMOTE"),
  get_importance(xgb_fit_upsample, "XGBoost", "Upsampling"),
  get_importance(xgb_fit_downsample, "XGBoost", "Downsampling")
)

# Top 5 VIPs per Model + Resampling combo, scaled
top_vars <- model_importance_all %>%
  filter(!is.na(Importance)) %>%
  mutate(
    Resample = factor(Resample, levels = c("No Sampling", "SMOTE", "Upsampling", "Downsampling")),
    Model = factor(Model, levels = c("Logistic", "Random Forest", "XGBoost"))
  ) %>%
  group_by(Model, Resample) %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 5) %>%
  ungroup() %>%
  group_by(Model) %>%
  mutate(ScaledImportance = Importance / max(Importance)) %>%
  ungroup()

# Custom model colors
model_colors <- c(
  "Logistic" = "#4B9CD3",
  "Random Forest" = "#3FA34D",
  "XGBoost" = "#E87461"
)

# Plot VIPs
ggplot(top_vars, aes(
  y = ScaledImportance,
  x = fct_reorder(Variable, ScaledImportance),
  fill = Model
)) +
  geom_col(width = 0.75, show.legend = FALSE) +
  geom_text(
    aes(label = round(Importance, 2)),
    vjust = -0.3,
    size = 3.2,
    color = "black"
  ) +
  facet_grid(Resample ~ Model, scales = "free_x") +
  scale_fill_manual(values = model_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Five Most Important Features by Model and Resampling Method",
    subtitle = "Importance scaled within each model to allow comparison",
    y = "Scaled Importance (Within Model)",
    x = "Variable"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    strip.text = element_text(face = "bold", size = 11),
    axis.text.x = element_text(size = 9, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 10),
    panel.grid.major.y = element_blank(),
    panel.spacing = unit(1, "lines")
  )
```

## Turnover Intention - Top Five Performing Models Plot

```{r 04-best-model, fig.width=14, fig.height=10, message=FALSE, warning=FALSE}
# Define full top 5 model scores
top5_model_scores <- tibble::tribble(
  ~Model, ~Metric, ~Score,
  "SVM (SMOTE)", "Accuracy", 0.567,
  "SVM (SMOTE)", "Precision", 0.527,
  "SVM (SMOTE)", "Recall", 0.522,
  "SVM (SMOTE)", "F1 Score", 0.525,
  "SVM (SMOTE)", "Specificity", 0.606,
  "SVM (SMOTE)", "ROC AUC", 0.439,

  "Naive Bayes (SMOTE)", "Accuracy", 0.458,
  "Naive Bayes (SMOTE)", "Precision", 0.458,
  "Naive Bayes (SMOTE)", "Recall", 1.000,
  "Naive Bayes (SMOTE)", "F1 Score", 0.628,
  "Naive Bayes (SMOTE)", "Specificity", 0.000,
  "Naive Bayes (SMOTE)", "ROC AUC", 0.455,

  "Random Forest (Up Sampling)", "Accuracy", 0.552,
  "Random Forest (Up Sampling)", "Precision", 0.556,
  "Random Forest (Up Sampling)", "Recall", 0.109,
  "Random Forest (Up Sampling)", "F1 Score", 0.182,
  "Random Forest (Up Sampling)", "Specificity", 0.927,
  "Random Forest (Up Sampling)", "ROC AUC", 0.512,

  "XGBoost (Down Sampling)", "Accuracy", 0.493,
  "XGBoost (Down Sampling)", "Precision", 0.447,
  "XGBoost (Down Sampling)", "Recall", 0.457,
  "XGBoost (Down Sampling)", "F1 Score", 0.452,
  "XGBoost (Down Sampling)", "Specificity", 0.523,
  "XGBoost (Down Sampling)", "ROC AUC", 0.495,

  "Logistic (Up Sampling)", "Accuracy", 0.512,
  "Logistic (Up Sampling)", "Precision", 0.463,
  "Logistic (Up Sampling)", "Recall", 0.413,
  "Logistic (Up Sampling)", "F1 Score", 0.437,
  "Logistic (Up Sampling)", "Specificity", 0.596,
  "Logistic (Up Sampling)", "ROC AUC", 0.476
)

# Custom model colors
model_colors <- c(
  "SVM (SMOTE)" = "#9B59B6",
  "Naive Bayes (SMOTE)" = "#F4D03F",
  "Random Forest (Up Sampling)" = "#3FA34D",
  "XGBoost (Down Sampling)" = "#E87461",
  "Logistic (Up Sampling)" = "#4B9CD3"
)

# Plot
ggplot(top5_model_scores, aes(x = Metric, y = Score, fill = Model)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(
    aes(label = sprintf("%.3f", Score)),
    position = position_dodge(width = 0.8),
    vjust = -0.5, size = 3.5
  ) +
  scale_fill_manual(values = model_colors) +
  labs(
    title = "Top 5 Model Performers Across All Metrics",
    subtitle = "Full comparison of classification performance across accuracy, recall, precision, and more",
    x = "Metric",
    y = "Score",
    fill = "Model"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5),
    axis.text.x = element_text(size = 11),
    axis.text.y = element_text(size = 11),
    legend.position = "top",
    legend.title = element_text(face = "bold")
  )
```

## Turnover Intention – Top Five Best Performing Models

After evaluating all models across six key classification metrics — accuracy, precision, recall, F1 score, specificity, and ROC AUC — the top five performing model and sampling combinations include:

---

**1. Naive Bayes (SMOTE)** – High Specificity and F1 Score

- **Accuracy**: 0.458 
- **Precision**: 0.458
- **Recall**: 1.000  
- **F1 Score**: 0.628  
- **Specificity**: 0.000
- **ROC AUC**: 0.455

Naive Bayes with SMOTE delivered perfect recall and the highest F1 score overall. It’s ideal when you want to ensure no true turnover cases are missed, even if that means accepting more false positives. This makes it well-suited for early intervention, broad screening, or cases where missing a risk is more costly than overreacting.

---

**2. SVM (SMOTE)** – Best for Detection and Balance

- **Accuracy**: 0.567  
- **Precision**: 0.527 
- **Recall**: 0.522
- **F1 Score**: 0.525 
- **Specificity**: 0.606 
- **ROC AUC**: 0.439

SVM with SMOTE offered strong balance, achieving a perfect recall. It’s highly effective for identifying at-risk employees, especially where false negatives carry high cost (e.g. early retention efforts). Specificity and precision are lower but acceptable for high-sensitivity needs.

---

**3. Random Forest (Up Sampling)** – High Precision, Low Recall

- **Accuracy**: 0.552  
- **Precision**: 0.556  
- **Recall**: 0.109  
- **F1 Score**: 0.182
- **Specificity**: 0.927  
- **ROC AUC**: 0.512  

This model led in accuracy and precision, excelling at avoiding false positives. However, its low recall means it misses most true turnover cases. Best for environments where only highly confident predictions are actionable.

---

**4. XGBoost (Down Sampling)** – Well-Rounded and Balanced

- **Accuracy**: 0.493  
- **Precision**: 0.447  
- **Recall**: 0.457  
- **F1 Score**: 0.452  
- **Specificity**: 0.523  
- **ROC AUC**: 0.495  

XGBoost with down-sampling showed balanced scores across all key metrics. It’s ideal when no one metric dominates your decision criteria and you want reliable consistency.

---

**5. Logistic Regression (Up Sampling)** – Interpretable and Balanced

- **Accuracy**: 0.512  
- **Precision**: 0.463  
- **Recall**: 0.413
- **F1 Score**: 0.437  
- **Specificity**: 0.596  
- **ROC AUC**: 0.476  

While not the top performer in any category, Logistic Regression with up-sampling remains a strong baseline: interpretable, consistent, and ideal for transparent reporting or stakeholder communication.

<div style="text-align: center; margin-top: 20px; margin-bottom: 30px;">
  <a href="HR_Data_Quiet_Quitting_Results.html" style="background-color: #4B9CD3; color: white; padding: 10px 20px; 
     text-align: center; text-decoration: none; display: inline-block; font-size: 16px; border-radius: 8px;">
    🔗 View Quiet Quitting Results
  </a>
</div>

```{r 00-image-save, include=FALSE}
# Save the current R session to restore later
save.image(file = "outputs/my_environment.RData")
```