---
title: "Workforce Analytics Using Synthetic HR Data: Modeling and Predicting Turnover Intention"
author: "Kevin Allen Cauley"
date: "`r format(Sys.Date(), '%B %d, %Y')`"

output:
  html_document:
    always_allow_html: true
    self_contained: false
    theme: flatly
    highlight: null
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 3
    df_print: paged
    css: "assets/styles.css"
    output_file: HR_Data_Turnover_Intention_Results.html
    output_dir: ../docs

number_sections: true
---

```{r 00-setup, include=FALSE}
# Function Helpers
source(here::here("analysis", "helper.R"))

# Standard Metric Plot Function
plot_model_metrics <- function(df, title_text, filename) {
  
  df <- df %>%
    mutate(
      Resample = recode(Resample,
        "No Sampling" = "None",
        "Up Sampling" = "Up",
        "Down Sampling" = "Down"
      )
    ) %>%
    group_by(.metric) %>%
    mutate(best = .estimate == max(.estimate)) %>%
    ungroup()
  
  p <- df %>%
    filter(.metric %in% c("accuracy","precision","recall","f_meas","spec","roc_auc")) %>%
    mutate(.metric = factor(
      .metric,
      levels = c("accuracy","precision","recall","f_meas","spec","roc_auc"),
      labels = c("Accuracy","Precision","Recall","F1 Score","Specificity","ROC AUC")
    )) %>%
    ggplot(aes(x = Resample, y = .estimate, fill = .metric)) +
    
    geom_col(width = 0.65) +
    
    geom_col(
      data = ~ dplyr::filter(.x, best),
      fill = NA,
      color = "black",
      linewidth = 1.2,
      width = 0.65
    ) +
    
    facet_wrap(~ .metric, scales = "free_y") +
    
    scale_fill_manual(values = c(
      "Accuracy" = "#E76F51",
      "Precision" = "#B59F3B",
      "Recall" = "#4CAF50",
      "F1 Score" = "#5DA5A4",
      "Specificity" = "#6C8CD5",
      "ROC AUC" = "#D36AC2"
    )) +
    
    labs(
      title = title_text,
      subtitle = "Black outline indicates best performance",
      x = "Resampling Method",
      y = "Score"
    ) +
    
    theme_hr(base_size = 13) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5),
      plot.subtitle = element_text(hjust = 0.5),
      axis.text.x = element_text(face = "bold"),
      strip.text = element_text(face = "bold", size = 11),
      legend.position = "right"
    )
  
  print(p)
  save_plot(p, filename, 12, 8, plot_dir)
}

# Reproducibility
set.seed(123)

# Ensure all paths resolve from project root
library(rprojroot)

knitr::opts_knit$set(
  root.dir = rprojroot::find_rstudio_root_file()
)

# Output Directory
plot_dir <- here::here("docs", "figures", "turnover")
dir.create(plot_dir, recursive = TRUE, showWarnings = FALSE)

# Global chunk options
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  out.width = "100%",
  fig.retina = 2,
  fig.width = 8,
  fig.height = 5,
  dpi = 300,
  cache = FALSE,
  fig.path = "figures/turnover/"
)
```

```{r 00-load-packages, include=FALSE}
required_packages <- c(
  "car", "cowplot", "caret", "corrplot", "discrim", "dplyr",
  "e1071", "fastDummies", "forcats", "gbm", "GGally",
  "ggcorrplot", "ggplot2", "ggthemes", "glmnet", "grid",
  "gridExtra", "iml", "kableExtra", "knitr", "mgcv",
  "naivebayes", "naniar", "patchwork", "pheatmap",
  "plotly", "psych", "randomForest", "ranger",
  "RColorBrewer", "readxl", "reshape2", "skimr",
  "themis", "tidymodels", "tidyverse", "vip",
  "viridis", "xgboost", "janitor", "pROC"
)

missing_packages <- required_packages[
  !required_packages %in% rownames(installed.packages())
]

if (length(missing_packages) > 0) {
  stop(
    paste0(
      "Missing required packages: ",
      paste(missing_packages, collapse = ", "),
      ".\nPlease install them before rendering."
    )
  )
}

invisible(lapply(required_packages, library, character.only = TRUE))
```

```{r 00-load-data}
# Load
hr_df <- readRDS(here::here("data", "processed", "hr_dataset_clean.rds"))

# Validate First
required_cols <- c("turnover_intention")

missing_cols <- setdiff(required_cols, names(hr_df))

if (length(missing_cols) > 0) {
  stop(
    paste0(
      "Missing required columns: ",
      paste(missing_cols, collapse = ", ")
    )
  )
}

# Then Mutate
hr_df <- hr_df %>%
  mutate(turnover_intention = as.factor(turnover_intention))
```

```{=html}
<div class="main-container">

<div style="
  display:grid;
  grid-template-columns: 1fr auto 1fr;
  align-items:center;
  margin:30px 0;
">

  <div style="text-align:left;">
    <a href="HR_Data_EDA.html" style="
      display:inline-block;
      padding:12px 28px;
      border-radius:999px;
      font-weight:600;
      color:white;
      text-decoration:none;
      background:linear-gradient(135deg, #74B9FF, #2E86C1);
    ">
      ← Back
    </a>
  </div>

  <div style="text-align:center;">
    <a href="index.html" style="
      display:inline-block;
      padding:12px 28px;
      border-radius:999px;
      font-weight:600;
      color:white;
      text-decoration:none;
      background:linear-gradient(135deg, #2A9D8F, #52B788);
      box-shadow:0 6px 16px rgba(42,157,143,0.35);
    ">
      Home
    </a>
  </div>

  <div style="text-align:right;">
    <a href="HR_Data_Quiet_Quitting_Results.html" style="
      display:inline-block;
      padding:12px 28px;
      border-radius:999px;
      font-weight:600;
      color:white;
      text-decoration:none;
      background:linear-gradient(135deg, #FF7675, #922B21);
    ">
      Next →
    </a>
  </div>

</div>
```

# Turnover Intention Modeling

```{=html}
<div class="hero-section">
  <h1>Predicting Employee Turnover Intention</h1>
  <p class="hero-subtitle">
    This section develops predictive models to identify key drivers of employee turnover intention, 
    using statistical and machine learning techniques to uncover actionable insights for workforce retention.
  </p>
</div>

<hr>

<div class="section-block">

  <h2>Modeling Overview</h2>

  <div class="section-intro">
    Following exploratory analysis, predictive modeling was conducted to identify the most important drivers of employee turnover intention. 
    Multiple machine learning algorithms were evaluated to capture both linear relationships and more complex, non-linear workforce dynamics.
  </div>

  <h3>Data Preparation</h3>
  <p>
    Prior to modeling, numeric predictors were standardized where appropriate and categorical variables were encoded to ensure compatibility 
    with machine learning algorithms. The target variable for this analysis is <strong>Turnover Intention</strong>.
  </p>

  <h3>Models Evaluated</h3>
  <ul>
    <li>Logistic Regression</li>
    <li>Regularized Logistic Regression</li>
    <li>Random Forest</li>
    <li>Gradient Boosted Trees</li>
    <li>Support Vector Machines</li>
  </ul>

  <h3>Class Balancing Strategies</h3>
  <ul>
    <li>No Sampling</li>
    <li>Up Sampling</li>
    <li>Down Sampling</li>
    <li>SMOTE</li>
  </ul>

  <h3>Evaluation Metrics</h3>
  <ul>
    <li>Accuracy</li>
    <li>Precision</li>
    <li>Recall</li>
    <li>F1 Score</li>
    <li>Specificity</li>
    <li>ROC AUC</li>
  </ul>

  <p>
    Special emphasis is placed on recall and F1 score, as correctly identifying at-risk employees is more critical than maximizing overall accuracy.
  </p>

</div>
```

## Modeling Approach and Evaluation Framework

The objective of this section is to predict <strong>Turnover Intention</strong> using behavioral, demographic, and engagement-related predictors. 
Models are evaluated across multiple class balancing strategies to ensure robust and fair performance comparisons.

This framework provides a structured approach to identifying the most effective predictive model while maintaining interpretability and practical relevance.

### Model Setup and Validation Strategy

```{r 04-logit-shared-model-setup}
set.seed(123)

turnover_split <- initial_split(hr_df, prop = 0.8, strata = turnover_intention)
turnover_train_data <- training(turnover_split)
turnover_test_data  <- testing(turnover_split)

turnover_folds <- vfold_cv(turnover_train_data, v = 10, strata = turnover_intention)

base_recipe <- recipe(turnover_intention ~ ., data = turnover_train_data) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_numeric_predictors())

logit_model <- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")
```

```{r 04-logit-turnover-shared-metrics-setup}
all_metrics <- metric_set(accuracy, precision, recall, f_meas, yardstick::spec, roc_auc)
```

<div class="section-block">

<h2>Baseline Model: Logistic Regression</h2>

<div class="section-intro">
Logistic regression is used as a baseline model to establish a benchmark for performance. Its interpretability allows for clear understanding of relationships between predictors and turnover intention.
</div>

## Logistic Regression - No Resampling

```{r 04-logit-none-workflow, message=FALSE}
workflow_logit_none <- workflow() %>%
  add_model(logit_model) %>%
  add_recipe(base_recipe)

fit_logit_none <- fit(workflow_logit_none, data = turnover_train_data)

logit_preds_none <- predict(fit_logit_none, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(fit_logit_none, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(turnover_intention))
```

### Performance Metrics

```{r 04-logit-none-metrics, echo=TRUE}
logit_none_metrics <- all_metrics(
  logit_preds_none,
  truth = turnover_intention,
  estimate = .pred_class,
  .pred_Yes
)

p <- logit_none_metrics %>%
  filter(.metric %in% c("accuracy","precision","recall","f_meas","spec","roc_auc")) %>%
  mutate(.metric = factor(
    .metric,
    levels = c("accuracy","precision","recall","f_meas","spec","roc_auc"),
    labels = c("Accuracy","Precision","Recall","F1 Score","Specificity","ROC AUC")
  )) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#4B9CD3", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0,1), expand = expansion(mult = c(0,0.15))) +
  labs(title = "Logistic Regression — No Sampling", x = "Score") +
  theme_hr(base_size = 13) +
  theme(plot.title = element_text(face="bold", hjust=.5))

print(p)
save_plot(p, "logit_none_metrics.png", 8, 5, plot_dir)
```

### Confusion Matrix

```{r 04-logit-none-confmat, echo=TRUE}
cm_df <- conf_mat(logit_preds_none, truth = turnover_intention, estimate = .pred_class) %>%
  pluck("table") %>%
  as.data.frame()

cm_df$Truth <- factor(cm_df$Truth, levels = c("No","Yes"))
cm_df$Prediction <- factor(cm_df$Prediction, levels = c("No","Yes"))

p <- ggplot(cm_df, aes(Prediction, Truth, fill = Freq)) +
  geom_tile(color="white") +
  geom_text(aes(label=Freq), fontface="bold") +
  scale_fill_gradient(low="#EAF2FB", high="#4B9CD3") +
  labs(title="Logistic Regression — No Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "logit_none_confusion.png", 6, 5, plot_dir)
```

### ROC Curve

```{r 04-logit-none-roc-curve, echo=TRUE}
roc_df_none <- roc_curve(
  logit_preds_none,
  truth = turnover_intention,
  .pred_Yes
)

p <- ggplot(roc_df_none, aes(x = 1 - specificity, y = sensitivity)) +
  geom_line(color = "#4B9CD3", linewidth = 1.2) +
  geom_abline(linetype = "dashed", color = "gray60") +
  labs(
    title = "Logistic Regression — No Sampling",
    x = "False Positive Rate",
    y = "True Positive Rate"
  ) +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "logit_none_roc.png", 6, 5, plot_dir)
```

### Feature Importance

```{r 04-logit-none-vip, echo=TRUE}
vip_logit_none <- fit_logit_none %>%
  extract_fit_parsnip() %>%
  vi() %>%
  slice_max(Importance, n=15)

p <- ggplot(vip_logit_none,
            aes(Importance, fct_reorder(Variable, Importance))) +
  geom_col(fill="#4B9CD3") +
  labs(title="Logistic Regression — No Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "logit_none_vip.png", 8, 6, plot_dir)
```

## SMOTE

```{r 04-logit-smote-workflow, message=FALSE}
# SMOTE recipe
recipe_logit_smote <- base_recipe %>%
  step_smote(turnover_intention)

workflow_logit_smote <- workflow() %>%
  add_model(logit_model) %>%
  add_recipe(recipe_logit_smote)

# 10-Fold CV
set.seed(123)
logit_cv_smote <- fit_resamples(
  workflow_logit_smote,
  resamples = turnover_folds,
  metrics = metric_set(accuracy, precision, recall, f_meas, yardstick::spec, roc_auc)
)

# Final model fit and prediction on test data for SMOTE
fit_logit_smote <- fit(workflow_logit_smote, data = turnover_train_data)

logit_preds_smote <- predict(fit_logit_smote, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(fit_logit_smote, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(turnover_intention))
```

### Performance Metrics

```{r 04-logit-smote-finalfit, echo=TRUE}
logit_smote_metrics <- all_metrics(
  logit_preds_smote,
  truth = turnover_intention,
  estimate = .pred_class,
  .pred_Yes
)

p <- logit_smote_metrics %>%
  filter(.metric %in% c("accuracy","precision","recall","f_meas","spec","roc_auc")) %>%
  mutate(.metric = factor(
    .metric,
    levels = c("accuracy","precision","recall","f_meas","spec","roc_auc"),
    labels = c("Accuracy","Precision","Recall","F1 Score","Specificity","ROC AUC")
  )) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#4B9CD3", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0,1), expand = expansion(mult = c(0,0.15))) +
  labs(title = "Logistic Regression — SMOTE", x = "Score") +
  theme_hr(base_size = 13) +
  theme(plot.title = element_text(face="bold", hjust=.5))

print(p)
save_plot(p, "logit_smote_metrics.png", 8, 5, plot_dir)
```

### Confusion Matrix

```{r 04-logit-smote-confmat, echo=TRUE}
cm_df_smote <- conf_mat(
  logit_preds_smote,
  truth = turnover_intention,
  estimate = .pred_class
) %>%
  pluck("table") %>%
  as.data.frame()

cm_df_smote$Truth <- factor(cm_df_smote$Truth, levels = c("No","Yes"))
cm_df_smote$Prediction <- factor(cm_df_smote$Prediction, levels = c("No","Yes"))

p <- ggplot(cm_df_smote, aes(Prediction, Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#e6f2ff", high = "#4B9CD3", name = "Frequency") +
  labs(title = "Logistic Regression — SMOTE") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "logit_smote_confusion.png", 6, 5, plot_dir)
```

### ROC Curve

```{r 04-logit-smote-roc, echo=TRUE}
roc_df_smote <- roc_curve(
  logit_preds_smote,
  truth = turnover_intention,
  .pred_Yes
)

p <- ggplot(roc_df_smote, aes(x = 1 - specificity, y = sensitivity)) +
  geom_line(color = "#4B9CD3", linewidth = 1.2) +
  geom_abline(linetype = "dashed", color = "gray60") +
  labs(
    title = "Logistic Regression — SMOTE",
    x = "False Positive Rate",
    y = "True Positive Rate"
  ) +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "logit_smote_roc.png", 6, 5, plot_dir)
```

### Feature Importance

```{r 04-logit-smote-vip, echo=TRUE}
vip_logit_smote <- fit_logit_smote %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15)

p <- ggplot(
  vip_logit_smote,
  aes(x = Importance, y = fct_reorder(Variable, Importance))
) +
  geom_col(fill = "#4B9CD3") +
  geom_text(
    aes(label = round(Importance, 1)),
    hjust = -0.1,
    size = 3.5
  ) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Logistic Regression — SMOTE") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "logit_smote_vip.png", 8, 6, plot_dir)
```

## Up Sampling

```{r 04-logit-upsample-workflow, message=FALSE}
# Upsampling Recipe
recipe_logit_upsample <- base_recipe %>%
  step_upsample(turnover_intention)

# Workflow 
workflow_logit_upsample <- workflow() %>%
  add_model(logit_model) %>%
  add_recipe(recipe_logit_upsample)

# Fit
fit_logit_upsample <- fit(workflow_logit_upsample, data = turnover_train_data)

# Predict
logit_preds_upsample <- predict(fit_logit_upsample, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(fit_logit_upsample, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(turnover_intention))
```

### Performance Metrics

```{r 04-logit-upsample-metrics, echo=TRUE}
logit_upsample_metrics <- all_metrics(
  logit_preds_upsample,
  truth = turnover_intention,
  estimate = .pred_class,
  .pred_Yes
)

p <- logit_upsample_metrics %>%
  filter(.metric %in% c("accuracy","precision","recall","f_meas","spec","roc_auc")) %>%
  mutate(.metric = factor(
    .metric,
    levels = c("accuracy","precision","recall","f_meas","spec","roc_auc"),
    labels = c("Accuracy","Precision","Recall","F1 Score","Specificity","ROC AUC")
  )) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#4B9CD3", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0,1), expand = expansion(mult = c(0,0.15))) +
  labs(title = "Logistic Regression — Up Sampling", x = "Score") +
  theme_hr(base_size = 13) +
  theme(plot.title = element_text(face="bold", hjust=.5))

print(p)
save_plot(p, "logit_up_metrics.png", 8, 5, plot_dir)
```

### Confusion Matrix

```{r 04-logit-upsample-confmat, echo=TRUE}
cm_df_upsample <- conf_mat(
  logit_preds_upsample,
  truth = turnover_intention,
  estimate = .pred_class
) %>%
  pluck("table") %>%
  as.data.frame()

cm_df_upsample$Truth <- factor(cm_df_upsample$Truth, levels = c("No","Yes"))
cm_df_upsample$Prediction <- factor(cm_df_upsample$Prediction, levels = c("No","Yes"))

p <- ggplot(cm_df_upsample, aes(Prediction, Truth, fill = Freq)) +
  geom_tile(color = "white") +
  geom_text(aes(label = Freq), fontface = "bold") +
  scale_fill_gradient(low = "#e6f2ff", high = "#4B9CD3") +
  scale_y_discrete(limits = rev) +
  labs(title = "Logistic Regression — Up Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "logit_up_confusion.png", 6, 5, plot_dir)
```

### ROC Curve

```{r 04-logit-upsample-roc-curve, echo=TRUE}
roc_df_upsample <- roc_curve(
  logit_preds_upsample,
  truth = turnover_intention,
  .pred_Yes
)

p <- ggplot(roc_df_upsample, aes(x = 1 - specificity, y = sensitivity)) +
  geom_line(color = "#4B9CD3", linewidth = 1.2) +
  geom_abline(linetype = "dashed", color = "gray60") +
  labs(
    title = "Logistic Regression — Up Sampling",
    x = "False Positive Rate",
    y = "True Positive Rate"
  ) +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "logit_up_roc.png", 6, 5, plot_dir)
```

### Feature Importance

```{r 04-logit-upsample-vip, echo=TRUE}
vip_logit_upsample <- fit_logit_upsample %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15)

p <- ggplot(
  vip_logit_upsample,
  aes(x = Importance, y = fct_reorder(Variable, Importance))
) +
  geom_col(fill="#4B9CD3") +
  geom_text(aes(label = round(Importance,1)), hjust=-0.1, size=3.5) +
  scale_x_continuous(expand = expansion(mult=c(0,0.15))) +
  labs(title="Logistic Regression — Up Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "logit_up_vip.png", 8, 6, plot_dir)
```

## Down Sampling

```{r 04-logit-downsample-workflow, message=FALSE}
# Down Sampling Recipe
recipe_logit_downsample <- base_recipe %>%
  step_downsample(turnover_intention)

# Workflow 
workflow_logit_downsample <- workflow() %>%
  add_model(logit_model) %>%
  add_recipe(recipe_logit_downsample)

# Fit on training data
fit_logit_downsample <- fit(workflow_logit_downsample, data = turnover_train_data)

# Predict on test data
logit_preds_downsample <- predict(fit_logit_downsample, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(fit_logit_downsample, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(turnover_intention))
```

### Performance Metrics

```{r 04-logit-downsample-metrics, echo=TRUE}
logit_downsample_metrics <- all_metrics(
  logit_preds_downsample,
  truth = turnover_intention,
  estimate = .pred_class,
  .pred_Yes
)

p <- logit_downsample_metrics %>%
  filter(.metric %in% c("accuracy","precision","recall","f_meas","spec","roc_auc")) %>%
  mutate(.metric = factor(
    .metric,
    levels = c("accuracy","precision","recall","f_meas","spec","roc_auc"),
    labels = c("Accuracy","Precision","Recall","F1 Score","Specificity","ROC AUC")
  )) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill="#4B9CD3", width=0.6) +
  geom_text(aes(label=sprintf("%.3f", .estimate)), hjust=-0.1, fontface="bold") +
  scale_x_continuous(limits=c(0,1), expand=expansion(mult=c(0,0.15))) +
  labs(title = "Logistic Regression — Down Sampling", x = "Score") +
  theme_hr(base_size = 13) +
  theme(plot.title = element_text(face="bold", hjust=.5))

print(p)
save_plot(p, "logit_down_metrics.png", 8, 5, plot_dir)
```

### Confusion Matrix

```{r 04-logit-downsample-confmat, echo=TRUE}
cm_df_down <- conf_mat(
  logit_preds_downsample,
  truth = turnover_intention,
  estimate = .pred_class
) %>%
  pluck("table") %>%
  as.data.frame()

cm_df_down$Truth <- factor(cm_df_down$Truth, levels = c("No","Yes"))
cm_df_down$Prediction <- factor(cm_df_down$Prediction, levels = c("No","Yes"))

p <- ggplot(cm_df_down, aes(Prediction, Truth, fill = Freq)) +
  geom_tile(color="white") +
  geom_text(aes(label=Freq), fontface="bold") +
  scale_fill_gradient(low = "#e6f2ff", high = "#4B9CD3") +
  labs(title = "Logistic Regression — Down Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "logit_down_confusion.png", 6, 5, plot_dir)
```

### ROC Curve

```{r 04-logit-downsample-roc-curve, echo=TRUE}
roc_df_downsample <- roc_curve(
  logit_preds_downsample,
  truth = turnover_intention,
  .pred_Yes
)

p <- ggplot(roc_df_downsample, aes(x = 1 - specificity, y = sensitivity)) +
  geom_line(color = "#4B9CD3", linewidth = 1.2) +
  geom_abline(linetype = "dashed", color = "gray60") +
  labs(
    title = "Logistic Regression — Down Sampling",
    x = "False Positive Rate",
    y = "True Positive Rate"
  ) +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "logit_down_roc.png", 6, 5, plot_dir)
```

### Feature Importance

```{r 04-logit-downsample-vip, echo=TRUE}
vip_logit_down <- fit_logit_downsample %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15)

p <- ggplot(
  vip_logit_down,
  aes(x = Importance, y = fct_reorder(Variable, Importance))
) +
  geom_col(fill = "#4B9CD3") +
  labs(title = "Logistic Regression — Down Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "logit_down_vip.png", 8, 6, plot_dir)
```

## Results Table for Logistic Regression (Turnover Intention)

```{r 04-logit-all-metrics, echo=TRUE}
logit_all_metrics <- bind_rows(
  logit_none_metrics %>% mutate(Resample = "No Sampling"),
  logit_smote_metrics %>% mutate(Resample = "SMOTE"),
  logit_upsample_metrics %>% mutate(Resample = "Up Sampling"),
  logit_downsample_metrics %>% mutate(Resample = "Down Sampling")
)

plot_model_metrics(
  logit_all_metrics,
  "Logistic Regression Performance by Resampling Method",
  "logit_all_metrics.png"
)
```

Logistic Regression served as a transparent and interpretable baseline model across all resampling strategies. While it did not outperform more complex models in any single metric, it provided **consistent and stable performance**, making it valuable for benchmarking and stakeholder communication.

- **SMOTE** delivered the strongest balance across metrics, achieving the highest **recall (0.478)** and **F1 score (0.484)**, making it the most effective configuration for identifying at-risk employees.

- **Up Sampling** produced the highest **precision (0.463)** and **specificity (0.596)**, indicating improved performance in reducing false positives.

- **Down Sampling** demonstrated consistent performance across all metrics, matching SMOTE in **recall, precision, and F1 score (0.478)**, though with slightly lower overall discrimination (**ROC AUC = 0.437**).

- **No Sampling** achieved the highest **specificity (0.670)** and strong **accuracy (0.527)**, but under performed in **recall (0.359)**, highlighting the impact of class imbalance.

Overall, Logistic Regression provides a reliable baseline model, but its limitations in capturing complex patterns suggest that more advanced models may offer improved predictive performance.

```{=html}
<div class="takeaway-box">
<strong>Key Insight:</strong> Logistic regression provides a stable and interpretable baseline, 
but struggles to capture complex workforce patterns, indicating the need for more advanced models.
</div>
```

```{=html}
<div class="section-block">

<h2>Random Forest Model</h2>

<div class="section-intro">
Random Forest captures complex, non-linear relationships and interactions across workforce variables, making it a strong candidate for predictive modeling.
</div>
```

## Random Forest - No Resample

```{=html}
<div class="section-block">
  <p>
    Random Forest is an ensemble learning method that builds multiple decision trees and aggregates
    their predictions to improve classification performance and reduce overfitting.
  </p>
  <p>
    In this analysis, Random Forest is evaluated across multiple sampling strategies to assess how
    class balancing impacts its ability to detect employees with turnover intention.
  </p>
</div>
```

```{r 04-rf-no-resampling-workflow, message=FALSE}
# Define Random Forest model
rf_model <- rand_forest(mtry = tune(), trees = 500, min_n = tune()) %>%
  set_engine("ranger", importance = "impurity") %>%
  set_mode("classification")

# Workflow using base recipe
recipe_rf_none <- base_recipe

# Combine into Workflow
workflow_rf_none <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(recipe_rf_none)

# Tune the RF model with cross-validation
set.seed(123)
rf_tuned <- tune_grid(
  workflow_rf_none,
  resamples = turnover_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_rf <- select_best(rf_tuned, metric = "roc_auc")

# Final workflow with best RF model
final_rf_workflow <- finalize_workflow(workflow_rf_none, best_rf)

# Fit to training data
rf_fit_none <- fit(final_rf_workflow, data = turnover_train_data)

# Predict on test data
rf_preds_none <- predict(rf_fit_none, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(rf_fit_none, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(turnover_intention))
```

### Performance Metrics

```{r 04-rf-none-metrics, echo=TRUE}
rf_none_metrics <- all_metrics(
  rf_preds_none,
  truth = turnover_intention,
  estimate = .pred_class,
  .pred_Yes
)

p <- rf_none_metrics %>%
  filter(.metric %in% c("accuracy","precision","recall","f_meas","spec","roc_auc")) %>%
  mutate(.metric = factor(
    .metric,
    levels = c("accuracy","precision","recall","f_meas","spec","roc_auc"),
    labels = c("Accuracy","Precision","Recall","F1 Score","Specificity","ROC AUC")
  )) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#3FA34D", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0,1), expand = expansion(mult = c(0,0.15))) +
  labs(title = "Random Forest — No Sampling", x = "Score") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "rf_none_metrics.png", 8, 5, plot_dir)
```

### Confusion Matrix

```{r 04-rf-none-confmat, echo=TRUE}
cm_df_rf_none <- conf_mat(
  rf_preds_none,
  truth = turnover_intention,
  estimate = .pred_class
) %>%
  pluck("table") %>%
  as.data.frame()

cm_df_rf_none$Truth <- factor(cm_df_rf_none$Truth, levels = c("No","Yes"))
cm_df_rf_none$Prediction <- factor(cm_df_rf_none$Prediction, levels = c("No","Yes"))

p <- ggplot(cm_df_rf_none, aes(Prediction, Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(
    low = "#E8F5E9",
    high = "#3FA34D",
    name = "Frequency"
  ) +
  labs(title = "Random Forest — No Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "rf_none_confusion.png", 6, 5, plot_dir)
```

### ROC Curve

```{r 04-rf-none-roc, echo=TRUE}
roc_df_rf_none <- roc_curve(
  rf_preds_none,
  truth = turnover_intention,
  .pred_Yes
)

p <- ggplot(roc_df_rf_none, aes(x = 1 - specificity, y = sensitivity)) +
  geom_line(color = "#3FA34D", linewidth = 1.2) +
  geom_abline(linetype = "dashed", color = "gray60") +
  labs(
    title = "Random Forest — No Sampling",
    x = "False Positive Rate",
    y = "True Positive Rate"
  ) +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "rf_none_roc.png", 6, 5, plot_dir)
```

### Feature Importance

```{r 04-rf-none-vip, echo=TRUE}
vip_rf_none <- rf_fit_none %>%
  extract_fit_parsnip() %>%
  vi() %>%
  slice_max(Importance, n = 15)

p <- ggplot(vip_rf_none,
            aes(Importance, fct_reorder(Variable, Importance))) +
  geom_col(fill = "#3FA34D") +
  labs(title = "Random Forest — No Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "rf_none_vip.png", 8, 6, plot_dir)
```

## Random Forest - SMOTE

```{r 04-rf-smote-workflow, message=FALSE}
# Create Recipe from Base
recipe_rf_smote <- base_recipe %>%
  step_smote(turnover_intention)

# Workflow with Random Forest
workflow_rf_smote <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(recipe_rf_smote)

# Tune the RF model with cross-validation
set.seed(123)
rf_tuned_smote <- tune_grid(
  workflow_rf_smote,
  resamples = turnover_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_rf_smote <- select_best(rf_tuned_smote, metric = "roc_auc")

# Final workflow with best RF model
final_rf_workflow_smote <- finalize_workflow(workflow_rf_smote, best_rf_smote)

# Fit to training data
rf_fit_smote <- fit(final_rf_workflow_smote, data = turnover_train_data)

# Predict on test data
rf_preds_smote <- predict(rf_fit_smote, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(rf_fit_smote, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(turnover_intention))
```

### Performance Metrics

```{r 04-rf-smote-metrics, echo=TRUE}
rf_smote_metrics <- all_metrics(
  rf_preds_smote,
  truth = turnover_intention,
  estimate = .pred_class,
  .pred_Yes
)

p <- rf_smote_metrics %>%
  filter(.metric %in% c("accuracy","precision","recall","f_meas","spec","roc_auc")) %>%
  mutate(.metric = factor(
    .metric,
    levels = c("accuracy","precision","recall","f_meas","spec","roc_auc"),
    labels = c("Accuracy","Precision","Recall","F1 Score","Specificity","ROC AUC")
  )) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#3FA34D", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, size = 4.2, fontface = "bold") +
  labs(title = "Random Forest — SMOTE") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "rf_smote_metrics.png", 8, 5, plot_dir)
```

### Confusion Matrix

```{r 04-rf-smote-confmat, echo=TRUE}
cm_df_rf_smote <- conf_mat(
  rf_preds_smote,
  truth = turnover_intention,
  estimate = .pred_class
) %>%
  pluck("table") %>%
  as.data.frame()

cm_df_rf_smote$Truth <- factor(cm_df_rf_smote$Truth, levels = c("No", "Yes"))
cm_df_rf_smote$Prediction <- factor(cm_df_rf_smote$Prediction, levels = c("No", "Yes"))

p <- ggplot(cm_df_rf_smote, aes(Prediction, Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(
    low = "#E8F5E9",
    high = "#3FA34D",
    name = "Frequency"
  ) +
  labs(title = "Random Forest — SMOTE") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "rf_smote_confusion.png", 6, 5, plot_dir)
```

### ROC Curve

```{r 04-rf-smote-roc, echo=TRUE}
roc_df_rf_smote <- roc_curve(
  rf_preds_smote,
  truth = turnover_intention,
  .pred_Yes
)

p <- ggplot(roc_df_rf_smote, aes(x = 1 - specificity, y = sensitivity)) +
  geom_line(color = "#3FA34D", linewidth = 1.2) +
  geom_abline(linetype = "dashed", color = "gray60") +
  labs(
    title = "Random Forest — SMOTE",
    x = "False Positive Rate",
    y = "True Positive Rate"
  ) +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "rf_smote_roc.png", 6, 5, plot_dir)
```

### Feature Importance

```{r 04-rf-smote-vip, echo=TRUE}
vip_rf_smote <- rf_fit_smote %>%
  extract_fit_parsnip() %>%
  vi() %>%
  slice_max(Importance, n = 15)

p <- ggplot(vip_rf_smote,
            aes(Importance, fct_reorder(Variable, Importance))) +
  geom_col(fill = "#3FA34D") +
  labs(title = "Random Forest — SMOTE") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "rf_smote_vip.png", 8, 6, plot_dir)
```

## Random Forest - Up Sampling

```{r 04-rf-upsample-workflow, message=FALSE}
# Create Recipe from Base
recipe_rf_upsample <- base_recipe %>%
  step_upsample(turnover_intention)

# Workflow with Random Forest
workflow_rf_upsample <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(recipe_rf_upsample)

# Tune the RF model with cross-validation
set.seed(123)
rf_tuned_upsample <- tune_grid(
  workflow_rf_upsample,
  resamples = turnover_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_rf_upsample <- select_best(rf_tuned_upsample, metric = "roc_auc")

# Final workflow with best RF model
final_rf_workflow_upsample <- finalize_workflow(workflow_rf_upsample, best_rf_upsample)

# Fit to training data
rf_fit_upsample <- fit(final_rf_workflow_upsample, data = turnover_train_data)

# Predict on test data
rf_preds_upsample <- predict(rf_fit_upsample, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(rf_fit_upsample, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(turnover_intention))
```

### Performance Metrics

```{r 04-rf-upsample-metrics, echo=TRUE}
rf_up_metrics <- all_metrics(
  rf_preds_upsample,
  truth = turnover_intention,
  estimate = .pred_class,
  .pred_Yes
)

p <- rf_up_metrics %>%
  filter(.metric %in% c("accuracy","precision","recall","f_meas","spec","roc_auc")) %>%
  mutate(.metric = factor(
    .metric,
    levels = c("accuracy","precision","recall","f_meas","spec","roc_auc"),
    labels = c("Accuracy","Precision","Recall","F1 Score","Specificity","ROC AUC")
  )) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#3FA34D", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, size = 4.2, fontface = "bold") +
  labs(title = "Random Forest — Up Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "rf_up_metrics.png", 8, 5, plot_dir)
```

### Confusion Matrix

```{r 04-rf-upsample-confmat, echo=TRUE}
cm_df_rf_up <- conf_mat(
  rf_preds_upsample,
  truth = turnover_intention,
  estimate = .pred_class
) %>%
  pluck("table") %>%
  as.data.frame()

cm_df_rf_up$Truth <- factor(cm_df_rf_up$Truth, levels = c("No","Yes"))
cm_df_rf_up$Prediction <- factor(cm_df_rf_up$Prediction, levels = c("No","Yes"))

p <- ggplot(cm_df_rf_up, aes(Prediction, Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#E8F5E9", high = "#3FA34D") +
  labs(title = "Random Forest — Up Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "rf_up_confusion.png", 6, 5, plot_dir)
```

### ROC Curve

```{r 04-rf-upsample-roc, echo=TRUE}
roc_df_rf_up <- roc_curve(
  rf_preds_upsample,
  truth = turnover_intention,
  .pred_Yes
)

p <- ggplot(roc_df_rf_up, aes(x = 1 - specificity, y = sensitivity)) +
  geom_line(color = "#3FA34D", linewidth = 1.2) +
  geom_abline(linetype = "dashed", color = "gray60") +
  labs(
    title = "Random Forest — Up Sampling",
    x = "False Positive Rate",
    y = "True Positive Rate"
  ) +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "rf_up_roc.png", 6, 5, plot_dir)
```

### Feature Importance

```{r 04-rf-upsample-vip, echo=TRUE}
vip_rf_up <- rf_fit_upsample %>%
  extract_fit_parsnip() %>%
  vi() %>%
  slice_max(Importance, n = 15)

p <- ggplot(
  vip_rf_up,
  aes(x = Importance, y = fct_reorder(Variable, Importance))
) +
  geom_col(fill = "#3FA34D") +
  labs(title = "Random Forest — Up Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "rf_up_vip.png", 8, 6, plot_dir)
```

## Random Forest - Down Sampling

```{r 04-rf-downsample-workflow, message=FALSE}
# Create Recipe from Base
recipe_rf_downsample <- base_recipe %>%
  step_downsample(turnover_intention)

# Workflow with Random Forest
workflow_rf_downsample <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(recipe_rf_downsample)

# Tune the RF model with cross-validation
set.seed(123)
rf_tuned_downsample <- tune_grid(
  workflow_rf_downsample,
  resamples = turnover_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_rf_downsample <- select_best(rf_tuned_downsample, metric = "roc_auc")

# Final workflow with best RF model
final_rf_workflow_downsample <- finalize_workflow(workflow_rf_downsample, best_rf_downsample)

# Fit to training data
rf_fit_downsample <- fit(final_rf_workflow_downsample, data = turnover_train_data)

# Predict on test data
rf_preds_downsample <- predict(rf_fit_downsample, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(rf_fit_downsample, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(turnover_intention))
```

### Performance Metrics

```{r 04-rf-down-metrics, echo=TRUE}
rf_down_metrics <- all_metrics(
  rf_preds_downsample,
  truth = turnover_intention,
  estimate = .pred_class,
  .pred_Yes
)

p <- rf_down_metrics %>%
  filter(.metric %in% c("accuracy","precision","recall","f_meas","spec","roc_auc")) %>%
  mutate(.metric = factor(
    .metric,
    levels = c("accuracy","precision","recall","f_meas","spec","roc_auc"),
    labels = c("Accuracy","Precision","Recall","F1 Score","Specificity","ROC AUC")
  )) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#3FA34D", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, size = 4.2, fontface = "bold") +
  labs(title = "Random Forest — Down Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "rf_down_metrics.png", 8, 5, plot_dir)
```

### Confusion Matrix

```{r 04-rf-down-confmat, echo=TRUE}
cm_df_rf_down <- conf_mat(
  rf_preds_downsample,
  truth = turnover_intention,
  estimate = .pred_class
) %>%
  pluck("table") %>%
  as.data.frame()

cm_df_rf_down$Truth <- factor(cm_df_rf_down$Truth, levels = c("No","Yes"))
cm_df_rf_down$Prediction <- factor(cm_df_rf_down$Prediction, levels = c("No","Yes"))

p <- ggplot(cm_df_rf_down, aes(Prediction, Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#E8F5E9", high = "#3FA34D") +
  labs(title = "Random Forest — Down Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "rf_down_confusion.png", 6, 5, plot_dir)
```

### ROC Curve

```{r 04-rf-down-roc, echo=TRUE}
roc_df_rf_down <- roc_curve(
  rf_preds_downsample,
  truth = turnover_intention,
  .pred_Yes
)

p <- ggplot(roc_df_rf_down, aes(x = 1 - specificity, y = sensitivity)) +
  geom_line(color = "#3FA34D", linewidth = 1.2) +
  geom_abline(linetype = "dashed", color = "gray60") +
  labs(
    title = "Random Forest — Down Sampling",
    x = "False Positive Rate",
    y = "True Positive Rate"
  ) +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "rf_down_roc.png", 6, 5, plot_dir)
```

### Feature Importance

```{r 04-rf-down-vip, echo=TRUE}
vip_rf_down <- rf_fit_downsample %>%
  extract_fit_parsnip() %>%
  vi() %>%
  slice_max(Importance, n = 15)

p <- ggplot(
  vip_rf_down,
  aes(x = Importance, y = fct_reorder(Variable, Importance))
) +
  geom_col(fill = "#3FA34D") +
  labs(title = "Random Forest — Down Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "rf_down_vip.png", 8, 6, plot_dir)
```

<div class="section-intro">
The following visualization summarizes Random Forest performance across all resampling strategies, 
highlighting the strongest configuration for each evaluation metric.
</div>

### Results Table for Random Forest (Turnover Intention)

```{r 04-rf-all-metrics, echo=TRUE}
rf_all_metrics <- bind_rows(
  rf_none_metrics %>% mutate(Resample = "No Sampling"),
  rf_smote_metrics %>% mutate(Resample = "SMOTE"),
  rf_up_metrics %>% mutate(Resample = "Up Sampling"),
  rf_down_metrics %>% mutate(Resample = "Down Sampling")
)

plot_model_metrics(
  rf_all_metrics,
  "Random Forest Performance by Resampling Method",
  "rf_all_metrics.png"
)
```

```{=html}
<div class="section-block">
  <h4>Interpretation — Random Forest with Down Sampling</h4>
  <p>
    Down sampling was used to reduce class imbalance by limiting the number of majority-class observations during training.
    This approach often helps the model become more responsive to the minority class, which is valuable when the primary goal
    is identifying employees at risk of turnover intention.
  </p>
  <p>
    In this Random Forest configuration, the results should be interpreted by balancing sensitivity against specificity.
    If sensitivity improves meaningfully compared to the baseline model, the down-sampled version may be more useful for
    early identification purposes, even if overall accuracy declines slightly. The confusion matrix and ROC curve together
    help determine whether this tradeoff is acceptable within the project’s workforce analytics context.
  </p>
</div>
```

### Executive Summary for Random Forest (Turnover Intention)

Random Forest significantly outperformed the baseline logistic regression model, demonstrating strong capability in capturing complex workforce patterns. It performed especially well on precision, specificity, and ROC AUC across multiple resampling strategies, confirming its robustness as a turnover prediction model.

- **Up Sampling** yielded the best **precision** (**0.556**), **specificity** (**0.927**), and **accuracy** (**0.552**), making it well-suited for minimizing false positives and identifying confident cases of turnover risk.

- **Down Sampling** balanced metrics across the board with solid **recall** (**0.478**), **F1 score** (**0.463**), and **ROC AUC** (**0.495**), offering a steady and dependable configuration.

- **SMOTE** improved **recall** (**0.380**) and **F1 score** (**0.417**) over the baseline, helpful when capturing at-risk employees is critical.

- **No Sampling** still performed reasonably well, especially in **specificity** (**0.651**) and **ROC AUC** (**0.499**), showing Random Forest’s natural ability to handle imbalance without adjustments.

---

## XGBoost - No Sampling

```{r 04-xgb-no-resampling-workflow, message=FALSE}
# Define XGBoost model
xgb_model <- boost_tree(
  trees = 1000,
  tree_depth = tune(),
  min_n = tune(),
  loss_reduction = tune(),        
  sample_size = tune(),           
  mtry = tune(),                  
  learn_rate = tune()             
) %>%
  set_engine("xgboost") %>%
  set_mode("classification")

# Recipe with No Sampling
recipe_xgb_none <- base_recipe

# Combine into Workflow
workflow_xgb_none <- workflow() %>%
  add_model(xgb_model) %>%
  add_recipe(recipe_xgb_none)

# Tune the XGBoost model with cross-validation
set.seed(123)
xgb_tuned <- tune_grid(
  workflow_xgb_none,
  resamples = turnover_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_xgb <- select_best(xgb_tuned, metric = "roc_auc")

# Final workflow with best XGBoost model
final_xgb_workflow <- finalize_workflow(workflow_xgb_none, best_xgb)

# Fit to training data
xgb_fit_none <- fit(final_xgb_workflow, data = turnover_train_data)

# Predict on test data
xgb_preds_none <- predict(xgb_fit_none, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(xgb_fit_none, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(turnover_intention))
```

### Performance Metrics

```{r 04-xgb-no-resampling-metrics, echo=TRUE}
xgb_none_metrics <- all_metrics(
  xgb_preds_none,
  truth = turnover_intention,
  estimate = .pred_class,
  .pred_Yes
)

p <- xgb_none_metrics %>%
  filter(.metric %in% c("accuracy","precision","recall","f_meas","spec","roc_auc")) %>%
  mutate(.metric = factor(
    .metric,
    levels = c("accuracy","precision","recall","f_meas","spec","roc_auc"),
    labels = c("Accuracy","Precision","Recall","F1 Score","Specificity","ROC AUC")
  )) %>%
  ggplot(aes(.estimate, fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#E87461", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, fontface = "bold") +
  scale_x_continuous(limits = c(0,1), expand = expansion(mult = c(0,0.15))) +
  labs(title = "XGBoost — No Sampling", x = "Score") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "xgb_none_metrics.png", 8, 5, plot_dir)
```

### Confusion Matrix

```{r 04-xgb-no-resampling-confmat, echo=TRUE}
cm_df_xgb_none <- conf_mat(
  xgb_preds_none,
  truth = turnover_intention,
  estimate = .pred_class
) %>%
  pluck("table") %>%
  as.data.frame()

cm_df_xgb_none$Truth <- factor(cm_df_xgb_none$Truth, levels = c("No","Yes"))
cm_df_xgb_none$Prediction <- factor(cm_df_xgb_none$Prediction, levels = c("No","Yes"))

p <- ggplot(cm_df_xgb_none, aes(Prediction, Truth, fill = Freq)) +
  geom_tile(color="white", linewidth = 0.5) +
  geom_text(aes(label=Freq), color="black", fontface="bold") +
  scale_fill_gradient(low="#FDEDEC", high="#E87461") +
  labs(title="XGBoost — No Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "xgb_none_confusion.png", 6, 5, plot_dir)
```

### ROC Curve

```{r 04-xgb-no-resampling-roc-curve, echo=TRUE}
roc_df_xgb_none <- roc_curve(
  xgb_preds_none,
  truth = turnover_intention,
  .pred_Yes
)

p <- ggplot(roc_df_xgb_none, aes(x = 1 - specificity, y = sensitivity)) +
  geom_line(color = "#E87461", linewidth = 1.2) +
  geom_abline(linetype = "dashed", color = "gray60") +
  labs(title = "XGBoost — No Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "xgb_none_roc.png", 6, 5, plot_dir)
```

### Feature Importance

```{r 04-xgb-no-resampling-vip, echo=TRUE}
vip_xgb_none <- xgb_fit_none %>%
  extract_fit_parsnip() %>%
  vi() %>%
  slice_max(Importance, n=15)

p <- ggplot(vip_xgb_none,
            aes(Importance, fct_reorder(Variable, Importance))) +
  geom_col(fill="#E87461") +
  labs(title="XGBoost — No Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "xgb_none_vip.png", 8, 6, plot_dir)
```

## XGBoost - SMOTE

```{r 04-xgb-smote-workflow, message=FALSE}
# Create SMOTE Recipe
recipe_xgb_smote <- base_recipe %>%
  step_smote(turnover_intention)

# Build XGBoost Workflow with SMOTE
workflow_xgb_smote <- workflow() %>%
  add_model(xgb_model) %>%
  add_recipe(recipe_xgb_smote)

# Tune Hyperparameters with Cross-Validation
set.seed(123)
xgb_tuned_smote <- tune_grid(
  workflow_xgb_smote,
  resamples = turnover_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
)

# Select Best Parameters Based on ROC AUC
best_xgb_smote <- select_best(xgb_tuned_smote, metric = "roc_auc")

# Finalize Workflow with Best Parameters
final_xgb_workflow_smote <- finalize_workflow(workflow_xgb_smote, best_xgb_smote)

# Fit Final Model to Training Data
xgb_fit_smote <- fit(final_xgb_workflow_smote, data = turnover_train_data)

# Predict on Test Data
xgb_preds_smote <- predict(xgb_fit_smote, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(xgb_fit_smote, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(turnover_intention))
```

### Peformance Metrics

```{r 04-xgb-smote-metrics, echo=TRUE}
xgb_smote_metrics <- all_metrics(
  xgb_preds_smote,
  truth = turnover_intention,
  estimate = .pred_class,
  .pred_Yes
)

p <- xgb_smote_metrics %>%
  filter(.metric %in% c("accuracy","precision","recall","f_meas","spec","roc_auc")) %>%
  ggplot(aes(.estimate, fct_reorder(.metric, .estimate))) +
  geom_col(fill="#E87461") +
  labs(title="XGBoost — SMOTE") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "xgb_smote_metrics.png", 8, 5, plot_dir)
```

### Confusion Matrix

```{r 04-xgb-smote-confmat, echo=TRUE}
cm_df_xgb_smote <- conf_mat(
  xgb_preds_smote,
  truth = turnover_intention,
  estimate = .pred_class
) %>%
  pluck("table") %>%
  as.data.frame()

cm_df_xgb_smote$Truth <- factor(cm_df_xgb_smote$Truth, levels = c("No","Yes"))
cm_df_xgb_smote$Prediction <- factor(cm_df_xgb_smote$Prediction, levels = c("No","Yes"))

p <- ggplot(cm_df_xgb_smote, aes(Prediction, Truth, fill = Freq)) +
  geom_tile(color="white", linewidth = 0.5) +
  geom_text(aes(label=Freq), fontface="bold") +
  scale_fill_gradient(low="#FDEDEC", high="#E87461") +
  labs(title="XGBoost — SMOTE") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "xgb_smote_confusion.png", 6, 5, plot_dir)
```

### ROC Curve

```{r 04-xgb-smote-roc-curve, echo=TRUE}
roc_df_xgb_smote <- roc_curve(
  xgb_preds_smote,
  truth = turnover_intention,
  .pred_Yes
)

p <- ggplot(roc_df_xgb_smote, aes(x = 1 - specificity, y = sensitivity)) +
  geom_line(color = "#E87461", linewidth = 1.2) +
  geom_abline(linetype = "dashed", color = "gray60") +
  labs(title = "XGBoost — SMOTE") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "xgb_smote_roc.png", 6, 5, plot_dir)
```

### Feature Importance

```{r 04-xgb-smote-vip, echo=TRUE}
vip_xgb_smote <- xgb_fit_smote %>%
  extract_fit_parsnip() %>%
  vi() %>%
  slice_max(Importance, n = 15)

p <- ggplot(
  vip_xgb_smote,
  aes(x = Importance, y = fct_reorder(Variable, Importance))
) +
  geom_col(fill = "#E87461") +
  labs(title = "XGBoost — SMOTE") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "xgb_smote_vip.png", 8, 6, plot_dir)
```

### XGBoost - Up Sampling

```{r 04-xgb-upsample-workflow, message=FALSE}
# Create Recipe from Base
recipe_xgb_upsample <- base_recipe %>%
  step_upsample(turnover_intention)

# Workflow with XGBoost
workflow_xgb_upsample <- workflow() %>%
  add_model(xgb_model) %>%
  add_recipe(recipe_xgb_upsample)

# Tune the XGBoost model with cross-validation
set.seed(123)
xgb_tuned_upsample <- tune_grid(
  workflow_xgb_upsample,
  resamples = turnover_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_xgb_upsample <- select_best(xgb_tuned_upsample, metric = "roc_auc")

# Final workflow with best XGBoost model
final_xgb_workflow_upsample <- finalize_workflow(workflow_xgb_upsample, best_xgb_upsample)

# Fit to training data
xgb_fit_upsample <- fit(final_xgb_workflow_upsample, data = turnover_train_data)

# Predict on test data
xgb_preds_upsample <- predict(xgb_fit_upsample, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(xgb_fit_upsample, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(turnover_intention))
```

### Performance Metrics

```{r 04-xgb-upsample-metrics, echo=TRUE}
xgb_upsample_metrics <- all_metrics(
  xgb_preds_upsample,
  truth = turnover_intention,
  estimate = .pred_class,
  .pred_Yes
)

p <- xgb_upsample_metrics %>%
  filter(.metric %in% c("accuracy","precision","recall","f_meas","spec","roc_auc")) %>%
  mutate(.metric = factor(
    .metric,
    levels = c("accuracy","precision","recall","f_meas","spec","roc_auc"),
    labels = c("Accuracy","Precision","Recall","F1 Score","Specificity","ROC AUC")
  )) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#E87461", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0,1), expand = expansion(mult = c(0,0.15))) +
  labs(title = "XGBoost — Up Sampling", x = "Score") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "xgb_up_metrics.png", 8, 5, plot_dir)
```

### Confusion Matrix

```{r 04-xgb-upsample-confmat, echo=TRUE}
cm_df_xgb_up <- conf_mat(
  xgb_preds_upsample,
  truth = turnover_intention,
  estimate = .pred_class
) %>%
  pluck("table") %>%
  as.data.frame()

cm_df_xgb_up$Truth <- factor(cm_df_xgb_up$Truth, levels = c("No","Yes"))
cm_df_xgb_up$Prediction <- factor(cm_df_xgb_up$Prediction, levels = c("No","Yes"))

p <- ggplot(cm_df_xgb_up, aes(Prediction, Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), fontface = "bold") +
  scale_fill_gradient(low = "#FDEDEC", high = "#E87461") +
  labs(title = "XGBoost — Up Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "xgb_up_confusion.png", 6, 5, plot_dir)
```

### ROC Curve

```{r 04-xgb-upsample-roc-curve, echo=TRUE}
roc_df_xgb_up <- roc_curve(
  xgb_preds_upsample, 
  truth = turnover_intention, 
  .pred_Yes)

p <- ggplot(roc_df_xgb_up, aes(x = 1 - specificity, y = sensitivity)) +
  geom_line(color = "#E87461", linewidth = 1.2) +
  geom_abline(linetype = "dashed", color = "gray60") +
  labs(title = "XGBoost — Up Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "xgb_up_roc.png", 6, 5, plot_dir)
```

## Feature Importance

```{r 04-xgb-upsample-vip, echo=TRUE}
vip_xgb_up <- xgb_fit_upsample %>%
  extract_fit_parsnip() %>%
  vi() %>%
  slice_max(Importance, n = 15)

p <- ggplot(
  vip_xgb_up,
  aes(x = Importance, y = fct_reorder(Variable, Importance))
) +
  geom_col(fill = "#E87461") +
  labs(title = "XGBoost — Up Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "xgb_up_vip.png", 8, 6, plot_dir)
```

## XGBoost - Down Sampling

```{r 04-xgb-downsample-workflow, message=FALSE}
# Create Recipe from Base
recipe_xgb_downsample <- base_recipe %>%
  step_downsample(turnover_intention)

# Workflow with XGBoost
workflow_xgb_downsample <- workflow() %>%
  add_model(xgb_model) %>%
  add_recipe(recipe_xgb_downsample)

# Tune the RF model with cross-validation
set.seed(123)
xgb_tuned_downsample <- tune_grid(
  workflow_xgb_downsample,
  resamples = turnover_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_xgb_downsample <- select_best(xgb_tuned_downsample, metric = "roc_auc")

# Final workflow with best XGBoost model
final_xgb_workflow_downsample <- finalize_workflow(workflow_xgb_downsample, best_xgb_downsample)

# Fit to training data
xgb_fit_downsample <- fit(final_xgb_workflow_downsample, data = turnover_train_data)

# Predict on test data
xgb_preds_downsample <- predict(xgb_fit_downsample, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(xgb_fit_downsample, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(turnover_intention))
```

### Performance Metrics

```{r 04-xgb-downsample-metrics, echo=TRUE}
xgb_downsample_metrics <- all_metrics(
  xgb_preds_downsample,
  truth = turnover_intention,
  estimate = .pred_class,
  .pred_Yes
)

p <- xgb_downsample_metrics %>%
  filter(.metric %in% c("accuracy","precision","recall","f_meas","spec","roc_auc")) %>%
  mutate(.metric = factor(
    .metric,
    levels = c("accuracy","precision","recall","f_meas","spec","roc_auc"),
    labels = c("Accuracy","Precision","Recall","F1 Score","Specificity","ROC AUC")
  )) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#E87461", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0,1), expand = expansion(mult = c(0,0.15))) +
  labs(title = "XGBoost — Down Sampling", x = "Score") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "xgb_down_metrics.png", 8, 5, plot_dir)
```

### Confusion Matrix

```{r 04-xgb-down-sample-confmat, echo=TRUE}
cm_df_xgb_down <- conf_mat(
  xgb_preds_downsample,
  truth = turnover_intention,
  estimate = .pred_class
) %>%
  pluck("table") %>%
  as.data.frame()

cm_df_xgb_down$Truth <- factor(cm_df_xgb_down$Truth, levels = c("No","Yes"))
cm_df_xgb_down$Prediction <- factor(cm_df_xgb_down$Prediction, levels = c("No","Yes"))

p <- ggplot(cm_df_xgb_down, aes(Prediction, Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), fontface = "bold") +
  scale_fill_gradient(low = "#FDEDEC", high = "#E87461") +
  labs(title = "XGBoost — Down Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "xgb_down_confusion.png", 6, 5, plot_dir)
```

### ROC Curve

```{r 04-xgb-downsample-roc-curve, echo=TRUE}
roc_df_xgb_down <- roc_curve(
  xgb_preds_downsample,
  truth = turnover_intention,
  .pred_Yes
)

p <- ggplot(roc_df_xgb_down, aes(x = 1 - specificity, y = sensitivity)) +
  geom_line(color = "#E87461", linewidth = 1.2) +
  geom_abline(linetype = "dashed", color = "gray60") +
  labs(title = "XGBoost — Down Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "xgb_down_roc.png", 6, 5, plot_dir)
```

### Feature Importance

```{r 04-xgb-downsample-vip, echo=TRUE}
vip_xgb_down <- xgb_fit_downsample %>%
  extract_fit_parsnip() %>%
  vi() %>%
  slice_max(Importance, n = 15)

p <- ggplot(
  vip_xgb_down,
  aes(x = Importance, y = fct_reorder(Variable, Importance))
) +
  geom_col(fill = "#E87461") +
  labs(title = "XGBoost — Down Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "xgb_down_vip.png", 8, 6, plot_dir)
```

### Results Table for XGBoost (Turnover Intention)

```{r 04-xgb-all-metrics, echo=TRUE}
xgb_all_metrics <- bind_rows(
  xgb_none_metrics %>% mutate(Resample = "No Sampling"),
  xgb_smote_metrics %>% mutate(Resample = "SMOTE"),
  xgb_upsample_metrics %>% mutate(Resample = "Up Sampling"),
  xgb_downsample_metrics %>% mutate(Resample = "Down Sampling")
)

plot_model_metrics(
  xgb_all_metrics,
  "XGBoost Performance by Resampling Method",
  "xgb_all_metrics.png"
)
```

### Executive Summary for XGBoost (Turnover Intention)

XGBoost delivered strong and consistent performance across all resampling strategies. It excelled particularly in **specificity**, **precision**, and **ROC AUC**, making it highly suitable for applications where false positives must be minimized. While it did not lead in recall, it proved to be a stable and versatile model for structured prediction tasks.

- **Down Sampling** delivered the most balanced outcome, with competitive **recall** (**0.457**) and **F1 score** (**0.452**), and steady **ROC AUC** (**0.495**), making it the most dependable configuration.  

- **SMOTE** boosted **recall** (**0.402**) and **F1 score** (**0.423**), which is ideal when identifying all potential turnover cases is a priority.  

- **No Sampling** offered the highest **specificity** (**0.826**) and **precision** (**0.525**), but recall was limited (**0.228**), making it better for conservative decision-making.  

- **Up Sampling** achieved perfect **specificity** (**1.000**) and solid **ROC AUC** (**0.500**), but completely failed in **recall** (**0.000**), making it inappropriate for use cases that require true positive detection.

---

## Naive Bayes - No Sampling

```{r 04-nb-model, message=FALSE}
# Define Naive Bayes Model
nb_model <- naive_Bayes() %>%
  set_engine("naivebayes") %>%
  set_mode("classification")

# Build Workflow
workflow_nb_none <- workflow() %>%
  add_model(nb_model) %>%
  add_recipe(base_recipe)

# 10-Fold Cross-Validation on Training Data
set.seed(123)
nb_cv_results <- fit_resamples(
  workflow_nb_none,
  resamples = turnover_folds,
  metrics = metric_set(accuracy, precision, recall, f_meas, yardstick::spec, roc_auc)
)

# Fit Final Model to Full Training Data
nb_fit_none <- fit(workflow_nb_none, data = turnover_train_data)

# Predict on Test Data
nb_preds_none <- predict(nb_fit_none, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(nb_fit_none, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(turnover_intention))
```

### Performance Metrics

```{r 04-nb-metrics, echo=TRUE}
nb_none_metrics <- all_metrics(
  nb_preds_none,
  truth = turnover_intention,
  estimate = .pred_class,
  .pred_Yes
)

p <- nb_none_metrics %>%
  filter(.metric %in% c("accuracy","precision","recall","f_meas","spec","roc_auc")) %>%
  mutate(.metric = factor(
    .metric,
    levels = c("accuracy","precision","recall","f_meas","spec","roc_auc"),
    labels = c("Accuracy","Precision","Recall","F1 Score","Specificity","ROC AUC")
  )) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#F4D03F", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0,1), expand = expansion(mult = c(0,0.15))) +
  labs(title = "Naive Bayes — No Sampling", x = "Score") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "nb_none_metrics.png", 8, 5, plot_dir)
```

### Confusion Matrix

```{r 05-nb-none-confmat, echo=TRUE}
cm_df_nb_none <- conf_mat(
  nb_preds_none,
  truth = turnover_intention,
  estimate = .pred_class
) %>%
  pluck("table") %>%
  as.data.frame()

cm_df_nb_none$Truth <- factor(cm_df_nb_none$Truth, levels = c("No","Yes"))
cm_df_nb_none$Prediction <- factor(cm_df_nb_none$Prediction, levels = c("No","Yes"))

p <- ggplot(cm_df_nb_none, aes(Prediction, Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#FFF9C4", high = "#F4D03F") +
  labs(title = "Naive Bayes — No Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "nb_none_confusion.png", 6, 5, plot_dir)
```

### ROC Curve

```{r 05-nb-none-roc, echo=TRUE}
roc_df_nb_none <- roc_curve(
  nb_preds_none, 
  truth = turnover_intention, 
  .pred_Yes)

p <- ggplot(roc_df_nb_none, aes(x = 1 - specificity, y = sensitivity)) +
  geom_line(color = "#F4D03F", linewidth = 1.2) +
  geom_abline(linetype = "dashed", color = "gray60") +
  labs(title = "Naive Bayes — No Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "nb_none_roc.png", 6, 5, plot_dir)
```

### Naive Bayes - Smote

```{r 04-nb-smote, message=FALSE}
# Create Upsampling Recipe
recipe_nb_smote <- base_recipe %>%
  step_smote(turnover_intention)

# Build Workflow with Upsampling
workflow_nb_smote <- workflow() %>%
  add_model(nb_model) %>%
  add_recipe(recipe_nb_smote)

# 10-Fold Cross-Validation
set.seed(123)
nb_cv_smote <- fit_resamples(
  workflow_nb_smote,
  resamples = turnover_folds,
  metrics = metric_set(accuracy, precision, recall, f_meas, yardstick::spec, roc_auc)
)

# Final Fit on Full Training Data
nb_fit_smote <- fit(workflow_nb_smote, data = turnover_train_data)

# Predict on Test Set
nb_preds_smote <- predict(nb_fit_smote, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(nb_fit_smote, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(turnover_intention))
```

### Performance Metrics

```{r 04-nb-smote-metrics, echo=TRUE}
nb_smote_metrics <- all_metrics(
  nb_preds_smote,
  truth = turnover_intention,
  estimate = .pred_class,
  .pred_Yes
)

p <- nb_smote_metrics %>%
  filter(.metric %in% c("accuracy","precision","recall","f_meas","spec","roc_auc")) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#F4D03F") +
  labs(title = "Naive Bayes — SMOTE") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "nb_smote_metrics.png", 8, 5, plot_dir)
```

### Confusion Matrix

```{r 04-nb-smote-matrix, echo=TRUE}
cm_df_nb_smote <- conf_mat(
  nb_preds_smote,
  truth = turnover_intention,
  estimate = .pred_class
) %>%
  pluck("table") %>%
  as.data.frame()

cm_df_nb_smote$Truth <- factor(cm_df_nb_smote$Truth, levels = c("No","Yes"))
cm_df_nb_smote$Prediction <- factor(cm_df_nb_smote$Prediction, levels = c("No","Yes"))

p <- ggplot(cm_df_nb_smote, aes(Prediction, Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), fontface = "bold") +
  scale_fill_gradient(low = "#FFF9C4", high = "#F4D03F") +
  labs(title = "Naive Bayes — SMOTE") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "nb_smote_confusion.png", 6, 5, plot_dir)
```

### ROC Curve

```{r 05-nb-smote-roc, echo=TRUE}
roc_df_nb_smote <- roc_curve(
  nb_preds_smote, 
  truth = turnover_intention, 
  .pred_Yes)

p <- ggplot(roc_df_nb_smote, aes(x = 1 - specificity, y = sensitivity)) +
  geom_line(color = "#F4D03F", linewidth = 1.2) +
  geom_abline(linetype = "dashed") +
  labs(title = "Naive Bayes — SMOTE") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "nb_smote_roc.png", 6, 5, plot_dir)
```

## Naive Bayes - Up Sampling

```{r 04-nb-upsample}
# Create Up Sampling Recipe
recipe_nb_upsample <- base_recipe %>%
  step_upsample(turnover_intention)

# Build Workflow with Up Sampling
workflow_nb_upsample <- workflow() %>%
  add_model(nb_model) %>%
  add_recipe(recipe_nb_upsample)

# 10-Fold Cross-Validation
set.seed(123)
nb_cv_upsample <- fit_resamples(
  workflow_nb_upsample,
  resamples = turnover_folds,
  metrics = metric_set(accuracy, precision, recall, f_meas, yardstick::spec, roc_auc)
)

# Final Fit on Full Training Data
nb_fit_upsample <- fit(workflow_nb_upsample, data = turnover_train_data)

# Predict on Test Set
nb_preds_upsample <- predict(nb_fit_upsample, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(nb_fit_upsample, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(turnover_intention))
```

### Performance Metrics

```{r 04-nb-upsample-metrics, echo=TRUE}
nb_upsample_metrics <- all_metrics(
  nb_preds_upsample,
  truth = turnover_intention,
  estimate = .pred_class,
  .pred_Yes
)

p <- nb_upsample_metrics %>%
  filter(.metric %in% c("accuracy","precision","recall","f_meas","spec","roc_auc")) %>%
  mutate(.metric = factor(
    .metric,
    levels = c("accuracy","precision","recall","f_meas","spec","roc_auc"),
    labels = c("Accuracy","Precision","Recall","F1 Score","Specificity","ROC AUC")
  )) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#F4D03F", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0,1), expand = expansion(mult = c(0,0.15))) +
  labs(title = "Naive Bayes — Up Sampling", x = "Score") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "nb_up_metrics.png", 8, 5, plot_dir)
```

### Confusion Matrix

```{r 04-nb-upsample-confmat, echo=TRUE}
cm_df_nb_up <- conf_mat(
  nb_preds_upsample,
  truth = turnover_intention,
  estimate = .pred_class
) %>%
  pluck("table") %>%
  as.data.frame()

cm_df_nb_up$Truth <- factor(cm_df_nb_up$Truth, levels = c("No","Yes"))
cm_df_nb_up$Prediction <- factor(cm_df_nb_up$Prediction, levels = c("No","Yes"))

p <- ggplot(cm_df_nb_up, aes(Prediction, Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), fontface = "bold") +
  scale_fill_gradient(low = "#FFF9C4", high = "#F4D03F") +
  labs(title = "Naive Bayes — Up Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "nb_up_confusion.png", 6, 5, plot_dir)
```

### ROC Curve

```{r 04-nb-upsample-roc-curve, echo=TRUE}
roc_df_nb_up <- roc_curve(
  nb_preds_upsample, 
  truth = turnover_intention, 
  .pred_Yes)

p <- ggplot(roc_df_nb_up, aes(x = 1 - specificity, y = sensitivity)) +
  geom_line(color = "#F4D03F", linewidth = 1.2) +
  geom_abline(linetype = "dashed", color = "gray60") +
  labs(title = "Naive Bayes — Up Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "nb_up_roc.png", 6, 5, plot_dir)
```

### Naive Bayes - Down Sampling

```{r 04-nb-down-sample, message=FALSE}
# Create Down Sampling Recipe
recipe_nb_downsample <- base_recipe %>%
  step_downsample(turnover_intention)

# Build Workflow with Down Sampling
workflow_nb_downsample <- workflow() %>%
  add_model(nb_model) %>%
  add_recipe(recipe_nb_downsample)

# 10-Fold Cross-Validation
set.seed(123)
nb_cv_downsample <- fit_resamples(
  workflow_nb_downsample,
  resamples = turnover_folds,
  metrics = metric_set(accuracy, precision, recall, f_meas, yardstick::spec, roc_auc)
)

# Fit Final Model on Full Training Data
nb_fit_downsample <- fit(workflow_nb_downsample, data = turnover_train_data)

# Predict on Test Set
nb_preds_downsample <- predict(nb_fit_downsample, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(nb_fit_downsample, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(turnover_intention))
```

### Performance Metrics

```{r 04-nb-down-sample-metrics, echo=TRUE}
nb_downsample_metrics <- all_metrics(
  nb_preds_downsample,
  truth = turnover_intention,
  estimate = .pred_class,
  .pred_Yes
)

p <- nb_downsample_metrics %>%
  filter(.metric %in% c("accuracy","precision","recall","f_meas","spec","roc_auc")) %>%
  mutate(.metric = factor(
    .metric,
    levels = c("accuracy","precision","recall","f_meas","spec","roc_auc"),
    labels = c("Accuracy","Precision","Recall","F1 Score","Specificity","ROC AUC")
  )) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#F4D03F", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0,1), expand = expansion(mult = c(0,0.15))) +
  labs(title = "Naive Bayes — Down Sampling", x = "Score") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "nb_down_metrics.png", 8, 5, plot_dir)
```

### Confusion Matrix

```{r 04-nb-down-sample-confmat, echo=TRUE}
cm_df_nb_down <- conf_mat(
  nb_preds_downsample,
  truth = turnover_intention,
  estimate = .pred_class
) %>%
  pluck("table") %>%
  as.data.frame()

cm_df_nb_down$Truth <- factor(cm_df_nb_down$Truth, levels = c("No","Yes"))
cm_df_nb_down$Prediction <- factor(cm_df_nb_down$Prediction, levels = c("No","Yes"))

p <- ggplot(cm_df_nb_down, aes(Prediction, Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), fontface = "bold") +
  scale_fill_gradient(low = "#FFF9C4", high = "#F4D03F") +
  labs(title = "Naive Bayes — Down Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "nb_down_confusion.png", 6, 5, plot_dir)
```

### ROC Curve

```{r 04-nb-down-sample-roc-curve, echo=TRUE}
roc_df_nb_down <- roc_curve(
  nb_preds_downsample, 
  truth = turnover_intention, 
  .pred_Yes)

p <- ggplot(roc_df_nb_down, aes(x = 1 - specificity, y = sensitivity)) +
  geom_line(color = "#F4D03F", linewidth = 1.2) +
  geom_abline(linetype = "dashed", color = "gray60") +
  labs(title = "Naive Bayes — Down Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "nb_down_roc.png", 6, 5, plot_dir)
```

### Results Table for Naive Bayes

```{r 04-nb-all-metrics, echo=TRUE}
nb_all_metrics <- bind_rows(
  nb_none_metrics %>% mutate(Resample = "No Sampling"),
  nb_smote_metrics %>% mutate(Resample = "SMOTE"),
  nb_upsample_metrics %>% mutate(Resample = "Up Sampling"),
  nb_downsample_metrics %>% mutate(Resample = "Down Sampling")
)

plot_model_metrics(
  nb_all_metrics,
  "Naive Bayes Performance by Resampling Method",
  "nb_all_metrics.png"
)
```

### Executive Summary for Naive Bayes (Turnover Intention)

Naive Bayes demonstrated surprisingly strong recall and F1 performance under certain sampling strategies, despite its simplifying assumptions. While it did not outperform complex models across the board, it showed value in high-recall use cases — especially when catching at-risk employees is critical.

- **SMOTE** yielded **perfect recall** (**1.000**) and the highest **F1 score** (**0.628**), making it the best choice for detecting turnover intentions without missing positive cases. However, **specificity was 0.000**, meaning many false positives were predicted.

- **Up Sampling** closely followed with high **recall (0.957)** and **F1 score (0.618)**, while offering the **highest ROC AUC** (**0.519**) of all sampling techniques — a balanced choice when both sensitivity and ranking performance matter.

- **Down Sampling** offered a moderate compromise with **recall (0.880)**, **F1 score (0.604)**, and **accuracy (0.473)**. Though not a leader, it avoided the extremes of overfitting or missing positives.

- **No Sampling** performed the worst, with **recall (0.033)** and **F1 score (0.061)** nearly zero. Despite its **high specificity (0.972)**, it failed to identify turnover risk effectively.

---

## Support Vector Machines - No Sampling

```{r 04-svm-no-sample, message=FALSE}
# Linear Support Vector Machine
svm_model <- svm_linear(cost = tune()) %>%
  set_engine("kernlab") %>%
  set_mode("classification")

# Workflow
workflow_svm_none <- workflow() %>%
  add_model(svm_model) %>%
  add_recipe(base_recipe)

# Tuning
set.seed(123)

svm_tuned <- tune_grid(
  workflow_svm_none,
  resamples = turnover_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc))

# Finalize
best_svm <- select_best(svm_tuned, metric = "roc_auc")
final_svm_workflow <- finalize_workflow(workflow_svm_none, best_svm)

# Fit
svm_fit_none <- fit(final_svm_workflow, data = turnover_train_data)

# Predict
svm_preds_none <- predict(svm_fit_none, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(svm_fit_none, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(turnover_intention))
```

### Performance Metrics

```{r 04-svm-no-sampling-metrics, echo=TRUE}
svm_none_metrics <- all_metrics(
  svm_preds_none,
  truth = turnover_intention,
  estimate = .pred_class,
  .pred_Yes
)

p <- svm_none_metrics %>%
  filter(.metric %in% c("accuracy","precision","recall","f_meas","spec","roc_auc")) %>%
  mutate(.metric = factor(
    .metric,
    levels = c("accuracy","precision","recall","f_meas","spec","roc_auc"),
    labels = c("Accuracy","Precision","Recall","F1 Score","Specificity","ROC AUC")
  )) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#9370DB", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, fontface = "bold") +
  scale_x_continuous(limits = c(0,1), expand = expansion(mult = c(0,0.15))) +
  labs(title = "SVM — No Sampling", x = "Score") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "svm_none_metrics.png", 8, 5, plot_dir)
```

### Confusion Matrix

```{r 04-svm-matrix-no-sampling, echo=TRUE}
cm_df_svm_none <- conf_mat(
  svm_preds_none,
  truth = turnover_intention,
  estimate = .pred_class
) %>%
  pluck("table") %>%
  as.data.frame()

cm_df_svm_none$Truth <- factor(cm_df_svm_none$Truth, levels = c("No","Yes"))
cm_df_svm_none$Prediction <- factor(cm_df_svm_none$Prediction, levels = c("No","Yes"))

p <- ggplot(cm_df_svm_none, aes(Prediction, Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), fontface = "bold") +
  scale_fill_gradient(low = "#F3E5F5", high = "#9370DB") +
  labs(title = "SVM — No Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "svm_none_confusion.png", 6, 5, plot_dir)
```

### ROC Curve

```{r 04-svm-roc-curve-no-sampling, echo=TRUE}
roc_df_svm_none <- roc_curve(
  svm_preds_none, 
  truth = turnover_intention, 
  .pred_Yes)

p <- ggplot(roc_df_svm_none, aes(x = 1 - specificity, y = sensitivity)) +
  geom_line(color = "#9370DB", linewidth = 1.2) +
  geom_abline(linetype = "dashed", color = "gray60") +
  labs(title = "SVM — No Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "svm_none_roc.png", 6, 5, plot_dir)
```

## Support Vector Machines - Smote

```{r 04-svm-smote, message=FALSE}
# Define the SVM Model
svm_model <- svm_linear(cost = tune()) %>%
  set_engine("kernlab") %>%
  set_mode("classification")

# Create SMOTE Recipe
recipe_svm_smote <- base_recipe %>%
  step_smote(turnover_intention)

# Build Workflow
workflow_svm_smote <- workflow() %>%
  add_model(svm_model) %>%
  add_recipe(recipe_svm_smote)

# Tune with 10-Fold Cross-Validation
set.seed(123)
svm_tuned_smote <- tune_grid(
  workflow_svm_smote,
  resamples = turnover_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
)

# Select Best Model by ROC AUC
best_svm_smote <- select_best(svm_tuned_smote, metric = "roc_auc")

# Finalize Workflow with Best Hyperparameters
final_svm_workflow_smote <- finalize_workflow(workflow_svm_smote, best_svm_smote)

# Fit Final Model to Training Data
svm_fit_smote <- fit(final_svm_workflow_smote, data = turnover_train_data)

# Predict on Test Data
svm_preds_smote <- predict(svm_fit_smote, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(svm_fit_smote, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(turnover_intention))
```

### Performance Metrics

```{r 04-svm-smote-metrics, echo=TRUE}
svm_smote_metrics <- all_metrics(
  svm_preds_smote,
  truth = turnover_intention,
  estimate = .pred_class,
  .pred_Yes
)

p <- svm_smote_metrics %>%
  filter(.metric %in% c("accuracy","precision","recall","f_meas","spec","roc_auc")) %>%
  mutate(.metric = factor(
    .metric,
    levels = c("accuracy","precision","recall","f_meas","spec","roc_auc"),
    labels = c("Accuracy","Precision","Recall","F1 Score","Specificity","ROC AUC")
  )) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#9370DB", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, fontface = "bold") +
  labs(title = "SVM — SMOTE") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "svm_smote_metrics.png", 8, 5, plot_dir)
```

### Confusion Matrix

```{r 04-svm-matrix-smote, echo=TRUE}
cm_df_svm_smote <- conf_mat(
  svm_preds_smote,
  truth = turnover_intention,
  estimate = .pred_class
) %>%
  pluck("table") %>%
  as.data.frame()

cm_df_svm_smote$Truth <- factor(cm_df_svm_smote$Truth, levels = c("No","Yes"))
cm_df_svm_smote$Prediction <- factor(cm_df_svm_smote$Prediction, levels = c("No","Yes"))

p <- ggplot(cm_df_svm_smote, aes(Prediction, Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), fontface = "bold") +
  scale_fill_gradient(low = "#F3E5F5", high = "#9370DB") +
  labs(title = "SVM — SMOTE") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "svm_smote_confusion.png", 6, 5, plot_dir)
```

### ROC Curve

```{r 04-svm-roc-curve-smote, echo=TRUE}
roc_df_svm_smote <- roc_curve(
  svm_preds_smote,
  truth = turnover_intention,
  .pred_Yes
)

p <- ggplot(roc_df_svm_smote, aes(x = 1 - specificity, y = sensitivity)) +
  geom_line(color = "#9370DB", linewidth = 1.2) +
  geom_abline(linetype = "dashed", color = "gray60") +
  labs(title = "SVM — SMOTE") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "svm_smote_roc.png", 6, 5, plot_dir)
```

## Support Vector Machines - Up Sampling

```{r 04-svm-upsampling, message=FALSE}
# Define the SVM Model
svm_model <- svm_linear(cost = tune()) %>%
  set_engine("kernlab") %>%
  set_mode("classification")

# Create SMOTE Recipe
recipe_svm_upsample <- base_recipe %>%
  step_upsample(turnover_intention)

# Build Workflow
workflow_svm_upsample <- workflow() %>%
  add_model(svm_model) %>%
  add_recipe(recipe_svm_upsample)

# Tune with 10-Fold Cross-Validation
set.seed(123)
svm_tuned_upsample <- tune_grid(
  workflow_svm_upsample,
  resamples = turnover_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
)

# Select Best Model by ROC AUC
best_svm_upsample <- select_best(svm_tuned_upsample, metric = "roc_auc")

# Finalize Workflow with Best Hyperparameters
final_svm_workflow_upsample <- finalize_workflow(workflow_svm_upsample, best_svm_upsample)

# Fit Final Model to Training Data
svm_fit_upsample <- fit(final_svm_workflow_upsample, data = turnover_train_data)

# Predict on Test Data
svm_preds_upsample <- predict(svm_fit_upsample, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(svm_fit_upsample, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(turnover_intention))
```

### Performance Metrics

```{r 04-svm-upsample-metrics, echo=TRUE}
svm_upsample_metrics <- all_metrics(
  svm_preds_upsample,
  truth = turnover_intention,
  estimate = .pred_class,
  .pred_Yes
)

p <- svm_upsample_metrics %>%
  filter(.metric %in% c("accuracy","precision","recall","f_meas","spec","roc_auc")) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#9370DB") +
  labs(title = "SVM — Up Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "svm_up_metrics.png", 8, 5, plot_dir)
```

### Confusion Matrix

```{r 04-svm-matrix-upsample, echo=TRUE}
cm_df_svm_up <- conf_mat(
  svm_preds_upsample,
  truth = turnover_intention,
  estimate = .pred_class
) %>%
  pluck("table") %>%
  as.data.frame()

cm_df_svm_up$Truth <- factor(cm_df_svm_up$Truth, levels = c("No","Yes"))
cm_df_svm_up$Prediction <- factor(cm_df_svm_up$Prediction, levels = c("No","Yes"))

p <- ggplot(cm_df_svm_up, aes(Prediction, Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), fontface = "bold") +
  scale_fill_gradient(low = "#F3E5F5", high = "#9370DB") +
  labs(title = "SVM — Up Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "svm_up_confusion.png", 6, 5, plot_dir)
```

### ROC Curve

```{r 04-svm-roc-curve-upsample, echo=TRUE}
roc_df_svm_up <- roc_curve(
  svm_preds_upsample, 
  truth = turnover_intention, 
  .pred_Yes)

p <- ggplot(roc_df_svm_up, aes(x = 1 - specificity, y = sensitivity)) +
  geom_line(color = "#9370DB", linewidth = 1.2) +
  geom_abline(linetype = "dashed") +
  labs(title = "SVM — Up Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "svm_up_roc.png", 6, 5, plot_dir)
```

## Support Vector Machines - Down Sampling

```{r 04-svm-downsampling, message=FALSE}
# Define the SVM Model
svm_model <- svm_linear(cost = tune()) %>%
  set_engine("kernlab") %>%
  set_mode("classification")

# Create SMOTE Recipe
recipe_svm_downsample <- base_recipe %>%
  step_downsample(turnover_intention)

# Build Workflow
workflow_svm_downsample <- workflow() %>%
  add_model(svm_model) %>%
  add_recipe(recipe_svm_downsample)

# Tune with 10-Fold Cross-Validation
set.seed(123)
svm_tuned_downsample <- tune_grid(
  workflow_svm_downsample,
  resamples = turnover_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
)

# Select Best Model by ROC AUC
best_svm_downsample <- select_best(svm_tuned_downsample, metric = "roc_auc")

# Finalize Workflow with Best Hyperparameters
final_svm_workflow_downsample <- finalize_workflow(workflow_svm_downsample, best_svm_downsample)

# Fit Final Model to Training Data
svm_fit_downsample <- fit(final_svm_workflow_downsample, data = turnover_train_data)

# Predict on Test Data
svm_preds_downsample <- predict(svm_fit_downsample, new_data = turnover_test_data, type = "prob") %>%
  bind_cols(predict(svm_fit_downsample, new_data = turnover_test_data)) %>%
  bind_cols(turnover_test_data %>% select(turnover_intention))
```

### Performance Metrics

```{r 04-svm-downsampling-metrics, echo=TRUE}
svm_downsample_metrics <- all_metrics(
  svm_preds_downsample,
  truth = turnover_intention,
  estimate = .pred_class,
  .pred_Yes
)

p <- svm_downsample_metrics %>%
  filter(.metric %in% c("accuracy","precision","recall","f_meas","spec","roc_auc")) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#9370DB") +
  labs(title = "SVM — Down Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "svm_down_metrics.png", 8, 5, plot_dir)
```

### Confusion Matrix

```{r 04-svm-matrix-downsample, echo=TRUE}
cm_df_svm_down <- conf_mat(
  svm_preds_downsample,
  truth = turnover_intention,
  estimate = .pred_class
) %>%
  pluck("table") %>%
  as.data.frame()

cm_df_svm_down$Truth <- factor(cm_df_svm_down$Truth, levels = c("No","Yes"))
cm_df_svm_down$Prediction <- factor(cm_df_svm_down$Prediction, levels = c("No","Yes"))

p <- ggplot(cm_df_svm_down, aes(Prediction, Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), fontface = "bold") +
  scale_fill_gradient(low = "#F3E5F5", high = "#9370DB") +
  labs(title = "SVM — Down Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "svm_down_confusion.png", 6, 5, plot_dir)
```

### ROC Curve

```{r 04-svm-roc-curve-downsample, echo=TRUE}
roc_df_svm_down <- roc_curve(
  svm_preds_downsample, 
  truth = turnover_intention, 
  .pred_Yes)

p <- ggplot(roc_df_svm_down, aes(x = 1 - specificity, y = sensitivity)) +
  geom_line(color = "#9370DB", linewidth = 1.2) +
  geom_abline(linetype = "dashed") +
  labs(title = "SVM — Down Sampling") +
  theme_hr(base_size = 13)

print(p)
save_plot(p, "svm_down_roc.png", 6, 5, plot_dir)
```

### Results Table for Support Vector Machines

```{r 04-svm-all-metrics, echo=TRUE}
svm_all_metrics <- bind_rows(
  svm_none_metrics %>% mutate(Resample = "No Sampling"),
  svm_smote_metrics %>% mutate(Resample = "SMOTE"),
  svm_upsample_metrics %>% mutate(Resample = "Up Sampling"),
  svm_downsample_metrics %>% mutate(Resample = "Down Sampling")
)

plot_model_metrics(
  svm_all_metrics,
  "Support Vector Machine Performance by Resampling Method",
  "svm_all_metrics.png"
)
```

### Executive Summary for Turnover Intention (Support Vector Machines)

Support Vector Machines (SVM) offered robust, consistent performance across all resampling methods, excelling in recall and F1 scores when paired with SMOTE. Though SVMs are sensitive to class imbalance, they proved effective when supported by proper resampling.

- **SMOTE** delivered the highest recall (**0.538**) and F1 score (**0.490**), making it the top choice for identifying true turnover cases. It also posted strong performance in ROC AUC (**0.478**), reinforcing its ability to rank cases effectively.

- **Up Sampling** achieved balanced results, with precision (**0.478**) and recall (**0.478**) closely trailing SMOTE. Its F1 score (**0.478**) confirmed a reliable overall trade-off between precision and recall.

- **Down Sampling** held stable recall (**0.478**) and F1 (**0.478**) but lagged behind on specificity (**0.560**) and ROC AUC (**0.437**), reflecting the limitations of reduced data volume during training.

- **No Sampling** underperformed relative to other strategies, with the lowest recall (**0.380**) and F1 score (**0.432**). Although specificity remained high (**0.679**), the model struggled to detect minority class instances due to imbalance.

---

```{=html}
<div class="takeaway-box">
<strong>Model Selection Insight:</strong> While multiple models demonstrated strong performance, SVM (SMOTE) and Random Forest (Up Sampling) emerged as the most effective depending on business priorities—highlighting the importance of aligning model selection with organizational risk tolerance.
</div>
```

```{r 04-all-models-metrics-data, echo=FALSE}
final_metrics <- bind_rows(
  logit_all_metrics %>% mutate(Model = "Logistic"),
  rf_all_metrics %>% mutate(Model = "Random Forest"),
  xgb_all_metrics %>% mutate(Model = "XGBoost"),
  nb_all_metrics %>% mutate(Model = "Naive Bayes"),
  svm_all_metrics %>% mutate(Model = "SVM")
)
```

### Model Performance Comparison

The following visualization compares all model configurations across key evaluation metrics. 
Highlighted bars indicate the best-performing model for each metric, emphasizing trade-offs between precision, recall, and overall classification performance.

```{r 04-derived-comparison-objects, echo=FALSE}
model_colors <- c(
  "Logistic" = "#5FA8D3",
  "Random Forest" = "#2A9D8F",
  "XGBoost" = "#F4A261",
  "Naive Bayes" = "#D7263D",
  "SVM" = "#6C8CD5"
)

all_vip <- bind_rows(
  vip_logit_none %>% mutate(Model = "Logistic", Resample = "No Sampling"),
  vip_logit_smote %>% mutate(Model = "Logistic", Resample = "SMOTE"),
  vip_logit_upsample %>% mutate(Model = "Logistic", Resample = "Up Sampling"),
  vip_logit_down %>% mutate(Model = "Logistic", Resample = "Down Sampling"),

  vip_rf_none %>% mutate(Model = "Random Forest", Resample = "No Sampling"),
  vip_rf_smote %>% mutate(Model = "Random Forest", Resample = "SMOTE"),
  vip_rf_up %>% mutate(Model = "Random Forest", Resample = "Up Sampling"),
  vip_rf_down %>% mutate(Model = "Random Forest", Resample = "Down Sampling"),

  vip_xgb_none %>% mutate(Model = "XGBoost", Resample = "No Sampling"),
  vip_xgb_smote %>% mutate(Model = "XGBoost", Resample = "SMOTE"),
  vip_xgb_up %>% mutate(Model = "XGBoost", Resample = "Up Sampling"),
  vip_xgb_down %>% mutate(Model = "XGBoost", Resample = "Down Sampling")
)

top_vars <- all_vip %>%
  group_by(Model, Resample) %>%
  slice_max(order_by = Importance, n = 5, with_ties = FALSE) %>%
  mutate(ScaledImportance = Importance / max(Importance)) %>%
  ungroup()

top5_model_scores <- final_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(
    Metric = recode(
      .metric,
      "accuracy" = "Accuracy",
      "precision" = "Precision",
      "recall" = "Recall",
      "f_meas" = "F1 Score",
      "spec" = "Specificity",
      "roc_auc" = "ROC AUC"
    ),
    Model_Config = paste(Model, Resample, sep = " — ")
  ) %>%
  group_by(Metric) %>%
  slice_max(order_by = .estimate, n = 5, with_ties = FALSE) %>%
  ungroup() %>%
  rename(Score = .estimate)
```

```{r 04-model-accuracy-comparison, fig.width=14, fig.height=8}
p <- final_metrics %>%
  
  filter(.metric %in% c("accuracy","precision","recall","f_meas","spec","roc_auc")) %>%
  
  mutate(
    .metric = factor(
      .metric,
      levels = c("accuracy","precision","recall","f_meas","spec","roc_auc"),
      labels = c("Accuracy","Precision","Recall","F1 Score","Specificity","ROC AUC")
    ),
    Model_Config = paste(Model, Resample, sep = " — ")
  ) %>%
  
  group_by(.metric) %>%
  mutate(best = .estimate == max(.estimate)) %>%
  ungroup() %>%
  
  ggplot(aes(x = Model_Config, y = .estimate, fill = Model)) +
  
  geom_col(width = 0.65) +
  
  geom_col(
    data = ~ dplyr::filter(.x, best),
    fill = NA,
    color = "black",
    linewidth = 1.1,
    width = 0.65
  ) +
  
  facet_wrap(~ .metric, scales = "free_y") +
  
  scale_fill_manual(values = model_colors) +
  
  labs(
    title = "Final Model Comparison Across All Methods",
    subtitle = "Best-performing model per metric highlighted",
    x = "Model Configuration",
    y = "Score"
  ) +
  
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    strip.text = element_text(face = "bold"),
    axis.text.x = element_text(angle = 30, hjust = 1),
    legend.position = "none"
  )

print(p)
save_plot(p, "final_model_comparison.png", 14, 8, plot_dir)
```

## Turnover Intention - Master Variable Importance Plot

```{r 04-vip-imp-clean-horizontal, fig.width=14, fig.height=10}
p <- ggplot(top_vars, aes(
  x = ScaledImportance,
  y = fct_reorder(Variable, ScaledImportance),
  fill = Model
)) +
  
  geom_col(width = 0.7, show.legend = FALSE) +
  
  geom_text(
    aes(label = round(Importance, 2)),
    hjust = -0.2,
    size = 3
  ) +
  
  facet_grid(Resample ~ Model, scales = "free_y", space = "free") +
  
  scale_fill_manual(values = model_colors) +
  
  coord_cartesian(clip = "off") +
  
  labs(
    title = "Top Five Most Important Features by Model and Resampling Method",
    x = "Scaled Importance",
    y = "Variable"
  ) +
  
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    strip.text = element_text(face = "bold"),
    panel.spacing = unit(1, "lines")
  )

print(p)
save_plot(p, "master_variable_importance.png", 14, 10, plot_dir)
```

## Turnover Intention - Top Five Performing Models Plot

```{r 04-best-model, fig.width=14, fig.height=10}
  p <- top5_model_scores %>%
  
  mutate(
    Metric = factor(
      Metric,
      levels = c("Accuracy","Precision","Recall","F1 Score","Specificity","ROC AUC")
    ),
    
    Model_Base = gsub(" — .*", "", Model_Config)
  ) %>%
  
  ggplot(aes(x = Model_Config, y = Score, fill = Model_Base)) +
  
  geom_col(width = 0.65) +
  
  geom_text(
    aes(label = sprintf("%.3f", Score)),
    vjust = -0.3,
    size = 3
  ) +
  
  facet_wrap(~ Metric, scales = "free_y") +
  
  scale_fill_manual(values = model_colors) +
  
  labs(
    title = "Top Performing Model Configurations by Metric",
    x = "Model Configuration",
    y = "Score"
  ) +
  
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text.x = element_text(angle = 30, hjust = 1),
    strip.text = element_text(face = "bold"),
    legend.position = "none"
  )

print(p)
save_plot(p, "top5_model_comparison.png", 14, 10, plot_dir)
```

```{=html}
<div class="takeaway-box">

<strong>Top Model Insights:</strong>

- <strong>Naive Bayes (SMOTE)</strong> achieved the highest recall, making it highly effective for identifying potential turnover cases, though with an increased risk of false positives.

- <strong>SVM (SMOTE)</strong> provided the strongest overall balance across performance metrics, making it the most reliable general-purpose model for turnover prediction.

- <strong>Random Forest (Up Sampling)</strong> excelled in precision and specificity, making it ideal when minimizing false positives is a priority.

- <strong>XGBoost (Down Sampling)</strong> demonstrated stable, well-balanced performance across all evaluation metrics.

- <strong>Logistic Regression (Up Sampling)</strong> remains a strong, interpretable baseline model, particularly valuable for stakeholder communication.

</div>
```

```{=html}
<hr>

<div style="
  display:grid;
  grid-template-columns: 1fr auto 1fr;
  align-items:center;
  margin:30px 0;
">

  <!-- LEFT -->
  <div style="text-align:left;">
    <a href="HR_Data_EDA.html" style="
      display:inline-block;
      padding:12px 28px;
      border-radius:999px;
      font-weight:600;
      color:white;
      text-decoration:none;
      background:linear-gradient(135deg, #74B9FF, #2E86C1);
    ">
      ← Back
    </a>
  </div>

  <!-- CENTER -->
  <div style="text-align:center;">
    <a href="index.html" style="
      display:inline-block;
      padding:12px 28px;
      border-radius:999px;
      font-weight:600;
      color:white;
      text-decoration:none;
      background:linear-gradient(135deg, #2A9D8F, #52B788);
      box-shadow:0 6px 16px rgba(42,157,143,0.35);
    ">
      🏠 Home
    </a>
  </div>

  <!-- RIGHT -->
  <div style="text-align:right;">
    <a href="HR_Data_Quiet_Quitting_Results.html" style="
      display:inline-block;
      padding:12px 28px;
      border-radius:999px;
      font-weight:600;
      color:white;
      text-decoration:none;
      background:linear-gradient(135deg, #FF7675, #922B21);
    ">
      Next →
    </a>
  </div>

</div>

</div>
```