---
title: "Insights and Recommendations | HR Workforce Analytics"
output:
  html_document:
    always_allow_html: true
    self_contained: false
    toc: false
    toc_float: false
    theme: null
    highlight: null
    df_print: paged
    fig_caption: true
    css: assets/styles.css
number_sections: false
---

```{=html}
<div class="report-container">

  <div class="page-layout">

    <!-- SIDEBAR -->
    <div class="sidebar">
      <div class="sidebar-card">
        <div class="sidebar-label">On this page</div>
        <nav class="sidebar-nav">
          <a href="#executive-summary">Executive Summary</a>
          <a href="#modeling-results">Portfolio-Level Modeling Results</a>
          <a href="#workforce-patterns">Cross-Model Workforce Patterns</a>
          <a href="#strategic-recommendations">Strategic Workforce Recommendations</a>
          <a href="#responsible-use">Responsible Use & Governance</a>
          <a href="#executive-action-roadmap">Roadmap Timeline View</a>
          <a href="#final-takeaway">Final Takeaway</a>
        </nav>
      </div>
    </div>

    <!-- MAIN CONTENT -->
    <div class="page-content">

      <!-- MATCH EDA STRUCTURE -->
      <div class="hero-section">
        <p class="hero-meta">
          Kevin Allen Cauley • `r format(Sys.Date(), '%B %d, %Y')`
        </p>

        <h1 class="hero-title">Executive Insights: Workforce Risk Modeling Results</h1>

        <p class="hero-subtitle">
        This section synthesizes the final modeling results across turnover intention, quiet quitting, and monitoring-related turnover risk, translating model performance into executive-level workforce insights, governance guidance, and practical HR action priorities.
        </p>
      </div>

      <!-- REPLACED page-nav WITH cta-row -->
      <div class="cta-row">
        <a href="hr_data_monitoring_privacy.html" class="cta-btn btn-blue">← Back</a>
        <a href="index.html" class="cta-btn btn-teal">Home</a>
      </div>
```

```{r 00-setup, include=FALSE}
source(here::here("analysis", "helper.R"))

PLOT_SUBFOLDER <- "insights"

cat("Using plot folder:", PLOT_SUBFOLDER, "\n")

# Required Packages
required_pkgs <- c(
  "conflicted", "tidyverse", "here", "gt", "scales", "glue"
)

for (pkg in required_pkgs) {
  if (!pkg %in% rownames(installed.packages())) {
    install.packages(pkg)
  }
}

library(conflicted)
library(tidyverse)
library(here)
library(gt)
library(scales)
library(glue)

conflicted::conflicts_prefer(
  dplyr::filter,
  dplyr::select,
  dplyr::mutate,
  dplyr::summarise,
  dplyr::arrange,
  dplyr::slice,
  dplyr::slice_head,
  dplyr::slice_max,
  dplyr::slice_min,
  dplyr::recode,
  dplyr::rename,
  dplyr::lag
)

# Global Options
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  results = "hide",
  fig.align = "center",
  out.width = "100%",
  fig.width = 9.5,
  fig.height = 5.8,
  fig.retina = 2,
  dpi = 300,
  cache = FALSE,
  dev = "png"
)

set.seed(123)

# Load final model summaries exported from each modeling report
summary_dir <- here::here("data", "processed", "model_summaries")

summary_files <- c(
  "turnover_model_summary.rds",
  "quiet_quitting_model_summary.rds",
  "monitoring_privacy_model_summary.rds"
)

missing_files <- summary_files[!file.exists(file.path(summary_dir, summary_files))]

if (length(missing_files) > 0) {
  stop(
    "Missing model summary file(s): ",
    paste(missing_files, collapse = ", "),
    ". Re-render the modeling reports before rendering Insights Gleaned."
  )
}

model_summaries <- purrr::map_dfr(
  file.path(summary_dir, summary_files),
  readRDS
) %>%
  mutate(
    Outcome = factor(
      Outcome,
      levels = c(
        "Turnover Intention",
        "Quiet Quitting",
        "Monitoring & Privacy Risk"
      )
    ),
    Model_Config = paste(Model, Sampling, sep = " + "),
    across(
      c(
        CV_F1,
        Holdout_Accuracy,
        Holdout_Precision,
        Holdout_Recall,
        Holdout_F1,
        Holdout_ROC_AUC
      ),
      as.numeric
    )
  ) %>%
  arrange(Outcome)

best_holdout_model <- model_summaries %>%
  dplyr::arrange(dplyr::desc(Holdout_F1)) %>%
  utils::head(1)

highest_precision_model <- model_summaries %>%
  dplyr::arrange(dplyr::desc(Holdout_Precision)) %>%
  utils::head(1)

highest_recall_model <- model_summaries %>%
  dplyr::arrange(dplyr::desc(Holdout_Recall)) %>%
  utils::head(1)

lowest_holdout_model <- model_summaries %>%
  dplyr::arrange(Holdout_F1) %>%
  utils::head(1)

outcome_colors <- c(
  "Turnover Intention" = "#4B9CD3",
  "Quiet Quitting" = "#2A9D8F",
  "Monitoring & Privacy Risk" = "#E87461"
)

outcome_light_colors <- c(
  "Turnover Intention" = "#D6EAF8",
  "Quiet Quitting" = "#D5F5E3",
  "Monitoring & Privacy Risk" = "#FADBD8"
)

metric_order <- c(
  "Accuracy",
  "Precision",
  "Recall",
  "F1 Score",
  "ROC AUC"
)

model_summaries <- model_summaries %>%
  mutate(
    Outcome_Label = as.character(Outcome),
    Outcome_Label = factor(
      Outcome_Label,
      levels = c(
        "Turnover Intention",
        "Quiet Quitting",
        "Monitoring & Privacy Risk"
      )
    )
  )

# Consistent portfolio colors
portfolio_colors <- c(
  "Turnover Intention" = "#5AA0D1",
  "Quiet Quitting" = "#45A79C",
  "Monitoring & Privacy Risk" = "#E67E6B"
)

outline_color <- "#223247"
subtitle_color <- "#5a6b7c"
grid_color <- "#D5DBDB"

# -------------------------------------------------------------------
# Executive summary helper objects
# -------------------------------------------------------------------

required_summary_cols <- c(
  "Outcome",
  "Model_Config",
  "Holdout_F1",
  "Holdout_Precision",
  "Holdout_Recall"
)

missing_summary_cols <- setdiff(required_summary_cols, names(model_summaries))

if (length(missing_summary_cols) > 0) {
  stop(
    "Missing required model_summaries columns: ",
    paste(missing_summary_cols, collapse = ", ")
  )
}

strongest_holdout_model <- model_summaries %>%
  arrange(desc(Holdout_F1)) %>%
  slice(1)

highest_precision_model <- model_summaries %>%
  arrange(desc(Holdout_Precision)) %>%
  slice(1)

highest_recall_model <- model_summaries %>%
  arrange(desc(Holdout_Recall)) %>%
  slice(1)

lowest_holdout_model <- model_summaries %>%
  arrange(Holdout_F1) %>%
  slice(1)
```

# Executive Summary {#executive-summary}

This final section brings the HR workforce analytics portfolio together by translating the modeling results into executive-level insight.

Across the three modeling tracks, the project evaluates **turnover intention**, **quiet quitting**, and **monitoring-related turnover risk** using a consistent modeling framework. Together, these outcomes show that workforce risk is not one-dimensional. Each target reflects a different business problem, requires a different interpretation strategy, and supports a different type of decision.

The strongest holdout model was **`r strongest_holdout_model$Model_Config`** for **`r strongest_holdout_model$Outcome`**, with a holdout F1 Score of **`r round(strongest_holdout_model$Holdout_F1, 3)`**. The strongest precision result was **`r highest_precision_model$Model_Config`** for **`r highest_precision_model$Outcome`**, indicating the most reliable review queue among flagged cases.

The biggest project insight is not that one model “wins.” The more important takeaway is that each model supports a different business use case:

- **Turnover Intention** is best used for retention-risk prioritization.
- **Quiet Quitting** is best used as a cautious early engagement signal.
- **Monitoring & Privacy Risk** is best used for privacy-aware governance and workforce trust review.

These models should be used as **risk prioritization tools**, not automated decision systems. Their value is in helping HR, leadership, and governance teams focus attention earlier, identify patterns across workforce signals, and guide more targeted human review.

## Portfolio-Level Takeaways

```{=html}
<div class="model-highlight-card">

  <div class="model-badge">Executive Portfolio Summary</div>

  <h3 class="model-title">
    Three Workforce Risk Models, Three Different Decision Uses
  </h3>

  <p class="model-subtitle">
    The final portfolio shows how predictive modeling can support retention, engagement, and governance review without replacing human judgment.
  </p>

  <div class="model-metrics">
    <div class="metric-box">
      <span class="metric-label">Most Balanced</span>
      <span class="metric-value">
        `r strongest_holdout_model$Outcome`
      </span>
    </div>

    <div class="metric-box">
      <span class="metric-label">Most Reliable Flags</span>
      <span class="metric-value">
        `r highest_precision_model$Outcome`
      </span>
    </div>

    <div class="metric-box">
      <span class="metric-label">Use Case</span>
      <span class="metric-value">
        Human-in-the-Loop Risk Prioritization
      </span>
    </div>
  </div>

</div>
```

```{=html}
<div class="decision-box">

<h3>How to Interpret the Three Outcomes</h3>

<p>
<strong>Turnover intention</strong> is most useful for identifying employees or workforce groups that may require retention conversations, manager follow-up, workload review, or career-pathing support.
</p>

<p>
<strong>Quiet quitting</strong> is more ambiguous because disengagement can be behavioral, emotional, or contextual. It is better suited for engagement monitoring and qualitative follow-up than direct intervention.
</p>

<p>
<strong>Monitoring-related risk</strong> is the clearest governance signal because it connects workforce risk to privacy, trust, surveillance, and policy design.
</p>

</div>
```

# Final Modeling Results {#modeling-results}

The final modeling results provide a portfolio-level view of how each workforce-risk outcome performed after model development, cross-validation, and holdout confirmation.

These outcomes should not be compared as if they represent the same type of problem. Turnover intention, quiet quitting, and monitoring-related turnover risk differ in signal strength, business sensitivity, interpretability, and practical use.

The goal of this section is to translate model performance into executive decision guidance: which outcomes are most reliable, which require caution, and how each model should be used responsibly.

```{r 01-final-model-summary-table, echo=FALSE}
model_summaries %>%
  mutate(
    `Selected Model` = Model,
    `Sampling Strategy` = Sampling,
    `Model Configuration` = Model_Config,
    `Selection Strategy` = Selection_Strategy
  ) %>%
  select(
    Outcome,
    `Selected Model`,
    `Sampling Strategy`,
    `Selection Strategy`,
    CV_F1,
    Holdout_Accuracy,
    Holdout_Precision,
    Holdout_Recall,
    Holdout_F1,
    Holdout_ROC_AUC
  ) %>%
  gt() %>%
  tab_header(
    title = "Final Selected Models Across Workforce Outcomes",
    subtitle = "Cross-validated selection with holdout performance confirmation"
  ) %>%
  cols_label(
    CV_F1 = "CV F1",
    Holdout_Accuracy = "Accuracy",
    Holdout_Precision = "Precision",
    Holdout_Recall = "Recall",
    Holdout_F1 = "F1",
    Holdout_ROC_AUC = "ROC AUC"
  ) %>%
  fmt_number(
    columns = c(
      CV_F1,
      Holdout_Accuracy,
      Holdout_Precision,
      Holdout_Recall,
      Holdout_F1,
      Holdout_ROC_AUC
    ),
    decimals = 3
  ) %>%
  cols_align("center") %>%
  data_color(
    columns = c(
      CV_F1,
      Holdout_Accuracy,
      Holdout_Precision,
      Holdout_Recall,
      Holdout_F1,
      Holdout_ROC_AUC
    ),
    colors = scales::col_numeric(
      palette = c("#EAF2F8", "#A9CCE3", "#2E86C1"),
      domain = c(0, 1)
    )
  ) %>%
  tab_options(
    table.width = pct(95),
    table.font.size = px(12)
  )
```

## Executive Model Scorecard

```{r 01b-executive-model-scorecard, echo=FALSE}
scorecard_long <- model_summaries %>%
  mutate(
    Outcome = factor(
      as.character(Outcome),
      levels = c(
        "Turnover Intention",
        "Quiet Quitting",
        "Monitoring & Privacy Risk"
      )
    )
  ) %>%
  transmute(
    Outcome,
    `F1 Score` = Holdout_F1,
    Precision = Holdout_Precision,
    Recall = Holdout_Recall
  ) %>%
  tidyr::pivot_longer(
    cols = c(`F1 Score`, Precision, Recall),
    names_to = "Metric",
    values_to = "Score"
  ) %>%
  mutate(
    Metric = factor(
      Metric,
      levels = c("F1 Score", "Precision", "Recall")
    ),
    Outcome = forcats::fct_rev(Outcome),
    Score_Label = scales::number(Score, accuracy = 0.001)
  )

p <- ggplot(
  scorecard_long,
  aes(
    x = Score,
    y = Outcome,
    color = Outcome
  )
) +
  geom_segment(
    aes(
      x = 0,
      xend = Score,
      yend = Outcome
    ),
    linewidth = 5.2,
    alpha = 0.24,
    show.legend = FALSE
  ) +
  geom_point(
    size = 5.4,
    show.legend = FALSE
  ) +
  geom_text(
    aes(label = Score_Label),
    hjust = -0.22,
    size = 3.8,
    fontface = "bold",
    color = outline_color,
    show.legend = FALSE
  ) +
  facet_wrap(
    ~ Metric,
    nrow = 1
  ) +
  scale_color_manual(values = portfolio_colors) +
  scale_x_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, 1.10),
    breaks = c(0, 0.25, 0.50, 0.75, 1.00),
    expand = expansion(mult = c(0.01, 0.02))
  ) +
  labs(
    title = "Executive Model Scorecard",
    subtitle = "Selected models compared by balance, reliability, and detection coverage",
    x = "Holdout score",
    y = NULL
  ) +
  coord_cartesian(clip = "off") +
  theme_hr(11) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(color = subtitle_color, size = 11.5),
    strip.text = element_text(face = "bold", size = 11.5),
    axis.text.y = element_text(face = "bold", size = 10.5),
    axis.text.x = element_text(size = 9.5),
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    plot.margin = margin(12, 38, 16, 12)
  )

print(p)

save_plot(
  p,
  "executive_model_scorecard.png",
  width = 11.2,
  height = 5.8
)
```

The scorecard gives a compact executive view of how the selected models behave in practice. **F1 Score** represents overall balance, **precision** represents reliability of flagged cases, and **recall** represents detection coverage.

Monitoring & Privacy Risk shows the strongest portfolio profile, especially on precision and overall balance. This makes it most appropriate for a focused governance review queue where false positives could create unnecessary privacy or trust concerns.

Turnover Intention remains a usable retention-prioritization signal because its precision and recall are relatively balanced. It should be used to help HR and managers identify employees or groups that may warrant earlier retention review.

Quiet Quitting is the most caution-sensitive model. Its weaker precision and F1 Score mean it should be treated as an early engagement signal, not a standalone intervention trigger.

## Final Model Comparison

```{r 02-final-model-comparison-heatmap, echo=FALSE}
comparison_long <- model_summaries %>%
  select(
    Outcome,
    Model_Config,
    Holdout_Accuracy,
    Holdout_Precision,
    Holdout_Recall,
    Holdout_F1,
    Holdout_ROC_AUC
  ) %>%
  pivot_longer(
    cols = starts_with("Holdout"),
    names_to = "Metric",
    values_to = "Score"
  ) %>%
  mutate(
    Metric = recode(
      Metric,
      Holdout_Accuracy = "Accuracy",
      Holdout_Precision = "Precision",
      Holdout_Recall = "Recall",
      Holdout_F1 = "F1 Score",
      Holdout_ROC_AUC = "ROC AUC"
    ),
    Metric = factor(Metric, levels = metric_order),
    Outcome = factor(
      Outcome,
      levels = c(
        "Turnover Intention",
        "Quiet Quitting",
        "Monitoring & Privacy Risk"
      )
    ),
    Outcome_Label = paste0(as.character(Outcome), "\n", Model_Config),
    Plot_Label = ifelse(
      is.na(Score),
      "N/A",
      scales::number(Score, accuracy = 0.001)
    )
  )

p <- ggplot(
  comparison_long,
  aes(x = Metric, y = fct_rev(Outcome_Label), fill = Score)
) +
  geom_tile(
    color = "white",
    linewidth = 1.1,
    width = 0.95,
    height = 0.88
  ) +
  geom_text(
    aes(label = Plot_Label),
    fontface = "bold",
    size = 4.1,
    color = "#1F2D3D"
  ) +
  scale_fill_gradient(
    low = "#F4F8FB",
    high = "#1F618D",
    limits = c(0, 1),
    na.value = "#F2F4F4",
    name = "Score"
  ) +
  labs(
    title = "Final Holdout Performance Across Workforce Risk Models",
    subtitle = "Selected models compared using the same final evaluation metrics",
    x = NULL,
    y = NULL
  ) +
  theme_hr(11) +
  theme(
    plot.title = element_text(face = "bold", size = 17),
    plot.subtitle = element_text(color = "#5a6b7c", size = 12, margin = margin(b = 12)),
    axis.text.x = element_text(face = "bold", size = 10),
    axis.text.y = element_text(size = 9.5, lineheight = 0.95),
    legend.position = "right",
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank(),
    plot.margin = margin(12, 22, 12, 10)
  )

print(p)
save_plot(p, "portfolio_holdout_metric_heatmap.png", 11, 6)
```

The portfolio comparison shows that monitoring-related turnover risk produced the strongest overall holdout profile, especially in precision and F1 Score. This suggests the monitoring/privacy model is the most reliable for identifying a focused review queue.

Turnover intention produced moderate but usable holdout performance, making it appropriate for retention-risk prioritization rather than definitive classification. Quiet quitting showed weaker transfer from cross-validation to holdout performance, meaning it should be interpreted more cautiously as an early-warning engagement screen.

```{=html}
<div class="decision-box">

<h3>Portfolio Interpretation</h3>

<p>
The three outcomes should not be ranked as if they represent the same business problem. Each model answers a different workforce question.
</p>

<ul>
  <li><strong>Turnover Intention:</strong> Who may need earlier retention review?</li>
  <li><strong>Quiet Quitting:</strong> Where might disengagement risk be emerging?</li>
  <li><strong>Monitoring & Privacy Risk:</strong> Where might monitoring practices, privacy concerns, or trust signals require governance review?</li>
</ul>

<p>
The strongest portfolio insight is that predictive modeling is most useful when the model output is matched to the right decision context.
</p>

</div>
```

## Prediction Confidence & Practical Use

```{r 03-precision-recall-plot, echo=FALSE}
tradeoff_df <- model_summaries %>%
  mutate(
    Outcome = factor(
      as.character(Outcome),
      levels = c(
        "Turnover Intention",
        "Quiet Quitting",
        "Monitoring & Privacy Risk"
      )
    )
  ) %>%
  transmute(
    Outcome,
    Recall = Holdout_Recall,
    Precision = Holdout_Precision,
    Holdout_F1
  )

label_df <- tradeoff_df %>%
  mutate(
    label_x = case_when(
      Outcome == "Turnover Intention" ~ 0.50,
      Outcome == "Quiet Quitting" ~ 0.37,
      Outcome == "Monitoring & Privacy Risk" ~ 0.48
    ),
    label_y = case_when(
      Outcome == "Turnover Intention" ~ 0.51,
      Outcome == "Quiet Quitting" ~ 0.24,
      Outcome == "Monitoring & Privacy Risk" ~ 0.91
    ),
    label_text = paste0(
      Outcome,
      "\nF1 = ",
      scales::number(Holdout_F1, accuracy = 0.001)
    )
  )

zone_df <- tibble::tribble(
  ~xmin, ~xmax, ~ymin, ~ymax, ~fill,
  0.00, 0.50, 0.50, 1.00, "#EDF4FB",
  0.50, 1.00, 0.50, 1.00, "#EEF7F2",
  0.50, 1.00, 0.00, 0.50, "#F8F2E3",
  0.00, 0.50, 0.00, 0.50, "#F8EEEE"
)

p <- ggplot(tradeoff_df, aes(x = Recall, y = Precision)) +
  geom_rect(
    data = zone_df,
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
    inherit.aes = FALSE,
    fill = zone_df$fill,
    color = NA,
    alpha = 0.95
  ) +
  geom_vline(
    xintercept = 0.50,
    linetype = "dashed",
    color = "#AAB7B8",
    linewidth = 0.7
  ) +
  geom_hline(
    yintercept = 0.50,
    linetype = "dashed",
    color = "#AAB7B8",
    linewidth = 0.7
  ) +
  annotate(
    "label",
    x = 0.22, y = 0.93,
    label = "High-Confidence\nReview Queue",
    fill = "white",
    color = "#24588B",
    label.size = 0.35,
    size = 3.6,
    fontface = "bold"
  ) +
  annotate(
    "label",
    x = 0.79, y = 0.93,
    label = "Balanced\nPrioritization",
    fill = "white",
    color = "#25523B",
    label.size = 0.35,
    size = 3.6,
    fontface = "bold"
  ) +
  annotate(
    "label",
    x = 0.79, y = 0.13,
    label = "Broad Screening\nSignal",
    fill = "white",
    color = "#8A6A00",
    label.size = 0.35,
    size = 3.6,
    fontface = "bold"
  ) +
  annotate(
    "label",
    x = 0.22, y = 0.16,
    label = "Cautious\nValidation Needed",
    fill = "white",
    color = "#7A4E4E",
    label.size = 0.35,
    size = 3.5,
    fontface = "bold"
  ) +
  geom_point(
    aes(size = Holdout_F1, fill = Outcome),
    shape = 21,
    color = "white",
    stroke = 1.2,
    alpha = 0.95
  ) +
  geom_segment(
    data = label_df,
    aes(
      x = Recall,
      y = Precision,
      xend = label_x,
      yend = label_y
    ),
    color = "#AAB7B8",
    linewidth = 0.5
  ) +
  geom_label(
    data = label_df,
    aes(
      x = label_x,
      y = label_y,
      label = label_text
    ),
    fill = "white",
    color = outline_color,
    fontface = "bold",
    size = 3.6,
    label.size = 0.35,
    lineheight = 1.05
  ) +
  scale_fill_manual(values = portfolio_colors) +
  scale_size_continuous(
    name = "Holdout F1",
    range = c(8, 14),
    breaks = c(0.30, 0.40, 0.50, 0.60),
    labels = scales::number_format(accuracy = 0.001)
  ) +
  scale_x_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, 1.02),
    expand = expansion(mult = c(0.01, 0.02))
  ) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, 1.08),
    breaks = c(0, 0.25, 0.50, 0.75, 1.00),
    expand = expansion(mult = c(0, 0.02))
  ) +
  labs(
    title = "Prediction Confidence and Detection Tradeoffs",
    subtitle = "Precision indicates reliability of flagged cases; recall indicates detection coverage",
    x = "Recall: Share of actual risk cases detected",
    y = "Precision: Reliability of positive predictions",
    fill = "Outcome"
  ) +
  coord_cartesian(clip = "off") +
  theme_hr(11) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(color = subtitle_color, size = 11.5),
    legend.position = "right",
    legend.title = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    plot.margin = margin(12, 30, 18, 12)
  )

print(p)
save_plot(p, "precision_recall_tradeoff.png", 10.8, 6.6)
```

This view shows why each model should be used differently. A model with higher precision is better suited for focused review because flagged employees are more likely to represent meaningful risk. A model with higher recall is better suited for broader screening because it captures more potential cases, but it may also generate more false positives.

The monitoring/privacy model shows the strongest evidence for confident prioritization because it produces the highest reliability among positive predictions. The turnover model provides a more balanced retention-risk signal. The quiet quitting model should be interpreted more cautiously as an early engagement signal that requires additional survey, manager, or qualitative context before action is taken.

## Risk Segmentation Across Outcomes

```{r 04-model-use-profile-table, echo=FALSE}
use_profile_table <- model_summaries %>%
  mutate(
    Practical_Profile = case_when(
      Holdout_Precision >= 0.80 & Holdout_Recall < 0.60 ~
        "High-confidence review queue",
      Holdout_Recall >= 0.80 & Holdout_Precision < 0.40 ~
        "Broad screening signal",
      Holdout_F1 >= 0.45 ~
        "Balanced prioritization signal",
      TRUE ~
        "Cautious early-warning signal"
    ),
    Recommended_Use = case_when(
      Outcome == "Turnover Intention" ~
        "Use for retention-risk prioritization and targeted manager/HR review.",
      Outcome == "Quiet Quitting" ~
        "Use as an early engagement-risk signal supported by survey and manager context.",
      Outcome == "Monitoring & Privacy Risk" ~
        "Use for privacy, trust, and monitoring-governance review prioritization.",
      TRUE ~
        "Use as a decision-support signal."
    )
  ) %>%
  select(
    Outcome,
    Model_Config,
    Practical_Profile,
    Recommended_Use
  )

use_profile_table %>%
  gt() %>%
  tab_header(
    title = "Practical Use Profile by Workforce Risk Outcome",
    subtitle = "How each selected model should be interpreted in decision-making"
  ) %>%
  cols_label(
    Model_Config = "Selected Configuration",
    Practical_Profile = "Practical Profile",
    Recommended_Use = "Recommended Use"
  ) %>%
  cols_width(
    Outcome ~ px(190),
    Model_Config ~ px(240),
    Practical_Profile ~ px(245),
    Recommended_Use ~ px(440)
  ) %>%
  cols_align("center", columns = c(Outcome, Model_Config, Practical_Profile)) %>%
  tab_style(
    style = list(
      cell_fill(color = "#EBF5FB"),
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = Practical_Profile,
      rows = Practical_Profile == "High-confidence review queue"
    )
  ) %>%
  tab_style(
    style = list(
      cell_fill(color = "#EAF7EF"),
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = Practical_Profile,
      rows = Practical_Profile == "Balanced prioritization signal"
    )
  ) %>%
  tab_style(
    style = list(
      cell_fill(color = "#FFF4DA"),
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      columns = Practical_Profile,
      rows = Practical_Profile == "Cautious early-warning signal"
    )
  ) %>%
  tab_options(
    table.width = pct(95),
    table.font.size = px(12),
    data_row.padding = px(9)
  )
```

## Portfolio Risk Use Matrix

```{r 04-portfolio-risk-use-profile, echo=FALSE}
use_profile_long <- model_summaries %>%
  mutate(
    Outcome = factor(
      as.character(Outcome),
      levels = c(
        "Turnover Intention",
        "Quiet Quitting",
        "Monitoring & Privacy Risk"
      )
    )
  ) %>%
  transmute(
    Outcome,
    Reliability = Holdout_Precision,
    Coverage = Holdout_Recall,
    Balance = Holdout_F1
  ) %>%
  pivot_longer(
    cols = c(Reliability, Coverage, Balance),
    names_to = "Metric",
    values_to = "Score"
  ) %>%
  mutate(
    Metric = factor(
      Metric,
      levels = c("Reliability", "Coverage", "Balance")
    ),
    Outcome = forcats::fct_rev(Outcome),
    Score_Label = scales::number(Score, accuracy = 0.001)
  )

p <- ggplot(use_profile_long, aes(x = Score, y = Outcome, color = Outcome)) +
  geom_segment(
    aes(x = 0, xend = Score, yend = Outcome),
    linewidth = 4,
    alpha = 0.35,
    show.legend = FALSE
  ) +
  geom_point(
    size = 5.2,
    show.legend = FALSE
  ) +
  geom_text(
    aes(label = Score_Label),
    hjust = -0.20,
    size = 3.8,
    fontface = "bold",
    color = outline_color,
    show.legend = FALSE
  ) +
  facet_wrap(~ Metric, nrow = 1) +
  scale_color_manual(values = portfolio_colors) +
  scale_x_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, 1.08),
    expand = expansion(mult = c(0.01, 0.02))
  ) +
  labs(
    title = "Portfolio Risk Use Profile",
    subtitle = "Reliability = precision, coverage = recall, and balance = holdout F1",
    x = "Score",
    y = NULL
  ) +
  coord_cartesian(clip = "off") +
  theme_hr(10.8) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(color = subtitle_color, size = 11.5),
    strip.text = element_text(face = "bold", size = 11.5),
    axis.text.y = element_text(face = "bold", size = 10.5),
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    plot.margin = margin(12, 30, 16, 12)
  )

print(p)
save_plot(p, "portfolio_risk_use_profile.png", 11.2, 5.8)
```

This profile converts technical metrics into decision language. **Reliability** reflects how trustworthy a positive flag is, **coverage** reflects how much risk the model detects, and **balance** summarizes the overall usefulness of the selected model.

The comparison reinforces the broader portfolio story. Monitoring & Privacy Risk is strongest on reliability and overall balance, Turnover Intention is moderate but usable across all three dimensions, and Quiet Quitting is comparatively weaker and should be used as a supporting engagement signal rather than a standalone decision tool.

---

# Cross-Model Workforce Patterns {#cross-model-patterns}

The modeling results point to a common theme across the project: workforce risk rarely emerges from a single variable. Instead, risk appears through overlapping behavioral, organizational, wellbeing, trust, and employee-experience signals.

The three outcomes represent distinct forms of workforce vulnerability:

- **Turnover intention** reflects potential exit risk.
- **Quiet quitting** reflects potential disengagement while remaining employed.
- **Monitoring-related turnover risk** reflects workforce trust and privacy pressure connected to surveillance or monitoring practices.

Together, these outcomes show that employee risk should be interpreted as a portfolio of signals rather than a single prediction.

```{=html}
<div class="decision-box">

<h3>Shared Workforce Risk Pattern</h3>

<p>
Across the portfolio, the strongest theme is that workforce risk is not isolated. Retention, disengagement, wellbeing, trust, manager support, workload pressure, and monitoring perceptions interact with each other.
</p>

<p>
This means HR analytics should not stop at prediction. The more valuable step is translating model outputs into better questions for leaders, managers, HR partners, and governance stakeholders.
</p>

</div>
```

## Shared Risk Themes

Across the project, the strongest workforce-risk story is not that one model is universally best. Instead, the results show that each outcome captures a different part of the employee experience.

Common themes include:

- Trust and support shape whether employees stay connected to the organization.
- Workload and wellbeing influence both withdrawal risk and retention risk.
- Role clarity and manager support affect how employees interpret their work environment.
- Monitoring and privacy perceptions add a governance layer to workforce analytics.

## Outcome-Specific Risk Differences

Although the models share common workforce themes, each outcome requires a different business response.

```{=html}
<div class="model-highlight-card">

  <div class="model-badge">Outcome-Specific Decision Use</div>

  <h3 class="model-title">
    Different Signals Require Different Responses
  </h3>

  <p class="model-subtitle">
    Each model should be interpreted according to the type of workforce decision it supports.
  </p>

  <div class="model-metrics">
    <div class="metric-box">
      <span class="metric-label">Turnover Intention</span>
      <span class="metric-value">
        Retention Review
      </span>
    </div>

    <div class="metric-box">
      <span class="metric-label">Quiet Quitting</span>
      <span class="metric-value">
        Engagement Review
      </span>
    </div>

    <div class="metric-box">
      <span class="metric-label">Monitoring & Privacy Risk</span>
      <span class="metric-value">
        Governance Review
      </span>
    </div>
  </div>

</div>
```

```{=html}
<div class="decision-box">

<h3>Decision Context Matters</h3>

<p>
<strong>Turnover intention</strong> supports retention review by helping HR and managers identify where earlier follow-up, workload review, or career-pathing support may be useful.
</p>

<p>
<strong>Quiet quitting</strong> should be treated as a cautious engagement signal because disengagement can be behavioral, emotional, or contextual.
</p>

<p>
<strong>Monitoring-related risk</strong> is best suited for governance review because it connects workforce risk to privacy, trust, surveillance, and policy design.
</p>

</div>
```

# Strategic Workforce Recommendations {#strategic-recommendations}

The modeling results support a practical executive strategy: use predictive outputs to prioritize attention, then combine those signals with human context, qualitative information, and organizational judgment.

The recommendations below translate the modeling results into workforce action areas.

```{r 06-strategic-recommendations-table, echo=FALSE}
recommendations <- tibble::tribble(
  ~Risk_Area, ~Primary_Model_Signal, ~Recommended_Action,
  "Retention Risk",
  "Turnover intention model",
  "Use risk scores to prioritize retention conversations, manager check-ins, workload review, and career-development discussions.",

  "Quiet Quitting / Engagement Risk",
  "Quiet quitting model",
  "Treat predictions as early engagement signals. Pair model output with pulse surveys, manager observations, role clarity review, and discretionary-effort indicators.",

  "Monitoring & Privacy Risk",
  "Monitoring/privacy model",
  "Use flagged cases and segments to guide monitoring governance review, privacy communication, trust-building, and proportionality checks.",

  "Workforce Planning",
  "All models",
  "Compare risk patterns across departments, roles, tenure groups, and work arrangements to identify structural workforce pressure points.",

  "Leadership Decision Support",
  "All models",
  "Use model outputs to focus executive attention, not to automate employment decisions or replace manager/HR judgment."
)

recommendations %>%
  gt() %>%
  tab_header(
    title = "Strategic Workforce Recommendations",
    subtitle = "Translating model outputs into executive action"
  ) %>%
  cols_label(
    Risk_Area = "Risk Area",
    Primary_Model_Signal = "Primary Model Signal",
    Recommended_Action = "Recommended Action"
  ) %>%
  cols_width(
    Risk_Area ~ px(210),
    Primary_Model_Signal ~ px(230),
    Recommended_Action ~ px(560)
  ) %>%
  tab_style(
    style = list(
      cell_fill(color = "#EBF5FB"),
      cell_text(weight = "bold")
    ),
    locations = cells_body(columns = Risk_Area)
  ) %>%
  tab_options(
    table.width = pct(95),
    table.font.size = px(12),
    data_row.padding = px(9)
  )
```

# Responsible Use & Governance {#responsible-use}

Given that these models deal with workforce risk, responsible use is central to the project. The models should support better questions, earlier review, and more consistent prioritization. They should not be used as automated decision systems.

The strongest governance principle is simple: **model outputs should inform human review, not replace it.**

```{=html}
<div class="decision-box">

<h3>Responsible Use Principle</h3>

<p>
The appropriate use case is human-in-the-loop prioritization. Model outputs can help identify where review may be needed, but final interpretation should include employee context, manager knowledge, organizational conditions, privacy considerations, and ethical review.
</p>

<p>
This is especially important for monitoring-related risk, where model outputs may reflect employee trust and privacy concerns. In that context, the goal is not to increase surveillance, but to improve transparency, governance, and workforce support.
</p>

</div>
```

## What the Models Can Do

These models can help organizations:

- Identify where workforce risk may be concentrated
- Compare risk patterns across outcomes
- Prioritize HR, manager, leadership, or governance review
- Support earlier intervention before risk becomes visible through turnover, disengagement, or employee complaints
- Create a repeatable analytics framework for workforce decision support

## What the Models Should Not Do

These models should not be used to:

- Make automatic employment decisions
- Label employees as problems or flight risks without context
- Replace manager judgment, HR review, or employee voice
- Justify monitoring expansion without governance review
- Trigger punitive action without human validation

## Human-in-the-Loop Decision Guidance

The recommended workflow is:

1. Use model output to identify where review may be useful.
2. Add manager, HR, survey, workload, and employee-context evidence.
3. Evaluate whether the signal reflects an individual issue, a team-level pattern, or a broader organizational condition.
4. Decide whether action should involve support, communication, governance review, or further analysis.

This approach keeps the models in the right role: supporting better decision-making without replacing human judgment.

```{=html}
<div class="model-highlight-card">

  <div class="model-badge">From Insight to Action</div>

  <h3 class="model-title">
    The Next Step Is Operationalizing Responsible Review
  </h3>

  <p class="model-subtitle">
    The portfolio results are most useful when translated into a repeatable review process, not a one-time report.
  </p>

  <div class="model-metrics">
    <div class="metric-box">
      <span class="metric-label">Immediate</span>
      <span class="metric-value">
        Clarify Use Cases
      </span>
    </div>

    <div class="metric-box">
      <span class="metric-label">30–60 Days</span>
      <span class="metric-value">
        Build Dashboard
      </span>
    </div>

    <div class="metric-box">
      <span class="metric-label">Long Term</span>
      <span class="metric-value">
        Monitor Drift
      </span>
    </div>
  </div>

</div>
```

## Roadmap Timeline View {#executive-action-roadmap}

```{r 08-roadmap-timeline-view, echo=FALSE}
roadmap_plot_df <- tibble::tribble(
  ~step, ~stage, ~y, ~label_y, ~label,
  1, "Immediate Actions", 1, 1.34, "Clarify model use cases",
  2, "Immediate Actions", 1, 0.78, "Review monitoring and privacy signals",
  3, "30–60 Day Actions", 2, 2.34, "Build recurring risk dashboard",
  4, "30–60 Day Actions", 2, 1.78, "Add qualitative review inputs",
  5, "Long-Term Strategy", 3, 3.34, "Formalize governance standards",
  6, "Long-Term Strategy", 3, 2.78, "Monitor drift and recalibrate"
) %>%
  mutate(
    stage = factor(
      stage,
      levels = c(
        "Immediate Actions",
        "30–60 Day Actions",
        "Long-Term Strategy"
      )
    ),
    label = stringr::str_wrap(label, width = 22)
  )

p <- ggplot(roadmap_plot_df, aes(x = step, y = y, color = stage)) +
  geom_path(
    linewidth = 2.2,
    color = "#C9D3D3",
    show.legend = FALSE
  ) +
  geom_segment(
    aes(
      x = step,
      xend = step,
      y = y,
      yend = label_y
    ),
    color = "#B7C4C9",
    linewidth = 0.7,
    show.legend = FALSE
  ) +
  geom_point(
    size = 14,
    show.legend = FALSE
  ) +
  geom_text(
    aes(label = step),
    color = "white",
    fontface = "bold",
    size = 5
  ) +
  geom_label(
    aes(y = label_y, label = label),
    fill = "white",
    color = outline_color,
    fontface = "bold",
    size = 3.6,
    label.size = 0.35,
    label.padding = grid::unit(0.16, "lines"),
    label.r = grid::unit(0.14, "lines"),
    lineheight = 1.0,
    show.legend = FALSE
  ) +
  scale_color_manual(values = c(
    "Immediate Actions" = "#5AA0D1",
    "30–60 Day Actions" = "#45A79C",
    "Long-Term Strategy" = "#E67E6B"
  )) +
  scale_x_continuous(
    breaks = 1:6,
    limits = c(0.5, 6.6)
  ) +
  scale_y_continuous(
    breaks = c(1, 2, 3),
    labels = c(
      "Immediate Actions",
      "30–60 Day Actions",
      "Long-Term Strategy"
    ),
    limits = c(0.45, 3.60)
  ) +
  labs(
    title = "Executive Action Roadmap Timeline",
    subtitle = "A staged path from model outputs to responsible workforce analytics operations",
    x = "Implementation Sequence",
    y = NULL
  ) +
  coord_cartesian(clip = "off") +
  theme_hr(10.8) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(color = subtitle_color, size = 11.5),
    axis.text.y = element_text(face = "bold", size = 10.5),
    panel.grid.minor = element_blank(),
    legend.position = "none",
    plot.margin = margin(16, 34, 30, 18)
  )

print(p)
save_plot(p, "executive_action_roadmap_timeline.png", 11.0, 6.6)
```

# Final Takeaway {#final-takeaway}

This project demonstrates how synthetic HR data can support a realistic workforce analytics pipeline, from data preparation and exploratory analysis to predictive modeling and executive decision support.

The final value of the project is not just predictive performance. It is the end-to-end demonstration of how synthetic data, responsible modeling, executive visualization, and human-in-the-loop governance can work together in a realistic HR analytics workflow.

```{=html}
<div class="model-highlight-card">

  <div class="model-badge">Final Portfolio Takeaway</div>

  <h3 class="model-title">
    Predictive HR Analytics Works Best as Decision Support
  </h3>

  <p class="model-subtitle">
    The models help leaders prioritize attention earlier, but they do not replace human judgment, employee context, or responsible governance.
  </p>

  <div class="model-metrics">
    <div class="metric-box">
      <span class="metric-label">Retention</span>
      <span class="metric-value">
        Prioritize Review
      </span>
    </div>

    <div class="metric-box">
      <span class="metric-label">Engagement</span>
      <span class="metric-value">
        Add Context
      </span>
    </div>

    <div class="metric-box">
      <span class="metric-label">Monitoring</span>
      <span class="metric-value">
        Govern Responsibly
      </span>
    </div>
  </div>

</div>
```

Turnover intention is useful for retention-risk prioritization. Quiet quitting is best treated as an early engagement signal that requires additional context. Monitoring and privacy risk provides the clearest governance-focused review signal because it connects workforce risk to trust, transparency, surveillance, and policy design.

Used responsibly, this framework helps organizations move from reactive reporting to proactive workforce intelligence. The models do not replace human judgment; they help leaders ask better questions, prioritize attention, and respond earlier to patterns that may affect retention, engagement, trust, and organizational effectiveness.

```{=html}
<hr class="section-divider">

<div class="page-nav">

  <div class="nav-left">
    <a href="hr_data_monitoring_privacy.html" class="cta-btn btn-blue">← Back</a>
  </div>

  <div class="nav-center">
    <a href="index.html" class="cta-btn btn-teal">Home</a>
  </div>

</div>

</div> <!-- page-content -->
</div> <!-- page-layout -->
</div> <!-- report-container -->
```