---
title: ""

output:
  html_document:
    always_allow_html: true
    self_contained: false
    toc: false
    toc_float: false
    theme: null
    highlight: null
    df_print: paged
    fig_caption: true
    css: assets/styles.css
    includes:
      in_header: ../docs/assets/header.html

number_sections: true
---

```{=html}
<div class="report-container">

  <div class="page-layout">

    <!-- SIDEBAR -->
    <div class="sidebar">
      <div class="sidebar-card">
        <div class="sidebar-label">On this page</div>
        <nav class="sidebar-nav">
          <a href="#executive-summary">Executive Summary</a>
          <a href="#modeling-overview">Modeling Overview</a>
          <a href="#data-prep-framework">Data Preparation</a>
          <a href="#model-development">Model Development</a>
          <a href="#model-evaluation">Model Evaluation</a>
          <a href="#model-evaluation-results">Evaluation Results</a>
          <a href="#sampling-strategy-framework">Sampling Strategy Framework</a>
          <a href="#precision-recall">Precision–Recall</a>
          <a href="#model-selection">Model Selection</a>
          <a href="#model-summary">Model Summary</a>
          <a href="#driver-interpretation">Driver Interpretation</a>
          <a href="#feature-action">Feature → Action</a>
          <a href="#business-impact">Business Impact</a>
        </nav>
      </div>
    </div>

    <!-- MAIN CONTENT -->
    <div class="page-content">

      <!-- MATCH EDA STRUCTURE -->
      <div class="hero-section">
        <p class="hero-meta">
          Kevin Allen Cauley • `r format(Sys.Date(), '%B %d, %Y')`
        </p>

        <h1 class="hero-title">Monitoring & Privacy Risk Modeling</h1>

        <p class="hero-subtitle">
  This section develops and evaluates predictive models to identify employees at risk of monitoring-related turnover, interpret key privacy and surveillance risk patterns, and translate model outputs into executive-level workforce trust recommendations.
        </p>
      </div>

      <!-- REPLACED page-nav WITH cta-row -->
      <div class="cta-row">
        <a href="HR_Data_Quiet_Quitting.html" class="cta-btn btn-blue">← Back</a>
        <a href="index.html" class="cta-btn btn-teal">Home</a>
        <a href="HR_Data_Insights_Gleaned.html" class="cta-btn btn-red">Next →</a>
      </div>
```

```{r 00-setup, include=FALSE}
source(here::here("analysis", "helper.R"))

PLOT_SUBFOLDER <- "monitoring_privacy"

cat("Using plot folder:", PLOT_SUBFOLDER, "\n")

# Required Packages
required_pkgs <- c(
  "conflicted", "tidyverse", "tidymodels", "themis", "here",
  "klaR", "discrim", "ranger", "kernlab", "xgboost", "gt", "scales"
)

for (pkg in required_pkgs) {
  if (!pkg %in% rownames(installed.packages())) {
    install.packages(pkg)
  }
}

library(conflicted)
library(tidyverse)
library(tidymodels)
library(themis)
library(here)
library(klaR)
library(discrim)
library(gt)
library(scales)

conflicted::conflicts_prefer(
  dplyr::filter,
  dplyr::select,
  dplyr::mutate,
  dplyr::summarise,
  dplyr::arrange,
  dplyr::slice,
  dplyr::slice_head,
  dplyr::slice_max,
  dplyr::slice_min,
  dplyr::recode,
  dplyr::rename,
  dplyr::lag
)

# Global Options
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  results = "hide",
  fig.align = "center",
  out.width = "100%",
  fig.retina = 2,
  fig.width = 9,
  fig.height = 5.75,
  dpi = 300,
  cache = FALSE,
  dev = "png"
)

# Reproducibility
set.seed(123)

# Load CLEAN dataset from EDA (DO NOT TOUCH THIS)
df <- readRDS(here::here("data", "processed", "hr_dataset_clean.rds"))

# Base dataset
base_df <- df %>%
  mutate(across(where(is.character), as.factor))
```

# Executive Summary {#executive-summary}

This analysis develops a predictive modeling framework to identify employees at risk of monitoring-related turnover using a synthetic, research-informed HR dataset.

Multiple machine learning models—including Logistic Regression, Random Forest, XGBoost, Support Vector Machines, and Naive Bayes—were evaluated across multiple sampling strategies to address class imbalance and compare predictive performance under different modeling conditions.

Results show that **model effectiveness is shaped by both algorithm choice and data strategy**. Sampling strategies such as up-sampling, down-sampling, and SMOTE changed model behavior in meaningful ways, demonstrating that class-balancing choices can materially affect precision, recall, F1 Score, and overall usability.

The final model-selection process uses **cross-validated F1 Score** as the primary ranking metric, while holdout test performance is used as a stability check to confirm that the selected configuration remains practically usable on unseen data.

Rather than relying on a single predictive signal, monitoring-related turnover risk is influenced by a **combination of privacy perceptions, monitoring intensity, workplace trust, workload pressure, wellbeing, and organizational support factors**.

From a business perspective, this framework enables a shift from **reactive privacy concern response to proactive workforce trust and monitoring-risk management**, supporting earlier intervention, better governance, and improved retention outcomes.

# Modeling Overview {#modeling-overview}

This section builds on the exploratory analysis by applying predictive modeling techniques to identify employees at risk of monitoring-related turnover.

Rather than relying on a single model, multiple algorithms are evaluated and compared to determine the most effective approach for workforce decision-making.

Each model is assessed using cross-validation and evaluated across multiple sampling strategies to address class imbalance.

The objective is not only to identify the most accurate model, but to determine which approach best aligns with organizational priorities.

# Data Preparation & Modeling Framework {#data-prep-framework}

## Target Definition

The target variable for this analysis is **monitoring-related turnover risk**, representing whether an employee is likely to show elevated turnover risk associated with workplace monitoring, privacy concerns, or surveillance-related strain.

This binary outcome is structured as a classification problem, with levels defined as **"No"** and **"Yes"**, where **"Yes"** represents employees flagged as higher risk for monitoring-related turnover in the synthetic dataset.

Within this framework, the target should not be interpreted as a directly observed resignation event. Instead, it reflects a workforce-risk signal tied to monitoring perceptions, privacy concerns, trust, workload strain, and employee experience.

This distinction is important because monitoring-related risk is both a workforce analytics issue and a governance issue. The goal is not to label employees, but to identify patterns that may warrant earlier HR, leadership, or monitoring-governance review.

```{r 01-target, echo=FALSE}
# Ensure correct factor levels
base_df <- base_df %>%
  mutate(
    monitoring_turnover_risk = factor(monitoring_turnover_risk, levels = c("No", "Yes"))
  )
```

## Data Partitioning

To evaluate model performance objectively, the dataset is split into training and testing subsets.

The training set is used to fit and tune models, while the test set is reserved for final evaluation. Stratified sampling is applied to preserve the original class distribution of monitoring-based turnover risk, ensuring that model performance reflects realistic conditions.

```{r 02-split, echo=FALSE}
split <- initial_split(base_df, prop = 0.8, strata = monitoring_turnover_risk)

train_data <- training(split)
test_data  <- testing(split)
```

## Feature Engineering Pipeline

A standardized preprocessing workflow was applied to ensure consistent evaluation across all modeling approaches.

Workforce variables representing organizational context, monitoring exposure, privacy perceptions, employee wellbeing, trust, managerial support, workload conditions, and behavioral indicators were transformed into model-compatible formats while preserving interpretability for downstream HR and governance analysis.

Categorical workforce attributes were converted into numeric representations using dummy encoding, while continuous variables were normalized where required for scale-sensitive algorithms such as Logistic Regression and Support Vector Machines.

Predictors with zero variance or limited informational value were removed to reduce noise and improve model stability across resampling conditions.

This preprocessing framework ensures that observed performance differences are driven primarily by modeling behavior and sampling strategy rather than inconsistent data preparation.

```{r 03-recipe, echo=FALSE}
rec <- recipe(monitoring_turnover_risk ~ ., data = train_data) %>%
  
  # Handle factors
  step_dummy(all_nominal_predictors()) %>%
  
  # Normalize numeric features (important for SVM, Logistic)
  step_normalize(all_numeric_predictors()) %>%
  
  # Remove zero variance
  step_zv(all_predictors())
```

## Cross-Validation Strategy

Cross-validation is used to evaluate model performance during training and reduce the risk of overfitting.

By partitioning the training data into multiple folds, each model is evaluated across different subsets of the data. This provides a more robust estimate of model performance compared to a single train-test split.

```{r 04-resampling, echo=FALSE}
cv_folds <- vfold_cv(train_data, v = 5, strata = monitoring_turnover_risk)
```

# Model Development {#model-development}

Multiple predictive modeling approaches were evaluated to determine how different algorithms interpret monitoring-related workforce-risk patterns within a synthetic HR environment.

The goal was not simply to identify the highest-performing model, but to compare how various modeling strategies balance interpretability, detection capability, stability, and operational usefulness within a privacy-preserving analytics framework.

Each model was trained using a standardized preprocessing workflow to ensure that performance differences reflected modeling behavior rather than inconsistent data preparation.

---

```{r 05-logistic, echo=FALSE}
log_model <- logistic_reg() %>%
  set_engine("glm")
```

```{r 06-rf, echo=FALSE}
rf_model <- rand_forest(mtry = tune(), trees = 500, min_n = tune()) %>%
  set_engine("ranger") %>%
  set_mode("classification")
```

```{r 07-xgb, echo=FALSE}
xgb_model <- boost_tree(
  trees = 500,
  mtry = tune(),
  learn_rate = tune(),
  min_n = tune()
) %>%
  set_engine("xgboost") %>%
  set_mode("classification")
```

```{r 08-nb, echo=FALSE}
nb_model <- naive_Bayes() %>%
  set_engine("klaR")
```

```{r 09-svm, echo=FALSE}
svm_model <- svm_rbf(cost = tune(), rbf_sigma = tune()) %>%
  set_engine("kernlab") %>%
  set_mode("classification")
```

## Model Pipelines

Workflows are used to combine preprocessing steps and model specifications into a single pipeline.

This ensures that all transformations are applied consistently during both training and prediction, reducing the risk of data leakage and simplifying model management.

```{r 10-workflows, echo=FALSE}
log_wf <- workflow() %>% add_model(log_model) %>% add_recipe(rec)
rf_wf  <- workflow() %>% add_model(rf_model)  %>% add_recipe(rec)
xgb_wf <- workflow() %>% add_model(xgb_model) %>% add_recipe(rec)
svm_wf <- workflow() %>% add_model(svm_model) %>% add_recipe(rec)
nb_wf  <- workflow() %>% add_model(nb_model)  %>% add_recipe(rec)
```

## Evaluation Metrics

Model performance is evaluated using a combination of classification metrics.

Accuracy provides an overall performance measure, while precision and recall capture trade-offs between false positives and false negatives. The F1 Score balances these two metrics, and ROC AUC evaluates the model’s ability to distinguish between classes.

Using multiple metrics ensures a comprehensive evaluation aligned with business priorities.

```{r 11-metrics, echo=FALSE}
# Classification metrics
# IMPORTANT:
# monitoring_turnover_risk levels are c("No", "Yes")
# event_level = "second" ensures the positive class is "Yes" = at risk of monitoring-related turnover

options(yardstick.event_first = FALSE)

class_metrics <- metric_set(
  accuracy,
  precision,
  recall,
  f_meas,
  roc_auc
)

target_event_level <- "second"

model_colors <- c(
  "Logistic" = "#4B9CD3",
  "Random Forest" = "#3FA34D",
  "XGBoost" = "#E87461",
  "Naive Bayes" = "#F4D03F",
  "SVM" = "#9370DB"
)

sampling_order <- c(
  "No Sampling",
  "Up Sampling",
  "Down Sampling",
  "SMOTE"
)

metric_labels <- c(
  accuracy = "Accuracy",
  precision = "Precision",
  recall = "Recall",
  f_meas = "F1 Score",
  roc_auc = "ROC AUC"
)
```

## Hyperparameter Tuning

For models with tunable parameters, grid search is used to identify optimal configurations.

This process evaluates multiple parameter combinations using cross-validation, selecting those that maximize predictive performance. Tuning helps improve model accuracy while maintaining generalizability.

## Model Training

Multiple supervised learning algorithms were evaluated, including Logistic Regression, Random Forest, XGBoost, Support Vector Machines, and Naive Bayes.

Each model was trained under multiple class-balancing conditions to evaluate how resampling strategies influence monitoring-related risk detection behavior. This is particularly important because employees flagged for monitoring-related turnover risk typically represent a smaller portion of the workforce population.

Rather than focusing on a single performance statistic, the modeling framework evaluates how each configuration balances detection capability, prediction reliability, and operational practicality.

This comparative approach reinforces the broader objective of the project: demonstrating how research-informed synthetic workforce data can support realistic HR analytics workflows while preserving employee privacy and organizational confidentiality.

```{r 12-modeling-framework, include=FALSE, results='hide', message=FALSE, warning=FALSE}
# This framework evaluates each model + sampling strategy combination together using cross-validation. Tunable models are tuned separately under each sampling condition.

# Final model selection is based on cross-validated F1 Score with a holdout viability check before reporting the selected model.

# Sampling recipes
sampling_recipes <- list(
  "No Sampling" = rec,
  "Up Sampling" = rec %>% step_upsample(monitoring_turnover_risk),
  "Down Sampling" = rec %>% step_downsample(monitoring_turnover_risk),
  "SMOTE" = rec %>% step_smote(monitoring_turnover_risk)
)

# Model specifications
model_specs <- list(
  "Logistic" = log_model,
  "Naive Bayes" = nb_model,
  "Random Forest" = rf_model,
  "XGBoost" = xgb_model,
  "SVM" = svm_model
)

# Tunable model names
tunable_models <- c("Random Forest", "XGBoost", "SVM")

# Unified prediction function
get_preds <- function(model, data) {
  predict(model, data, type = "prob") %>%
    bind_cols(predict(model, data)) %>%
    bind_cols(data %>% select(monitoring_turnover_risk))
}

# Unified scoring function
score_predictions <- function(preds, model_name, sampling_name) {
  class_metrics(
    preds,
    truth = monitoring_turnover_risk,
    estimate = .pred_class,
    .pred_Yes,
    event_level = target_event_level
  ) %>%
    mutate(
      Model = model_name,
      Sampling = sampling_name
    )
}

# Evaluate one model + sampling combination using CV
run_cv_model_sampling <- function(model_name, sampling_name) {
  
  model_spec <- model_specs[[model_name]]
  recipe_used <- sampling_recipes[[sampling_name]]
  
  wf <- workflow() %>%
    add_model(model_spec) %>%
    add_recipe(recipe_used)
  
  set.seed(123)
  
  if (model_name %in% tunable_models) {
    
    cv_fit <- tune_grid(
      wf,
      resamples = cv_folds,
      grid = 10,
      metrics = class_metrics,
      control = control_grid(save_pred = TRUE)
    )
    
    best_params <- select_best(cv_fit, metric = "f_meas")
    
    # Keep only metrics for the best tuned parameter combination.
    # This prevents duplicate model + sampling rows from appearing later.
    cv_metrics <- collect_metrics(cv_fit) %>%
      inner_join(best_params, by = names(best_params)) %>%
      mutate(
        Model = model_name,
        Sampling = sampling_name,
        Model_Config = paste(model_name, sampling_name, sep = " + ")
      )
    
    final_wf <- finalize_workflow(wf, best_params)
    
  } else {
    
    cv_fit <- fit_resamples(
      wf,
      resamples = cv_folds,
      metrics = class_metrics,
      control = control_resamples(save_pred = TRUE)
    )
    
    cv_metrics <- collect_metrics(cv_fit) %>%
      mutate(
        Model = model_name,
        Sampling = sampling_name,
        Model_Config = paste(model_name, sampling_name, sep = " + ")
      )
    
    final_wf <- wf
  }
  
  # Fit finalized workflow to full training data
  set.seed(123)
  final_fit <- fit(final_wf, data = train_data)
  
  # Evaluate finalized workflow on holdout test data
  test_preds <- get_preds(final_fit, test_data)
  
  test_metrics <- score_predictions(
    preds = test_preds,
    model_name = model_name,
    sampling_name = sampling_name
  ) %>%
    mutate(Model_Config = paste(model_name, sampling_name, sep = " + "))
  
  list(
    cv_fit = cv_fit,
    cv_metrics = cv_metrics,
    final_workflow = final_wf,
    final_fit = final_fit,
    test_preds = test_preds,
    test_metrics = test_metrics
  )
}

# Run all model + sampling combinations
model_grid <- tidyr::expand_grid(
  Model = names(model_specs),
  Sampling = names(sampling_recipes)
)

set.seed(123)

model_training_output <- capture.output({
  model_results <- purrr::pmap(
    list(model_grid$Model, model_grid$Sampling),
    run_cv_model_sampling
  )
})

names(model_results) <- paste(model_grid$Model, model_grid$Sampling, sep = " + ")

# Cross-validation results for all best model + sampling configurations
cv_results_all <- purrr::map_dfr(model_results, "cv_metrics") %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "roc_auc")) %>%
  mutate(
    Metric = dplyr::recode(.metric, !!!metric_labels),
    Sampling = factor(Sampling, levels = sampling_order),
    Model = factor(Model, levels = names(model_colors))
  )

# Holdout test results for all finalized configurations
results_all <- purrr::map_dfr(model_results, "test_metrics") %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "roc_auc")) %>%
  mutate(
    Metric = dplyr::recode(.metric, !!!metric_labels),
    Sampling = factor(Sampling, levels = sampling_order),
    Model = factor(Model, levels = names(model_colors))
  )
```

---

# Model Evaluation {#model-evaluation}

This evaluation framework examines how different modeling strategies perform on **unseen workforce data** while emphasizing practical HR and monitoring-governance decision-support considerations rather than purely statistical optimization.

Because workforce intervention and governance-review resources are finite, the analysis prioritizes models that balance early monitoring-related risk detection with prediction reliability. The objective is to support proactive privacy, trust, and retention review processes while minimizing unnecessary intervention burden.

Evaluation therefore focuses not only on predictive accuracy, but also on how effectively each configuration translates into operationally useful workforce prioritization signals.

Rather than focusing solely on accuracy, evaluation emphasizes the **precision–recall tradeoff**, balancing the ability to detect monitoring-related risk with the cost of unnecessary review.

This ensures model selection aligns with operational priorities, where both detection capability and decision efficiency matter.

Key metrics include:

- **Accuracy** → Overall correctness  
- **Precision** → Reliability of positive monitoring-risk predictions  
- **Recall** → Ability to detect monitoring-related risk cases  
- **F1 Score** → Balance between precision and recall  
- **ROC AUC** → Overall classification strength  

This comparison highlights the trade-offs between detection, prediction reliability, and practical review burden.

## Test Set Performance

Final model assessment is conducted using the holdout test dataset to evaluate how each configuration performs on previously unseen workforce observations.

This stage acts as a practical stability check rather than the primary model-selection mechanism. Cross-validation identifies the strongest generalized configurations during training, while holdout evaluation confirms whether those patterns remain operationally useful under unseen conditions.

This distinction is especially important for monitoring-related turnover risk because privacy, trust, and surveillance-related concerns may vary across organizational context and employee groups.

The evaluation framework therefore separates model discovery from final operational validation, creating a more defensible and realistic assessment process for workforce-risk modeling.

---

# Sampling Strategy Framework {#sampling-strategy-framework}

Sampling techniques are applied within the preprocessing pipeline to ensure consistency across models.

By integrating sampling directly into the workflow, transformations are applied only to the training data, preventing data leakage and preserving the integrity of model evaluation.

## Sampling Recipes and Evaluation Setup

Each sampling strategy is embedded directly into the modeling recipe. This ensures that class balancing occurs only within the training folds during cross-validation, preventing leakage into validation or test data.

# Model Evaluation Results & Insights {#model-evaluation-results}

This section synthesizes model performance across sampling strategies and evaluation metrics, providing a comparative view of how different approaches perform under realistic conditions.

The focus shifts from individual model behavior to cross-model insights, highlighting trade-offs that directly impact workforce decision-making.

The goal is to identify a model that balances detection capability with operational efficiency—minimizing both missed risk and unnecessary intervention.

```{r 13-validation-check, include=FALSE}
stopifnot(exists("cv_results_all"))
stopifnot(exists("results_all"))
stopifnot(nrow(cv_results_all) > 0)
stopifnot(nrow(results_all) > 0)
stopifnot(all(c("Model", "Sampling", ".metric", ".estimate") %in% names(results_all)))
stopifnot(all(c("Model", "Sampling", ".metric", "mean") %in% names(cv_results_all)))
```

```{r 14-final-model-selection, include=FALSE}
# -------------------------------------------------------------------
# Final model selection
# -------------------------------------------------------------------
# Gold-standard selection is based primarily on cross-validated F1 Score,
# but the selected model must also meet minimum holdout stability checks.
#
# This prevents a model from being selected if it performs well in CV
# but fails to produce usable positive-class predictions on the holdout set.
# -------------------------------------------------------------------

# Minimum holdout requirements for a model to be considered usable
min_holdout_f1 <- 0.40
min_holdout_recall <- 0.30
min_holdout_precision <- 0.35

# Create holdout metric screen
holdout_screen <- results_all %>%
  filter(.metric %in% c("precision", "recall", "f_meas", "roc_auc")) %>%
  select(Model, Sampling, .metric, .estimate) %>%
  pivot_wider(names_from = .metric, values_from = .estimate) %>%
  mutate(
    Model = as.character(Model),
    Sampling = as.character(Sampling),
    viable_holdout = !is.na(precision) &
      !is.na(recall) &
      !is.na(f_meas) &
      precision > 0 &
      recall > 0 &
      f_meas > 0,
    stable_holdout = viable_holdout &
      f_meas >= min_holdout_f1 &
      recall >= min_holdout_recall &
      precision >= min_holdout_precision
  )

# Prefer models that meet the stronger holdout stability screen.
eligible_configs <- holdout_screen %>%
  dplyr::filter(stable_holdout) %>%
  dplyr::select(Model, Sampling)

# Fallback: if no model meets the stronger screen, use basic viability.
if (nrow(eligible_configs) == 0) {
  eligible_configs <- holdout_screen %>%
    dplyr::filter(viable_holdout) %>%
    dplyr::select(Model, Sampling)
}

# Stop if no model is even viable on the holdout set.
if (nrow(eligible_configs) == 0) {
  stop(
    "No model configuration produced viable positive-class holdout predictions. ",
    "Review class balance, sampling methods, thresholding, and model specifications before selecting a final model."
  )
}

# Cross-validated F1 table
cv_f1_table <- cv_results_all %>%
  filter(.metric == "f_meas") %>%
  mutate(
    Model = as.character(Model),
    Sampling = as.character(Sampling)
  ) %>%
  select(Model, Sampling, cv_f1 = mean)

# Select the best cross-validated model among eligible holdout configurations
top_model_cv <- cv_f1_table %>%
  inner_join(eligible_configs, by = c("Model", "Sampling")) %>%
  left_join(
    holdout_screen %>%
      dplyr::select(
        Model,
        Sampling,
        holdout_precision = precision,
        holdout_recall = recall,
        holdout_f1 = f_meas,
        holdout_roc_auc = roc_auc,
        stable_holdout,
        viable_holdout
      ),
    by = c("Model", "Sampling")
  ) %>%
  dplyr::arrange(dplyr::desc(cv_f1), dplyr::desc(holdout_f1)) %>%
  utils::head(1)

top_model_name <- as.character(top_model_cv$Model)
top_sampling_method <- as.character(top_model_cv$Sampling)
top_model_config <- paste(top_model_name, top_sampling_method, sep = " + ")

top_f1 <- round(top_model_cv$cv_f1, 3)
top_holdout_f1 <- round(top_model_cv$holdout_f1, 3)
top_holdout_precision <- round(top_model_cv$holdout_precision, 3)
top_holdout_recall <- round(top_model_cv$holdout_recall, 3)
top_holdout_roc_auc <- round(top_model_cv$holdout_roc_auc, 3)

selection_strategy_label <- ifelse(
  isTRUE(top_model_cv$stable_holdout),
  "Holdout-Stable Risk Prioritization",
  "Holdout-Viable Risk Prioritization"
)

selection_rule_note <- ifelse(
  isTRUE(top_model_cv$stable_holdout),
  "The selected model met the holdout stability screen and ranked strongest by cross-validated F1 Score among stable candidates.",
  "No configuration met the full holdout stability screen, so the selected model is the strongest viable holdout performer among cross-validated candidates."
)

# Retrieve selected model object from saved model results
selected_result <- model_results[[top_model_config]]

final_wf <- selected_result$final_workflow
final_fit <- selected_result$final_fit
final_preds <- selected_result$test_preds

# Holdout test metrics for selected model
final_model_metrics <- results_all %>%
  filter(
    as.character(Model) == top_model_name,
    as.character(Sampling) == top_sampling_method,
    .metric %in% c("accuracy", "precision", "recall", "f_meas", "roc_auc")
  ) %>%
  mutate(
    Metric = dplyr::recode(.metric, !!!metric_labels),
    Score = round(.estimate, 3)
  )
```

```{r 14b-selected-model-consistency-check, include=FALSE}
# Final selected model consistency check
stopifnot(exists("top_model_name"))
stopifnot(exists("top_sampling_method"))
stopifnot(exists("top_model_config"))
stopifnot(exists("final_model_metrics"))
stopifnot(exists("final_preds"))

stopifnot(length(top_model_name) == 1)
stopifnot(length(top_sampling_method) == 1)
stopifnot(nrow(final_model_metrics) == 5)

stopifnot(
  all(c("Accuracy", "Precision", "Recall", "F1 Score", "ROC AUC") %in% final_model_metrics$Metric)
)
```

## Sampling Strategy Comparison {#sampling-strategy-comparison}

Class imbalance is a major challenge in monitoring-related turnover-risk modeling because employees flagged for monitoring-related risk typically represent a smaller portion of the workforce population.

To better understand how class-balancing strategies influence workforce-risk detection, multiple resampling approaches were evaluated across all algorithms. The goal is not to artificially inflate performance, but to examine how different balancing strategies affect the tradeoff between detection capability and prediction reliability.

This comparison helps demonstrate how modeling decisions influence practical workforce prioritization outcomes within HR analytics and monitoring-governance environments.

The models include:

- No Sampling (baseline)
- Up Sampling (increase minority class)
- Down Sampling (reduce majority class)
- SMOTE (synthetic minority generation)

**The goal is to understand how balancing the dataset impacts model performance.**

```{r 15-plot-sampling-comparison, echo=FALSE}
plot_df <- results_all %>%
  filter(.metric %in% c("precision", "recall", "f_meas", "roc_auc")) %>%
  mutate(
    Metric = factor(
      Metric,
      levels = c("Precision", "Recall", "F1 Score", "ROC AUC")
    ),
    Model_Sampling = paste(Model, Sampling, sep = " + ")
  )

f1_order <- plot_df %>%
  filter(.metric == "f_meas") %>%
  arrange(desc(.estimate)) %>%
  pull(Model_Sampling)

plot_df <- plot_df %>%
  mutate(
    Model_Sampling = factor(Model_Sampling, levels = rev(f1_order)),
    Plot_Label = ifelse(
      is.na(.estimate),
      "",
      scales::number(.estimate, accuracy = 0.001)
    )
  )

p <- ggplot(plot_df, aes(x = Metric, y = Model_Sampling, fill = .estimate)) +
  geom_tile(color = "white", linewidth = 0.6) +
  geom_text(
    aes(label = Plot_Label),
    fontface = "bold",
    size = 3.5,
    color = "#243447"
  ) +
  scale_fill_gradient(
    low = "#EAF2F8",
    high = "#2E86C1",
    limits = c(0, 1),
    na.value = "#F2F4F4",
    name = "Score"
  ) +
  labs(
    title = "Holdout Performance Across Model Configurations",
    subtitle = "Rows are ranked by holdout F1 Score; higher values indicate stronger performance",
    x = NULL,
    y = NULL
  ) +
  theme_hr(10) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(size = 12, color = "#5a6b7c"),
    axis.text.y = element_text(size = 10, face = "bold"),
    axis.text.x = element_text(size = 10, face = "bold"),
    legend.position = "right",
    panel.grid = element_blank()
  )

print(p)
save_plot(p, "sampling_strategy_performance_heatmap.png", 10.5, 8)
```

The heatmap shows holdout test performance across model and sampling combinations. It is used as a confirmation view rather than the sole model-selection mechanism. The final model is selected using cross-validated F1 Score among configurations that also produce usable holdout performance.

This distinction matters because a model can look strong on one holdout metric while being less stable during cross-validation. In this report, cross-validation provides the primary model-selection evidence, while holdout results confirm whether the selected configuration remains practically useful on unseen data.

The results show that no single configuration dominates every metric. Some approaches improve recall but reduce precision, while others are more conservative. This reinforces that monitoring-related turnover modeling should be interpreted as a decision-support layer, not as an automatic classification system.

## Precision–Recall Curve Analysis {#precision-recall}

```{r 16-precision-recall, echo=FALSE}
pr_df <- results_all %>%
  filter(.metric %in% c("precision", "recall", "f_meas")) %>%
  select(Model, Sampling, .metric, .estimate) %>%
  pivot_wider(names_from = .metric, values_from = .estimate) %>%
  filter(!is.na(precision), !is.na(recall), !is.na(f_meas)) %>%
  mutate(
    Model = factor(as.character(Model), levels = names(model_colors)),
    Sampling = factor(as.character(Sampling), levels = sampling_order),
    Model_Config = paste(Model, Sampling, sep = " + "),
    Is_Selected = Model_Config == top_model_config
  )

selected_point <- pr_df %>%
  dplyr::filter(Is_Selected) %>%
  utils::head(1)

p <- ggplot(
  pr_df,
  aes(
    x = recall,
    y = precision,
    color = Model,
    shape = Sampling
  )
) +
  geom_vline(
    xintercept = 0.50,
    linetype = "dashed",
    color = "#D5DBDB",
    linewidth = 0.5
  ) +
  geom_hline(
    yintercept = 0.50,
    linetype = "dashed",
    color = "#D5DBDB",
    linewidth = 0.5
  ) +
  annotate(
    "rect",
    xmin = 0.45,
    xmax = 0.65,
    ymin = 0.45,
    ymax = 0.65,
    alpha = 0.08,
    fill = "#2A9D8F"
  ) +
  annotate(
    "text",
    x = 0.55,
    y = 0.67,
    label = "Balanced Operational Zone",
    size = 4.2,
    fontface = "bold",
    color = "#1B4332"
  ) +
  geom_point(
    aes(size = f_meas),
    alpha = 0.72
  ) +
  geom_point(
    data = selected_point,
    aes(x = recall, y = precision),
    inherit.aes = FALSE,
    shape = 21,
    size = 7,
    stroke = 1.3,
    fill = "white",
    color = "#1F2D3D"
  ) +
  geom_label(
    data = selected_point,
    aes(
      x = pmin(recall + 0.06, 0.88),
      y = pmin(precision + 0.07, 0.92),
      label = paste0(
        "Selected Model\n",
        Model_Config,
        "\nF1 = ",
        scales::number(f_meas, accuracy = 0.001)
      )
    ),
    inherit.aes = FALSE,
    size = 3.1,
    fontface = "bold",
    color = "#243447",
    fill = "white",
    label.size = 0.25
  ) +
  scale_color_manual(values = model_colors) +
  scale_shape_manual(
    values = c(
      "No Sampling" = 16,
      "Up Sampling" = 17,
      "Down Sampling" = 15,
      "SMOTE" = 3
    )
  ) +
  scale_size_continuous(
    range = c(3.2, 6.8),
    guide = "none"
  ) +
  scale_x_continuous(
    limits = c(0, 1),
    labels = scales::percent_format(accuracy = 1)
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    labels = scales::percent_format(accuracy = 1)
  ) +
  labs(
    title = "Precision–Recall Tradeoff by Model Configuration",
    subtitle = "Holdout results compare detection capability against prediction reliability",
    x = "Recall: Share of monitoring-related risk cases detected",
    y = "Precision: Reliability of monitoring-risk predictions",
    color = "Model",
    shape = "Sampling"
  ) +
  guides(
    color = guide_legend(order = 1, override.aes = list(size = 4, alpha = 1)),
    shape = guide_legend(order = 2, override.aes = list(size = 4, alpha = 1))
  ) +
  theme_hr(11) +
  theme(
    plot.title = element_text(face = "bold", size = 15),
    plot.subtitle = element_text(
  size = 12,
  color = "#5a6b7c",
  margin = margin(b = 12)
),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 9),
    legend.position = "right",
    legend.box = "vertical",
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 8.5),
    panel.grid.minor = element_blank(),
    plot.margin = margin(10, 20, 15, 10)
  )

print(p)
save_plot(p, "precision_recall_tradeoff.png", 10.5, 5.8)
```

The precision–recall view shows how each model configuration balances detection capability against prediction reliability. Configurations farther to the right detect more employees with monitoring-related turnover risk, while configurations higher on the chart produce more reliable positive predictions.

The selected configuration is highlighted with an outlined marker. It is not chosen because it maximizes a single metric in isolation; rather, it provides the strongest cross-validated F1 Score among configurations that also produced usable holdout performance.

This reinforces the business logic behind the model-selection process: the goal is not to flag as many employees as possible or eliminate all false positives, but to create a practical monitoring-risk screening layer that can guide earlier review without replacing human judgment.

## Cross-Validated Model Comparison

```{r 17-model-decision-breakdown, echo=FALSE}
decision_table <- cv_results_all %>%
  group_by(Model, Sampling) %>%
  summarise(
    Accuracy = mean(mean[.metric == "accuracy"], na.rm = TRUE),
    Precision = mean(mean[.metric == "precision"], na.rm = TRUE),
    Recall = mean(mean[.metric == "recall"], na.rm = TRUE),
    F1 = mean(mean[.metric == "f_meas"], na.rm = TRUE),
    ROC_AUC = mean(mean[.metric == "roc_auc"], na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(F1))

decision_table %>%
  dplyr::slice_head(n = 12) %>%
  gt() %>%
  tab_header(
    title = "Cross-Validated Model Comparison",
    subtitle = "Top configurations ranked by cross-validated F1 Score"
  ) %>%
  fmt_number(
    columns = c(Accuracy, Precision, Recall, F1, ROC_AUC),
    decimals = 3
  ) %>%
  data_color(
    columns = c(Accuracy, Precision, Recall, F1, ROC_AUC),
    colors = scales::col_numeric(
      palette = c("#EAF2F8", "#A9CCE3", "#2E86C1"),
      domain = c(0, 1)
    )
  ) %>%
  cols_align("center") %>%
  tab_options(
    table.width = pct(92),
    table.font.size = px(12),
    data_row.padding = px(8)
  )
```

### Model Selection {#model-selection}

Model selection was guided by a two-step rule designed to balance statistical rigor with business usefulness.

First, candidate configurations were ranked by cross-validated F1 Score, which serves as the primary model-selection metric because it balances precision and recall under class imbalance. Second, the leading candidates were reviewed on the holdout test set to confirm that they still produced usable positive-class predictions on unseen data.

This second step matters. A configuration can rank highly during resampling yet behave too aggressively or too weakly on the holdout set. For this reason, the final model is selected as the strongest cross-validated configuration that also satisfies minimum holdout stability requirements.

In this run, the recommended configuration was **`r top_model_name` with `r top_sampling_method`**. The model should be interpreted as a **risk prioritization tool**, meaning it helps identify employees or workforce segments that may warrant earlier privacy, trust, or monitoring-governance review, while leaving final decisions to HR leaders, managers, governance stakeholders, and organizational context.

## Selected Model Holdout Performance

```{r 18-final-model-metrics-plot, echo=FALSE}
final_model_plot_df <- final_model_metrics %>%
  mutate(
    Metric = factor(
      Metric,
      levels = c("Accuracy", "Precision", "Recall", "F1 Score", "ROC AUC")
    ),
    Highlight = ifelse(Metric == "F1 Score", "Primary Selection Metric", "Supporting Metric"),
    Plot_Label = ifelse(is.na(Score), "N/A", scales::number(Score, accuracy = 0.001))
  )

y_max <- max(0.70, max(final_model_plot_df$Score, na.rm = TRUE) + 0.10)

p <- ggplot(final_model_plot_df, aes(x = Metric, y = Score, fill = Highlight)) +
  geom_col(width = 0.62, na.rm = TRUE) +
  geom_text(
    aes(label = Plot_Label),
    vjust = -0.4,
    fontface = "bold",
    size = 4
  ) +
  scale_fill_manual(values = c(
    "Primary Selection Metric" = "#2E86C1",
    "Supporting Metric" = "#AAB7B8"
  )) +
  scale_y_continuous(
    limits = c(0, y_max),
    labels = scales::percent_format(accuracy = 1),
    expand = expansion(mult = c(0, 0.04))
  ) +
  labs(
    title = paste("Final Holdout Performance:", top_model_name),
    subtitle = paste("Sampling Strategy:", top_sampling_method),
    x = NULL,
    y = "Score"
  ) +
  theme_hr(11) +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(color = "#5a6b7c"),
    axis.text.x = element_text(face = "bold")
  )

print(p)
save_plot(p, "final_model_metrics.png", 8, 4.6)
```

## Risk Distribution

The risk distribution shows how the selected model assigns monitoring-related turnover-risk probabilities across the holdout test set. These scores should be interpreted as **review-priority signals**, not definitive employee labels.

Because monitoring-related risk is tied to privacy perceptions, trust, workload pressure, and governance context, probability scores should be interpreted as prioritization signals that help identify where additional HR or monitoring-governance review may be warranted.

```{r 19-risk-distribution, echo=FALSE}
risk_distribution <- final_preds %>%
  mutate(
    predicted_probability = .pred_Yes,
    risk_percentile = percent_rank(predicted_probability),
    review_priority = case_when(
      risk_percentile >= 0.90 ~ "High Review Priority",
      risk_percentile >= 0.60 ~ "Moderate Review Priority",
      TRUE ~ "Lower Review Priority"
    ),
    review_priority = factor(
      review_priority,
      levels = c(
        "Lower Review Priority",
        "Moderate Review Priority",
        "High Review Priority"
      )
    )
  )

p <- ggplot(
  risk_distribution,
  aes(x = predicted_probability, fill = review_priority)
) +
  geom_histogram(
    bins = 28,
    color = "white",
    alpha = 0.90
  ) +
  scale_fill_manual(
    values = c(
      "Lower Review Priority" = "#D5E8D4",
      "Moderate Review Priority" = "#F9E79F",
      "High Review Priority" = "#F5B7B1"
    )
  ) +
  scale_x_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, 1),
    expand = expansion(mult = c(0.01, 0.01))
  ) +
  labs(
    title = "Predicted Monitoring-Related Turnover Review-Priority Distribution",
    subtitle = "Relative workforce-risk signals across the holdout population",
    x = "Predicted Probability of Monitoring-Related Turnover Risk",
    y = "Employee Count",
    fill = "Review Priority"
  ) +
  theme_hr(11) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(color = "#5a6b7c", size = 11),
    legend.position = "top",
    panel.grid.minor = element_blank(),
    plot.margin = margin(10, 20, 15, 10)
  )

print(p)
save_plot(p, "risk_distribution.png", 8.6, 5.2)
```

## Review-Priority Segmentation

The segmentation view translates model probabilities into practical review-priority tiers. These tiers are not automatic decisions; they organize attention for HR, leadership, governance, or workforce planning review.

Because monitoring-related risk is sensitive and context-dependent, relative ranking is more useful than rigid fixed thresholds. In practice, HR and governance teams could adjust review queues based on available capacity, privacy sensitivity, intervention goals, and tolerance for missed risk.

```{r 20-risk-segmentation, echo=FALSE, message=FALSE, warning=FALSE}
risk_segment_summary <- risk_distribution %>%
  count(review_priority, .drop = FALSE) %>%
  mutate(
    percent = n / sum(n),
    label = paste0(n, "\n", scales::percent(percent, accuracy = 1))
  )

p <- ggplot(
  risk_segment_summary,
  aes(
    x = review_priority,
    y = n,
    fill = review_priority
  )
) +
  geom_col(width = 0.65) +
  geom_text(
    aes(label = label),
    vjust = -0.45,
    fontface = "bold",
    size = 4
  ) +
  scale_fill_manual(
    values = c(
      "Lower Review Priority" = "#7FBF7F",
      "Moderate Review Priority" = "#F39C12",
      "High Review Priority" = "#D35454"
    )
  ) +
  scale_y_continuous(
    limits = c(0, max(risk_segment_summary$n, na.rm = TRUE) * 1.18),
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(
    title = "Employee Monitoring-Risk Review-Priority Segmentation",
    subtitle = "Risk tiers translate model scores into practical privacy and trust review priorities",
    x = NULL,
    y = "Employee Count"
  ) +
  theme_hr(11) +
  theme(
    plot.title = element_text(face = "bold", size = 15),
    plot.subtitle = element_text(color = "#5a6b7c", size = 11),
    axis.text.x = element_text(size = 10, face = "bold"),
    legend.position = "none",
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    plot.margin = margin(10, 18, 18, 10)
  )

print(p)
save_plot(p, "risk_segments.png", 7.8, 5)
```

For monitoring-related turnover risk, these tiers should be treated as privacy, trust, and governance prioritization signals. Lower-priority employees may require normal monitoring-governance communication, moderate-priority employees may warrant trend monitoring, and high-priority employees may warrant earlier HR or governance review.

# Final Model Summary & Key Takeaways {#model-summary}

The selected model represents the strongest cross-validated F1 Score among configurations that also met minimum holdout stability requirements. This means the final model was not selected only because it performed well during resampling; it also had to produce usable positive-class predictions on unseen data.

While the absolute performance values remain moderate, the model is useful as a **risk prioritization tool** because it helps rank employees by relative likelihood of monitoring-related turnover risk. This distinction is important: the model should not be used to make automatic employment decisions. Instead, it supports earlier privacy-aware review, monitoring-governance conversations, and more focused workforce trust management.

## 📊 Key Takeaways

```{=html}
<div class="model-highlight-card">

  <div class="model-badge">Recommended Model Selected</div>

  <h3 class="model-title">
    `r top_model_name`
  </h3>

  <p class="model-subtitle">
    Optimized using <strong>`r top_sampling_method`</strong> with a focus on <strong>F1 Score</strong>
  </p>

  <div class="model-metrics">
    <div class="metric-box">
      <span class="metric-label">CV F1 Score</span>
      <span class="metric-value">
        `r top_f1`
      </span>
    </div>

    <div class="metric-box">
      <span class="metric-label">Strategy</span>
      <span class="metric-value">
        `r selection_strategy_label`
      </span>
    </div>

    <div class="metric-box">
      <span class="metric-label">Use Case</span>
      <span class="metric-value">
        Monitoring & Privacy Risk Targeting
      </span>
    </div>
  </div>

</div>
```

## Why This Model Was Selected

```{=html}
<div class="decision-box">

<h3>Why This Model Was Selected</h3>

<p>
The final model was selected using a two-step process: cross-validation identified strong candidate configurations, and the holdout test set confirmed whether those candidates remained practically usable on unseen data.
</p>

<p>
In this run, the selected configuration was <strong>`r top_model_name` with `r top_sampling_method`</strong>. It achieved a cross-validated F1 Score of <strong>`r top_f1`</strong>, while its holdout test results produced precision of <strong>`r top_holdout_precision`</strong>, recall of <strong>`r top_holdout_recall`</strong>, and F1 Score of <strong>`r top_holdout_f1`</strong>.
</p>

<p>
`r selection_rule_note`
</p>

<p>
This selection rule balances model-development rigor with practical business usefulness. Cross-validation identifies the strongest generalizable configuration, while the holdout stability check ensures the final model still produces usable predictions on unseen data.
</p>

<p>
Because the model is designed for workforce risk screening, predictions should be treated as prompts for further HR or manager review, not as standalone conclusions about employee behavior.
</p>

</div>
```

## Selected Model Performance Metrics

The table below reports holdout test performance for the selected configuration. These values are not used alone to select the model; instead, they confirm whether the cross-validated winner remains practically usable on unseen data.

```{r 21-final-model-performance-table, echo=FALSE}
final_model_metrics %>%
  select(Metric, Score) %>%
  mutate(
    Score = ifelse(is.na(Score), NA_real_, Score)
  ) %>%
  gt() %>%
  tab_header(
    title = "Selected Model Performance Summary",
    subtitle = paste("Holdout test performance for", top_model_name, "with", top_sampling_method)
  ) %>%
  fmt_number(columns = Score, decimals = 3) %>%
  sub_missing(columns = Score, missing_text = "N/A") %>%
  cols_align("center") %>%
  tab_options(
    table.width = pct(55),
    table.font.size = px(13)
  )
```

## Top Model Configurations

```{r 22-top-models-table, echo=FALSE}
top_config_table <- cv_results_all %>%
  filter(.metric == "f_meas") %>%
  transmute(
    Model = as.character(Model),
    Sampling = as.character(Sampling),
    `CV F1 Score` = round(mean, 3)
  ) %>%
  left_join(
    holdout_screen %>%
      transmute(
        Model,
        Sampling,
        `Holdout Precision` = round(precision, 3),
        `Holdout Recall` = round(recall, 3),
        `Holdout F1` = round(f_meas, 3),
        Status = case_when(
          Model == top_model_name & Sampling == top_sampling_method ~ "Selected Final Model",
          stable_holdout ~ "Holdout-Stable Alternative",
          viable_holdout ~ "Viable but Below Stability Screen",
          TRUE ~ "Not Selected"
        )
      ),
    by = c("Model", "Sampling")
  ) %>%
  arrange(desc(`CV F1 Score`), desc(`Holdout F1`)) %>%
  slice_head(n = 10)

top_config_table %>%
  gt() %>%
  tab_header(
    title = "Top Model Configurations",
    subtitle = "Cross-validated ranking with holdout confirmation"
  ) %>%
  fmt_number(
    columns = c(`CV F1 Score`, `Holdout Precision`, `Holdout Recall`, `Holdout F1`),
    decimals = 3
  ) %>%
  sub_missing(
    columns = c(`Holdout Precision`, `Holdout Recall`, `Holdout F1`),
    missing_text = "N/A"
  ) %>%
  cols_align("center") %>%
  data_color(
    columns = c(`CV F1 Score`, `Holdout Precision`, `Holdout Recall`, `Holdout F1`),
    colors = scales::col_numeric(
      palette = c("#EAF2F8", "#A9CCE3", "#2E86C1"),
      domain = c(0, 1),
      na.color = "#F2F4F4"
    )
  ) %>%
  cols_width(
    Model ~ px(130),
    Sampling ~ px(150),
    `CV F1 Score` ~ px(120),
    `Holdout Precision` ~ px(130),
    `Holdout Recall` ~ px(120),
    `Holdout F1` ~ px(110),
    Status ~ px(220)
  ) %>%
  tab_style(
    style = list(
      cell_fill(color = "#EBF5FB"),
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      rows = Status == "Selected Final Model"
    )
  ) %>%
  tab_options(
    table.width = pct(92),
    table.font.size = px(12)
  )
```

Performance metrics indicate which configurations screen monitoring-related turnover risk most effectively. The next step is to translate those outputs into practical workforce trust, privacy, and monitoring-governance insight by connecting model behavior back to broader organizational patterns.

This section interprets the likely drivers of monitoring-related turnover risk and links those patterns to actionable workforce response areas.

# Driver Interpretation {#driver-interpretation}

Because the selected predictive workflow does not provide a straightforward, model-agnostic feature-importance ranking, driver interpretation is based on the broader analytical pattern observed across the modeling and exploratory analysis.

The results suggest that monitoring-based turnover risk is not driven by one isolated factor. Instead, risk appears to reflect a combination of **privacy perceptions, monitoring intensity, trust, workload, wellbeing, and organizational support factors**.

This is important from a business perspective because monitoring risk should not be treated only as a technology or compliance issue. Effective monitoring governance requires coordinated action across transparency, fairness, privacy protection, manager communication, workload expectations, and employee trust.

## Feature → Action Mapping {#feature-action}

To translate model insights into practical action, key risk areas are mapped to targeted workforce interventions. This bridges the gap between predictive analytics and real-world decision-making by connecting model interpretation to practical HR response strategies.

```{r 23-driver-action-table, echo=FALSE}
driver_actions <- tibble::tribble(
  ~Risk_Area, ~What_It_Suggests, ~Recommended_Action,
  "Monitoring intensity and perceived surveillance",
  "Employees exposed to higher monitoring intensity may experience reduced autonomy, increased pressure, or heightened surveillance concerns.",
  "Review monitoring scope, necessity, transparency, and proportionality before expanding monitoring practices.",

  "Privacy concerns and trust",
  "Higher privacy concern may reduce trust in leadership and increase monitoring-related turnover risk.",
  "Improve communication about data use, access controls, retention policies, and employee privacy protections.",

  "Workload, stress, and wellbeing",
  "Monitoring may compound existing strain when employees already feel overloaded, unsupported, or closely scrutinized.",
  "Pair monitoring programs with workload review, wellbeing resources, and manager support practices.",

  "Manager and organizational support",
  "Employees may tolerate monitoring better when they perceive fair treatment, support, and a clear business purpose.",
  "Train managers to explain monitoring practices, respond to concerns, and avoid punitive use of productivity data.",

  "Governance and fairness",
  "Unclear or inconsistent monitoring policies may increase perceived unfairness and risk.",
  "Establish transparent governance, audit monitoring impacts, and evaluate whether tools affect employee groups differently."
)

driver_actions %>%
  gt() %>%
  tab_header(
    title = "Monitoring & Privacy Risk Driver-to-Action Map",
    subtitle = "Translating model insights into workforce trust and monitoring governance strategy"
  ) %>%
  cols_label(
    Risk_Area = "Risk Area",
    What_It_Suggests = "What It Suggests",
    Recommended_Action = "Recommended Action"
  ) %>%
  cols_width(
    Risk_Area ~ px(190),
    What_It_Suggests ~ px(360),
    Recommended_Action ~ px(430)
  ) %>%
  tab_options(
    table.width = pct(90),
    table.font.size = px(13)
  )
```

# Business Impact & Decision Guidance {#business-impact}

The modeling framework is designed to support workforce decision-making at a strategic level while preserving employee privacy through synthetic data simulation.

Rather than functioning as an automated employment-decision system, the selected model should be interpreted as a workforce intelligence layer that helps HR, leadership, and governance stakeholders prioritize where additional privacy, trust, monitoring-policy, or retention review may be beneficial.

The practical value of the framework lies in its ability to:

- Identify workforce segments that may warrant earlier monitoring-governance review
- Support proactive privacy and trust risk management
- Improve allocation of HR, leadership, and governance review resources
- Enhance organizational visibility into monitoring-related workforce-risk patterns
- Demonstrate privacy-preserving HR analytics workflows

Importantly, the model is intended to support human decision-making, not replace it. Monitoring-related workforce outcomes remain dependent on organizational context, leadership interpretation, employee communication, governance standards, and ethical review processes.

This project demonstrates how predictive analytics can be integrated into HR environments responsibly using synthetic, research-informed data that protects employee confidentiality while still enabling advanced workforce analysis.

## Strategic Decision Guidance {#decision-guidance}

```{=html}
<div class="decision-box">

<p>
This model is best deployed as a <strong>risk prioritization tool</strong>, not as a standalone decision system.
</p>

<ul>
  <li><strong>High-priority employees or groups</strong> → Prioritize for HR, manager, privacy, or monitoring-governance review</li>
  <li><strong>Moderate-priority employees or groups</strong> → Monitor privacy concerns, workload, trust, and support trends</li>
  <li><strong>Lower-priority employees or groups</strong> → Maintain baseline monitoring communication, trust-building, and governance practices</li>
</ul>

<p>
The value of the model is not in replacing human judgment, but in helping decision-makers focus attention earlier and more consistently.
</p>

</div>
```

## Model Limitations {#model-limitations}

```{=html}
<div class="decision-box">

<p>
The model should be interpreted within the context of synthetic data, simulated workforce patterns, and organizational variability.
</p>

<ul>
  <li>Predictions reflect patterns in synthetic data, not observed outcomes from a real organization</li>
  <li>Performance may shift when applied to different employee populations, industries, or monitoring environments</li>
  <li>Risk scores should support review and prioritization, not automatic employment decisions</li>
  <li>Human context, privacy governance, and ethical review remain essential before any intervention is taken</li>
</ul>

</div>
```

## Final Takeaway

This project demonstrates how synthetic, research-informed workforce data can support realistic HR analytics workflows without exposing sensitive employee information.

Monitoring-related turnover risk is not driven by a single variable. It reflects a network of privacy, trust, monitoring, workload, wellbeing, and organizational support signals that collectively shape how employees experience workplace surveillance.

The modeling framework shows how predictive analytics can move beyond static reporting and contribute to proactive workforce trust management, early risk visibility, and operational prioritization while maintaining strong privacy-preserving principles.

Although predictive performance remains intentionally moderate and realistic, the selected workflow successfully demonstrates how monitoring-related workforce-risk modeling can function as a practical decision-support capability within HR and governance environments.

Used appropriately, this type of framework can help organizations identify emerging monitoring-related workforce patterns earlier, allocate review resources more effectively, strengthen transparency and trust, and support more informed strategic HR decision-making.

```{r 24-export-final-model-summary, include=FALSE}
summary_dir <- here::here("data", "processed", "model_summaries")
dir.create(summary_dir, recursive = TRUE, showWarnings = FALSE)

top_holdout_accuracy <- final_model_metrics %>%
  filter(Metric == "Accuracy") %>%
  pull(Score) %>%
  round(3)

final_model_summary <- tibble::tibble(
  Outcome = "Monitoring & Privacy Risk",
  Model = top_model_name,
  Sampling = top_sampling_method,
  Model_Config = top_model_config,
  Selection_Strategy = selection_strategy_label,
  Selection_Rule = selection_rule_note,
  CV_F1 = top_f1,
  Holdout_Accuracy = top_holdout_accuracy,
  Holdout_Precision = top_holdout_precision,
  Holdout_Recall = top_holdout_recall,
  Holdout_F1 = top_holdout_f1,
  Holdout_ROC_AUC = top_holdout_roc_auc
)

saveRDS(
  final_model_summary,
  here::here("data", "processed", "model_summaries", "monitoring_privacy_model_summary.rds")
)
```

```{=html}
<hr class="section-divider">

<div class="page-nav">

  <div class="nav-left">
    <a href="HR_Data_Quiet_Quitting.html" class="cta-btn btn-blue">← Back</a>
  </div>

  <div class="nav-center">
    <a href="index.html" class="cta-btn btn-teal">Home</a>
  </div>

  <div class="nav-right">
    <a href="HR_Data_Insights_Gleaned.html" class="cta-btn btn-red">Next →</a>
  </div>

</div>

</div> <!-- page-content -->
</div> <!-- page-layout -->
</div> <!-- main-container -->
```