---
title: "HR Predictive Analytics Project – Quiet Quitting"
author: "Kevin Allen Cauley"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    theme: flatly
    highlight: tango
    df_print: paged
---

```{r 00-setup, include=FALSE}
# Set working directory to the project root
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

# Global chunk options
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.path = "figures/",
  fig.align = "center",
  cache = TRUE
)

# Load previous session
if (file.exists("outputs/my_environment.RData")) {
  load("outputs/my_environment.RData")
}
```

```{r 00-setup-packages, include=FALSE}
# Load packages
library(car)
library(cowplot)
library(caret)
library(corrplot)
library(discrim)
library(dplyr)
library(e1071)
library(fastDummies)
library(forcats)
library(gbm)
library(GGally)
library(ggcorrplot)
library(ggplot2)
library(ggthemes)
library(glmnet)
library(grid)
library(gridExtra)
library(iml)
library(kableExtra)
library(keras3)
library(keras)
library(knitr)
library(mgcv)
library(naivebayes)
library(naniar)
library(patchwork)
library(pheatmap)
library(plotly)
library(psych)
library(randomForest)
library(ranger)
library(RColorBrewer)
library(readxl)
library(reshape2)
library(skimr)
library(tensorflow)
library(themis)
library(tidymodels)
library(tidyverse)
library(vip)
library(viridis)
library(xgboost)
```


## Predictive Modeling on Quiet Quitting

The goal of this modeling section is to predict **Quiet Quitting** using a set of behavioral, demographic, and engagement-related predictors.

We will evaluate a range of classification models using multiple **resampling strategies** (No Sampling, Up-Sampling, Down-Sampling, and SMOTE). Models will be evaluated using six classification metrics:

-   Accuracy
-   Precision
-   Recall
-   F1 Score
-   Specificity
-   ROC AUC

This approach ensures fairness, robustness, and clarity in identifying the most effective model and configuration.

### Shared Model Setup - Quiet Quitting

```{r 04-logit-shared-model-setup}
# Ensure Quiet Quitting is a Factor
hr_df <- hr_df %>%
  mutate(quiet_quitting = as.factor(quiet_quitting))

# Set Seed
set.seed(42)

# 80/20 train/test split with stratification
quiet_split <- initial_split(hr_df, prop = 0.8, strata = Quiet_Quitting)
quiet_train_data <- training(quiet_split)
quiet_test_data  <- testing(quiet_split)

# 10-Fold Cross Validation
quiet_folds <- vfold_cv(quiet_train_data, v = 10, strata = Quiet_Quitting)

# Shared Pre-processing (no resampling)
base_recipe <- recipe(Quiet_Quitting ~ ., data = quiet_train_data) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_numeric_predictors())

# Basic Logistic Model
logit_model <- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")
```

```{r 04-logit-quiet-quitting-shared-metrics-setup}
# Shared metric set for all models
all_metrics <- metric_set(accuracy, precision, recall, f_meas, yardstick::spec, roc_auc)
```

## Quiet Quitting with Logistic Regression

### Logistic Regression - No Resampling

```{r 04-logit-none-workflow, message=FALSE}
# Workflow with Base Recipe
workflow_logit_none <- workflow() %>%
  add_model(logit_model) %>%
  add_recipe(base_recipe)

# Fit to training data
fit_logit_none <- fit(workflow_logit_none, data = quiet_train_data)

# Predict on test data
logit_preds_none <- predict(fit_logit_none, new_data = quiet_test_data, type = "prob") %>%
  bind_cols(predict(fit_logit_none, new_data = quiet_test_data)) %>%
  bind_cols(quiet_test_data %>% select(Quiet_Quitting))
```

### Logistic Regression - No Resampling - Performance Metrics

```{r 04-logit-none-metrics, echo=TRUE}
# Metrics
logit_none_metrics <- all_metrics(logit_preds_none, truth = Quiet_Quitting, estimate = .pred_class, .pred_Yes)

# Plot
logit_none_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#4B9CD3", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – LR (No Resampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

```{r 04-logit-none-confmat, echo=TRUE}
library(ggplot2)
library(yardstick)
library(forcats)

# Confusion matrix to dataframe
cm_logit_none <- conf_mat(logit_preds_none, truth = Quiet_Quitting, estimate = .pred_class)
cm_df <- as.data.frame(cm_logit_none$table)

# Ensure proper order for plotting
cm_df$Truth <- factor(cm_df$Truth, levels = c("Yes", "No"))         
cm_df$Prediction <- factor(cm_df$Prediction, levels = c("No", "Yes")) 

# Plot
ggplot(cm_df, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#e6f2ff", high = "#1A5276", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – LR (No Resampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )
```

```{r 04-logit-none-roc-curve, echo=TRUE}
# Generate ROC curve
roc_curve(logit_preds_none, truth = Quiet_Quitting, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – LR (No Resampling)",
    subtitle = "Model performance on predicting Quiet Quitting",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )
```

```{r 04-logit-none-vip, echo=TRUE}
# Extract full importance and then limit to top 15
vip_logit_none <- fit_logit_none %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance))

# Plot
ggplot(vip_logit_none, aes(x = Importance, y = Variable)) +
  geom_col(fill = "#1A5276", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – LR (No Resampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

### Logistic Regression – SMOTE

```{r 04-logit-smote-workflow, message=FALSE}
# SMOTE recipe
recipe_logit_smote <- base_recipe %>%
  step_smote(Quiet_Quitting)

workflow_logit_smote <- workflow() %>%
  add_model(logit_model) %>%
  add_recipe(recipe_logit_smote)

# 10-Fold CV
set.seed(42)
logit_cv_smote <- fit_resamples(
  workflow_logit_smote,
  resamples = quiet_folds,
  metrics = metric_set(accuracy, precision, recall, f_meas, yardstick::spec, roc_auc)
)

# Final model fit and prediction on test data for SMOTE
fit_logit_smote <- fit(workflow_logit_smote, data = quiet_train_data)

logit_preds_smote <- predict(fit_logit_smote, new_data = quiet_test_data, type = "prob") %>%
  bind_cols(predict(fit_logit_smote, new_data = quiet_test_data)) %>%
  bind_cols(quiet_test_data %>% select(Quiet_Quitting))
```

### Logistic Regression – SMOTE - Performance Metrics

```{r 04-logit-smote-finalfit, message=FALSE}
# Metrics
logit_smote_metrics <- all_metrics(logit_preds_smote, truth = Quiet_Quitting, estimate = .pred_class, .pred_Yes)

# Plot
logit_smote_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#4B9CD3", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – LR (SMOTE)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

```{r 04-logit-smote-confmat, echo=TRUE}
# Create confusion matrix
cm_logit_smote <- conf_mat(logit_preds_smote, truth = Quiet_Quitting, estimate = .pred_class)
cm_df_smote <- as.data.frame(cm_logit_smote$table)

# Order axes
cm_df_smote$Truth <- factor(cm_df_smote$Truth, levels = c("Yes", "No"))
cm_df_smote$Prediction <- factor(cm_df_smote$Prediction, levels = c("No", "Yes"))

# Plot heatmap
ggplot(cm_df_smote, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#e6f2ff", high = "#1A5276", name = "Frequency") +
  scale_y_discrete(limits = rev) +  # YES on top
  labs(
    title = "Confusion Matrix – LR (SMOTE)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )
```

```{r 04-logit-smote-roc, echo=TRUE}
# ROC Curve
roc_curve(logit_preds_smote, truth = Quiet_Quitting, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – LR (SMOTE)",
    subtitle = "Model performance on predicting Quiet Quitting",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )
```

```{r 04-logit-smote-vip, echo=TRUE}
# Top 15 features only
fit_logit_smote %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#1A5276", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – LR (SMOTE)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

### Logistic Regression - Upsampling

```{r 04-logit-upsample-workflow, message=FALSE}
# Upsampling Recipe
recipe_logit_upsample <- base_recipe %>%
  step_upsample(Quiet_Quitting)

# Workflow 
workflow_logit_upsample <- workflow() %>%
  add_model(logit_model) %>%
  add_recipe(recipe_logit_upsample)

# Fit
fit_logit_upsample <- fit(workflow_logit_upsample, data = quiet_train_data)

# Predict
logit_preds_upsample <- predict(fit_logit_upsample, new_data = quiet_test_data, type = "prob") %>%
  bind_cols(predict(fit_logit_upsample, new_data = quiet_test_data)) %>%
  bind_cols(quiet_test_data %>% select(Quiet_Quitting))
```

### Logistic Regression - Upsampling - Performance Metrics

```{r 04-logit-upsample-metrics, echo=TRUE}
# Metrics
logit_upsample_metrics <- all_metrics(logit_preds_upsample, truth = Quiet_Quitting, estimate = .pred_class, .pred_Yes)

# Plot
logit_upsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#4B9CD3", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – LR (Upsampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

```{r 04-logit-upsample-confmat, echo=TRUE}
# Confusion matrix data prep
cm_logit_upsample <- conf_mat(logit_preds_upsample, truth = Quiet_Quitting, estimate = .pred_class)
cm_df_upsample <- as.data.frame(cm_logit_upsample$table)

# Set factor levels to match layout: Yes on top, No bottom; No left, Yes right
cm_df_upsample$Truth <- factor(cm_df_upsample$Truth, levels = c("Yes", "No"))
cm_df_upsample$Prediction <- factor(cm_df_upsample$Prediction, levels = c("No", "Yes"))

# Plot
ggplot(cm_df_upsample, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#e6f2ff", high = "#1A5276", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – LR (Upsampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )
```

```{r 04-logit-upsample-roc-curve, echo=TRUE}
# Roc
roc_curve(logit_preds_upsample, truth = Quiet_Quitting, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – LR (Upsampling)",
    subtitle = "Model performance on predicting Quiet Quitting",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )
```

```{r 04-logit-upsample-vip, echo=TRUE}
# Styled VIP for top 15 features
fit_logit_upsample %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#1A5276", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – LR (Upsampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

### Logistic Regression - Downsampling

```{r 04-logit-downsample-workflow, message=FALSE}
# Downsampling Recipe
recipe_logit_downsample <- base_recipe %>%
  step_downsample(Quiet_Quitting)

# Workflow 
workflow_logit_downsample <- workflow() %>%
  add_model(logit_model) %>%
  add_recipe(recipe_logit_downsample)

# Fit on training data
fit_logit_downsample <- fit(workflow_logit_downsample, data = quiet_train_data)

# Predict on test data
logit_preds_downsample <- predict(fit_logit_downsample, new_data = quiet_test_data, type = "prob") %>%
  bind_cols(predict(fit_logit_downsample, new_data = quiet_test_data)) %>%
  bind_cols(quiet_test_data %>% select(Quiet_Quitting))
```

### Logistic Regression - Downsampling - Performance Metrics

```{r 04-logit-downsample-metrics, echo=TRUE}
# Metrics
logit_downsample_metrics <- all_metrics(logit_preds_downsample, truth = Quiet_Quitting, estimate = .pred_class, .pred_Yes)

# Plot
logit_downsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#4B9CD3", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – LR (Downsampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

```{r 04-logit-downsample-confmat, echo=TRUE}
cm_logit_down <- conf_mat(logit_preds_downsample, truth = Quiet_Quitting, estimate = .pred_class)
cm_df_down <- as.data.frame(cm_logit_down$table)

cm_df_down$Truth <- factor(cm_df_down$Truth, levels = c("Yes", "No"))
cm_df_down$Prediction <- factor(cm_df_down$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_down, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#e6f2ff", high = "#1A5276", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – LR (Downsampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )
```

```{r 04-logit-downsample-roc-curve, echo=TRUE}
roc_curve(logit_preds_downsample, truth = Quiet_Quitting, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – LR (Downsampling)",
    subtitle = "Model performance on predicting Quiet Quitting",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )
```

```{r 04-logit-downsample-vip, echo=TRUE}
fit_logit_downsample %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#1A5276", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – LR (Downsampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

### Results Table for Logistic Regression (Quiet Quitting)

```{r 04-logit-results-table, results='asis', echo=FALSE}
logit_results_long <- bind_rows(
  logit_none_metrics %>% mutate(Resample = "No Sampling"),
  logit_smote_metrics %>% mutate(Resample = "SMOTE"),
  logit_upsample_metrics %>% mutate(Resample = "Up Sampling"),
  logit_downsample_metrics %>% mutate(Resample = "Down Sampling")
) %>%
  mutate(
    Metric = factor(.metric,
                    levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                    labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC")),
    Score = .estimate,
    Resample = factor(Resample, levels = c("No Sampling", "SMOTE", "Up Sampling", "Down Sampling"))
  ) %>%
  filter(!is.na(Score))

highlight_logit <- logit_results_long %>%
  group_by(Metric) %>%
  filter(Score == max(Score, na.rm = TRUE)) %>%
  mutate(is_best = TRUE)

logit_results_highlight <- left_join(logit_results_long, highlight_logit %>% select(Resample, Metric, is_best),
                                     by = c("Resample", "Metric")) %>%
  mutate(is_best = ifelse(is.na(is_best), FALSE, is_best))

ggplot(logit_results_highlight, aes(x = Resample, y = Score, fill = Metric)) +
  geom_col(width = 0.7) +
  geom_col(data = subset(logit_results_highlight, is_best),
           fill = NA, color = "black", linewidth = 1.2, width = 0.7) +
  facet_wrap(~ Metric, nrow = 2, scales = "free_y") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_manual(values = rep("#4B9CD3", 6)) +
  labs(
    title = "LR Performance by Metric and Resampling Method",
    subtitle = "Black outline shows the best result for each metric",
    x = "Resampling Strategy", y = "Score"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 30, hjust = 1),
    strip.text = element_text(face = "bold"),
    legend.position = "none"
  )
```

### Executive Summary for Logistic Regression (Quiet Quitting)

Logistic Regression provided a consistent and interpretable benchmark across all resampling strategies. It excelled under *No Resampling*, producing the best recall and F1 score — making it a strong candidate for early detection of quiet quitters. While more advanced models may outperform it in specific areas, Logistic Regression’s transparency makes it ideal for communicating insights to stakeholders.

- **No Resampling** achieved the **highest recall (0.924)** and **F1 score (0.854)**, indicating strong performance in capturing most disengaged employees — but at the cost of **low specificity (0.116)**, which may result in many false positives.

- **SMOTE** and **Upsampling** produced identical results, with a solid **precision (0.800)** and **F1 score (0.707)**. These strategies strike a better balance between over- and under-identification, making them suitable when precision matters more.

- **Downsampling** performed the weakest, with a drop in **recall (0.449)** and **F1 score (0.561)**, despite its relatively better **ROC AUC (0.503)** and **specificity (0.442)**. This makes it more conservative but less useful for broad detection efforts.

## Quiet Quitting with Random Forest

### Random Forest - No Resampling

```{r 04-rf-no-resampling-workflow, message=FALSE}
# Define Random Forest model
rf_model <- rand_forest(mtry = tune(), trees = 500, min_n = tune()) %>%
  set_engine("ranger", importance = "impurity") %>%
  set_mode("classification")

# Workflow using base recipe
recipe_rf_none <- base_recipe

# Combine into Workflow
workflow_rf_none <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(recipe_rf_none)

# Tune the RF model with cross-validation
set.seed(42)
rf_tuned <- tune_grid(
  workflow_rf_none,
  resamples = quiet_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_rf <- select_best(rf_tuned, metric = "roc_auc")

# Final workflow with best RF model
final_rf_workflow <- finalize_workflow(workflow_rf_none, best_rf)

# Fit to training data
rf_fit_none <- fit(final_rf_workflow, data = quiet_train_data)

# Predict on test data
rf_preds_none <- predict(rf_fit_none, new_data = quiet_test_data, type = "prob") %>%
  bind_cols(predict(rf_fit_none, new_data = quiet_test_data)) %>%
  bind_cols(quiet_test_data %>% select(Quiet_Quitting))
```

### Random Forest - No Resampling - Performance Metrics

```{r 04-rf-no-resampling-metrics, echo=TRUE}
# Metrics
rf_none_metrics <- all_metrics(rf_preds_none, truth = Quiet_Quitting, estimate = .pred_class, .pred_Yes)

# Plot
rf_none_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#3FA34D", width = 0.7) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – RF (No Resampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-rf-no-resampling-confmat, echo=TRUE}
rf_cm_none <- conf_mat(rf_preds_none, truth = Quiet_Quitting, estimate = .pred_class)
rf_cm_df_none <- as.data.frame(rf_cm_none$table)

rf_cm_df_none$Truth <- factor(rf_cm_df_none$Truth, levels = c("Yes", "No"))
rf_cm_df_none$Prediction <- factor(rf_cm_df_none$Prediction, levels = c("No", "Yes"))

ggplot(rf_cm_df_none, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#D5F5E3", high = "#145A32", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – RF (No Resampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )
```

```{r 04-rf-no-resampling-roc-curve, echo=TRUE}
# ROC Curve
roc_curve(rf_preds_none, truth = Quiet_Quitting, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray70") +
  labs(
    title = "ROC Curve – RF (No Resampling)",
    subtitle = "Model performance on predicting Quiet Quitting",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

```{r 04-rf-no-resampling-vip, echo=TRUE}
# Variable Importance
rf_fit_none %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#3FA34D", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – RF (No Resampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

### Random Forest - SMOTE

```{r 04-rf-smote-workflow, message=FALSE}
# Create Recipe from Base
recipe_rf_smote <- base_recipe %>%
  step_smote(Quiet_Quitting)

# Workflow with Random Forest
workflow_rf_smote <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(recipe_rf_smote)

# Tune the RF model with cross-validation
set.seed(42)
rf_tuned_smote <- tune_grid(
  workflow_rf_smote,
  resamples = quiet_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_rf_smote <- select_best(rf_tuned_smote, metric = "roc_auc")

# Final workflow with best RF model
final_rf_workflow_smote <- finalize_workflow(workflow_rf_smote, best_rf_smote)

# Fit to training data
rf_fit_smote <- fit(final_rf_workflow_smote, data = quiet_train_data)

# Predict on test data
rf_preds_smote <- predict(rf_fit_smote, new_data = quiet_test_data, type = "prob") %>%
  bind_cols(predict(rf_fit_smote, new_data = quiet_test_data)) %>%
  bind_cols(quiet_test_data %>% select(Quiet_Quitting))
```

### Random Forest - SMOTE - Performance Metrics

```{r 04-rf-smote-metrics, echo=TRUE}
# Metrics
rf_smote_metrics <- all_metrics(rf_preds_smote, truth = Quiet_Quitting, estimate = .pred_class, .pred_Yes)

# Plot
rf_smote_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#3FA34D", width = 0.7) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – RF (SMOTE)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-rf-smote-confmat, echo=TRUE}
# Confusion Matrix
rf_cm_smote <- conf_mat(rf_preds_smote, truth = Quiet_Quitting, estimate = .pred_class)
rf_cm_df_smote <- as.data.frame(rf_cm_smote$table)

rf_cm_df_smote$Truth <- factor(rf_cm_df_smote$Truth, levels = c("Yes", "No"))
rf_cm_df_smote$Prediction <- factor(rf_cm_df_smote$Prediction, levels = c("No", "Yes"))

ggplot(rf_cm_df_smote, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#D5F5E3", high = "#145A32", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – RF (SMOTE)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-rf-smote-roc-curve, echo=TRUE}
# ROC Curve
roc_curve(rf_preds_smote, truth = Quiet_Quitting, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray70") +
  labs(
    title = "ROC Curve: RF on Quiet Quitting (SMOTE)",
    subtitle = "Model performance on predicting Quiet Quitting",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

```{r 04-rf-smote-vip, echo=TRUE}
# Variable Importance
rf_fit_smote %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#3FA34D", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – RF (SMOTE)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

### Random Forest - Upsampling

```{r 04-rf-upsample-workflow, message=FALSE}
# Create Recipe from Base
recipe_rf_upsample <- base_recipe %>%
  step_upsample(Quiet_Quitting)

# Workflow with Random Forest
workflow_rf_upsample <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(recipe_rf_upsample)

# Tune the RF model with cross-validation
set.seed(42)
rf_tuned_upsample <- tune_grid(
  workflow_rf_upsample,
  resamples = quiet_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_rf_upsample <- select_best(rf_tuned_upsample, metric = "roc_auc")

# Final workflow with best RF model
final_rf_workflow_upsample <- finalize_workflow(workflow_rf_upsample, best_rf_upsample)

# Fit to training data
rf_fit_upsample <- fit(final_rf_workflow_upsample, data = quiet_train_data)

# Predict on test data
rf_preds_upsample <- predict(rf_fit_upsample, new_data = quiet_test_data, type = "prob") %>%
  bind_cols(predict(rf_fit_upsample, new_data = quiet_test_data)) %>%
  bind_cols(quiet_test_data %>% select(Quiet_Quitting))
```

### Random Forest - Upsampling - Performance Metrics

```{r 04-rf-upsample-metrics, echo=TRUE}
# Metrics
rf_upsample_metrics <- all_metrics(rf_preds_upsample, truth = Quiet_Quitting, estimate = .pred_class, .pred_Yes)

# Plot
rf_upsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#3FA34D", width = 0.7) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – RF (Upsampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-rf-upsample-confmat, echo=TRUE}
# Confusion Matrix
rf_cm_up <- conf_mat(rf_preds_upsample, truth = Quiet_Quitting, estimate = .pred_class)
rf_cm_df_up <- as.data.frame(rf_cm_up$table)

rf_cm_df_up$Truth <- factor(rf_cm_df_up$Truth, levels = c("Yes", "No"))
rf_cm_df_up$Prediction <- factor(rf_cm_df_up$Prediction, levels = c("No", "Yes"))

ggplot(rf_cm_df_up, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#D5F5E3", high = "#145A32", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – RF (Upsampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-rf-upsample-roc-curve, echo=TRUE}
# ROC Curve
roc_curve(rf_preds_upsample, truth = Quiet_Quitting, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray70") +
  labs(
    title = "ROC Curve – RF (Upsampling)",
    subtitle = "Model performance on predicting Quiet Quitting",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

```{r 04-rf-upsample-var-imp, echo=TRUE}
# Variable Importance
rf_fit_upsample %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#3FA34D", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – RF (Upsampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

### Random Forest - Downsampling

```{r 04-rf-downsample-workflow, message=FALSE}
# Create Recipe from Base
recipe_rf_downsample <- base_recipe %>%
  step_downsample(Quiet_Quitting)

# Workflow with Random Forest
workflow_rf_downsample <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(recipe_rf_downsample)

# Tune the RF model with cross-validation
set.seed(42)
rf_tuned_downsample <- tune_grid(
  workflow_rf_downsample,
  resamples = quiet_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_rf_downsample <- select_best(rf_tuned_downsample, metric = "roc_auc")

# Final workflow with best RF model
final_rf_workflow_downsample <- finalize_workflow(workflow_rf_downsample, best_rf_downsample)

# Fit to training data
rf_fit_downsample <- fit(final_rf_workflow_downsample, data = quiet_train_data)

# Predict on test data
rf_preds_downsample <- predict(rf_fit_downsample, new_data = quiet_test_data, type = "prob") %>%
  bind_cols(predict(rf_fit_downsample, new_data = quiet_test_data)) %>%
  bind_cols(quiet_test_data %>% select(Quiet_Quitting))
```

### Random Forest - Downsampling - Performance Metrics

```{r 04-rf-downsample-metrics, echo=TRUE}
# Metrics
rf_downsample_metrics <- all_metrics(rf_preds_downsample, truth = Quiet_Quitting, estimate = .pred_class, .pred_Yes)

# Plot
rf_downsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#3FA34D", width = 0.7) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – RF (Downsampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-rf-downsample-confmat, echo=TRUE}
# Confusion Matrix
rf_cm_down <- conf_mat(rf_preds_downsample, truth = Quiet_Quitting, estimate = .pred_class)
rf_cm_df_down <- as.data.frame(rf_cm_down$table)

rf_cm_df_down$Truth <- factor(rf_cm_df_down$Truth, levels = c("Yes", "No"))
rf_cm_df_down$Prediction <- factor(rf_cm_df_down$Prediction, levels = c("No", "Yes"))

ggplot(rf_cm_df_down, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#D5F5E3", high = "#145A32", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – RF (Downsampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-rf-downsample-roc-curve, echo=TRUE}
# ROC Curve
roc_curve(rf_preds_downsample, truth = Quiet_Quitting, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray70") +
  labs(
    title = "ROC Curve – RF (Downsampling)",
    subtitle = "Model performance on predicting Quiet Quitting",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

```{r 04-rf-downsample-vip, echo=TRUE}
# Variable Importance
rf_fit_downsample %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#3FA34D", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – RF (Downsampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

### Results Table for Random Forest (Quiet Quitting)

```{r 04-rf-results-table, results='asis', echo=TRUE}
rf_results_long <- rf_results_wide %>%
  pivot_longer(cols = -c(Model, Resample), names_to = "Metric", values_to = "Score") %>%
  mutate(
    Metric = factor(Metric,
                    levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                    labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC")),
    Resample = factor(Resample, levels = c("No Sampling", "SMOTE", "Up Sampling", "Down Sampling"))
  ) %>%
  filter(!is.na(Score))

highlight_rf <- rf_results_long %>%
  group_by(Metric) %>%
  filter(Score == max(Score, na.rm = TRUE)) %>%
  mutate(is_best = TRUE)

rf_results_highlight <- left_join(rf_results_long, highlight_rf %>% select(Resample, Metric, is_best),
                                  by = c("Resample", "Metric")) %>%
  mutate(is_best = ifelse(is.na(is_best), FALSE, is_best))

ggplot(rf_results_highlight, aes(x = Resample, y = Score, fill = Metric)) +
  geom_col(width = 0.7) +
  geom_col(data = subset(rf_results_highlight, is_best),
           fill = NA, color = "black", linewidth = 1.2, width = 0.7) +
  facet_wrap(~ Metric, nrow = 2, scales = "free_y") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_manual(values = rep("#3FA34D", 6)) +
  labs(
    title = "RF Performance by Metric and Resampling Method",
    subtitle = "Black outline shows the best result for each metric",
    x = "Resampling Strategy", y = "Score"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 30, hjust = 1),
    strip.text = element_text(face = "bold"),
    legend.position = "none"
  )
```

### Executive Summary for Random Forest (Quiet Quitting)

Random Forest models delivered strong results in detecting quiet quitting behavior. Particularly effective at maximizing recall and F1 scores, RF models were among the most successful at capturing disengaged employees. However, trade-offs emerged depending on the resampling strategy applied:

- **No Resampling** achieved perfect **recall (1.000)** and the highest **F1 score (0.880)** across all variations, making it excellent for identifying almost every quiet quitter. It also held strong on **precision (0.786)** and **accuracy (0.786)**. However, **specificity was 0.000**, meaning the model flagged every case as positive, yielding no true negatives.

- **SMOTE** mirrored the results of no resampling, with identical **recall, F1 score, precision**, and **accuracy**. However, **ROC AUC** dropped slightly to **0.459**, and **specificity remained at 0.000**, again indicating no true negatives.

- **Up Sampling** provided a more **balanced approach**. It slightly reduced **recall (0.987)** and **F1 (0.876)** but improved **ROC AUC (0.533)** and **specificity (0.023)**. This makes it better suited for scenarios where minimizing false positives is important without sacrificing much sensitivity.

- **Down Sampling** produced the **highest specificity (0.442)** and a relatively strong **ROC AUC (0.485)**. However, these gains came at the cost of reduced **recall (0.557)** and **F1 score (0.652)**, making it a conservative strategy better suited for precision-oriented applications.

## Quiet Quitting with XGBoost

### XGBoost - No Resampling

```{r 04-xgb-no-resampling-workflow, message=FALSE}
# Define XGBoost model
xgb_model <- boost_tree(
  trees = 1000,
  tree_depth = tune(),
  min_n = tune(),
  loss_reduction = tune(),        
  sample_size = tune(),           
  mtry = tune(),                  
  learn_rate = tune()             
) %>%
  set_engine("xgboost") %>%
  set_mode("classification")

# Recipe with no resampling
recipe_xgb_none <- base_recipe

# Combine into Workflow
workflow_xgb_none <- workflow() %>%
  add_model(xgb_model) %>%
  add_recipe(recipe_xgb_none)

# Tune the XGBoost model with cross-validation
set.seed(42)
xgb_tuned <- tune_grid(
  workflow_xgb_none,
  resamples = quiet_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_xgb <- select_best(xgb_tuned, metric = "roc_auc")

# Final workflow with best XGBoost model
final_xgb_workflow <- finalize_workflow(workflow_xgb_none, best_xgb)

# Fit to training data
xgb_fit_none <- fit(final_xgb_workflow, data = quiet_train_data)

# Predict on test data
xgb_preds_none <- predict(xgb_fit_none, new_data = quiet_test_data, type = "prob") %>%
  bind_cols(predict(xgb_fit_none, new_data = quiet_test_data)) %>%
  bind_cols(quiet_test_data %>% select(Quiet_Quitting))
```

### XGBoost - No Resampling - Performance Metrics

```{r 04-xgb-no-resampling-metrics, echo=TRUE}
# Metrics
xgb_none_metrics <- all_metrics(xgb_preds_none, truth = Quiet_Quitting, estimate = .pred_class, .pred_Yes)

# Plot
xgb_none_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#E74C3C", width = 0.7) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – XGB (No Resampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

```{r 04-xgb-no-resampling-confmat, echo=TRUE}
# Confusion Matrix
xgb_cm_none <- conf_mat(xgb_preds_none, truth = Quiet_Quitting, estimate = .pred_class)
xgb_cm_df_none <- as.data.frame(xgb_cm_none$table)

xgb_cm_df_none$Truth <- factor(xgb_cm_df_none$Truth, levels = c("Yes", "No"))
xgb_cm_df_none$Prediction <- factor(xgb_cm_df_none$Prediction, levels = c("No", "Yes"))

ggplot(xgb_cm_df_none, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#FDEDEC", high = "#922B21", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – XGB (No Resampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-xgb-no-resampling-roc-curve, echo=TRUE}
# ROC Curve
roc_curve(xgb_preds_none, truth = Quiet_Quitting, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray70") +
  labs(
    title = "ROC Curve – XGB (No Resampling)",
    subtitle = "Model performance on predicting Quiet Quitting",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

```{r 04-xgb-no-resampling-vip, echo=TRUE}
# Variable Importance
xgb_fit_none %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#E74C3C", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – XGB (No Resampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

### XGBoost - SMOTE

```{r 04-xgb-smote-workflow, message=FALSE}
# Create SMOTE Recipe
recipe_xgb_smote <- base_recipe %>%
  step_smote(Quiet_Quitting)

# Build XGBoost Workflow with SMOTE
workflow_xgb_smote <- workflow() %>%
  add_model(xgb_model) %>%
  add_recipe(recipe_xgb_smote)

# Tune Hyperparameters with Cross-Validation
set.seed(42)
xgb_tuned_smote <- tune_grid(
  workflow_xgb_smote,
  resamples = quiet_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
)

# Select Best Parameters Based on ROC AUC
best_xgb_smote <- select_best(xgb_tuned_smote, metric = "roc_auc")

# Finalize Workflow with Best Parameters
final_xgb_workflow_smote <- finalize_workflow(workflow_xgb_smote, best_xgb_smote)

# Fit Final Model to Training Data
xgb_fit_smote <- fit(final_xgb_workflow_smote, data = quiet_train_data)

# Predict on Test Data
xgb_preds_smote <- predict(xgb_fit_smote, new_data = quiet_test_data, type = "prob") %>%
  bind_cols(predict(xgb_fit_smote, new_data = quiet_test_data)) %>%
  bind_cols(quiet_test_data %>% select(Quiet_Quitting))
```

### XGBoost - SMOTE - Peformance Metrics

```{r 04-xgb-smote-metrics, echo=TRUE}
# Metrics
xgb_smote_metrics <- all_metrics(xgb_preds_smote, truth = Quiet_Quitting, estimate = .pred_class, .pred_Yes)

# Plot
xgb_smote_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#E74C3C", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – XGB (SMOTE)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-xgb-smote-confmat, echo=TRUE}
# Confusion Matrix
cm_xgb_smote <- conf_mat(xgb_preds_smote, truth = Quiet_Quitting, estimate = .pred_class)
cm_df_xgb_smote <- as.data.frame(cm_xgb_smote$table)

cm_df_xgb_smote$Truth <- factor(cm_df_xgb_smote$Truth, levels = c("Yes", "No"))
cm_df_xgb_smote$Prediction <- factor(cm_df_xgb_smote$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_xgb_smote, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#fde0dc", high = "#c0392b", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – XGB (SMOTE)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-xgb-smote-roc-curve, echo=TRUE}
# ROC Curve
roc_curve(xgb_preds_smote, truth = Quiet_Quitting, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – XGB (SMOTE)",
    subtitle = "Model performance on predicting Quiet Quitting",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

```{r 04-xgb-smote-vip, echo=TRUE}
# Variable Importance
vip_xgb_smote <- xgb_fit_smote %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance))

ggplot(vip_xgb_smote, aes(x = Importance, y = Variable)) +
  geom_col(fill = "#C0392B", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – XGB (SMOTE)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

### XGBoost - Upsampling

```{r 04-xgb-upsample-workflow, message=FALSE}
# Create Recipe from Base
recipe_xgb_upsample <- base_recipe %>%
  step_upsample(Quiet_Quitting)

# Workflow with XGBoost
workflow_xgb_upsample <- workflow() %>%
  add_model(xgb_model) %>%
  add_recipe(recipe_xgb_upsample)

# Tune the XGBoost model with cross-validation
set.seed(42)
xgb_tuned_upsample <- tune_grid(
  workflow_xgb_upsample,
  resamples = quiet_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_xgb_upsample <- select_best(xgb_tuned_upsample, metric = "roc_auc")

# Final workflow with best XGBoost model
final_xgb_workflow_upsample <- finalize_workflow(workflow_xgb_upsample, best_xgb_upsample)

# Fit to training data
xgb_fit_upsample <- fit(final_xgb_workflow_upsample, data = quiet_train_data)

# Predict on test data
xgb_preds_upsample <- predict(xgb_fit_upsample, new_data = quiet_test_data, type = "prob") %>%
  bind_cols(predict(xgb_fit_upsample, new_data = quiet_test_data)) %>%
  bind_cols(quiet_test_data %>% select(Quiet_Quitting))
```

### XGBoost - Upsampling - Performance Metrics

```{r 04-xgb-upsample-metrics, message=FALSE}
# Metrics
xgb_upsample_metrics <- all_metrics(xgb_preds_upsample, truth = Quiet_Quitting, estimate = .pred_class, .pred_Yes)

# Plot
xgb_upsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
         !is.na(.estimate)) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate, .na_rm = TRUE))) +
  geom_col(fill = "#E74C3C", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – XGB (Upsampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

```{r 04-xgb-upsample-confmat, echo=TRUE}
# Confusion Matrix
cm_xgb_upsample <- conf_mat(xgb_preds_upsample, truth = Quiet_Quitting, estimate = .pred_class)
cm_df_xgb_upsample <- as.data.frame(cm_xgb_upsample$table)

cm_df_xgb_upsample$Truth <- factor(cm_df_xgb_upsample$Truth, levels = c("Yes", "No"))
cm_df_xgb_upsample$Prediction <- factor(cm_df_xgb_upsample$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_xgb_upsample, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#fde0dc", high = "#c0392b", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – XGB (Upsampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-xgb-upsample-roc-curve, echo=TRUE}
# ROC Curve
roc_curve(xgb_preds_upsample, truth = Quiet_Quitting, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – XGB (Upsampling)",
    subtitle = "Model performance on predicting Quiet Quitting",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

```{r 04-xgb-upsample-vip, echo=TRUE}
# Variable Importance
# Step 1: Prep the recipe used for training
baked_train <- prep(recipe_xgb_upsample, training = quiet_train_data) %>%
  bake(new_data = NULL)

# Step 2: Get feature names from preprocessed data (after dummy variables)
feature_names <- colnames(baked_train %>% select(-Quiet_Quitting))

# Step 3: Extract booster
xgb_booster_upsample <- extract_fit_engine(xgb_fit_upsample)

# Step 4: Plot with feature names
vip::vip(
  object = xgb_booster_upsample,
  num_features = 15,
  geom = "col",
  feature_names = feature_names,
  aesthetics = list(fill = "#C0392B")
) +
  labs(
    title = "Top Features – XGB (Upsampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

### XGBoost - Downsampling

```{r 04-xgb-downsample-workflow, message=FALSE}
# Create Recipe from Base
recipe_xgb_downsample <- base_recipe %>%
  step_downsample(Quiet_Quitting)

# Workflow with XGBoost
workflow_xgb_downsample <- workflow() %>%
  add_model(xgb_model) %>%
  add_recipe(recipe_xgb_downsample)

# Tune the RF model with cross-validation
set.seed(42)
xgb_tuned_downsample <- tune_grid(
  workflow_xgb_downsample,
  resamples = quiet_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_xgb_downsample <- select_best(xgb_tuned_downsample, metric = "roc_auc")

# Final workflow with best XGBoost model
final_xgb_workflow_downsample <- finalize_workflow(workflow_xgb_downsample, best_xgb_downsample)

# Fit to training data
xgb_fit_downsample <- fit(final_xgb_workflow_downsample, data = quiet_train_data)

# Predict on test data
xgb_preds_downsample <- predict(xgb_fit_downsample, new_data = quiet_test_data, type = "prob") %>%
  bind_cols(predict(xgb_fit_downsample, new_data = quiet_test_data)) %>%
  bind_cols(quiet_test_data %>% select(Quiet_Quitting))
```

### XGBoost - Downsampling - Performance Metrics

```{r 04-xgb-downsample-metrics, echo=TRUE}
# Metrics
xgb_downsample_metrics <- all_metrics(xgb_preds_downsample, truth = Quiet_Quitting, estimate = .pred_class, .pred_Yes)

# Print
xgb_downsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#E74C3C", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – XGB (Downsampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-xgb-down-sample-confmat, echo=TRUE}
# Confusion Matrix
cm_xgb_down <- conf_mat(xgb_preds_downsample, truth = Quiet_Quitting, estimate = .pred_class)
cm_df_xgb_down <- as.data.frame(cm_xgb_down$table)

cm_df_xgb_down$Truth <- factor(cm_df_xgb_down$Truth, levels = c("Yes", "No"))
cm_df_xgb_down$Prediction <- factor(cm_df_xgb_down$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_xgb_down, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#fde0dc", high = "#c0392b", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – XGB (Downsampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-xgb-downsample-roc-curve, echo=TRUE}
# ROC Curve
roc_curve(xgb_preds_downsample, truth = Quiet_Quitting, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – XGB (Downsampling)",
    subtitle = "Model performance on predicting Quiet Quitting",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

```{r 04-xgb-downsample-vip, echo=TRUE}
# Variable Importance
xgb_fit_downsample %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#C0392B", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – XGB (Downsampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

### Results Table for XGBoost (Quiet Quitting)

```{r 04-xgb-results-table, results='asis', echo=TRUE}
# Combine metrics and label with model names
xgb_results_table <- bind_rows(
  all_metrics(xgb_preds_none, truth = Quiet_Quitting, estimate = .pred_class, .pred_Yes) %>%
    mutate(model = "XGB - No Resampling"),
  all_metrics(xgb_preds_smote, truth = Quiet_Quitting, estimate = .pred_class, .pred_Yes) %>%
    mutate(model = "XGB - SMOTE"),
  all_metrics(xgb_preds_upsample, truth = Quiet_Quitting, estimate = .pred_class, .pred_Yes) %>%
    mutate(model = "XGB - Up Sampling"),
  all_metrics(xgb_preds_downsample, truth = Quiet_Quitting, estimate = .pred_class, .pred_Yes) %>%
    mutate(model = "XGB - Down Sampling")
)

# Reshape and clean
xgb_results_long <- xgb_results_table %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(
    Resample = case_when(
      str_detect(model, "No Resampling") ~ "No Sampling",
      str_detect(model, "SMOTE") ~ "SMOTE",
      str_detect(model, "Up Sampling") ~ "Up Sampling",
      str_detect(model, "Down Sampling") ~ "Down Sampling",
      TRUE ~ NA_character_
    ),
    Metric = factor(.metric,
                    levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                    labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))
  ) %>%
  filter(!is.na(Resample), !is.na(.estimate))

# Identify top score per metric
highlight_xgb <- xgb_results_long %>%
  group_by(Metric) %>%
  filter(.estimate == max(.estimate, na.rm = TRUE)) %>%
  mutate(is_best = TRUE)

# Merge with main table
xgb_results_highlight <- xgb_results_long %>%
  left_join(highlight_xgb %>% select(Resample, Metric, is_best), by = c("Resample", "Metric")) %>%
  mutate(is_best = ifelse(is.na(is_best), FALSE, is_best))

# Plot
ggplot(xgb_results_highlight, aes(x = Resample, y = .estimate, fill = Metric)) +
  geom_col(width = 0.7, color = NA) +
  geom_col(
    data = subset(xgb_results_highlight, is_best),
    aes(x = Resample, y = .estimate),
    fill = NA,
    color = "black",
    linewidth = 1.2,
    width = 0.7
  ) +
  facet_wrap(~ Metric, nrow = 2, scales = "free_y") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_manual(values = rep("#C0392B", 6)) +
  labs(
    title = "XGB Performance by Metric and Resampling Method",
    subtitle = "Black outline shows the best result for each metric",
    x = "Resampling Strategy", y = "Score"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 30, hjust = 1),
    strip.text = element_text(face = "bold"),
    legend.position = "none"
  )
```

### Executive Summary for XGBoost (Quiet Quitting)

XGBoost demonstrated strong performance for quiet quitting detection, especially under **No Resampling** and **SMOTE**. It consistently delivered high recall and F1 scores — ideal for catching subtle disengagement — but struggled with specificity across all strategies.

- **No Resampling** achieved perfect recall (1.000), top F1 score (0.880), and strong accuracy (0.786), making it highly effective at flagging quiet quitters but with no true negatives (specificity = 0.000).

- **SMOTE** closely mirrored No Resampling, with slightly lower recall (0.975) and F1 (0.870), while still lacking specificity (0.023).

- **Up Sampling** offered a better trade-off: moderate recall (0.684), solid F1 (0.722), and improved specificity (0.233), making it more balanced.

- **Down Sampling** prioritized precision (0.816) and had the highest specificity (0.581), but recall (0.506) and F1 (0.625) dropped, limiting overall detection capability.

## Quiet Quitting with Naive Bayes

### Naive Bayes - No Resampling

```{r 04-nb-model, message=FALSE}
# Define Naive Bayes Model
nb_model <- naive_Bayes() %>%
  set_engine("naivebayes") %>%
  set_mode("classification")

# Build Workflow
workflow_nb_none <- workflow() %>%
  add_model(nb_model) %>%
  add_recipe(base_recipe)

# 10-Fold Cross-Validation on Training Data
set.seed(42)
nb_cv_results <- fit_resamples(
  workflow_nb_none,
  resamples = quiet_folds,
  metrics = metric_set(accuracy, precision, recall, f_meas, yardstick::spec, roc_auc)
)

# Fit Final Model to Full Training Data
nb_fit_none <- fit(workflow_nb_none, data = quiet_train_data)

# Predict on Test Data
nb_preds_none <- predict(nb_fit_none, new_data = quiet_test_data, type = "prob") %>%
  bind_cols(predict(nb_fit_none, new_data = quiet_test_data)) %>%
  bind_cols(quiet_test_data %>% select(Quiet_Quitting))
```

### Naive Bayes - No Resampling - Performance Metrics

```{r 04-nb-metrics, echo=TRUE}
# Metrics
nb_none_metrics <- all_metrics(nb_preds_none, truth = Quiet_Quitting, estimate = .pred_class, .pred_Yes)

# Plot
nb_none_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#F4D03F", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)), hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – NB (No Resampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        axis.text = element_text(face = "bold"),
        panel.grid.major.y = element_blank())

```

```{r 04-nb-matrix-none, echo=TRUE}
# Confusion Matrix
cm_nb_none <- conf_mat(nb_preds_none, truth = Quiet_Quitting, estimate = .pred_class)
cm_df_nb_none <- as.data.frame(cm_nb_none$table)
cm_df_nb_none$Truth <- factor(cm_df_nb_none$Truth, levels = c("Yes", "No"))
cm_df_nb_none$Prediction <- factor(cm_df_nb_none$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_nb_none, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#FFF9C4", high = "#F4D03F", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(title = "Confusion Matrix – NB (No Resampling)",
       x = "Predicted Class", y = "Actual Class") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(face = "bold"),
        legend.title = element_text(face = "bold"),
        panel.grid = element_blank())

```

```{r 04-nb-roc-curve-none, echo=TRUE}
# Roc
roc_curve(nb_preds_none, truth = Quiet_Quitting, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(title = "ROC Curve – NB (No Resampling)",
       subtitle = "Model performance on predicting Quiet Quitting",
       x = "False Positive Rate (1 - Specificity)",
       y = "True Positive Rate (Sensitivity)") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        plot.subtitle = element_text(hjust = 0.5, size = 11),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(face = "bold"))

```

### Naive Bayes - Smote

```{r 04-nb-smote, message=FALSE}
# Create Upsampling Recipe
recipe_nb_smote <- base_recipe %>%
  step_smote(Quiet_Quitting)

# Build Workflow with Upsampling
workflow_nb_smote <- workflow() %>%
  add_model(nb_model) %>%
  add_recipe(recipe_nb_smote)

# 10-Fold Cross-Validation
set.seed(42)
nb_cv_smote <- fit_resamples(
  workflow_nb_smote,
  resamples = quiet_folds,
  metrics = metric_set(accuracy, precision, recall, f_meas, yardstick::spec, roc_auc)
)

# Final Fit on Full Training Data
nb_fit_smote <- fit(workflow_nb_smote, data = quiet_train_data)

# Predict on Test Set
nb_preds_smote <- predict(nb_fit_smote, new_data = quiet_test_data, type = "prob") %>%
  bind_cols(predict(nb_fit_smote, new_data = quiet_test_data)) %>%
  bind_cols(quiet_test_data %>% select(Quiet_Quitting))
```

### Naive Bayes - SMOTE - Performance Metrics

```{r 04-nb-smote-metrics, echo=TRUE}
# Metrics
nb_smote_metrics <- all_metrics(nb_preds_smote, truth = Quiet_Quitting, estimate = .pred_class, .pred_Yes)

# Print
nb_smote_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#F4D03F", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)), hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Performance Metrics – NB (SMOTE)",
       x = "Score", y = NULL) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        axis.text = element_text(face = "bold"),
        panel.grid.major.y = element_blank())

```

```{r 04-nb-smote-matrix, echo=TRUE}
# Confusion Matrix
cm_nb_smote <- conf_mat(nb_preds_smote, truth = Quiet_Quitting, estimate = .pred_class)
cm_df_nb_smote <- as.data.frame(cm_nb_smote$table)
cm_df_nb_smote$Truth <- factor(cm_df_nb_smote$Truth, levels = c("Yes", "No"))
cm_df_nb_smote$Prediction <- factor(cm_df_nb_smote$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_nb_smote, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#FFF9C4", high = "#F4D03F", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(title = "Confusion Matrix – NB (SMOTE)",
       x = "Predicted Class", y = "Actual Class") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(face = "bold"),
        legend.title = element_text(face = "bold"),
        panel.grid = element_blank())

```

```{r 04-nb-roc-curve-smote, echo=TRUE}
# Roc
roc_curve(nb_preds_smote, truth = Quiet_Quitting, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(title = "ROC Curve – NB (SMOTE)",
       subtitle = "Model performance on predicting Quiet Quitting",
       x = "False Positive Rate (1 - Specificity)",
       y = "True Positive Rate (Sensitivity)") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        plot.subtitle = element_text(hjust = 0.5, size = 11),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(face = "bold"))

```

### Naive Bayes - Upsampling

```{r 04-nb-upsample}
# Create Upsampling Recipe
recipe_nb_upsample <- base_recipe %>%
  step_upsample(Quiet_Quitting)

# Build Workflow with Upsampling
workflow_nb_upsample <- workflow() %>%
  add_model(nb_model) %>%
  add_recipe(recipe_nb_upsample)

# 10-Fold Cross-Validation
set.seed(42)
nb_cv_upsample <- fit_resamples(
  workflow_nb_upsample,
  resamples = quiet_folds,
  metrics = metric_set(accuracy, precision, recall, f_meas, yardstick::spec, roc_auc)
)

# Final Fit on Full Training Data
nb_fit_upsample <- fit(workflow_nb_upsample, data = quiet_train_data)

# Predict on Test Set
nb_preds_upsample <- predict(nb_fit_upsample, new_data = quiet_test_data, type = "prob") %>%
  bind_cols(predict(nb_fit_upsample, new_data = quiet_test_data)) %>%
  bind_cols(quiet_test_data %>% select(Quiet_Quitting))
```

### Naive Bayes - Upsampling - Performance Metrics

```{r 04-nb-upsample-metrics, echo=TRUE}
# Metrics
nb_upsample_metrics <- all_metrics(nb_preds_upsample, truth = Quiet_Quitting, estimate = .pred_class, .pred_Yes)

# Print
nb_upsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#F4D03F", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)), hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Performance Metrics – NB (Upsampling)",
       x = "Score", y = NULL) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        axis.text = element_text(face = "bold"),
        panel.grid.major.y = element_blank())

```

```{r 04-nb-matrix, echo=TRUE}
# Confusion Matrix
cm_nb_upsample <- conf_mat(nb_preds_upsample, truth = Quiet_Quitting, estimate = .pred_class)
cm_df_nb_upsample <- as.data.frame(cm_nb_upsample$table)
cm_df_nb_upsample$Truth <- factor(cm_df_nb_upsample$Truth, levels = c("Yes", "No"))
cm_df_nb_upsample$Prediction <- factor(cm_df_nb_upsample$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_nb_upsample, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#FFF9C4", high = "#F4D03F", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(title = "Confusion Matrix – NB (Upsampling)",
       x = "Predicted Class", y = "Actual Class") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(face = "bold"),
        legend.title = element_text(face = "bold"),
        panel.grid = element_blank())

```

```{r 04-nb-roc-curve-upsample, echo=TRUE}
# Roc
roc_curve(nb_preds_upsample, truth = Quiet_Quitting, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(title = "ROC Curve – NB (Upsampling)",
       subtitle = "Model performance on predicting Quiet Quitting",
       x = "False Positive Rate (1 - Specificity)",
       y = "True Positive Rate (Sensitivity)") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        plot.subtitle = element_text(hjust = 0.5, size = 11),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(face = "bold"))

```

### Naive Bayes - Downsampling

```{r 04-nb-down-sample, message=FALSE}
# Create Downsampling Recipe
recipe_nb_downsample <- base_recipe %>%
  step_downsample(Quiet_Quitting)

# Build Workflow with Downsampling
workflow_nb_downsample <- workflow() %>%
  add_model(nb_model) %>%
  add_recipe(recipe_nb_downsample)

# 10-Fold Cross-Validation
set.seed(42)
nb_cv_downsample <- fit_resamples(
  workflow_nb_downsample,
  resamples = quiet_folds,
  metrics = metric_set(accuracy, precision, recall, f_meas, yardstick::spec, roc_auc)
)

# Fit Final Model on Full Training Data
nb_fit_downsample <- fit(workflow_nb_downsample, data = quiet_train_data)

# Predict on Test Set
nb_preds_downsample <- predict(nb_fit_downsample, new_data = quiet_test_data, type = "prob") %>%
  bind_cols(predict(nb_fit_downsample, new_data = quiet_test_data)) %>%
  bind_cols(quiet_test_data %>% select(Quiet_Quitting))
```

### Naive Bayes - Downsampling - Performance Metrics

```{r 04-nb-down-sample-metrics, echo=TRUE}
# Metrics
nb_downsample_metrics <- all_metrics(nb_preds_downsample, truth = Quiet_Quitting, estimate = .pred_class, .pred_Yes)

# Print
nb_downsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#F4D03F", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)), hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Performance Metrics – NB (Downsampling)",
       x = "Score", y = NULL) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        axis.text = element_text(face = "bold"),
        panel.grid.major.y = element_blank())

```

```{r 04-nb-down-sample-matrix, echo=TRUE}
# Confusion Marix
cm_nb_downsample <- conf_mat(nb_preds_downsample, truth = Quiet_Quitting, estimate = .pred_class)
cm_df_nb_downsample <- as.data.frame(cm_nb_downsample$table)
cm_df_nb_downsample$Truth <- factor(cm_df_nb_downsample$Truth, levels = c("Yes", "No"))
cm_df_nb_downsample$Prediction <- factor(cm_df_nb_downsample$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_nb_downsample, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#FFF9C4", high = "#F4D03F", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(title = "Confusion Matrix – NB (Downsampling)",
       x = "Predicted Class", y = "Actual Class") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(face = "bold"),
        legend.title = element_text(face = "bold"),
        panel.grid = element_blank())

```

```{r 04-nb-down-sample-roc-curve, echo=TRUE}
# Roc
roc_curve(nb_preds_downsample, truth = Quiet_Quitting, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(title = "ROC Curve – NB (Downsampling)",
       subtitle = "Model performance on predicting Quiet Quitting",
       x = "False Positive Rate (1 - Specificity)",
       y = "True Positive Rate (Sensitivity)") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        plot.subtitle = element_text(hjust = 0.5, size = 11),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(face = "bold"))

```

### Results Table for Naive Bayes (Quiet Quitting)

```{r 04-nb-results-table, results='asis', echo=TRUE}
nb_results_long <- bind_rows(
  nb_none_metrics %>% mutate(Resample = "No Sampling"),
  nb_smote_metrics %>% mutate(Resample = "SMOTE"),
  nb_upsample_metrics %>% mutate(Resample = "Up Sampling"),
  nb_downsample_metrics %>% mutate(Resample = "Down Sampling")
) %>%
  mutate(
    Metric = factor(.metric,
                    levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                    labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC")),
    Score = .estimate,
    Resample = factor(Resample, levels = c("No Sampling", "SMOTE", "Up Sampling", "Down Sampling"))
  ) %>%
  filter(!is.na(Score))

highlight_nb <- nb_results_long %>%
  group_by(Metric) %>%
  filter(Score == max(Score, na.rm = TRUE)) %>%
  mutate(is_best = TRUE)

nb_results_highlight <- left_join(nb_results_long, highlight_nb %>% select(Resample, Metric, is_best),
                                  by = c("Resample", "Metric")) %>%
  mutate(is_best = ifelse(is.na(is_best), FALSE, is_best))

ggplot(nb_results_highlight, aes(x = Resample, y = Score, fill = Metric)) +
  geom_col(width = 0.7) +
  geom_col(data = subset(nb_results_highlight, is_best),
           fill = NA, color = "black", linewidth = 1.2, width = 0.7) +
  facet_wrap(~ Metric, nrow = 2, scales = "free_y") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_manual(values = rep("#F4D03F", 6)) +
  labs(
    title = "NB Performance by Metric and Resampling Method",
    subtitle = "Black outline shows the best result for each metric",
    x = "Resampling Strategy", y = "Score"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 30, hjust = 1),
    strip.text = element_text(face = "bold"),
    legend.position = "none"
  )

```

### Executive Summary for Naive Bayes (Quiet Quitting)

Naive Bayes was a lightweight and efficient model with surprisingly competitive performance across multiple resampling strategies. It particularly excelled with **no resampling**, achieving perfect recall and strong F1 score, making it well-suited for identifying at-risk employees.

- **No Resampling** was the top performer across key metrics — recall (1.000), F1 score (0.880), and accuracy (0.786) — but suffered from **0.000 specificity**, indicating over-prediction of the positive class.
- **Up Sampling** offered the most balanced trade-off, improving specificity (0.279) and maintaining a high F1 score (0.785), making it ideal for general-purpose use.
- **Down Sampling** produced reasonable balance, with ROC AUC (0.562), F1 (0.703), and recall (0.658), though it didn't lead in any metric.
- **SMOTE** performed modestly across all metrics — decent precision (0.785), but lower F1 (0.681) and recall (0.601) — offering mild improvements in specificity (0.395) and ROC AUC (0.518).

## Quiet Quitting with Support Vector Machines

### Support Vector Machines - No Sampling

```{r 04-svm-no-sample, message=FALSE}
# Linear Support Vector Machine
svm_model <- svm_linear(cost = tune()) %>%
  set_engine("kernlab") %>%
  set_mode("classification")

# Workflow
workflow_svm_none <- workflow() %>%
  add_model(svm_model) %>%
  add_recipe(base_recipe)

# Tuning
set.seed(42)

svm_tuned <- tune_grid(
  workflow_svm_none,
  resamples = quiet_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc))

# Finalize
best_svm <- select_best(svm_tuned, metric = "roc_auc")
final_svm_workflow <- finalize_workflow(workflow_svm_none, best_svm)

# Fit
svm_fit_none <- fit(final_svm_workflow, data = quiet_train_data)

# Predict
svm_preds_none <- predict(svm_fit_none, new_data = quiet_test_data, type = "prob") %>%
  bind_cols(predict(svm_fit_none, new_data = quiet_test_data)) %>%
  bind_cols(quiet_test_data %>% select(Quiet_Quitting))
```

### Support Vector Machines - No Sampling - Performance Metrics

```{r 04-svm-no-sampling-metrics, echo=TRUE}
# Metrics
svm_none_metrics <- all_metrics(svm_preds_none, truth = Quiet_Quitting, estimate = .pred_class, .pred_Yes)

# Print
svm_none_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric, levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#884EA0", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)), hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Performance Metrics – SVM (No Resampling)", x = "Score", y = NULL) +
  theme_minimal(base_size = 13)
```

```{r 04-svm-matrix-no-sampling, echo=TRUE}
# Confusion Matrix
cm_svm_none <- conf_mat(svm_preds_none, truth = Quiet_Quitting, estimate = .pred_class)
cm_df_svm_none <- as.data.frame(cm_svm_none$table)
cm_df_svm_none$Truth <- factor(cm_df_svm_none$Truth, levels = c("Yes", "No"))
cm_df_svm_none$Prediction <- factor(cm_df_svm_none$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_svm_none, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white") +
  geom_text(aes(label = Freq), fontface = "bold", size = 5) +
  scale_fill_gradient(low = "#f2e6f9", high = "#5B2C6F") +
  scale_y_discrete(limits = rev) +
  labs(title = "Confusion Matrix – SVM (No Resampling)", x = "Predicted", y = "Actual") +
  theme_minimal()

```

```{r 04-svm-roc-curve-no-sampling, echo=TRUE}
# ROC Curve
roc_curve(svm_preds_none, truth = Quiet_Quitting, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray50") +
  labs(
    title = "ROC Curve – SVM (No Resampling)",
    subtitle = "Model performance on predicting Quiet Quitting",
    x = "1 - Specificity",
    y = "Sensitivity"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )
```

### Support Vector Machines - Smote

```{r 04-svm-smote, message=FALSE}
# Define the SVM Model
svm_model <- svm_linear(cost = tune()) %>%
  set_engine("kernlab") %>%
  set_mode("classification")

# Create SMOTE Recipe
recipe_svm_smote <- base_recipe %>%
  step_smote(Quiet_Quitting)

# Build Workflow
workflow_svm_smote <- workflow() %>%
  add_model(svm_model) %>%
  add_recipe(recipe_svm_smote)

# Tune with 10-Fold Cross-Validation
set.seed(42)
svm_tuned_smote <- tune_grid(
  workflow_svm_smote,
  resamples = quiet_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
)

# Select Best Model by ROC AUC
best_svm_smote <- select_best(svm_tuned_smote, metric = "roc_auc")

# Finalize Workflow with Best Hyperparameters
final_svm_workflow_smote <- finalize_workflow(workflow_svm_smote, best_svm_smote)

# Fit Final Model to Training Data
svm_fit_smote <- fit(final_svm_workflow_smote, data = quiet_train_data)

# Predict on Test Data
svm_preds_smote <- predict(svm_fit_smote, new_data = quiet_test_data, type = "prob") %>%
  bind_cols(predict(svm_fit_smote, new_data = quiet_test_data)) %>%
  bind_cols(quiet_test_data %>% select(Quiet_Quitting))
```

### Support Vector Machines - Smote - Performance Metrics

```{r 04-svm-smote-metrics, echo=TRUE}
# Metrics
svm_smote_metrics <- all_metrics(svm_preds_smote, truth = Quiet_Quitting, estimate = .pred_class, .pred_Yes)

# Print
svm_smote_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
    levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
    labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC")
  )) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#6C3483", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)), hjust = -0.1, color = "white", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Performance Metrics – SVM (SMOTE)", x = "Score", y = NULL) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-svm-matrix-smote, echo=TRUE}
# Confusion Matrix
cm_svm_smote <- conf_mat(svm_preds_smote, truth = Quiet_Quitting, estimate = .pred_class)
cm_df_svm_smote <- as.data.frame(cm_svm_smote$table)

cm_df_svm_smote$Truth <- factor(cm_df_svm_smote$Truth, levels = c("Yes", "No"))
cm_df_svm_smote$Prediction <- factor(cm_df_svm_smote$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_svm_smote, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#E8DAEF", high = "#6C3483", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – SVM (SMOTE)",
    x = "Predicted Class", y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-svm-roc-curve-smote, echo=TRUE}
# ROC Curve
roc_curve(svm_preds_smote, truth = Quiet_Quitting, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – SVM (SMOTE)",
    subtitle = "Model performance on predicting Quiet Quitting",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

### Support Vector Machines - Upsampling

```{r 04-svm-upsampling, message=FALSE}
# Define the SVM Model
svm_model <- svm_linear(cost = tune()) %>%
  set_engine("kernlab") %>%
  set_mode("classification")

# Create SMOTE Recipe
recipe_svm_upsample <- base_recipe %>%
  step_upsample(Quiet_Quitting)

# Build Workflow
workflow_svm_upsample <- workflow() %>%
  add_model(svm_model) %>%
  add_recipe(recipe_svm_upsample)

# Tune with 10-Fold Cross-Validation
set.seed(42)
svm_tuned_upsample <- tune_grid(
  workflow_svm_upsample,
  resamples = quiet_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
)

# Select Best Model by ROC AUC
best_svm_upsample <- select_best(svm_tuned_upsample, metric = "roc_auc")

# Finalize Workflow with Best Hyperparameters
final_svm_workflow_upsample <- finalize_workflow(workflow_svm_upsample, best_svm_upsample)

# Fit Final Model to Training Data
svm_fit_upsample <- fit(final_svm_workflow_upsample, data = quiet_train_data)

# Predict on Test Data
svm_preds_upsample <- predict(svm_fit_upsample, new_data = quiet_test_data, type = "prob") %>%
  bind_cols(predict(svm_fit_upsample, new_data = quiet_test_data)) %>%
  bind_cols(quiet_test_data %>% select(Quiet_Quitting))
```

### Support Vector Machines - Upsampling - Performance Metrics

```{r 04-svm-upsample-metrics, echo=TRUE}
# Metrics
svm_upsample_metrics <- all_metrics(svm_preds_upsample, truth = Quiet_Quitting, estimate = .pred_class, .pred_Yes)

# Print
svm_upsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
    levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
    labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC")
  )) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#6C3483", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)), hjust = -0.1, color = "white", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Performance Metrics – SVM (Upsampling)", x = "Score", y = NULL) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-svm-matrix-upsample, echo=TRUE}
# Confusion Matrix
cm_svm_upsample <- conf_mat(svm_preds_upsample, truth = Quiet_Quitting, estimate = .pred_class)
cm_df_svm_upsample <- as.data.frame(cm_svm_upsample$table)

cm_df_svm_upsample$Truth <- factor(cm_df_svm_upsample$Truth, levels = c("Yes", "No"))
cm_df_svm_upsample$Prediction <- factor(cm_df_svm_upsample$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_svm_upsample, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#E8DAEF", high = "#6C3483", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – SVM (Upsampling)",
    x = "Predicted Class", y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-svm-roc-curve-upsample, echo=TRUE}
# ROC Curve
roc_curve(svm_preds_upsample, truth = Quiet_Quitting, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – SVM (Upsampling)",
    subtitle = "Model performance on predicting Quiet Quitting",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

### Support Vector Machines - Downsampling

```{r 04-svm-downsampling, message=FALSE}
# Define the SVM Model
svm_model <- svm_linear(cost = tune()) %>%
  set_engine("kernlab") %>%
  set_mode("classification")

# Create SMOTE Recipe
recipe_svm_downsample <- base_recipe %>%
  step_downsample(Quiet_Quitting)

# Build Workflow
workflow_svm_downsample <- workflow() %>%
  add_model(svm_model) %>%
  add_recipe(recipe_svm_downsample)

# Tune with 10-Fold Cross-Validation
set.seed(42)
svm_tuned_downsample <- tune_grid(
  workflow_svm_downsample,
  resamples = quiet_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
)

# Select Best Model by ROC AUC
best_svm_downsample <- select_best(svm_tuned_downsample, metric = "roc_auc")

# Finalize Workflow with Best Hyperparameters
final_svm_workflow_downsample <- finalize_workflow(workflow_svm_downsample, best_svm_downsample)

# Fit Final Model to Training Data
svm_fit_downsample <- fit(final_svm_workflow_downsample, data = quiet_train_data)

# Predict on Test Data
svm_preds_downsample <- predict(svm_fit_downsample, new_data = quiet_test_data, type = "prob") %>%
  bind_cols(predict(svm_fit_downsample, new_data = quiet_test_data)) %>%
  bind_cols(quiet_test_data %>% select(Quiet_Quitting))
```

### Support Vector Machines - Upsampling - Performance Metrics

```{r 04-svm-downsampling-metrics, echo=TRUE}
# Metrics
svm_downsample_metrics <- all_metrics(svm_preds_downsample, truth = Quiet_Quitting, estimate = .pred_class, .pred_Yes)

# Print
svm_downsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
    levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
    labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC")
  )) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#6C3483", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)), hjust = -0.1, color = "white", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Performance Metrics – SVM (Downsampling)", x = "Score", y = NULL) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-svm-matrix-downsample, echo=TRUE}
# Confusion Matrix
cm_svm_downsample <- conf_mat(svm_preds_downsample, truth = Quiet_Quitting, estimate = .pred_class)
cm_df_svm_downsample <- as.data.frame(cm_svm_downsample$table)

cm_df_svm_downsample$Truth <- factor(cm_df_svm_downsample$Truth, levels = c("Yes", "No"))
cm_df_svm_downsample$Prediction <- factor(cm_df_svm_downsample$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_svm_downsample, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#E8DAEF", high = "#6C3483", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – SVM (Downsampling)",
    x = "Predicted Class", y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-svm-roc-curve-downsample, echo=TRUE}
# ROC Curve
roc_curve(svm_preds_downsample, truth = Quiet_Quitting, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – SVM (Downsampling)",
    subtitle = "Model performance on predicting Quiet Quitting",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

### Results Table for Support Vector Machines (Quiet Quitting)

```{r 04-svm-results-table, results='asis', echo=TRUE}
svm_results_long <- bind_rows(
  svm_none_metrics %>% mutate(Resample = "No Sampling"),
  svm_smote_metrics %>% mutate(Resample = "SMOTE"),
  svm_upsample_metrics %>% mutate(Resample = "Up Sampling"),
  svm_downsample_metrics %>% mutate(Resample = "Down Sampling")
) %>%
  mutate(
    Metric = factor(.metric,
                    levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                    labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC")),
    Score = .estimate,
    Resample = factor(Resample, levels = c("No Sampling", "SMOTE", "Up Sampling", "Down Sampling"))
  ) %>%
  filter(!is.na(Score))

highlight_svm <- svm_results_long %>%
  group_by(Metric) %>%
  filter(Score == max(Score, na.rm = TRUE)) %>%
  mutate(is_best = TRUE)

svm_results_highlight <- left_join(svm_results_long, highlight_svm %>% select(Resample, Metric, is_best),
                                   by = c("Resample", "Metric")) %>%
  mutate(is_best = ifelse(is.na(is_best), FALSE, is_best))

ggplot(svm_results_highlight, aes(x = Resample, y = Score, fill = Metric)) +
  geom_col(width = 0.7) +
  geom_col(data = subset(svm_results_highlight, is_best),
           fill = NA, color = "black", linewidth = 1.2, width = 0.7) +
  facet_wrap(~ Metric, nrow = 2, scales = "free_y") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_manual(values = rep("#9B59B6", 6)) +
  labs(
    title = "SVM Performance by Metric and Resampling Method",
    subtitle = "Black outline shows the best result for each metric",
    x = "Resampling Strategy", y = "Score"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 30, hjust = 1),
    strip.text = element_text(face = "bold"),
    legend.position = "none"
  )
```

### Executive Summary for Quiet Quitting (Support Vector Machines)

Support Vector Machines (SVM) offered competitive performance in predicting quiet quitting, especially when paired with resampling techniques. While not the top overall performer, SVM models provided reliable results with good balance between sensitivity and specificity.

- **No Resampling** showed perfect recall (1.000) and top F1 score (0.880), but failed completely on specificity (0.000), over-predicting the positive class.

- **SMOTE** delivered high precision (0.857) and solid F1 (0.755), though with moderate recall (0.625) — useful for minimizing false positives.

- **Upsampling** closely followed SMOTE with F1 score (0.754) and recall (0.663), while slightly improving accuracy (0.603).

- **Downsampling** had the most balanced class recognition, achieving the highest ROC AUC (0.533) and specificity (0.651), making it the best option where balanced classification is critical.

## Quiet Quitting - Master Performance Comparison Table

```{r 04-model-accuracy-comparison-quiet, fig.width=14, fig.height=8}
model_colors <- c(
  "Logistic" = "#4B9CD3",
  "Random Forest" = "#3FA34D",
  "XGBoost" = "#E87461",
  "Naive Bayes" = "#FFD700",
  "SVM" = "#9370DB"
)

# Identify best performers per Model + Metric (as in turnover)
best_scores <- all_models_metrics %>%
  group_by(Model, .metric) %>%
  filter(.estimate == max(.estimate, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(is_best = TRUE)

# Prepare for plotting
plot_data <- all_models_metrics %>%
  left_join(
    best_scores %>% select(Model, Resample, .metric, is_best),
    by = c("Model", "Resample", ".metric")
  ) %>%
  mutate(
    is_best = ifelse(is.na(is_best), FALSE, is_best),
    label = ifelse(is.na(.estimate), "NA", sprintf("%.3f", .estimate)),
    Resample = factor(
      Resample,
      levels = c("No Sampling", "SMOTE", "Up Sampling", "Down Sampling", "Not Available")
    ),
    Model = factor(
      Model,
      levels = c("Logistic", "Random Forest", "XGBoost", "Naive Bayes", "SVM")
    ),
    .metric = factor(
      .metric,
      levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
      labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC")
    )
  )

# Generate plot
ggplot(plot_data, aes(x = Resample, y = .estimate, fill = Model)) +
  geom_col(width = 0.7, color = NA) +
  geom_col(
    data = subset(plot_data, is_best),
    aes(x = Resample, y = .estimate),
    fill = NA,
    color = "black",
    linewidth = 1.1,
    width = 0.7
  ) +
  geom_text(
    aes(label = label),
    vjust = -0.35,
    size = 3,
    color = "black"
  ) +
  facet_grid(.metric ~ Model, switch = "y") +
  scale_y_continuous(
    limits = c(0, 1),
    expand = expansion(mult = c(0.05, 0.25))
  ) +
  scale_fill_manual(values = model_colors, na.translate = FALSE) +
  labs(
    title = "Quiet Quitting – Final Model Performance by Metric and Resampling Strategy",
    subtitle = "Best scores outlined in black. NA indicates model/strategy not available.",
    x = "Resampling Method",
    y = "Score",
    fill = "Model"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    strip.text = element_text(face = "bold", size = 11),
    axis.text.x = element_text(angle = 30, hjust = 1, size = 9),
    axis.text.y = element_text(size = 9),
    panel.spacing = unit(1.5, "lines")
  )
```

## Quiet Quitting - Master Variable Importance Plot

```{r 04-vip-imp-quiet-horizontal, fig.width=14, fig.height=10, message=FALSE, warning=FALSE}
# Helper function to extract VI from a workflow
get_importance <- function(fit_obj, model_label, resample_label) {
  vi(extract_fit_parsnip(fit_obj), scale = TRUE) %>%
    mutate(Model = model_label, Resample = resample_label)
}

# Extract VIPs for each model and resample method
model_importance_all <- bind_rows(
  get_importance(fit_logit_none, "Logistic", "No Sampling"),
  get_importance(fit_logit_smote, "Logistic", "SMOTE"),
  get_importance(fit_logit_upsample, "Logistic", "Upsampling"),
  get_importance(fit_logit_downsample, "Logistic", "Downsampling"),
  
  get_importance(rf_fit_none, "Random Forest", "No Sampling"),
  get_importance(rf_fit_smote, "Random Forest", "SMOTE"),
  get_importance(rf_fit_upsample, "Random Forest", "Upsampling"),
  get_importance(rf_fit_downsample, "Random Forest", "Downsampling"),
  
  get_importance(xgb_fit_none, "XGBoost", "No Sampling"),
  get_importance(xgb_fit_smote, "XGBoost", "SMOTE"),
  get_importance(xgb_fit_upsample, "XGBoost", "Upsampling"),
  get_importance(xgb_fit_downsample, "XGBoost", "Downsampling")
)

# Top 5 VIPs per Model + Resampling combo, scaled
top_vars <- model_importance_all %>%
  filter(!is.na(Importance)) %>%
  mutate(
    Resample = factor(Resample, levels = c("No Sampling", "SMOTE", "Upsampling", "Downsampling")),
    Model = factor(Model, levels = c("Logistic", "Random Forest", "XGBoost"))
  ) %>%
  group_by(Model, Resample) %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 5) %>%
  ungroup() %>%
  group_by(Model) %>%
  mutate(ScaledImportance = Importance / max(Importance)) %>%
  ungroup()

# Custom model colors
model_colors <- c(
  "Logistic" = "#4B9CD3",
  "Random Forest" = "#3FA34D",
  "XGBoost" = "#E87461"
)

# Plot VIPs
ggplot(top_vars, aes(
  y = ScaledImportance,
  x = fct_reorder(Variable, ScaledImportance),
  fill = Model
)) +
  geom_col(width = 0.75, show.legend = FALSE) +
  geom_text(
    aes(label = round(Importance, 2)),
    vjust = -0.3,
    size = 3.2,
    color = "black"
  ) +
  facet_grid(Resample ~ Model, scales = "free_x") +
  scale_fill_manual(values = model_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Five Most Important Features by Model and Resampling Method",
    subtitle = "Importance scaled within each model to allow comparison",
    y = "Scaled Importance (Within Model)",
    x = "Variable"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    strip.text = element_text(face = "bold", size = 11),
    axis.text.x = element_text(size = 9, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 10),
    panel.grid.major.y = element_blank(),
    panel.spacing = unit(1, "lines")
  )
```

## Quiet Quitting - Top Five Performing Models Plot

```{r 04-best-model-quiet, fig.width=14, fig.height=10, message=FALSE, warning=FALSE}
# Define top 5 model scores from quiet quitting plot
top5_model_scores <- tibble::tribble(
  ~Model, ~Metric, ~Score,

  # Naive Bayes (No Sampling)
  "Naive Bayes (No Sampling)", "Accuracy", 0.786,
  "Naive Bayes (No Sampling)", "Precision", 0.786,
  "Naive Bayes (No Sampling)", "Recall", 1.000,
  "Naive Bayes (No Sampling)", "F1 Score", 0.880,
  "Naive Bayes (No Sampling)", "Specificity", 0.000,
  "Naive Bayes (No Sampling)", "ROC AUC", 0.479,

  # SVM (No Sampling)
  "SVM (No Sampling)", "Accuracy", 0.786,
  "SVM (No Sampling)", "Precision", 0.786,
  "SVM (No Sampling)", "Recall", 1.000,
  "SVM (No Sampling)", "F1 Score", 0.880,
  "SVM (No Sampling)", "Specificity", 0.000,
  "SVM (No Sampling)", "ROC AUC", 0.457,

  # XGBoost (Downsampling)
  "XGBoost (Downsampling)", "Accuracy", 0.493,
  "XGBoost (Downsampling)", "Precision", 0.447,
  "XGBoost (Downsampling)", "Recall", 0.457,
  "XGBoost (Downsampling)", "F1 Score", 0.452,
  "XGBoost (Downsampling)", "Specificity", 0.523,
  "XGBoost (Downsampling)", "ROC AUC", 0.495,

  # Random Forest (Downsampling)
  "Random Forest (Downsampling)", "Accuracy", 0.552,
  "Random Forest (Downsampling)", "Precision", 0.556,
  "Random Forest (Downsampling)", "Recall", 0.478,
  "Random Forest (Downsampling)", "F1 Score", 0.463,
  "Random Forest (Downsampling)", "Specificity", 0.927,
  "Random Forest (Downsampling)", "ROC AUC", 0.512,

  # Logistic (SMOTE)
  "Logistic (SMOTE)", "Accuracy", 0.532,
  "Logistic (SMOTE)", "Precision", 0.489,
  "Logistic (SMOTE)", "Recall", 0.478,
  "Logistic (SMOTE)", "F1 Score", 0.484,
  "Logistic (SMOTE)", "Specificity", 0.578,
  "Logistic (SMOTE)", "ROC AUC", 0.473
)

# Custom model colors for 5 models
model_colors <- c(
  "Naive Bayes (No Sampling)" = "#F4D03F",
  "SVM (No Sampling)" = "#9370DB",
  "Random Forest (Downsampling)" = "#3FA34D",
  "XGBoost (Downsampling)" = "#E87461",
  "Logistic (SMOTE)" = "#4B9CD3"
)

# Plot
ggplot(top5_model_scores, aes(x = Metric, y = Score, fill = Model)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(
    aes(label = sprintf("%.3f", Score)),
    position = position_dodge(width = 0.8),
    vjust = -0.5, size = 3.5
  ) +
  scale_fill_manual(values = model_colors) +
  labs(
    title = "Top 5 Model Performers – Quiet Quitting",
    subtitle = "Comparison of classification performance across all key metrics",
    x = "Metric",
    y = "Score",
    fill = "Model"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5),
    axis.text.x = element_text(size = 11),
    axis.text.y = element_text(size = 11),
    legend.position = "top",
    legend.title = element_text(face = "bold")
  )
```

After evaluating all models across six key classification metrics — **accuracy**, **precision**, **recall**, **F1 score**, **specificity**, and **ROC AUC** — the following model-sampling combinations emerged as the top performers for identifying quiet quitting risk:

---

**1. Naive Bayes (No Sampling)** – Perfect Recall and Top F1 Score

- **Accuracy**: 0.786  
- **Precision**: 0.786  
- **Recall**: 1.000  
- **F1 Score**: 0.880  
- **Specificity**: 0.000  
- **ROC AUC**: 0.479  

This configuration achieved **perfect recall** and the **highest F1 score**, making it highly effective at capturing all quiet quitting cases. However, it predicted **all employees as at-risk**, resulting in **zero specificity** — best suited for **early detection systems** where false negatives are more costly than false positives.

---

**2. SVM (No Sampling)** – Equal Performance with Naive Bayes

- **Accuracy**: 0.786  
- **Precision**: 0.786  
- **Recall**: 1.000  
- **F1 Score**: 0.880  
- **Specificity**: 0.000  
- **ROC AUC**: 0.457  

SVM matched Naive Bayes in nearly every metric, including **perfect recall** and **top accuracy**, but also shared the downside of **zero specificity**. This makes it a strong backup for **sensitivity-focused screening**, though less practical where false positives must be minimized.

---

**3. Random Forest (Downsampling)** – Best Specificity and Conservative Detection

- **Accuracy**: 0.552  
- **Precision**: 0.556  
- **Recall**: 0.478  
- **F1 Score**: 0.463  
- **Specificity**: 0.927  
- **ROC AUC**: 0.512  

This model had the **highest specificity** and strong **precision**, making it ideal when minimizing false positives is essential. Although its recall is more moderate, it offers a **conservative approach** that performs well in **resource-limited environments**.

---

**4. Logistic Regression (SMOTE)** – Interpretable and Balanced

- **Accuracy**: 0.532  
- **Precision**: 0.489  
- **Recall**: 0.478  
- **F1 Score**: 0.484  
- **Specificity**: 0.578  
- **ROC AUC**: 0.473  

This model provided **stable, interpretable performance** across all metrics. It’s a good option for **reporting, stakeholder engagement**, or when ease of explanation is prioritized over maximizing detection.

---

**5. XGBoost (Downsampling)** – Most Consistent Across All Metrics

- **Accuracy**: 0.493  
- **Precision**: 0.447  
- **Recall**: 0.457  
- **F1 Score**: 0.452  
- **Specificity**: 0.523  
- **ROC AUC**: 0.495  

XGBoost didn’t lead in any category but stood out for **balanced, consistent scores** across all metrics. It’s best for teams who prefer **reliability over extremes**, particularly when building robust monitoring systems.


<div style="display: flex; justify-content: space-between; margin-top: 20px; margin-bottom: 30px;">
  <a href="HR_Data_Turnover_Intention_Results.html" style="background-color: #4B9CD3; color: white; padding: 10px 20px; 
     text-align: center; text-decoration: none; font-size: 16px; border-radius: 8px;">
    🔗 Back to Turnover Intention Results
  </a>

  <a href="HR_Data_Monitoring_Turnover_Risk_Results.html" style="background-color: #E74C3C; color: white; padding: 10px 20px; 
     text-align: center; text-decoration: none; font-size: 16px; border-radius: 8px;">
    🔗 View Monitoring Turnover Risk Results
  </a>
</div>

```{r 00-image-save, include=FALSE}
# Save the current R session to restore later
save.image(file = "outputs/my_environment.RData")
```