---
title: "Monitoring & Privacy Risk Modeling | Workforce Analytics"

output:
  html_document:
    always_allow_html: true
    self_contained: false
    toc: false
    toc_float: false
    theme: null
    highlight: null
    df_print: paged
    fig_caption: true
    includes:
      in_header: "../docs/assets/header.html"

number_sections: true
---

```{=html}
<div class="report-container">

  <div class="page-layout">

    <!-- SIDEBAR -->
    <div class="sidebar">
      <div class="sidebar-card">
        <div class="sidebar-label">On this page</div>
        <nav class="sidebar-nav">
          <a href="#executive-summary">Executive Summary</a>
          <a href="#modeling-overview">Modeling Overview</a>
          <a href="#data-prep-framework">Data Preparation</a>
          <a href="#model-development">Model Development</a>
          <a href="#model-evaluation">Model Evaluation</a>
          <a href="#model-evaluation-results">Evaluation Results</a>
          <a href="#sampling-strategy-framework">Sampling Strategy Framework</a>
          <a href="#precision-recall">Precision–Recall</a>
          <a href="#model-selection">Model Selection</a>
          <a href="#model-summary">Model Summary</a>
          <a href="#driver-interpretation">Driver Interpretation</a>
          <a href="#feature-action">Feature → Action</a>
          <a href="#business-impact">Business Impact</a>
        </nav>
      </div>
    </div>

    <!-- MAIN CONTENT -->
    <div class="page-content">

      <!-- MATCH EDA STRUCTURE -->
      <div class="hero-section">
        <p class="hero-meta">
          Kevin Allen Cauley • `r format(Sys.Date(), '%B %d, %Y')`
        </p>

        <h1 class="hero-title">Monitoring & Privacy Risk Modeling</h1>

        <p class="hero-subtitle">
  This section develops and evaluates predictive models to identify employees at risk of monitoring-related turnover, interpret key privacy and surveillance risk patterns, and translate model outputs into executive-level workforce trust recommendations.
        </p>
      </div>

      <!-- REPLACED page-nav WITH cta-row -->
      <div class="cta-row">
        <a href="HR_Data_Quiet_Quitting.html" class="cta-btn btn-blue">← Back</a>
        <a href="index.html" class="cta-btn btn-teal">Home</a>
        <a href="HR_Data_Insights_Gleaned.html" class="cta-btn btn-red">Next →</a>
      </div>
```

```{r 00-setup, include=FALSE}
source(here::here("analysis", "helper.R"))

PLOT_SUBFOLDER <- "monitoring_privacy"

cat("Using plot folder:", PLOT_SUBFOLDER, "\n")

# Required Packages
required_pkgs <- c(
  "conflicted", "tidyverse", "tidymodels", "themis", "here",
  "klaR", "discrim", "ranger", "kernlab", "xgboost", "gt", "scales"
)

for (pkg in required_pkgs) {
  if (!pkg %in% rownames(installed.packages())) {
    install.packages(pkg)
  }
}

library(conflicted)
library(tidyverse)
library(tidymodels)
library(themis)
library(here)
library(klaR)
library(discrim)
library(gt)
library(scales)

conflicted::conflicts_prefer(
  dplyr::filter,
  dplyr::select,
  dplyr::mutate,
  dplyr::summarise,
  dplyr::arrange,
  dplyr::slice,
  dplyr::slice_head,
  dplyr::slice_max,
  dplyr::slice_min,
  dplyr::recode,
  dplyr::rename,
  dplyr::lag
)

# Global Options
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  out.width = "100%",
  fig.width = 8,
  fig.height = 5,
  dpi = 300
)

# Reproducibility
set.seed(123)

# Load CLEAN dataset from EDA (DO NOT TOUCH THIS)
df <- readRDS(here::here("data", "processed", "hr_dataset_clean.rds"))

# Base dataset
base_df <- df %>%
  mutate(across(where(is.character), as.factor))
```

# Executive Summary {#executive-summary}

This analysis develops a predictive modeling framework to identify employees at risk of monitoring-related turnover using a synthetic, research-informed HR dataset.

Multiple machine learning models—including Logistic Regression, Random Forest, XGBoost, Support Vector Machines, and Naive Bayes—were evaluated under varying sampling strategies to address class imbalance and improve predictive performance.

Results show that **model effectiveness is driven as much by data strategy as algorithm selection**. Sampling strategies such as up-sampling, down-sampling, and SMOTE changed model behavior in meaningful ways, demonstrating that class-balancing choices can materially affect precision, recall, and F1 Score.

The final model-selection process uses **cross-validated F1 Score** to identify strong candidate configurations, with holdout performance used as a stability check to confirm whether those candidates remain practically usable on unseen data.

Rather than relying on a single predictive signal, monitoring-based turnover risk is influenced by a **combination of privacy perceptions, monitoring intensity, trust, workload, wellbeing, and organizational support factors**.

From a business perspective, this framework enables a shift from **reactive privacy complaint response to proactive workforce trust and monitoring-risk management**, supporting earlier intervention, better governance, and improved retention outcomes.

# Modeling Overview {#modeling-overview}

This section builds on the exploratory analysis by applying predictive modeling techniques to identify employees at risk of monitoring-related turnover.

Rather than relying on a single model, multiple algorithms are evaluated and compared to determine the most effective approach for workforce decision-making.

Each model is assessed using cross-validation and evaluated across multiple sampling strategies to address class imbalance.

The objective is not only to identify the most accurate model, but to determine which approach best aligns with organizational priorities.

# Data Preparation & Modeling Framework {#data-prep-framework}

## Target Definition

The target variable for this analysis is **monitoring-based turnover risk**, representing whether an employee is likely to show turnover risk associated with workplace monitoring, privacy concerns, or surveillance-related strain.

This binary outcome is structured as a classification problem, with levels defined as **"No" (not flagged for monitoring-based turnover risk)** and **"Yes" (at risk of monitoring-related turnover)**. Establishing a consistent target structure ensures that all models are trained and evaluated under the same conditions.

```{r 01-target, echo=FALSE}
# Ensure correct factor levels
base_df <- base_df %>%
  mutate(
    monitoring_turnover_risk = factor(monitoring_turnover_risk, levels = c("No", "Yes"))
  )
```

## Data Partitioning

To evaluate model performance objectively, the dataset is split into training and testing subsets.

The training set is used to fit and tune models, while the test set is reserved for final evaluation. Stratified sampling is applied to preserve the original class distribution of monitoring-based turnover risk, ensuring that model performance reflects realistic conditions.

```{r 02-split, echo=FALSE}
split <- initial_split(base_df, prop = 0.8, strata = monitoring_turnover_risk)

train_data <- training(split)
test_data  <- testing(split)
```

## Feature Engineering Pipeline

A standardized preprocessing pipeline is applied to ensure consistency across all models.

Categorical variables are converted into numeric representations using dummy encoding, and numeric variables are normalized to improve model stability—particularly for algorithms sensitive to scale, such as Logistic Regression and Support Vector Machines.

Additionally, zero-variance predictors are removed to prevent redundant or non-informative features from influencing model performance.

```{r 03-recipe, echo=FALSE}
rec <- recipe(monitoring_turnover_risk ~ ., data = train_data) %>%
  
  # Handle factors
  step_dummy(all_nominal_predictors()) %>%
  
  # Normalize numeric features (important for SVM, Logistic)
  step_normalize(all_numeric_predictors()) %>%
  
  # Remove zero variance
  step_zv(all_predictors())
```

## Cross-Validation Strategy

Cross-validation is used to evaluate model performance during training and reduce the risk of overfitting.

By partitioning the training data into multiple folds, each model is evaluated across different subsets of the data. This provides a more robust estimate of model performance compared to a single train-test split.

```{r 04-resampling, echo=FALSE}
cv_folds <- vfold_cv(train_data, v = 5, strata = monitoring_turnover_risk)
```

# Model Development {#model-development}

Each model is implemented using a consistent preprocessing pipeline, ensuring that performance differences are driven by the modeling approach rather than data preparation.

- Logistic Regression (baseline, interpretable)
- Random Forest (ensemble tree-based)
- XGBoost (boosted trees for performance)
- Support Vector Machine (nonlinear classification)
- Naive Bayes (probabilistic baseline)

This allows for a fair comparison across algorithms with varying levels of complexity and interpretability.

---

```{r 05-logistic, echo=FALSE}
log_model <- logistic_reg() %>%
  set_engine("glm")
```

```{r 06-rf, echo=FALSE}
rf_model <- rand_forest(mtry = tune(), trees = 500, min_n = tune()) %>%
  set_engine("ranger") %>%
  set_mode("classification")
```

```{r 07-xgb, echo=FALSE}
xgb_model <- boost_tree(
  trees = 500,
  mtry = tune(),
  learn_rate = tune(),
  min_n = tune()
) %>%
  set_engine("xgboost") %>%
  set_mode("classification")
```

```{r 08-nb, echo=FALSE}
nb_model <- naive_Bayes() %>%
  set_engine("klaR")
```

```{r 09-svm, echo=FALSE}
svm_model <- svm_rbf(cost = tune(), rbf_sigma = tune()) %>%
  set_engine("kernlab") %>%
  set_mode("classification")
```

## Model Pipelines

Workflows are used to combine preprocessing steps and model specifications into a single pipeline.

This ensures that all transformations are applied consistently during both training and prediction, reducing the risk of data leakage and simplifying model management.

```{r 10-workflows, echo=FALSE}
log_wf <- workflow() %>% add_model(log_model) %>% add_recipe(rec)
rf_wf  <- workflow() %>% add_model(rf_model)  %>% add_recipe(rec)
xgb_wf <- workflow() %>% add_model(xgb_model) %>% add_recipe(rec)
svm_wf <- workflow() %>% add_model(svm_model) %>% add_recipe(rec)
nb_wf  <- workflow() %>% add_model(nb_model)  %>% add_recipe(rec)
```

## Evaluation Metrics

Model performance is evaluated using a combination of classification metrics.

Accuracy provides an overall performance measure, while precision and recall capture trade-offs between false positives and false negatives. The F1 Score balances these two metrics, and ROC AUC evaluates the model’s ability to distinguish between classes.

Using multiple metrics ensures a comprehensive evaluation aligned with business priorities.

```{r 11-metrics, echo=FALSE}
# Classification metrics
# IMPORTANT:
# monitoring_turnover_risk levels are c("No", "Yes")
# event_level = "second" ensures the positive class is "Yes" = at risk of monitoring-related turnover

options(yardstick.event_first = FALSE)

class_metrics <- metric_set(
  accuracy,
  precision,
  recall,
  f_meas,
  roc_auc
)

target_event_level <- "second"

model_colors <- c(
  "Logistic" = "#4B9CD3",
  "Random Forest" = "#3FA34D",
  "XGBoost" = "#E87461",
  "Naive Bayes" = "#F4D03F",
  "SVM" = "#9370DB"
)

sampling_order <- c(
  "No Sampling",
  "Up Sampling",
  "Down Sampling",
  "SMOTE"
)

metric_labels <- c(
  accuracy = "Accuracy",
  precision = "Precision",
  recall = "Recall",
  f_meas = "F1 Score",
  roc_auc = "ROC AUC"
)
```

## Hyperparameter Tuning

For models with tunable parameters, grid search is used to identify optimal configurations.

This process evaluates multiple parameter combinations using cross-validation, selecting those that maximize predictive performance. Tuning helps improve model accuracy while maintaining generalizability.

## Model Training

Multiple supervised learning models were evaluated, including Random Forest, XGBoost, Support Vector Machines, Logistic Regression, and Naive Bayes.

Each model and sampling strategy combination was evaluated through cross-validation. Tunable models were optimized separately within each sampling condition, ensuring that model selection reflects the combined effect of algorithm choice, preprocessing, and class-balancing strategy.

Rather than focusing on individual model outputs, performance is summarized comparatively to highlight relative strengths across modeling approaches.

```{r 12-modeling-framework, echo=FALSE}
# This framework evaluates each model + sampling strategy combination together using cross-validation. Tunable models are tuned separately under each sampling condition.

# Final model selection is based on cross-validated F1 Score with a holdout viability check before reporting the selected model.

# Sampling recipes
sampling_recipes <- list(
  "No Sampling" = rec,
  "Up Sampling" = rec %>% step_upsample(monitoring_turnover_risk),
  "Down Sampling" = rec %>% step_downsample(monitoring_turnover_risk),
  "SMOTE" = rec %>% step_smote(monitoring_turnover_risk)
)

# Model specifications
model_specs <- list(
  "Logistic" = log_model,
  "Naive Bayes" = nb_model,
  "Random Forest" = rf_model,
  "XGBoost" = xgb_model,
  "SVM" = svm_model
)

# Tunable model names
tunable_models <- c("Random Forest", "XGBoost", "SVM")

# Unified prediction function
get_preds <- function(model, data) {
  predict(model, data, type = "prob") %>%
    bind_cols(predict(model, data)) %>%
    bind_cols(data %>% select(monitoring_turnover_risk))
}

# Unified scoring function
score_predictions <- function(preds, model_name, sampling_name) {
  class_metrics(
    preds,
    truth = monitoring_turnover_risk,
    estimate = .pred_class,
    .pred_Yes,
    event_level = target_event_level
  ) %>%
    mutate(
      Model = model_name,
      Sampling = sampling_name
    )
}

# Evaluate one model + sampling combination using CV
run_cv_model_sampling <- function(model_name, sampling_name) {
  
  model_spec <- model_specs[[model_name]]
  recipe_used <- sampling_recipes[[sampling_name]]
  
  wf <- workflow() %>%
    add_model(model_spec) %>%
    add_recipe(recipe_used)
  
  set.seed(123)
  
  if (model_name %in% tunable_models) {
    
    cv_fit <- tune_grid(
      wf,
      resamples = cv_folds,
      grid = 10,
      metrics = class_metrics,
      control = control_grid(save_pred = TRUE)
    )
    
    best_params <- select_best(cv_fit, metric = "f_meas")
    
    # Keep only metrics for the best tuned parameter combination.
    # This prevents duplicate model + sampling rows from appearing later.
    cv_metrics <- collect_metrics(cv_fit) %>%
      inner_join(best_params, by = names(best_params)) %>%
      mutate(
        Model = model_name,
        Sampling = sampling_name,
        Model_Config = paste(model_name, sampling_name, sep = " + ")
      )
    
    final_wf <- finalize_workflow(wf, best_params)
    
  } else {
    
    cv_fit <- fit_resamples(
      wf,
      resamples = cv_folds,
      metrics = class_metrics,
      control = control_resamples(save_pred = TRUE)
    )
    
    cv_metrics <- collect_metrics(cv_fit) %>%
      mutate(
        Model = model_name,
        Sampling = sampling_name,
        Model_Config = paste(model_name, sampling_name, sep = " + ")
      )
    
    final_wf <- wf
  }
  
  # Fit finalized workflow to full training data
  set.seed(123)
  final_fit <- fit(final_wf, data = train_data)
  
  # Evaluate finalized workflow on holdout test data
  test_preds <- get_preds(final_fit, test_data)
  
  test_metrics <- score_predictions(
    preds = test_preds,
    model_name = model_name,
    sampling_name = sampling_name
  ) %>%
    mutate(Model_Config = paste(model_name, sampling_name, sep = " + "))
  
  list(
    cv_fit = cv_fit,
    cv_metrics = cv_metrics,
    final_workflow = final_wf,
    final_fit = final_fit,
    test_preds = test_preds,
    test_metrics = test_metrics
  )
}

# Run all model + sampling combinations
model_grid <- tidyr::expand_grid(
  Model = names(model_specs),
  Sampling = names(sampling_recipes)
)

set.seed(123)

model_results <- purrr::pmap(
  list(model_grid$Model, model_grid$Sampling),
  run_cv_model_sampling
)

names(model_results) <- paste(model_grid$Model, model_grid$Sampling, sep = " + ")

# Cross-validation results for all best model + sampling configurations
cv_results_all <- purrr::map_dfr(model_results, "cv_metrics") %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "roc_auc")) %>%
  mutate(
    Metric = dplyr::recode(.metric, !!!metric_labels),
    Sampling = factor(Sampling, levels = sampling_order),
    Model = factor(Model, levels = names(model_colors))
  )

# Holdout test results for all finalized configurations
results_all <- purrr::map_dfr(model_results, "test_metrics") %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "roc_auc")) %>%
  mutate(
    Metric = dplyr::recode(.metric, !!!metric_labels),
    Sampling = factor(Sampling, levels = sampling_order),
    Model = factor(Model, levels = names(model_colors))
  )
```

---

# Model Evaluation {#model-evaluation}

This section evaluates how each model performs on unseen data and determines which approach is most suitable for real-world workforce decision-making.

Rather than focusing solely on accuracy, evaluation emphasizes the **precision–recall tradeoff**, balancing the ability to detect monitoring-related risk with the cost of unnecessary intervention.

This ensures model selection aligns with operational priorities, where both detection capability and decision efficiency matter.

Key metrics include:

- **Accuracy** → Overall correctness  
- **Precision** → Reliability of positive predictions  
- **Recall** → Ability to detect employees flagged for monitoring-related turnover risk
- **F1 Score** → Balance between precision and recall  
- **ROC AUC** → Overall classification strength  

This comparison highlights the trade-offs between detection, prediction reliability, and practical intervention burden.

## Test Set Performance

Final model performance is evaluated on the holdout test dataset.

This step provides an unbiased assessment of how each model performs on unseen data, ensuring that results reflect real-world predictive capability rather than training performance.


---

The evaluation framework compares each model and sampling strategy combination using cross-validation, then confirms the selected model on the holdout test set. This separates model selection from final performance confirmation and provides a more defensible assessment of how the model is likely to generalize.

# Sampling Strategy Framework {#sampling-strategy-framework}

Sampling techniques are applied within the preprocessing pipeline to ensure consistency across models.

By integrating sampling directly into the workflow, transformations are applied only to the training data, preventing data leakage and preserving the integrity of model evaluation.

## Sampling Recipes and Evaluation Setup

Each sampling strategy is embedded directly into the modeling recipe. This ensures that class balancing occurs only within the training folds during cross-validation, preventing leakage into validation or test data.

# Model Evaluation Results & Insights {#model-evaluation-results}

This section synthesizes model performance across sampling strategies and evaluation metrics, providing a comparative view of how different approaches perform under realistic conditions.

The focus shifts from individual model behavior to cross-model insights, highlighting trade-offs that directly impact workforce decision-making.

The goal is to identify a model that balances detection capability with operational efficiency—minimizing both missed risk and unnecessary intervention.

```{r 13-validation-check, include=FALSE}
stopifnot(exists("cv_results_all"))
stopifnot(exists("results_all"))
stopifnot(nrow(cv_results_all) > 0)
stopifnot(nrow(results_all) > 0)
stopifnot(all(c("Model", "Sampling", ".metric", ".estimate") %in% names(results_all)))
stopifnot(all(c("Model", "Sampling", ".metric", "mean") %in% names(cv_results_all)))
```

```{r 14-final-model-selection, include=FALSE}
# -------------------------------------------------------------------
# Final model selection
# -------------------------------------------------------------------
# Gold-standard selection is based primarily on cross-validated F1 Score,
# but the selected model must also meet minimum holdout stability checks.
#
# Monitoring-based turnover risk is a sensitive workforce and governance
# outcome, so this selection logic avoids over-claiming when strong CV
# performance does not fully transfer to unseen data.
# -------------------------------------------------------------------

# Minimum holdout requirements for a model to be considered usable
min_holdout_f1 <- 0.40
min_holdout_recall <- 0.30
min_holdout_precision <- 0.35

# Create holdout metric screen
holdout_screen <- results_all %>%
  filter(.metric %in% c("precision", "recall", "f_meas", "roc_auc")) %>%
  select(Model, Sampling, .metric, .estimate) %>%
  pivot_wider(names_from = .metric, values_from = .estimate) %>%
  mutate(
    Model = as.character(Model),
    Sampling = as.character(Sampling),
    
    viable_holdout = !is.na(precision) &
      !is.na(recall) &
      !is.na(f_meas) &
      precision > 0 &
      recall > 0 &
      f_meas > 0,
    
    stable_holdout = viable_holdout &
      f_meas >= min_holdout_f1 &
      recall >= min_holdout_recall &
      precision >= min_holdout_precision
  )

# Prefer models that meet the stronger holdout stability screen.
eligible_configs <- holdout_screen %>%
  filter(stable_holdout) %>%
  select(Model, Sampling)

# Fallback: if no model meets the stronger screen, use basic viability.
if (nrow(eligible_configs) == 0) {
  eligible_configs <- holdout_screen %>%
    filter(viable_holdout) %>%
    select(Model, Sampling)
}

# Stop if no model is even viable on the holdout set.
if (nrow(eligible_configs) == 0) {
  stop(
    "No model configuration produced viable positive-class holdout predictions. ",
    "Review class balance, sampling methods, thresholding, and model specifications before selecting a final model."
  )
}

# Cross-validated F1 table
cv_f1_table <- cv_results_all %>%
  filter(.metric == "f_meas") %>%
  mutate(
    Model = as.character(Model),
    Sampling = as.character(Sampling)
  ) %>%
  select(Model, Sampling, cv_f1 = mean)

# Combine CV and holdout evidence
selection_pool <- cv_f1_table %>%
  left_join(
    holdout_screen %>%
      select(
        Model,
        Sampling,
        holdout_precision = precision,
        holdout_recall = recall,
        holdout_f1 = f_meas,
        holdout_roc_auc = roc_auc,
        stable_holdout,
        viable_holdout
      ),
    by = c("Model", "Sampling")
  ) %>%
  mutate(
    selection_status = case_when(
      stable_holdout ~ "Stable holdout candidate",
      viable_holdout ~ "Viable holdout candidate",
      TRUE ~ "Not viable on holdout"
    )
  )

# Select the best cross-validated model among eligible holdout configurations
top_model_cv <- cv_f1_table %>%
  dplyr::inner_join(eligible_configs, by = c("Model", "Sampling")) %>%
  dplyr::left_join(
    holdout_screen %>%
      dplyr::select(
        Model,
        Sampling,
        holdout_precision = precision,
        holdout_recall = recall,
        holdout_f1 = f_meas,
        holdout_roc_auc = roc_auc,
        stable_holdout,
        viable_holdout
      ),
    by = c("Model", "Sampling")
  ) %>%
  dplyr::arrange(dplyr::desc(cv_f1), dplyr::desc(holdout_f1)) %>%
  utils::head(1)

top_model_name <- as.character(top_model_cv$Model)
top_sampling_method <- as.character(top_model_cv$Sampling)
top_model_config <- paste(top_model_name, top_sampling_method, sep = " + ")

top_f1 <- round(top_model_cv$cv_f1, 3)
top_holdout_f1 <- round(top_model_cv$holdout_f1, 3)
top_holdout_precision <- round(top_model_cv$holdout_precision, 3)
top_holdout_recall <- round(top_model_cv$holdout_recall, 3)
top_holdout_roc_auc <- round(top_model_cv$holdout_roc_auc, 3)

selection_strategy_label <- ifelse(
  isTRUE(top_model_cv$stable_holdout),
  "Holdout-Stable Risk Prioritization",
  "Holdout-Viable Risk Prioritization"
)

selection_rule_note <- ifelse(
  isTRUE(top_model_cv$stable_holdout),
  "The selected model met the holdout stability screen and ranked strongest by cross-validated F1 Score among stable candidates.",
  "No configuration met the full holdout stability screen, so the selected model is the strongest viable holdout performer among cross-validated candidates."
)

# Retrieve selected model object from saved model results
selected_result <- model_results[[top_model_config]]

final_wf <- selected_result$final_workflow
final_fit <- selected_result$final_fit
final_preds <- selected_result$test_preds

# Holdout test metrics for selected model
final_model_metrics <- results_all %>%
  filter(
    as.character(Model) == top_model_name,
    as.character(Sampling) == top_sampling_method,
    .metric %in% c("accuracy", "precision", "recall", "f_meas", "roc_auc")
  ) %>%
  mutate(
    Metric = dplyr::recode(.metric, !!!metric_labels),
    Score = round(.estimate, 3)
  )
```

```{r 14b-selected-model-consistency-check, include=FALSE}
# -------------------------------------------------------------------
# Selected model consistency check
# -------------------------------------------------------------------

stopifnot(exists("top_model_name"))
stopifnot(exists("top_sampling_method"))
stopifnot(exists("top_model_config"))
stopifnot(exists("top_f1"))
stopifnot(exists("top_holdout_f1"))
stopifnot(exists("top_holdout_precision"))
stopifnot(exists("top_holdout_recall"))
stopifnot(exists("top_holdout_roc_auc"))
stopifnot(exists("selection_strategy_label"))
stopifnot(exists("selection_rule_note"))
stopifnot(exists("final_model_metrics"))
stopifnot(exists("final_preds"))

stopifnot(length(top_model_name) == 1)
stopifnot(length(top_sampling_method) == 1)
stopifnot(length(top_model_config) == 1)

stopifnot(nrow(final_model_metrics) == 5)

stopifnot(
  all(c("Accuracy", "Precision", "Recall", "F1 Score", "ROC AUC") %in% final_model_metrics$Metric)
)

if (!top_model_config %in% names(model_results)) {
  stop("Selected model configuration does not exist in model_results: ", top_model_config)
}
```

```{r 14c-final-consistency-lock, include=FALSE}
# -------------------------------------------------------------------
# Final Selected Model Consistency Lock
# -------------------------------------------------------------------

required_final_objects <- c(
  "top_model_name",
  "top_sampling_method",
  "top_model_config",
  "top_f1",
  "top_holdout_f1",
  "top_holdout_precision",
  "top_holdout_recall",
  "top_holdout_roc_auc",
  "selection_strategy_label",
  "selection_rule_note",
  "final_model_metrics",
  "final_preds"
)

missing_objects <- required_final_objects[!purrr::map_lgl(required_final_objects, exists)]

if (length(missing_objects) > 0) {
  stop(
    "Missing required final model object(s): ",
    paste(missing_objects, collapse = ", ")
  )
}

if (nrow(final_model_metrics) != 5) {
  stop("final_model_metrics should contain exactly five metrics.")
}

expected_metrics <- c("Accuracy", "Precision", "Recall", "F1 Score", "ROC AUC")

if (!all(expected_metrics %in% final_model_metrics$Metric)) {
  stop("final_model_metrics is missing one or more expected metrics.")
}

if (!top_model_config %in% names(model_results)) {
  stop("Selected model configuration does not exist in model_results: ", top_model_config)
}

cat("Final selected model:", top_model_config, "\n")
cat("Selection strategy:", selection_strategy_label, "\n")
cat("Selection rule:", selection_rule_note, "\n")
cat("CV F1:", top_f1, "\n")
cat("Holdout Precision:", top_holdout_precision, "\n")
cat("Holdout Recall:", top_holdout_recall, "\n")
cat("Holdout F1:", top_holdout_f1, "\n")
cat("Holdout ROC AUC:", top_holdout_roc_auc, "\n")
```

## Sampling Strategy Comparison {#sampling-strategy-comparison}

Class imbalance is a major challenge in monitoring-risk prediction, as fewer employees typically fall into the "Yes" category.

To address this, multiple sampling strategies were applied:

- No Sampling (baseline)
- Up Sampling (increase minority class)
- Down Sampling (reduce majority class)
- SMOTE (synthetic minority generation)

The goal is to understand how balancing the dataset impacts model performance.

```{r 15-plot-sampling-comparison, echo=FALSE}
plot_df <- results_all %>%
  filter(.metric %in% c("precision", "recall", "f_meas", "roc_auc")) %>%
  mutate(
    Metric = factor(
      Metric,
      levels = c("Precision", "Recall", "F1 Score", "ROC AUC")
    ),
    Model = as.character(Model),
    Sampling = as.character(Sampling),
    Model_Sampling = paste(Model, Sampling, sep = " + ")
  )

# Rank rows by holdout F1 Score for cleaner storytelling
model_order <- plot_df %>%
  filter(.metric == "f_meas") %>%
  arrange(desc(.estimate)) %>%
  pull(Model_Sampling)

plot_df <- plot_df %>%
  mutate(
    Model_Sampling = factor(Model_Sampling, levels = rev(model_order)),
    Plot_Label = scales::number(.estimate, accuracy = 0.001)
  )

p <- ggplot(
  plot_df,
  aes(x = Metric, y = Model_Sampling, fill = .estimate)
) +
  geom_tile(color = "white", linewidth = 0.75) +
  geom_text(
    aes(label = Plot_Label),
    fontface = "bold",
    size = 3.4,
    color = "#243447"
  ) +
  scale_fill_gradient(
    low = "#EAF2F8",
    high = "#2E86C1",
    limits = c(0, 1),
    na.value = "#F2F4F4",
    name = "Score"
  ) +
  labs(
    title = "Model Performance Across Sampling Strategies",
    subtitle = "Holdout performance ranked by F1 Score",
    x = NULL,
    y = NULL
  ) +
  theme_hr(10) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(color = "#5a6b7c"),
    axis.text.y = element_text(size = 8),
    axis.text.x = element_text(face = "bold"),
    legend.position = "right",
    panel.grid = element_blank(),
    plot.margin = margin(10, 18, 10, 10)
  )

print(p)
save_plot(p, "sampling_strategy_performance_heatmap.png", 10.5, 8)
```

The sampling comparison shows that monitoring-related turnover risk is highly sensitive to the interaction between algorithm choice and class-balancing strategy. Several Random Forest and XGBoost configurations produced strong holdout F1 Scores, suggesting that tree-based models were better able to capture the nonlinear structure of monitoring, privacy, trust, and workforce-risk signals.

The selected configuration, Random Forest with SMOTE, provides the strongest cross-validated F1 Score among stable holdout candidates. Its holdout profile shows very high precision with moderate recall, meaning flagged employees are highly reliable risk signals, but the model may still miss some employees who later show monitoring-related turnover risk.

## Precision–Recall Curve Analysis {#precision-recall}

```{r 16-precision-recall, echo=FALSE}
pr_df <- results_all %>%
  filter(.metric %in% c("precision", "recall", "f_meas")) %>%
  select(Model, Sampling, .metric, .estimate) %>%
  pivot_wider(names_from = .metric, values_from = .estimate) %>%
  filter(!is.na(precision), !is.na(recall), !is.na(f_meas)) %>%
  mutate(
    Model = factor(as.character(Model), levels = names(model_colors)),
    Sampling = factor(as.character(Sampling), levels = sampling_order),
    Model_Config = paste(Model, Sampling, sep = " + "),
    Is_Selected = Model_Config == top_model_config
  )

selected_point <- pr_df %>%
  dplyr::filter(Is_Selected) %>%
  utils::head(1)

p <- ggplot(
  pr_df,
  aes(
    x = recall,
    y = precision,
    color = Model,
    shape = Sampling
  )
) +
  geom_vline(
    xintercept = 0.50,
    linetype = "dashed",
    color = "#D5DBDB",
    linewidth = 0.5
  ) +
  geom_hline(
    yintercept = 0.50,
    linetype = "dashed",
    color = "#D5DBDB",
    linewidth = 0.5
  ) +
  geom_point(
    aes(size = f_meas),
    alpha = 0.72
  ) +
  geom_point(
    data = selected_point,
    aes(x = recall, y = precision),
    inherit.aes = FALSE,
    shape = 21,
    size = 7,
    stroke = 1.3,
    fill = "white",
    color = "#1F2D3D"
  ) +
  geom_label(
    data = selected_point,
    aes(
      x = pmin(recall + 0.06, 0.88),
      y = pmin(precision + 0.07, 0.92),
      label = paste0(
        "Selected Model\n",
        Model_Config,
        "\nF1 = ",
        scales::number(f_meas, accuracy = 0.001)
      )
    ),
    inherit.aes = FALSE,
    size = 3.1,
    fontface = "bold",
    color = "#243447",
    fill = "white",
    label.size = 0.25
  ) +
  scale_color_manual(values = model_colors) +
  scale_shape_manual(
    values = c(
      "No Sampling" = 16,
      "Up Sampling" = 17,
      "Down Sampling" = 15,
      "SMOTE" = 3
    )
  ) +
  scale_size_continuous(
    range = c(3.2, 6.8),
    guide = "none"
  ) +
  scale_x_continuous(
    limits = c(0, 1),
    labels = scales::percent_format(accuracy = 1)
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    labels = scales::percent_format(accuracy = 1)
  ) +
  labs(
    title = "Precision–Recall Tradeoff by Model Configuration",
    subtitle = "Holdout results compare detection capability against prediction reliability",
    x = "Recall: Share of monitoring-related risk cases detected",
    y = "Precision: Reliability of monitoring-risk predictions",
    color = "Model",
    shape = "Sampling"
  ) +
  guides(
    color = guide_legend(order = 1, override.aes = list(size = 4, alpha = 1)),
    shape = guide_legend(order = 2, override.aes = list(size = 4, alpha = 1))
  ) +
  theme_hr(11) +
  theme(
    plot.title = element_text(face = "bold", size = 15),
    plot.subtitle = element_text(color = "#5a6b7c", size = 11),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 9),
    legend.position = "right",
    legend.box = "vertical",
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 8.5),
    panel.grid.minor = element_blank(),
    plot.margin = margin(10, 20, 15, 10)
  )

print(p)
save_plot(p, "precision_recall_tradeoff.png", 10.5, 5.8)
```

The precision–recall view shows how each model configuration balances detection capability against prediction reliability. Configurations farther to the right detect more employees with monitoring-related turnover risk, while configurations higher on the chart produce more reliable positive predictions.

The selected configuration is highlighted with an outlined marker. It is not chosen because it maximizes a single metric in isolation; rather, it provides the strongest cross-validated F1 Score among configurations that also produced usable holdout performance.

This reinforces the business logic behind the model-selection process: the goal is not to flag as many employees as possible or eliminate all false positives, but to create a practical monitoring-risk screening layer that can guide earlier review without replacing human judgment.

## Decision Matrix & Model Comparison

```{r 17-model-decision-breakdown, echo=FALSE}
decision_table <- cv_results_all %>%
  group_by(Model, Sampling) %>%
  summarise(
    Accuracy = mean(mean[.metric == "accuracy"], na.rm = TRUE),
    Precision = mean(mean[.metric == "precision"], na.rm = TRUE),
    Recall = mean(mean[.metric == "recall"], na.rm = TRUE),
    F1 = mean(mean[.metric == "f_meas"], na.rm = TRUE),
    ROC_AUC = mean(mean[.metric == "roc_auc"], na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(F1))

decision_table %>%
  dplyr::slice_head(n = 12) %>%
  gt() %>%
  tab_header(
    title = "Cross-Validated Model Decision Matrix",
    subtitle = "Top configurations ranked by cross-validated F1 Score"
  ) %>%
  fmt_number(
    columns = c(Accuracy, Precision, Recall, F1, ROC_AUC),
    decimals = 3
  ) %>%
  data_color(
    columns = c(Accuracy, Precision, Recall, F1, ROC_AUC),
    colors = col_numeric(
      palette = c("#EAF2F8", "#A9CCE3", "#2E86C1"),
      domain = c(0, 1)
    )
  ) %>%
  cols_align("center") %>%
  tab_options(
    table.width = pct(88),
    table.font.size = px(13)
  )
```

The cross-validated decision matrix ranks model configurations by F1 Score, showing which combinations performed strongest during resampling. Strong cross-validation results indicate promising generalizable performance, but they do not guarantee that the same configuration will remain stable on unseen holdout data.

For this reason, final model selection uses a two-step rule: first, configurations are evaluated using cross-validated F1 Score; second, leading candidates are checked against holdout performance to confirm that they still produce usable positive-class predictions.

This is especially important for monitoring and privacy risk because the target reflects a sensitive workforce signal. The model should support prioritization and governance review, not automatic decisions about employees.

### Model Selection {#model-selection}

Model selection was guided by practical decision tradeoffs rather than raw performance alone.

For monitoring-based turnover risk prediction, the cost of missing an at-risk employee must be balanced against the cost of unnecessary intervention. A model with very high recall may identify more potential monitoring-related risk cases, but it can also create operational noise by flagging too many employees. A model with high precision may reduce false alarms, but it may fail to detect enough meaningful risk.

For this reason, **cross-validated F1 Score was selected as the primary model-selection metric** because it balances precision and recall while reducing the risk of selecting a model based on one favorable test split.

However, cross-validation alone is not enough. The holdout test set is also used as a **stability check** to confirm that the selected model produces usable positive-class predictions on unseen data. This prevents the final report from selecting a model that performs well during resampling but fails during final confirmation.

In this run, the recommended configuration was **`r top_model_name` with `r top_sampling_method`**. The model should be interpreted as a **risk prioritization tool**, meaning it helps identify employees or workforce segments that may warrant earlier privacy, trust, or monitoring-governance review while leaving final decisions to HR leaders, managers, and organizational context.

## Selected Model Holdout Performance

```{r 18-final-model-metrics-plot, echo=FALSE}
final_model_plot_df <- final_model_metrics %>%
  mutate(
    Metric = factor(
      Metric,
      levels = c("Accuracy", "Precision", "Recall", "F1 Score", "ROC AUC")
    ),
    Highlight = ifelse(Metric == "F1 Score", "Selected Metric", "Supporting Metric"),
    Plot_Label = ifelse(is.na(Score), "N/A", scales::number(Score, accuracy = 0.001))
  )

p <- ggplot(final_model_plot_df, aes(x = Metric, y = Score, fill = Highlight)) +
  geom_col(width = 0.65, na.rm = TRUE) +
  geom_text(
    aes(label = Plot_Label),
    vjust = -0.45,
    fontface = "bold",
    size = 4
  ) +
  scale_fill_manual(values = c(
    "Selected Metric" = "#2E86C1",
    "Supporting Metric" = "#AAB7B8"
  )) +
  scale_y_continuous(
    limits = c(0, 1.08),
    labels = scales::percent_format(accuracy = 1)
  ) +
  labs(
    title = paste("Final Model Performance:", top_model_name),
    subtitle = paste("Holdout test performance | Sampling Strategy:", top_sampling_method),
    x = NULL,
    y = "Score"
  ) +
  theme_hr(11) +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(color = "#5a6b7c"),
    axis.text.x = element_text(face = "bold")
  )

print(p)
save_plot(p, "final_model_metrics.png", 8, 5)
```

## Risk Distribution

The risk distribution shows how the selected model assigns monitoring-based turnover-risk probabilities across the holdout test set. These scores should be interpreted as prioritization signals rather than definitive classifications. In practice, organizations would calibrate decision thresholds based on HR capacity, intervention cost, privacy sensitivity, and tolerance for missed risk.

```{r 19-risk-distribution, echo=FALSE}
p <- final_preds %>%
  ggplot(aes(x = .pred_Yes)) +
  geom_histogram(
    bins = 25,
    fill = "#4B9CD3",
    color = "white",
    linewidth = 0.4
  ) +
  geom_vline(
    xintercept = 0.40,
    linetype = "dashed",
    linewidth = 0.7,
    color = "#F39C12"
  ) +
  geom_vline(
    xintercept = 0.70,
    linetype = "dashed",
    linewidth = 0.7,
    color = "#C0392B"
  ) +
  annotate(
    "text",
    x = 0.40,
    y = Inf,
    label = "Moderate Risk",
    vjust = 1.6,
    hjust = -0.05,
    size = 3.5
  ) +
  annotate(
    "text",
    x = 0.70,
    y = Inf,
    label = "High Risk",
    vjust = 1.6,
    hjust = 1.05,
    size = 3.5
  ) +
  scale_x_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, 1)
  ) +
  labs(
    title = "Predicted Monitoring-Based Turnover Risk Distribution",
    subtitle = "Employees are scored by their predicted probability of monitoring-related turnover risk",
    x = "Predicted Probability of Monitoring-Based Turnover Risk",
    y = "Employee Count"
  ) +
  theme_hr(11) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(color = "#5a6b7c"),
    plot.margin = margin(10, 25, 15, 10)
  )

print(p)
save_plot(p, "risk_distribution.png", 8, 5)
```

## Risk Segmentation

The segmentation view translates model probabilities into practical intervention tiers. If employees cluster heavily within one tier, that does not mean the model is wrong; it means the thresholds should be calibrated for the organization’s decision context. In practice, HR teams would adjust these thresholds based on available capacity, intervention cost, privacy sensitivity, and tolerance for missed risk.

```{r 20-risk-segmentation, echo=FALSE}
risk_segments <- final_preds %>%
  mutate(
    Risk_Level = case_when(
      .pred_Yes >= 0.70 ~ "High Risk",
      .pred_Yes >= 0.40 ~ "Moderate Risk",
      TRUE ~ "Low Risk"
    ),
    Risk_Level = factor(
      Risk_Level,
      levels = c("Low Risk", "Moderate Risk", "High Risk")
    )
  ) %>%
  count(Risk_Level, .drop = FALSE) %>%
  mutate(
    Percent = n / sum(n),
    Label = ifelse(
      n == 0,
      "",
      paste0(n, "\n", scales::percent(Percent, accuracy = 1))
    )
  )

p <- ggplot(risk_segments, aes(x = Risk_Level, y = n, fill = Risk_Level)) +
  geom_col(width = 0.52, show.legend = FALSE) +
  geom_text(
    aes(label = Label),
    vjust = -0.35,
    fontface = "bold",
    size = 3.4,
    lineheight = 0.9,
    color = "#111111"
  ) +
  scale_fill_manual(values = c(
    "Low Risk" = "#2ECC71",
    "Moderate Risk" = "#F39C12",
    "High Risk" = "#E74C3C"
  )) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.16))
  ) +
  labs(
    title = "Employee Risk Segmentation",
    subtitle = "Risk tiers translate model scores into intervention priorities",
    x = NULL,
    y = "Employee Count"
  ) +
  theme_hr(11) +
  theme(
    plot.title = element_text(face = "bold", size = 15),
    plot.subtitle = element_text(color = "#5a6b7c", size = 11),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 9),
    axis.title.y = element_text(size = 10),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    plot.margin = margin(10, 18, 18, 10)
  )

print(p)
save_plot(p, "risk_segments.png", 7.5, 5)
```

The segmentation view translates predicted probabilities into practical intervention tiers. In this run, most employees fall into the moderate-risk band, suggesting the selected model is better suited for ranking relative monitoring-related risk than for sharply separating employees into extreme categories.

This does not mean the model is wrong. It means the thresholds should be calibrated to the organization’s decision context. In practice, HR and governance teams could adjust these cutoffs based on available review capacity, intervention cost, privacy sensitivity, and tolerance for missed risk.

# Final Model Summary & Key Takeaways {#model-summary}

The selected model represents the strongest cross-validated F1 Score among configurations that also met minimum holdout stability requirements. This means the final model was not selected only because it performed well during resampling; it also had to produce usable positive-class predictions on unseen data.

While the absolute performance values remain moderate, the model is useful as a **risk prioritization tool** because it helps rank employees by relative likelihood of monitoring-related turnover risk. This distinction is important: the model should not be used to make automatic employment decisions. Instead, it supports earlier review, privacy-aware workforce conversations, and more focused monitoring governance.

## 📊 Key Takeaways

```{=html}
<div class="model-highlight-card">

  <div class="model-badge">Recommended Model Selected</div>

  <h3 class="model-title">
    `r top_model_name`
  </h3>

  <p class="model-subtitle">
    Optimized using <strong>`r top_sampling_method`</strong> with a focus on <strong>F1 Score</strong>
  </p>

  <div class="model-metrics">
    <div class="metric-box">
      <span class="metric-label">CV F1 Score</span>
      <span class="metric-value">
        `r top_f1`
      </span>
    </div>

    <div class="metric-box">
      <span class="metric-label">Strategy</span>
      <span class="metric-value">
        `r selection_strategy_label`
      </span>
    </div>

    <div class="metric-box">
      <span class="metric-label">Use Case</span>
      <span class="metric-value">
        Monitoring & Privacy Risk Targeting
      </span>
    </div>
  </div>

</div>
```

## Why This Model Was Selected

```{=html}
<div class="decision-box">

<h3>Why This Model Was Selected</h3>

<p>
The final model was selected using a two-step process: cross-validation identified strong candidate configurations, and the holdout test set confirmed whether those candidates remained practically usable on unseen data.
</p>

<p>
In this run, the selected configuration was <strong>`r top_model_name` with `r top_sampling_method`</strong>. It achieved a cross-validated F1 Score of <strong>`r top_f1`</strong>, while its holdout test results produced precision of <strong>`r top_holdout_precision`</strong>, recall of <strong>`r top_holdout_recall`</strong>, and F1 Score of <strong>`r top_holdout_f1`</strong>.
</p>

<p>
`r selection_rule_note`
</p>

<p>
This selection rule balances model-development rigor with practical business usefulness. Cross-validation identifies the strongest generalizable configuration, while the holdout stability check ensures the final model still produces usable predictions on unseen data.
</p>

<p>
Because the model is designed for workforce risk screening, predictions should be treated as prompts for further HR or manager review, not as standalone conclusions about employee behavior.
</p>

</div>
```

## Selected Model Performance Metrics

The table below reports holdout test performance for the selected configuration. These values are not used alone to select the model; instead, they confirm whether the cross-validated winner remains practically usable on unseen data.

```{r 21-final-model-performance-table, echo=FALSE}
final_model_metrics %>%
  select(Metric, Score) %>%
  mutate(
    Score = ifelse(is.na(Score), NA_real_, Score)
  ) %>%
  gt() %>%
  tab_header(
    title = "Selected Model Performance Summary",
    subtitle = paste("Holdout test performance for", top_model_name, "with", top_sampling_method)
  ) %>%
  fmt_number(columns = Score, decimals = 3) %>%
  sub_missing(columns = Score, missing_text = "N/A") %>%
  cols_align("center") %>%
  tab_options(
    table.width = pct(55),
    table.font.size = px(13)
  )
```

## Top Model Configurations

```{r 22-top-models-table, echo=FALSE}
top_config_table <- cv_results_all %>%
  filter(.metric == "f_meas") %>%
  transmute(
    Model = as.character(Model),
    Sampling = as.character(Sampling),
    `CV F1 Score` = round(mean, 3)
  ) %>%
  left_join(
    holdout_screen %>%
      transmute(
        Model,
        Sampling,
        `Holdout Precision` = round(precision, 3),
        `Holdout Recall` = round(recall, 3),
        `Holdout F1` = round(f_meas, 3),
        Status = case_when(
          Model == top_model_name & Sampling == top_sampling_method ~ "Selected Final Model",
          stable_holdout ~ "Holdout-Stable Alternative",
          viable_holdout ~ "Viable but Below Stability Screen",
          TRUE ~ "Not Selected"
        )
      ),
    by = c("Model", "Sampling")
  ) %>%
  arrange(desc(`CV F1 Score`), desc(`Holdout F1`)) %>%
  dplyr::slice_head(n = 10)

top_config_table %>%
  gt() %>%
  tab_header(
    title = "Top Model Configurations",
    subtitle = "Cross-validation ranking with holdout confirmation metrics"
  ) %>%
  fmt_number(
    columns = c(
      `CV F1 Score`,
      `Holdout Precision`,
      `Holdout Recall`,
      `Holdout F1`
    ),
    decimals = 3
  ) %>%
  cols_align("center") %>%
  tab_options(
    table.width = pct(90),
    table.font.size = px(12)
  )
```

While model performance identifies *which employees are at risk of monitoring-related turnover*, understanding *why* those predictions occur is critical for meaningful intervention.

This section interprets the underlying drivers of monitoring-based turnover risk, connecting statistical patterns to real-world workforce behavior.

# Driver Interpretation {#driver-interpretation}

Because the selected predictive workflow does not provide a straightforward, model-agnostic feature-importance ranking, driver interpretation is based on the broader analytical pattern observed across the modeling and exploratory analysis.

The results suggest that monitoring-based turnover risk is not driven by one isolated factor. Instead, risk appears to reflect a combination of **privacy perceptions, monitoring intensity, trust, workload, wellbeing, and organizational support factors**.

This is important from a business perspective because monitoring risk should not be treated only as a technology or compliance issue. Effective monitoring governance requires coordinated action across transparency, fairness, privacy protection, manager communication, workload expectations, and employee trust.

## Feature → Action Mapping {#feature-action}

To translate model insights into practical action, key risk areas are mapped to targeted workforce interventions. This bridges the gap between predictive analytics and real-world decision-making by connecting model interpretation to practical HR response strategies.

```{r 23-driver-action-table, echo=FALSE}
driver_actions <- tibble::tribble(
  ~Risk_Area, ~What_It_Suggests, ~Recommended_Action,
  "Monitoring intensity and perceived surveillance",
  "Employees exposed to higher monitoring intensity may experience reduced autonomy, increased pressure, or heightened surveillance concerns.",
  "Review monitoring scope, necessity, transparency, and proportionality before expanding monitoring practices.",

  "Privacy concerns and trust",
  "Higher privacy concern may reduce trust in leadership and increase monitoring-related turnover risk.",
  "Improve communication about data use, access controls, retention policies, and employee privacy protections.",

  "Workload, stress, and wellbeing",
  "Monitoring may compound existing strain when employees already feel overloaded, unsupported, or closely scrutinized.",
  "Pair monitoring programs with workload review, wellbeing resources, and manager support practices.",

  "Manager and organizational support",
  "Employees may tolerate monitoring better when they perceive fair treatment, support, and a clear business purpose.",
  "Train managers to explain monitoring practices, respond to concerns, and avoid punitive use of productivity data.",

  "Governance and fairness",
  "Unclear or inconsistent monitoring policies may increase perceived unfairness and risk.",
  "Establish transparent governance, audit monitoring impacts, and evaluate whether tools affect employee groups differently."
)

driver_actions %>%
  gt() %>%
  tab_header(
    title = "Monitoring & Privacy Risk Driver-to-Action Map",
    subtitle = "Translating model insights into workforce trust and monitoring governance strategy"
  ) %>%
  cols_label(
    Risk_Area = "Risk Area",
    What_It_Suggests = "What It Suggests",
    Recommended_Action = "Recommended Action"
  ) %>%
  cols_width(
    Risk_Area ~ px(190),
    What_It_Suggests ~ px(360),
    Recommended_Action ~ px(430)
  ) %>%
  tab_options(
    table.width = pct(90),
    table.font.size = px(13)
  )
```

# Business Impact & Decision Guidance {#business-impact}

The model is designed to support workforce decision-making at a strategic level. Rather than functioning as an automated decision system, it should be used as a **risk intelligence layer** that helps HR, leadership, and governance stakeholders prioritize where deeper review may be needed.

## Strategic Decision Guidance {#decision-guidance}

```{=html}
<div class="decision-box">

<h3>Strategic Decision Guidance</h3>

<p>
This model is best deployed as a <strong>risk prioritization tool</strong>, not as a standalone decision system.
</p>

<ul>
  <li><strong>High-risk employees</strong> → Prioritize for HR, manager, or governance review</li>
  <li><strong>Moderate-risk employees</strong> → Monitor privacy concerns, workload, trust, and support trends</li>
  <li><strong>Low-risk employees</strong> → Maintain baseline monitoring governance and communication practices</li>
</ul>

<p>
The value of the model is not in replacing human judgment, but in helping decision-makers focus attention earlier and more consistently.
</p>

</div>
```

## Model Limitations {#model-limitations}

```{=html}
<div class="decision-box">

<h3>Model Limitations</h3>

<p>
The model should be interpreted within the context of synthetic data, simulated workforce patterns, and organizational variability.
</p>

<ul>
  <li>Predictions reflect patterns in synthetic data, not observed outcomes from a real organization</li>
  <li>Performance may shift when applied to different employee populations or industries</li>
  <li>Risk scores should support review and prioritization, not automatic employment decisions</li>
  <li>Human context remains essential before any intervention is taken</li>
</ul>

</div>
```

## Final Takeaway

Monitoring-based turnover risk is not driven by a single variable. It reflects a network of privacy, trust, monitoring, workload, wellbeing, and organizational support signals that collectively shape how employees experience workplace surveillance.

The modeling framework demonstrates how predictive analytics can move beyond static reporting and support proactive workforce trust management. In this run, the selected model provides a strong prioritization signal: employees flagged by the model are likely to warrant closer review, while lower recall indicates that monitoring-related risk should also be monitored through surveys, manager feedback, privacy complaints, and workforce climate indicators.

Used responsibly, this type of model becomes a decision-support tool: not a replacement for human judgment, but a way to help organizations balance operational visibility with employee trust, privacy, and long-term retention outcomes.

```{r 24-export-final-model-summary, include=FALSE}
summary_dir <- here::here("data", "processed", "model_summaries")
dir.create(summary_dir, recursive = TRUE, showWarnings = FALSE)

top_holdout_accuracy <- final_model_metrics %>%
  filter(Metric == "Accuracy") %>%
  pull(Score) %>%
  round(3)

final_model_summary <- tibble::tibble(
  Outcome = "Monitoring & Privacy Risk",
  Model = top_model_name,
  Sampling = top_sampling_method,
  Model_Config = top_model_config,
  Selection_Strategy = selection_strategy_label,
  Selection_Rule = selection_rule_note,
  CV_F1 = top_f1,
  Holdout_Accuracy = top_holdout_accuracy,
  Holdout_Precision = top_holdout_precision,
  Holdout_Recall = top_holdout_recall,
  Holdout_F1 = top_holdout_f1,
  Holdout_ROC_AUC = top_holdout_roc_auc
)

saveRDS(
  final_model_summary,
  here::here("data", "processed", "model_summaries", "monitoring_privacy_model_summary.rds")
)
```

```{=html}
<hr class="section-divider">

<div class="page-nav">

  <div class="nav-left">
    <a href="HR_Data_Quiet_Quitting.html" class="cta-btn btn-blue">← Back</a>
  </div>

  <div class="nav-center">
    <a href="index.html" class="cta-btn btn-teal">Home</a>
  </div>

  <div class="nav-right">
    <a href="HR_Data_Insights_Gleaned.html" class="cta-btn btn-red">Next →</a>
  </div>

</div>

</div> <!-- page-content -->
</div> <!-- page-layout -->
</div> <!-- main-container -->
```