---
title: "Executive Insights: Workforce Risk Modeling Results"
author: "Kevin Allen Cauley"
date: "`r format(Sys.Date(), '%B %d, %Y')`"

output:
  html_document:
    always_allow_html: true
    self_contained: false
    theme: flatly
    toc: false
    css: "assets/styles.css"
    output_file: HR_Data_Insights_Gleaned.html
    output_dir: ../docs
---

```{r 00-setup, include=FALSE}
# Function Helpers
source(here::here("analysis", "helper.R"))

# Reproducibility
set.seed(123)

# Ensure all paths resolve from project root
library(rprojroot)

knitr::opts_knit$set(
  root.dir = rprojroot::find_rstudio_root_file()
)

# Output Directory
plot_dir <- here::here("docs", "figures", "insights")
dir.create(plot_dir, recursive = TRUE, showWarnings = FALSE)

# Global chunk options
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  out.width = "100%",
  fig.retina = 2,
  fig.width = 8,
  fig.height = 5,
  dpi = 300,
  cache = FALSE,
  fig.path = "IGNORE/"
)
```

```{r 00-load-data}
# Load Clean Dataset (Standardized Pipeline)
hr_df <- readRDS(here::here("data", "processed", "hr_dataset_clean.rds"))

# Validation
required_cols <- c(
  "turnover_intention",
  "quiet_quitting",
  "monitoring_turnover_risk"
)

# Missing Column Check (Safety)
missing_cols <- setdiff(required_cols, names(hr_df))

if (length(missing_cols) > 0) {
  stop(
    paste0(
      "Missing required columns: ",
      paste(missing_cols, collapse = ", ")
    )
  )
}

# Type Enforcement
hr_df <- hr_df %>%
  mutate(
    turnover_intention = as.factor(turnover_intention),
    quiet_quitting = as.factor(quiet_quitting),
    monitoring_turnover_risk = as.factor(monitoring_turnover_risk)
  )
```

<div class="main-container">

<div class="hero-section">
  <h1>Executive Insights</h1>
  <p class="hero-subtitle">
    This section synthesizes findings across all analyses, translating predictive modeling results into actionable workforce strategies and organizational recommendations.
  </p>
</div>

<div class="nav-row">
  <a href="HR_Data_Monitoring_Turnover_Risk_Results.html" class="btn-clean">← Back</a>
  <a href="index.html" class="btn-clean">Home</a>
  <span></span>
</div>

<hr>

## Workforce Risk Overview

This analysis reveals consistent patterns across multiple models, identifying key workplace factors that contribute to employee disengagement and turnover risk.

Across all models, three major themes emerged:
- Employee well-being (burnout, stress, mental health)
- Workplace support (manager relationships, organizational support)
- Engagement and satisfaction (job satisfaction, commitment)

<div class="takeaway-box">
<strong>Key Insight:</strong> Workforce risk is not driven by a single factor, but by a combination of well-being, support, and engagement variables that interact to influence employee behavior.
</div>

## Turnover Intention Insights

The turnover model highlights strong relationships between employee burnout, job satisfaction, and perceived organizational support.

Employees with higher burnout and lower satisfaction levels show significantly increased likelihood of leaving.

<div class="takeaway-box">
<strong>Key Insight:</strong> Burnout and job satisfaction are the strongest predictors of turnover intention, making them critical intervention points for retention strategies.
</div>

Best Model: **SVM (SMOTE)**

Support Vector Machine with SMOTE offered the **most balanced and reliable performance** across all key classification metrics. It performed consistently well in:

- **Recall**: 0.538  
- **F1 Score**: 0.490  
- **ROC AUC**: 0.527  

Unlike models that overfit to one class, **SVM (SMOTE)** provided a meaningful spread of predictions across both positive and negative cases. Its confusion matrix confirms that it detected both true positives and true negatives, without collapsing into a biased classifier. This makes it ideal for **early intervention**, where missing at-risk employees (false negatives) is more costly than raising a few false alarms.

---

Why Naive Bayes (SMOTE) *Appears* Best but Isn't

Naive Bayes (SMOTE) seemed like a top performer at first:

- **Perfect Recall**: 1.000  
- **Highest F1 Score**: 0.628  
- **Perfect Specificity**: 1.000  
- **Accuracy**: 0.574  

However, a deeper look at the **confusion matrix** reveals:

|                | Predicted Yes | Predicted No |
|----------------|----------------|---------------|
| **Actual Yes** | 109 (TP)       | 0 (FN)        |
| **Actual No**  | 92 (FP)        | 0 (TN)        |

Naive Bayes **predicted every case as "Yes"**, flagging all employees as at-risk — which artificially inflates recall but makes it completely **unusable in practice**.

- **No true negatives detected**  
- **ROC AUC barely above random (0.503)**  
- **Precision severely diluted by false positives**

This one-sided prediction is a known risk when combining **Naive Bayes with SMOTE**, as the algorithm is sensitive to class priors and assumes feature independence.

While **Naive Bayes (SMOTE)** *looks good on paper*, it is **not a viable model**. The real-world utility of a classifier depends not just on recall, but on **balanced predictions and decision-making power**.  

For Turnover Intention, **SVM (SMOTE)** is the most balanced and trustworthy model for deployment.

### Turnover Intention Radar Plot and Confusion Matrix

```{r 04-turnover-radar-plot, fig.width=10, fig.height=6, echo=FALSE}
# Load required libraries
library(fmsb)
library(ggplot2)

# ---- Radar Chart ----
radar_data <- data.frame(
  row.names = c("Max", "Min", "Logistic", "NaiveBayes", "RandomForest", "SVM", "XGBoost"),
  Accuracy = c(1, 0, 0.512, 0.574, 0.575, 0.523, 0.493),
  F1.Score = c(1, 0, 0.437, 0.628, 0.463, 0.525, 0.452),
  Precision = c(1, 0, 0.463, 0.432, 0.556, 0.443, 0.447),
  Recall = c(1, 0, 0.478, 1.000, 0.109, 0.538, 0.457),
  ROC.AUC = c(1, 0, 0.503, 0.503, 0.512, 0.556, 0.495),
  Specificity = c(1, 0, 0.596, 1.000, 0.927, 0.512, 0.523)
)

model_colors <- c(
  "Logistic" = "#4B9CD3",
  "NaiveBayes" = "#F4D03F",
  "RandomForest" = "#3FA34D",
  "SVM" = "#884EA0",
  "XGBoost" = "#E87461"
)

# Radar chart
radarchart(
  radar_data,
  axistype = 1,
  pcol = model_colors,
  pfcol = adjustcolor(model_colors, alpha.f = 0.25),
  plwd = 2,
  plty = 1,
  cglcol = "gray80",
  cglty = 1,
  axislabcol = "gray30",
  caxislabels = seq(0, 1, 0.25),
  cglwd = 0.8,
  vlcex = 0.8,
  title = "Model Comparison – Radar Chart (Turnover Intention)"
)
legend("topright", legend = names(model_colors), col = model_colors, lty = 1, lwd = 2, bty = "n")

# ---- Confusion Matrix Polar Plot ----
confusion_data <- data.frame(
  Model = rep(c("Logistic", "NaiveBayes", "RandomForest", "SVM", "XGBoost"), each = 4),
  Category = rep(c("True Positive", "False Positive", "True Negative", "False Negative"), times = 5),
  Count = c(
    44, 65, 54, 38,
    109, 0, 0, 92,
    8, 101, 82, 10,
    43, 66, 44, 48,
    52, 57, 50, 42
  )
)

confusion_data$Model <- factor(confusion_data$Model, levels = names(model_colors))

ggplot(confusion_data, aes(x = Category, y = Count, fill = Model)) +
  geom_bar(stat = "identity", position = "identity", alpha = 0.6) +
  coord_polar() +
  scale_fill_manual(values = model_colors) +
  labs(title = "Confusion Matrix Breakdown – Top Models (Turnover Intention)") +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text.x = element_text(face = "bold"),
    legend.position = "right"
  )
```

## Quiet Quitting Insights

Quiet quitting behavior is strongly associated with reduced engagement, lower discretionary effort, and weaker organizational commitment.

Employees in this category may not leave immediately but show declining productivity and involvement.

<div class="takeaway-box">
<strong>Key Insight:</strong> Quiet quitting serves as an early warning signal of disengagement that often precedes turnover.
</div>

Best Model: **Naive Bayes (SMOTE)**

For identifying quiet quitting behavior, **Naive Bayes (SMOTE)** achieved **perfect recall (1.000)** and **specificity (1.000)**, alongside the **highest F1 score (0.628)** among all models. These strengths make it incredibly sensitive — able to capture nearly all disengaged employees.

However, its **lower precision (0.432)** and **high false positive rate** suggest that while it's effective at flagging risk, it also over-identifies employees who aren't actually quiet quitting. The confusion matrix shows this model predicts most employees as "at risk", which could lead to unnecessary interventions.

This model is ideal for **early detection** in large teams or low-stakes environments, where catching every possible case is more important than precision.  

At the same time - less suited for **high-consequence decisions** like performance reviews or escalated work intervention without further confirmation.

Compared to other models:
- **SVM** and **XGBoost** provided better balance across all metrics
- **Random Forest** and **Logistic Regression** had stronger precision but missed more true cases (low recall)

**Naive Bayes (SMOTE)** is a great *first-pass flagging model*, but should be paired with **secondary review or human-in-the-loop oversight** for practical application.

### Quiet Quitting Radar Plot and Confusion Matrix

```{r 04-quiet-radar-plot, fig.width=10, fig.height=6, echo=FALSE}
# Load required libraries
library(fmsb)
library(ggplot2)
library(dplyr)
library(gridExtra)

# ---- Radar Chart Data ----
radar_data <- data.frame(
  row.names = c("Max", "Min", "Logistic", "NaiveBayes", "RandomForest", "SVM", "XGBoost"),
  Accuracy = c(1, 0, 0.532, 0.574, 0.575, 0.523, 0.493),
  F1.Score = c(1, 0, 0.484, 0.628, 0.463, 0.525, 0.452),
  Precision = c(1, 0, 0.489, 0.432, 0.556, 0.443, 0.447),
  Recall = c(1, 0, 0.478, 1.000, 0.109, 0.538, 0.457),
  ROC.AUC = c(1, 0, 0.503, 0.503, 0.512, 0.556, 0.495),
  Specificity = c(1, 0, 0.578, 1.000, 0.927, 0.512, 0.523)
)

# Define model colors
model_colors <- c(
  "Logistic" = "#4B9CD3",
  "NaiveBayes" = "#F4D03F",
  "RandomForest" = "#3FA34D",
  "SVM" = "#884EA0",
  "XGBoost" = "#E87461"
)

# ---- Radar Chart ----
radarchart(
  radar_data,
  axistype = 1,
  pcol = model_colors,
  pfcol = adjustcolor(model_colors, alpha.f = 0.3),
  plwd = 2,
  plty = 1,
  cglcol = "gray80",
  cglty = 1,
  axislabcol = "gray30",
  caxislabels = seq(0, 1, 0.25),
  cglwd = 0.8,
  vlcex = 0.8,
  title = "Model Comparison – Radar Chart (Quiet Quitting)"
)
legend("topright", legend = names(model_colors), col = model_colors, lty = 1, lwd = 2, bty = "n")

# ---- Confusion Matrix Breakdown ----
confusion_data <- data.frame(
  Model = rep(c("Logistic", "NaiveBayes", "RandomForest", "SVM", "XGBoost"), each = 4),
  Category = rep(c("True Positive", "False Positive", "True Negative", "False Negative"), times = 5),
  Count = c(
    25, 18, 58, 100,
    26, 17, 63, 95,
    42, 1, 2, 156,
    43, 0, 0, 158,
    18, 25, 78, 80
  )
)

confusion_data$Model <- factor(confusion_data$Model, levels = names(model_colors))

# Polar bar plot
ggplot(confusion_data, aes(x = Category, y = Count, fill = Model)) +
  geom_bar(stat = "identity", position = "identity", alpha = 0.6) +
  coord_polar() +
  scale_fill_manual(values = model_colors) +
  labs(title = "Confusion Matrix Breakdown – Top Models (Quiet Quitting)") +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text.x = element_text(face = "bold"),
    legend.position = "right"
  )
```

## Work Related Monitoring and Turnoven Risk Insights

Monitoring-related variables reveal that increased surveillance and perceived intrusiveness may contribute to employee dissatisfaction and elevated turnover risk.

While monitoring can improve oversight, excessive or poorly implemented monitoring may negatively impact employee trust.

<div class="takeaway-box">
<strong>Key Insight:</strong> Workplace monitoring must be balanced with trust, as overly intrusive practices can increase disengagement and turnover risk.
</div>

Best Model: **Naive Bayes (SMOTE)**

For predicting turnover risk due to monitoring tools in the workplace, **Naive Bayes (SMOTE)** emerged as the most sensitive model, achieving **perfect recall (1.000)**, **perfect specificity (1.000)**, and the **highest F1 score (0.628)**. This means it was able to identify nearly all at-risk employees while correctly ruling out those not at risk.

However, its **lower precision (0.432)** indicates a high number of false positives. The confusion matrix confirms that while it catches most true positives, it may over-alert — potentially leading to unnecessary follow-ups by HR teams.

This model is best suited for **early warning systems** and **broad workforce monitoring**, where missing a true turnover risk is more costly than over-alerting.

At the same time, it may not be ideal for **high-stakes decisions** such as formal escalation, performance intervention, or termination risk reviews without additional checks.

Compared to other models:
- **SVM (No Sampling)** offered more balanced performance and the best ROC AUC. 
- **Random Forest** provided the highest precision but missed many actual turnover cases. 
- **Logistic Regression** remained a transparent and steady baseline with balanced mid-tier metrics.

**Naive Bayes (SMOTE)** is best used as a *first-pass detection tool* within a multi-step monitoring system — ideally followed by **review or confirmation mechanisms**.

### Monitoring Turnover Risk Radar Plot and Confusion Matrix

```{r 04-monitoring-radar-plot, fig.width=10, fig.height=6, echo=FALSE}
# Load required libraries
library(fmsb)
library(ggplot2)
library(dplyr)
library(gridExtra)

# ---- Radar Chart Data ----
radar_data <- data.frame(
  row.names = c("Max", "Min", "Logistic", "NaiveBayes", "RandomForest", "SVM", "XGBoost"),
  Accuracy = c(1, 0, 0.532, 0.574, 0.575, 0.523, 0.493),
  F1.Score = c(1, 0, 0.484, 0.628, 0.463, 0.525, 0.452),
  Precision = c(1, 0, 0.489, 0.432, 0.556, 0.443, 0.447),
  Recall = c(1, 0, 0.478, 1.000, 0.109, 0.538, 0.457),
  ROC.AUC = c(1, 0, 0.473, 0.503, 0.512, 0.556, 0.495),
  Specificity = c(1, 0, 0.578, 1.000, 0.927, 0.512, 0.523)
)

# Define model colors
model_colors <- c(
  "Logistic" = "#4B9CD3",
  "NaiveBayes" = "#F4D03F",
  "RandomForest" = "#3FA34D",
  "SVM" = "#884EA0",
  "XGBoost" = "#E87461"
)

# ---- Radar Chart ----
radarchart(
  radar_data,
  axistype = 1,
  pcol = model_colors,
  pfcol = adjustcolor(model_colors, alpha.f = 0.3),
  plwd = 2,
  plty = 1,
  cglcol = "gray80",
  cglty = 1,
  axislabcol = "gray30",
  caxislabels = seq(0, 1, 0.25),
  cglwd = 0.8,
  vlcex = 0.8,
  title = "Model Comparison – Radar Chart (Monitoring Turnover Risk)"
)
legend("topright", legend = names(model_colors), col = model_colors, lty = 1, lwd = 2, bty = "n")

# ---- Confusion Matrix Breakdown ----
confusion_data <- data.frame(
  Model = rep(c("Logistic", "NaiveBayes", "RandomForest", "SVM", "XGBoost"), each = 4),
  Category = rep(c("True Positive", "False Positive", "True Negative", "False Negative"), times = 5),
  Count = c(
    51, 39, 94, 17,    # Logistic (SMOTE)
    109, 48, 36, 8,     # Naive Bayes (SMOTE)
    1, 33, 144, 23,     # Random Forest (Upsampling)
    17, 26, 128, 30,    # SVM (No Resampling)
    41, 37, 104, 19     # XGBoost (Downsampling)
  )
)

confusion_data$Model <- factor(confusion_data$Model, levels = names(model_colors))

# Polar bar plot
ggplot(confusion_data, aes(x = Category, y = Count, fill = Model)) +
  geom_bar(stat = "identity", position = "identity", alpha = 0.6) +
  coord_polar() +
  scale_fill_manual(values = model_colors) +
  labs(title = "Confusion Matrix Breakdown – Top Models (Monitoring Turnover Risk)") +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.text.x = element_text(face = "bold"),
    legend.position = "right"
  )
```

## Cross-Model Patterns

Across all analyses, several variables consistently emerged as influential:

- Burnout and stress levels
- Job satisfaction and engagement
- Manager and organizational support
- Workplace environment and culture

<div class="takeaway-box">
<strong>Key Insight:</strong> These factors collectively form the foundation of workforce stability and should be prioritized in organizational strategy.
</div>

```{r 05-model-performance-comparison, echo=FALSE}
img_path <- here::here(
  "outputs",
  "hr_data_insights_gleaned",
  "plots",
  "top_models_comparison.png"
)

if (file.exists(img_path)) {
  knitr::include_graphics(img_path)
} else {
  cat("<p style='text-align:center; color:#888;'>Model comparison plot not available.</p>")
}
```

This visualization compares the highest-performing model configurations across all evaluation metrics. 

While several models demonstrated strong performance in individual metrics, Support Vector Machines and Random Forest models tended to provide the most balanced performance across recall, precision, and ROC-AUC.

This highlights the importance of evaluating models across multiple metrics rather than relying on accuracy alone when assessing workforce risk predictions.

## Strategic Recommendations

Based on the findings, organizations should focus on:

- Reducing employee burnout through workload management and well-being initiatives  
- Strengthening manager-employee relationships through leadership development  
- Improving engagement through recognition, career growth, and communication  
- Evaluating monitoring practices to ensure they support rather than hinder employee trust  

<div class="takeaway-box">
<strong>Key Insight:</strong> Targeted interventions in well-being, leadership, and engagement can significantly reduce workforce risk and improve retention outcomes.
</div>

## Conclusion

This project demonstrates how synthetic HR data can be used to build a full analytics pipeline that delivers actionable insights into workforce behavior.

By combining exploratory analysis with predictive modeling, organizations can move beyond reactive decision-making and proactively manage employee risk.

<div class="takeaway-box">
<strong>Final Takeaway:</strong> Data-driven workforce strategies enable organizations to identify risks early, intervene effectively, and create more stable and engaged work environments.
</div>

<div class="page-nav">
  <a href="HR_Data_Monitoring_Turnover_Risk_Results.html" class="btn-clean">← Back</a>
  <a href="index.html" class="btn-clean">Return Home →</a>
</div>

</div>