---
title: "HR Predictive Analytics Project – Monitoring Turnover Risk"
author: "Kevin Allen Cauley"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    theme: flatly
    highlight: tango
    df_print: paged
---

```{r 00-setup, include=FALSE}
# Set working directory to the project root
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

# Global chunk options
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.path = "figures/",
  fig.align = "center",
  cache = TRUE
)

# Load previous session
if (file.exists("outputs/my_environment.RData")) {
  load("outputs/my_environment.RData")
}
```

```{r 00-setup-packages, include=FALSE}
# Load packages
library(car)
library(cowplot)
library(caret)
library(corrplot)
library(discrim)
library(dplyr)
library(e1071)
library(fastDummies)
library(forcats)
library(gbm)
library(GGally)
library(ggcorrplot)
library(ggplot2)
library(ggthemes)
library(glmnet)
library(grid)
library(gridExtra)
library(gt)
library(iml)
library(kableExtra)
library(keras3)
library(keras)
library(knitr)
library(mgcv)
library(naivebayes)
library(naniar)
library(patchwork)
library(pheatmap)
library(plotly)
library(psych)
library(randomForest)
library(ranger)
library(RColorBrewer)
library(readxl)
library(reshape2)
library(skimr)
library(tensorflow)
library(themis)
library(tidymodels)
library(tidyverse)
library(vip)
library(viridis)
library(xgboost)
library(xfun)
```

## Predictive Modeling on Monitoring Turnover Risk

The goal of this modeling section is to predict **Monitoring Turnover Risk** using a set of behavioral, demographic, and engagement-related predictors.

We will evaluate a range of classification models using multiple **resampling strategies** (No Sampling, Up-Sampling, Down-Sampling, and SMOTE). Models will be evaluated using six classification metrics:

-   Accuracy
-   Precision
-   Recall
-   F1 Score
-   Specificity
-   ROC AUC

This approach ensures fairness, robustness, and clarity in identifying the most effective model and configuration.

### Shared Model Setup - Monitoring Turnover Risk

```{r 04-logit-shared-model-setup}
# Ensure Monitoring Turnover Risk is a Factor
hr_df <- hr_df %>%
  mutate(monitoring_turnover_risk = as.factor(monitoring_turnover_risk))

# Set Seed
set.seed(42)

# 80/20 train/test split with stratification
mtr_split <- initial_split(hr_df, prop = 0.8, strata = Monitoring_Turnover_Risk)
mtr_train_data <- training(mtr_split)
mtr_test_data  <- testing(mtr_split)

# 10-Fold Cross Validation
mtr_folds <- vfold_cv(mtr_train_data, v = 10, strata = Monitoring_Turnover_Risk)

# Shared Pre-processing (no resampling)
base_recipe <- recipe(Monitoring_Turnover_Risk ~ ., data = mtr_train_data) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_numeric_predictors())

# Basic Logistic Model
logit_model <- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")
```

```{r 04-logit-mtr-shared-metrics-setup}
# Shared metric set for all models
all_metrics <- metric_set(accuracy, precision, recall, f_meas, yardstick::spec, roc_auc)
```

## Monitoring Turnover Risk with Logistic Regression

### Logistic Regression - No Resampling

```{r 04-logit-none-workflow, message=FALSE}
# Workflow with Base Recipe
workflow_logit_none <- workflow() %>%
  add_model(logit_model) %>%
  add_recipe(base_recipe)

# Fit to training data
fit_logit_none <- fit(workflow_logit_none, data = mtr_train_data)

# Predict on test data
logit_preds_none <- predict(fit_logit_none, new_data = mtr_test_data, type = "prob") %>%
  bind_cols(predict(fit_logit_none, new_data = mtr_test_data)) %>%
  bind_cols(mtr_test_data %>% select(Monitoring_Turnover_Risk))
```

### Logistic Regression - No Resampling - Performance Metrics

```{r 04-logit-none-metrics, echo=TRUE}
# Metrics
logit_none_metrics <- all_metrics(logit_preds_none, truth = Monitoring_Turnover_Risk, estimate = .pred_class, .pred_Yes)

# Plot
logit_none_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#4B9CD3", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – LR (No Resampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

```{r 04-logit-none-confmat, echo=TRUE}
library(ggplot2)
library(yardstick)
library(forcats)

# Confusion matrix to dataframe
cm_logit_none <- conf_mat(logit_preds_none, truth = Monitoring_Turnover_Risk, estimate = .pred_class)
cm_df <- as.data.frame(cm_logit_none$table)

# Ensure proper order for plotting
cm_df$Truth <- factor(cm_df$Truth, levels = c("Yes", "No"))         
cm_df$Prediction <- factor(cm_df$Prediction, levels = c("No", "Yes")) 

# Plot
ggplot(cm_df, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#e6f2ff", high = "#1A5276", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – LR (No Resampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )
```

```{r 04-logit-none-roc-curve, echo=TRUE}
# Generate ROC curve
roc_curve(logit_preds_none, truth = Monitoring_Turnover_Risk, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – LR (No Resampling)",
    subtitle = "Model performance on predicting Monitoring Turnover Risk",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )
```

```{r 04-logit-none-vip, echo=TRUE}
# Extract full importance and then limit to top 15
vip_logit_none <- fit_logit_none %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance))

# Plot
ggplot(vip_logit_none, aes(x = Importance, y = Variable)) +
  geom_col(fill = "#1A5276", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – LR (No Resampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

### Logistic Regression – SMOTE

```{r 04-logit-smote-workflow, message=FALSE}
# SMOTE recipe
recipe_logit_smote <- base_recipe %>%
  step_smote(Monitoring_Turnover_Risk)

workflow_logit_smote <- workflow() %>%
  add_model(logit_model) %>%
  add_recipe(recipe_logit_smote)

# 10-Fold CV
set.seed(42)
logit_cv_smote <- fit_resamples(
  workflow_logit_smote,
  resamples = mtr_folds,
  metrics = metric_set(accuracy, precision, recall, f_meas, yardstick::spec, roc_auc)
)

# Final model fit and prediction on test data for SMOTE
fit_logit_smote <- fit(workflow_logit_smote, data = mtr_train_data)

logit_preds_smote <- predict(fit_logit_smote, new_data = mtr_test_data, type = "prob") %>%
  bind_cols(predict(fit_logit_smote, new_data = mtr_test_data)) %>%
  bind_cols(mtr_test_data %>% select(Monitoring_Turnover_Risk))
```

### Logistic Regression – SMOTE - Performance Metrics

```{r 04-logit-smote-finalfit, message=FALSE}
# Metrics
logit_smote_metrics <- all_metrics(logit_preds_smote, truth = Monitoring_Turnover_Risk, estimate = .pred_class, .pred_Yes)

# Plot
logit_smote_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#4B9CD3", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – LR (SMOTE)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

```{r 04-logit-smote-confmat, echo=TRUE}
# Create confusion matrix
cm_logit_smote <- conf_mat(logit_preds_smote, truth = Monitoring_Turnover_Risk, estimate = .pred_class)
cm_df_smote <- as.data.frame(cm_logit_smote$table)

# Order axes
cm_df_smote$Truth <- factor(cm_df_smote$Truth, levels = c("Yes", "No"))
cm_df_smote$Prediction <- factor(cm_df_smote$Prediction, levels = c("No", "Yes"))

# Plot heatmap
ggplot(cm_df_smote, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#e6f2ff", high = "#1A5276", name = "Frequency") +
  scale_y_discrete(limits = rev) +  # YES on top
  labs(
    title = "Confusion Matrix – LR (SMOTE)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )
```

```{r 04-logit-smote-roc, echo=TRUE}
# ROC Curve
roc_curve(logit_preds_smote, truth = Monitoring_Turnover_Risk, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – LR (SMOTE)",
    subtitle = "Model performance on predicting Monitoring Turnover Risk",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )
```

```{r 04-logit-smote-vip, echo=TRUE}
# Top 15 features only
fit_logit_smote %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#1A5276", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – LR (SMOTE)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

### Logistic Regression - Upsampling

```{r 04-logit-upsample-workflow, message=FALSE}
# Upsampling Recipe
recipe_logit_upsample <- base_recipe %>%
  step_upsample(Monitoring_Turnover_Risk)

# Workflow 
workflow_logit_upsample <- workflow() %>%
  add_model(logit_model) %>%
  add_recipe(recipe_logit_upsample)

# Fit
fit_logit_upsample <- fit(workflow_logit_upsample, data = mtr_train_data)

# Predict
logit_preds_upsample <- predict(fit_logit_upsample, new_data = mtr_test_data, type = "prob") %>%
  bind_cols(predict(fit_logit_upsample, new_data = mtr_test_data)) %>%
  bind_cols(mtr_test_data %>% select(Monitoring_Turnover_Risk))
```

### Logistic Regression - Upsampling - Performance Metrics

```{r 04-logit-upsample-metrics, echo=TRUE}
# Metrics
logit_upsample_metrics <- all_metrics(logit_preds_upsample, truth = Monitoring_Turnover_Risk, estimate = .pred_class, .pred_Yes)

# Plot
logit_upsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#4B9CD3", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – LR (Upsampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

```{r 04-logit-upsample-confmat, echo=TRUE}
# Confusion matrix data prep
cm_logit_upsample <- conf_mat(logit_preds_upsample, truth = Monitoring_Turnover_Risk, estimate = .pred_class)
cm_df_upsample <- as.data.frame(cm_logit_upsample$table)

# Set factor levels to match layout: Yes on top, No bottom; No left, Yes right
cm_df_upsample$Truth <- factor(cm_df_upsample$Truth, levels = c("Yes", "No"))
cm_df_upsample$Prediction <- factor(cm_df_upsample$Prediction, levels = c("No", "Yes"))

# Plot
ggplot(cm_df_upsample, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#e6f2ff", high = "#1A5276", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – LR (Upsampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )
```

```{r 04-logit-upsample-roc-curve, echo=TRUE}
# Roc
roc_curve(logit_preds_upsample, truth = Monitoring_Turnover_Risk, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – LR (Upsampling)",
    subtitle = "Model performance on predicting Monitoring Turnover Risk",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )
```

```{r 04-logit-upsample-vip, echo=TRUE}
# Styled VIP for top 15 features
fit_logit_upsample %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#1A5276", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – LR (Upsampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

### Logistic Regression - Downsampling

```{r 04-logit-downsample-workflow, message=FALSE}
# Downsampling Recipe
recipe_logit_downsample <- base_recipe %>%
  step_downsample(Monitoring_Turnover_Risk)

# Workflow 
workflow_logit_downsample <- workflow() %>%
  add_model(logit_model) %>%
  add_recipe(recipe_logit_downsample)

# Fit on training data
fit_logit_downsample <- fit(workflow_logit_downsample, data = mtr_train_data)

# Predict on test data
logit_preds_downsample <- predict(fit_logit_downsample, new_data = mtr_test_data, type = "prob") %>%
  bind_cols(predict(fit_logit_downsample, new_data = mtr_test_data)) %>%
  bind_cols(mtr_test_data %>% select(Monitoring_Turnover_Risk))
```

### Logistic Regression - Downsampling - Performance Metrics

```{r 04-logit-downsample-metrics, echo=TRUE}
# Metrics
logit_downsample_metrics <- all_metrics(logit_preds_downsample, truth = Monitoring_Turnover_Risk, estimate = .pred_class, .pred_Yes)

# Plot
logit_downsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#4B9CD3", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – LR (Downsampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

```{r 04-logit-downsample-confmat, echo=TRUE}
cm_logit_down <- conf_mat(logit_preds_downsample, truth = Monitoring_Turnover_Risk, estimate = .pred_class)
cm_df_down <- as.data.frame(cm_logit_down$table)

cm_df_down$Truth <- factor(cm_df_down$Truth, levels = c("Yes", "No"))
cm_df_down$Prediction <- factor(cm_df_down$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_down, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#e6f2ff", high = "#1A5276", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – LR (Downsampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )
```

```{r 04-logit-downsample-roc-curve, echo=TRUE}
roc_curve(logit_preds_downsample, truth = Monitoring_Turnover_Risk, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – LR (Downsampling)",
    subtitle = "Model performance on predicting Monitoring Turnover Risk",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )
```

```{r 04-logit-downsample-vip, echo=TRUE}
fit_logit_downsample %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#1A5276", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – LR (Downsampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

### Results Table for Logistic Regression (Monitoring Turnover Risk)

```{r 04-logit-results-table, results='asis', echo=FALSE}
logit_results_long <- bind_rows(
  logit_none_metrics %>% mutate(Resample = "No Sampling"),
  logit_smote_metrics %>% mutate(Resample = "SMOTE"),
  logit_upsample_metrics %>% mutate(Resample = "Up Sampling"),
  logit_downsample_metrics %>% mutate(Resample = "Down Sampling")
) %>%
  mutate(
    Metric = factor(.metric,
                    levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                    labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC")),
    Score = .estimate,
    Resample = factor(Resample, levels = c("No Sampling", "SMOTE", "Up Sampling", "Down Sampling"))
  ) %>%
  filter(!is.na(Score))

highlight_logit <- logit_results_long %>%
  group_by(Metric) %>%
  filter(Score == max(Score, na.rm = TRUE)) %>%
  mutate(is_best = TRUE)

logit_results_highlight <- left_join(logit_results_long, highlight_logit %>% select(Resample, Metric, is_best),
                                     by = c("Resample", "Metric")) %>%
  mutate(is_best = ifelse(is.na(is_best), FALSE, is_best))

ggplot(logit_results_highlight, aes(x = Resample, y = Score, fill = Metric)) +
  geom_col(width = 0.7) +
  geom_col(data = subset(logit_results_highlight, is_best),
           fill = NA, color = "black", linewidth = 1.2, width = 0.7) +
  facet_wrap(~ Metric, nrow = 2, scales = "free_y") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_manual(values = rep("#4B9CD3", 6)) +
  labs(
    title = "LR Performance by Metric and Resampling Method",
    subtitle = "Black outline shows the best result for each metric",
    x = "Resampling Strategy", y = "Score"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 30, hjust = 1),
    strip.text = element_text(face = "bold"),
    legend.position = "none"
  )
```

### Executive Summary for Logistic Regression (Monitoring Turnover Risk)

Logistic Regression offered a strong and interpretable baseline for turnover risk monitoring, especially under the no resampling strategy which led in multiple metrics.

- **No Resampling** delivered the top performance in **recall (0.828)**, **F1 score (0.819)**, and **accuracy (0.736)** — making it ideal for identifying at-risk employees, though its **ROC AUC (0.283)** and **specificity (0.500)** suggest limited class separation.

- **SMOTE** yielded the **highest precision (0.847)** while still maintaining good **F1 (0.734)** and **recall (0.648)**, offering a solid trade-off between minimizing false positives and detecting true positives.

- **Upsampling** produced very similar results to SMOTE — strong **precision (0.841)** and **F1 score (0.714)**, but with slightly lower **recall (0.621)**. Like SMOTE, ROC AUC remained modest (0.295).

- **Downsampling** emphasized **specificity (0.714)** and **precision (0.838)**, reducing false positives but at the expense of **recall (0.572)** and **F1 score (0.680)**. It’s useful when it's more important to avoid false alarms.

Overall, while Logistic Regression had limited ROC AUC (~0.28–0.30), it performed competitively across most other metrics and remains a reliable, interpretable choice.

## Monitoring Turnover Risk with Random Forest

### Random Forest - No Resampling

```{r 04-rf-no-resampling-workflow, message=FALSE}
# Define Random Forest model
rf_model <- rand_forest(mtry = tune(), trees = 500, min_n = tune()) %>%
  set_engine("ranger", importance = "impurity") %>%
  set_mode("classification")

# Workflow using base recipe
recipe_rf_none <- base_recipe

# Combine into Workflow
workflow_rf_none <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(recipe_rf_none)

# Tune the RF model with cross-validation
set.seed(42)
rf_tuned <- tune_grid(
  workflow_rf_none,
  resamples = mtr_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_rf <- select_best(rf_tuned, metric = "roc_auc")

# Final workflow with best RF model
final_rf_workflow <- finalize_workflow(workflow_rf_none, best_rf)

# Fit to training data
rf_fit_none <- fit(final_rf_workflow, data = mtr_train_data)

# Predict on test data
rf_preds_none <- predict(rf_fit_none, new_data = mtr_test_data, type = "prob") %>%
  bind_cols(predict(rf_fit_none, new_data = mtr_test_data)) %>%
  bind_cols(mtr_test_data %>% select(Monitoring_Turnover_Risk))
```

### Random Forest - No Resampling - Performance Metrics

```{r 04-rf-no-resampling-metrics, echo=TRUE}
# Metrics
rf_none_metrics <- all_metrics(rf_preds_none, truth = Monitoring_Turnover_Risk, estimate = .pred_class, .pred_Yes)

# Plot
rf_none_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#3FA34D", width = 0.7) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – RF (No Resampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-rf-no-resampling-confmat, echo=TRUE}
rf_cm_none <- conf_mat(rf_preds_none, truth = Monitoring_Turnover_Risk, estimate = .pred_class)
rf_cm_df_none <- as.data.frame(rf_cm_none$table)

rf_cm_df_none$Truth <- factor(rf_cm_df_none$Truth, levels = c("Yes", "No"))
rf_cm_df_none$Prediction <- factor(rf_cm_df_none$Prediction, levels = c("No", "Yes"))

ggplot(rf_cm_df_none, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#D5F5E3", high = "#145A32", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – RF (No Resampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )
```

```{r 04-rf-no-resampling-roc-curve, echo=TRUE}
# ROC Curve
roc_curve(rf_preds_none, truth = Monitoring_Turnover_Risk, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray70") +
  labs(
    title = "ROC Curve – RF (No Resampling)",
    subtitle = "Model performance on predicting Monitoring Turnover Risk",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

```{r 04-rf-no-resampling-vip, echo=TRUE}
# Variable Importance
rf_fit_none %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#3FA34D", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – RF (No Resampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

### Random Forest - SMOTE

```{r 04-rf-smote-workflow, message=FALSE}
# Create Recipe from Base
recipe_rf_smote <- base_recipe %>%
  step_smote(Monitoring_Turnover_Risk)

# Workflow with Random Forest
workflow_rf_smote <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(recipe_rf_smote)

# Tune the RF model with cross-validation
set.seed(42)
rf_tuned_smote <- tune_grid(
  workflow_rf_smote,
  resamples = mtr_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_rf_smote <- select_best(rf_tuned_smote, metric = "roc_auc")

# Final workflow with best RF model
final_rf_workflow_smote <- finalize_workflow(workflow_rf_smote, best_rf_smote)

# Fit to training data
rf_fit_smote <- fit(final_rf_workflow_smote, data = mtr_train_data)

# Predict on test data
rf_preds_smote <- predict(rf_fit_smote, new_data = mtr_test_data, type = "prob") %>%
  bind_cols(predict(rf_fit_smote, new_data = mtr_test_data)) %>%
  bind_cols(mtr_test_data %>% select(Monitoring_Turnover_Risk))
```

### Random Forest - SMOTE - Performance Metrics

```{r 04-rf-smote-metrics, echo=TRUE}
# Metrics
rf_smote_metrics <- all_metrics(rf_preds_smote, truth = Monitoring_Turnover_Risk, estimate = .pred_class, .pred_Yes)

# Plot
rf_smote_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#3FA34D", width = 0.7) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – RF (SMOTE)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-rf-smote-confmat, echo=TRUE}
# Confusion Matrix
rf_cm_smote <- conf_mat(rf_preds_smote, truth = Monitoring_Turnover_Risk, estimate = .pred_class)
rf_cm_df_smote <- as.data.frame(rf_cm_smote$table)

rf_cm_df_smote$Truth <- factor(rf_cm_df_smote$Truth, levels = c("Yes", "No"))
rf_cm_df_smote$Prediction <- factor(rf_cm_df_smote$Prediction, levels = c("No", "Yes"))

ggplot(rf_cm_df_smote, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#D5F5E3", high = "#145A32", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – RF (SMOTE)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-rf-smote-roc-curve, echo=TRUE}
# ROC Curve
roc_curve(rf_preds_smote, truth = Monitoring_Turnover_Risk, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray70") +
  labs(
    title = "ROC Curve: RF on Monitoring Turnover Risk (SMOTE)",
    subtitle = "Model performance on predicting Monitoring Turnover Risk",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

```{r 04-rf-smote-vip, echo=TRUE}
# Variable Importance
rf_fit_smote %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#3FA34D", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – RF (SMOTE)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

### Random Forest - Upsampling

```{r 04-rf-upsample-workflow, message=FALSE}
# Create Recipe from Base
recipe_rf_upsample <- base_recipe %>%
  step_upsample(Monitoring_Turnover_Risk)

# Workflow with Random Forest
workflow_rf_upsample <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(recipe_rf_upsample)

# Tune the RF model with cross-validation
set.seed(42)
rf_tuned_upsample <- tune_grid(
  workflow_rf_upsample,
  resamples = mtr_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_rf_upsample <- select_best(rf_tuned_upsample, metric = "roc_auc")

# Final workflow with best RF model
final_rf_workflow_upsample <- finalize_workflow(workflow_rf_upsample, best_rf_upsample)

# Fit to training data
rf_fit_upsample <- fit(final_rf_workflow_upsample, data = mtr_train_data)

# Predict on test data
rf_preds_upsample <- predict(rf_fit_upsample, new_data = mtr_test_data, type = "prob") %>%
  bind_cols(predict(rf_fit_upsample, new_data = mtr_test_data)) %>%
  bind_cols(mtr_test_data %>% select(Monitoring_Turnover_Risk))
```

### Random Forest - Upsampling - Performance Metrics

```{r 04-rf-upsample-metrics, echo=TRUE}
# Metrics
rf_upsample_metrics <- all_metrics(rf_preds_upsample, truth = Monitoring_Turnover_Risk, estimate = .pred_class, .pred_Yes)

# Plot
rf_upsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#3FA34D", width = 0.7) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – RF (Upsampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-rf-upsample-confmat, echo=TRUE}
# Confusion Matrix
rf_cm_up <- conf_mat(rf_preds_upsample, truth = Monitoring_Turnover_Risk, estimate = .pred_class)
rf_cm_df_up <- as.data.frame(rf_cm_up$table)

rf_cm_df_up$Truth <- factor(rf_cm_df_up$Truth, levels = c("Yes", "No"))
rf_cm_df_up$Prediction <- factor(rf_cm_df_up$Prediction, levels = c("No", "Yes"))

ggplot(rf_cm_df_up, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#D5F5E3", high = "#145A32", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – RF (Upsampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-rf-upsample-roc-curve, echo=TRUE}
# ROC Curve
roc_curve(rf_preds_upsample, truth = Monitoring_Turnover_Risk, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray70") +
  labs(
    title = "ROC Curve – RF (Upsampling)",
    subtitle = "Model performance on predicting Monitoring Turnover Risk",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

```{r 04-rf-upsample-var-imp, echo=TRUE}
# Variable Importance
rf_fit_upsample %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#3FA34D", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – RF (Upsampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

### Random Forest - Downsampling

```{r 04-rf-downsample-workflow, message=FALSE}
# Create Recipe from Base
recipe_rf_downsample <- base_recipe %>%
  step_downsample(Monitoring_Turnover_Risk)

# Workflow with Random Forest
workflow_rf_downsample <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(recipe_rf_downsample)

# Tune the RF model with cross-validation
set.seed(42)
rf_tuned_downsample <- tune_grid(
  workflow_rf_downsample,
  resamples = mtr_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_rf_downsample <- select_best(rf_tuned_downsample, metric = "roc_auc")

# Final workflow with best RF model
final_rf_workflow_downsample <- finalize_workflow(workflow_rf_downsample, best_rf_downsample)

# Fit to training data
rf_fit_downsample <- fit(final_rf_workflow_downsample, data = mtr_train_data)

# Predict on test data
rf_preds_downsample <- predict(rf_fit_downsample, new_data = mtr_test_data, type = "prob") %>%
  bind_cols(predict(rf_fit_downsample, new_data = mtr_test_data)) %>%
  bind_cols(mtr_test_data %>% select(Monitoring_Turnover_Risk))
```

### Random Forest - Downsampling - Performance Metrics

```{r 04-rf-downsample-metrics, echo=TRUE}
# Metrics
rf_downsample_metrics <- all_metrics(rf_preds_downsample, truth = Monitoring_Turnover_Risk, estimate = .pred_class, .pred_Yes)

# Plot
rf_downsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#3FA34D", width = 0.7) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – RF (Downsampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-rf-downsample-confmat, echo=TRUE}
# Confusion Matrix
rf_cm_down <- conf_mat(rf_preds_downsample, truth = Monitoring_Turnover_Risk, estimate = .pred_class)
rf_cm_df_down <- as.data.frame(rf_cm_down$table)

rf_cm_df_down$Truth <- factor(rf_cm_df_down$Truth, levels = c("Yes", "No"))
rf_cm_df_down$Prediction <- factor(rf_cm_df_down$Prediction, levels = c("No", "Yes"))

ggplot(rf_cm_df_down, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#D5F5E3", high = "#145A32", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – RF (Downsampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-rf-downsample-roc-curve, echo=TRUE}
# ROC Curve
roc_curve(rf_preds_downsample, truth = Monitoring_Turnover_Risk, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray70") +
  labs(
    title = "ROC Curve – RF (Downsampling)",
    subtitle = "Model performance on predicting Monitoring Turnover Risk",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

```{r 04-rf-downsample-vip, echo=TRUE}
# Variable Importance
rf_fit_downsample %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#3FA34D", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – RF (Downsampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

### Results Table for Random Forest (Monitoring Turnover Risk)

```{r 04-rf-results-table, results='asis', echo=TRUE}
rf_results_long <- rf_results_wide %>%
  pivot_longer(cols = -c(Model, Resample), names_to = "Metric", values_to = "Score") %>%
  mutate(
    Metric = factor(Metric,
                    levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                    labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC")),
    Resample = factor(Resample, levels = c("No Sampling", "SMOTE", "Up Sampling", "Down Sampling"))
  ) %>%
  filter(!is.na(Score))

highlight_rf <- rf_results_long %>%
  group_by(Metric) %>%
  filter(Score == max(Score, na.rm = TRUE)) %>%
  mutate(is_best = TRUE)

rf_results_highlight <- left_join(rf_results_long, highlight_rf %>% select(Resample, Metric, is_best),
                                  by = c("Resample", "Metric")) %>%
  mutate(is_best = ifelse(is.na(is_best), FALSE, is_best))

ggplot(rf_results_highlight, aes(x = Resample, y = Score, fill = Metric)) +
  geom_col(width = 0.7) +
  geom_col(data = subset(rf_results_highlight, is_best),
           fill = NA, color = "black", linewidth = 1.2, width = 0.7) +
  facet_wrap(~ Metric, nrow = 2, scales = "free_y") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_manual(values = rep("#3FA34D", 6)) +
  labs(
    title = "RF Performance by Metric and Resampling Method",
    subtitle = "Black outline shows the best result for each metric",
    x = "Resampling Strategy", y = "Score"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 30, hjust = 1),
    strip.text = element_text(face = "bold"),
    legend.position = "none"
  )
```

### Monitoring Turnover Risk – Random Forest Performance Summary

Random Forest models achieved top classification results across nearly all metrics, with consistent performance regardless of the resampling strategy. However, all variants exhibited **low ROC AUC scores**, indicating poor class probability ranking.

- **No Resampling** yielded **perfect recall (1.000)**, high **F1 score (0.927)**, and excellent **accuracy (0.886)**. **Precision (0.863)** was also strong, but **ROC AUC (0.159)** was weak, showing limited effectiveness in ranking true positives above false ones.

- **SMOTE** delivered identical results to no-resampling, with **recall (1.000)**, **F1 (0.927)**, **accuracy (0.886)**, and **precision (0.863)**. However, **ROC AUC dropped to 0.133**, indicating slight degradation in model confidence separation.

- **Upsampling** maintained high classification performance with **recall (0.993)** and **F1 score (0.923)**. **Accuracy (0.881)** and **precision (0.862)** were nearly on par. Still, **ROC AUC (0.148)** remained low, continuing the trend of poor class ranking.

- **Downsampling** stood out for the **highest specificity (0.661)**, balancing false positives better than other versions. While **recall (0.952)** and **F1 score (0.914)** were slightly reduced, it offered the **best ROC AUC (0.155)** of the group — albeit still low in absolute terms.

Despite outstanding performance on traditional classification metrics, the low ROC AUC values suggest caution: the model's confidence estimates may be unreliable for threshold-based decision-making or risk scoring purposes.

## Monitoring Turnover Risk with XGBoost

### XGBoost - No Resampling

```{r 04-xgb-no-resampling-workflow, message=FALSE}
# Define XGBoost model
xgb_model <- boost_tree(
  trees = 1000,
  tree_depth = tune(),
  min_n = tune(),
  loss_reduction = tune(),        
  sample_size = tune(),           
  mtry = tune(),                  
  learn_rate = tune()             
) %>%
  set_engine("xgboost") %>%
  set_mode("classification")

# Recipe with no resampling
recipe_xgb_none <- base_recipe

# Combine into Workflow
workflow_xgb_none <- workflow() %>%
  add_model(xgb_model) %>%
  add_recipe(recipe_xgb_none)

# Tune the XGBoost model with cross-validation
set.seed(42)
xgb_tuned <- tune_grid(
  workflow_xgb_none,
  resamples = mtr_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_xgb <- select_best(xgb_tuned, metric = "roc_auc")

# Final workflow with best XGBoost model
final_xgb_workflow <- finalize_workflow(workflow_xgb_none, best_xgb)

# Fit to training data
xgb_fit_none <- fit(final_xgb_workflow, data = mtr_train_data)

# Predict on test data
xgb_preds_none <- predict(xgb_fit_none, new_data = mtr_test_data, type = "prob") %>%
  bind_cols(predict(xgb_fit_none, new_data = mtr_test_data)) %>%
  bind_cols(mtr_test_data %>% select(Monitoring_Turnover_Risk))
```

### XGBoost - No Resampling - Performance Metrics

```{r 04-xgb-no-resampling-metrics, echo=TRUE}
# Metrics
xgb_none_metrics <- all_metrics(xgb_preds_none, truth = Monitoring_Turnover_Risk, estimate = .pred_class, .pred_Yes)

# Plot
xgb_none_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#E87461", width = 0.7) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – XGB (No Resampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

```{r 04-xgb-no-resampling-confmat, echo=TRUE}
# Confusion Matrix
xgb_cm_none <- conf_mat(xgb_preds_none, truth = Monitoring_Turnover_Risk, estimate = .pred_class)
xgb_cm_df_none <- as.data.frame(xgb_cm_none$table)

xgb_cm_df_none$Truth <- factor(xgb_cm_df_none$Truth, levels = c("Yes", "No"))
xgb_cm_df_none$Prediction <- factor(xgb_cm_df_none$Prediction, levels = c("No", "Yes"))

ggplot(xgb_cm_df_none, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#FDEDEC", high = "#922B21", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – XGB (No Resampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-xgb-no-resampling-roc-curve, echo=TRUE}
# ROC Curve
roc_curve(xgb_preds_none, truth = Monitoring_Turnover_Risk, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray70") +
  labs(
    title = "ROC Curve – XGB (No Resampling)",
    subtitle = "Model performance on predicting Monitoring Turnover Risk",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

```{r 04-xgb-no-resampling-vip, echo=TRUE}
# Variable Importance
xgb_fit_none %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#E87461", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – XGB (No Resampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

### XGBoost - SMOTE

```{r 04-xgb-smote-workflow, message=FALSE}
# Create SMOTE Recipe
recipe_xgb_smote <- base_recipe %>%
  step_smote(Monitoring_Turnover_Risk)

# Build XGBoost Workflow with SMOTE
workflow_xgb_smote <- workflow() %>%
  add_model(xgb_model) %>%
  add_recipe(recipe_xgb_smote)

# Tune Hyperparameters with Cross-Validation
set.seed(42)
xgb_tuned_smote <- tune_grid(
  workflow_xgb_smote,
  resamples = mtr_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
)

# Select Best Parameters Based on ROC AUC
best_xgb_smote <- select_best(xgb_tuned_smote, metric = "roc_auc")

# Finalize Workflow with Best Parameters
final_xgb_workflow_smote <- finalize_workflow(workflow_xgb_smote, best_xgb_smote)

# Fit Final Model to Training Data
xgb_fit_smote <- fit(final_xgb_workflow_smote, data = mtr_train_data)

# Predict on Test Data
xgb_preds_smote <- predict(xgb_fit_smote, new_data = mtr_test_data, type = "prob") %>%
  bind_cols(predict(xgb_fit_smote, new_data = mtr_test_data)) %>%
  bind_cols(mtr_test_data %>% select(Monitoring_Turnover_Risk))
```

### XGBoost - SMOTE - Peformance Metrics

```{r 04-xgb-smote-metrics, echo=TRUE}
# Metrics
xgb_smote_metrics <- all_metrics(xgb_preds_smote, truth = Monitoring_Turnover_Risk, estimate = .pred_class, .pred_Yes)

# Plot
xgb_smote_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#E74C3C", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – XGB (SMOTE)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-xgb-smote-confmat, echo=TRUE}
# Confusion Matrix
cm_xgb_smote <- conf_mat(xgb_preds_smote, truth = Monitoring_Turnover_Risk, estimate = .pred_class)
cm_df_xgb_smote <- as.data.frame(cm_xgb_smote$table)

cm_df_xgb_smote$Truth <- factor(cm_df_xgb_smote$Truth, levels = c("Yes", "No"))
cm_df_xgb_smote$Prediction <- factor(cm_df_xgb_smote$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_xgb_smote, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#fde0dc", high = "#c0392b", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – XGB (SMOTE)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-xgb-smote-roc-curve, echo=TRUE}
# ROC Curve
roc_curve(xgb_preds_smote, truth = Monitoring_Turnover_Risk, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – XGB (SMOTE)",
    subtitle = "Model performance on predicting Monitoring Turnover Risk",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

```{r 04-xgb-smote-vip, echo=TRUE}
# Variable Importance
vip_xgb_smote <- xgb_fit_smote %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance))

ggplot(vip_xgb_smote, aes(x = Importance, y = Variable)) +
  geom_col(fill = "#C0392B", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – XGB (SMOTE)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

### XGBoost - Upsampling

```{r 04-xgb-upsample-workflow, message=FALSE}
# Create Recipe from Base
recipe_xgb_upsample <- base_recipe %>%
  step_upsample(Monitoring_Turnover_Risk)

# Workflow with XGBoost
workflow_xgb_upsample <- workflow() %>%
  add_model(xgb_model) %>%
  add_recipe(recipe_xgb_upsample)

# Tune the XGBoost model with cross-validation
set.seed(42)
xgb_tuned_upsample <- tune_grid(
  workflow_xgb_upsample,
  resamples = mtr_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_xgb_upsample <- select_best(xgb_tuned_upsample, metric = "roc_auc")

# Final workflow with best XGBoost model
final_xgb_workflow_upsample <- finalize_workflow(workflow_xgb_upsample, best_xgb_upsample)

# Fit to training data
xgb_fit_upsample <- fit(final_xgb_workflow_upsample, data = mtr_train_data)

# Predict on test data
xgb_preds_upsample <- predict(xgb_fit_upsample, new_data = mtr_test_data, type = "prob") %>%
  bind_cols(predict(xgb_fit_upsample, new_data = mtr_test_data)) %>%
  bind_cols(mtr_test_data %>% select(Monitoring_Turnover_Risk))
```

### XGBoost - Upsampling - Performance Metrics

```{r 04-xgb-upsample-metrics, message=FALSE}
# Metrics
xgb_upsample_metrics <- all_metrics(xgb_preds_upsample, truth = Monitoring_Turnover_Risk, estimate = .pred_class, .pred_Yes)

# Plot
xgb_upsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
         !is.na(.estimate)) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate, .na_rm = TRUE))) +
  geom_col(fill = "#E74C3C", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – XGB (Upsampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )
```

```{r 04-xgb-upsample-confmat, echo=TRUE}
# Confusion Matrix
cm_xgb_upsample <- conf_mat(xgb_preds_upsample, truth = Monitoring_Turnover_Risk, estimate = .pred_class)
cm_df_xgb_upsample <- as.data.frame(cm_xgb_upsample$table)

cm_df_xgb_upsample$Truth <- factor(cm_df_xgb_upsample$Truth, levels = c("Yes", "No"))
cm_df_xgb_upsample$Prediction <- factor(cm_df_xgb_upsample$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_xgb_upsample, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#fde0dc", high = "#c0392b", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – XGB (Upsampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-xgb-upsample-roc-curve, echo=TRUE}
# ROC Curve
roc_curve(xgb_preds_upsample, truth = Monitoring_Turnover_Risk, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – XGB (Upsampling)",
    subtitle = "Model performance on predicting Monitoring Turnover Risk",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

```{r 04-xgb-upsample-vip, echo=TRUE}
# Variable Importance
# Step 1: Prep the recipe used for training
baked_train <- prep(recipe_xgb_upsample, training = mtr_train_data) %>%
  bake(new_data = NULL)

# Step 2: Get feature names from preprocessed data (after dummy variables)
feature_names <- colnames(baked_train %>% select(-Monitoring_Turnover_Risk))

# Step 3: Extract booster
xgb_booster_upsample <- extract_fit_engine(xgb_fit_upsample)

# Step 4: Plot with feature names
vip::vip(
  object = xgb_booster_upsample,
  num_features = 15,
  geom = "col",
  feature_names = feature_names,
  aesthetics = list(fill = "#C0392B")
) +
  labs(
    title = "Top Features – XGB (Upsampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

### XGBoost - Downsampling

```{r 04-xgb-downsample-workflow, message=FALSE}
# Create Recipe from Base
recipe_xgb_downsample <- base_recipe %>%
  step_downsample(Monitoring_Turnover_Risk)

# Workflow with XGBoost
workflow_xgb_downsample <- workflow() %>%
  add_model(xgb_model) %>%
  add_recipe(recipe_xgb_downsample)

# Tune the RF model with cross-validation
set.seed(42)
xgb_tuned_downsample <- tune_grid(
  workflow_xgb_downsample,
  resamples = mtr_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
  )

# Select Best model by AUC
best_xgb_downsample <- select_best(xgb_tuned_downsample, metric = "roc_auc")

# Final workflow with best XGBoost model
final_xgb_workflow_downsample <- finalize_workflow(workflow_xgb_downsample, best_xgb_downsample)

# Fit to training data
xgb_fit_downsample <- fit(final_xgb_workflow_downsample, data = mtr_train_data)

# Predict on test data
xgb_preds_downsample <- predict(xgb_fit_downsample, new_data = mtr_test_data, type = "prob") %>%
  bind_cols(predict(xgb_fit_downsample, new_data = mtr_test_data)) %>%
  bind_cols(mtr_test_data %>% select(Monitoring_Turnover_Risk))
```

### XGBoost - Downsampling - Performance Metrics

```{r 04-xgb-downsample-metrics, echo=TRUE}
# Metrics
xgb_downsample_metrics <- all_metrics(xgb_preds_downsample, truth = Monitoring_Turnover_Risk, estimate = .pred_class, .pred_Yes)

# Print
xgb_downsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#E74C3C", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)),
            hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – XGB (Downsampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-xgb-down-sample-confmat, echo=TRUE}
# Confusion Matrix
cm_xgb_down <- conf_mat(xgb_preds_downsample, truth = Monitoring_Turnover_Risk, estimate = .pred_class)
cm_df_xgb_down <- as.data.frame(cm_xgb_down$table)

cm_df_xgb_down$Truth <- factor(cm_df_xgb_down$Truth, levels = c("Yes", "No"))
cm_df_xgb_down$Prediction <- factor(cm_df_xgb_down$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_xgb_down, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#fde0dc", high = "#c0392b", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – XGB (Downsampling)",
    x = "Predicted Class",
    y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-xgb-downsample-roc-curve, echo=TRUE}
# ROC Curve
roc_curve(xgb_preds_downsample, truth = Monitoring_Turnover_Risk, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – XGB (Downsampling)",
    subtitle = "Model performance on predicting Monitoring Turnover Risk",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

```{r 04-xgb-downsample-vip, echo=TRUE}
# Variable Importance
xgb_fit_downsample %>%
  extract_fit_parsnip() %>%
  vi() %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 15) %>%
  mutate(Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable)) +
  geom_col(fill = "#C0392B", width = 0.7) +
  geom_text(aes(label = round(Importance, 1)), hjust = -0.1, color = "black", size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Features – XGB (Downsampling)",
    x = "Importance Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

### Results Table for XGBoost (Monitoring Turnover Risk)

```{r 04-xgb-results-table, results='asis', echo=TRUE}
# Combine metrics and label with model names
xgb_results_table <- bind_rows(
  all_metrics(xgb_preds_none, truth = Monitoring_Turnover_Risk, estimate = .pred_class, .pred_Yes) %>%
    mutate(model = "XGB - No Resampling"),
  all_metrics(xgb_preds_smote, truth = Monitoring_Turnover_Risk, estimate = .pred_class, .pred_Yes) %>%
    mutate(model = "XGB - SMOTE"),
  all_metrics(xgb_preds_upsample, truth = Monitoring_Turnover_Risk, estimate = .pred_class, .pred_Yes) %>%
    mutate(model = "XGB - Up Sampling"),
  all_metrics(xgb_preds_downsample, truth = Monitoring_Turnover_Risk, estimate = .pred_class, .pred_Yes) %>%
    mutate(model = "XGB - Down Sampling")
)

# Reshape and clean
xgb_results_long <- xgb_results_table %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(
    Resample = case_when(
      str_detect(model, "No Resampling") ~ "No Sampling",
      str_detect(model, "SMOTE") ~ "SMOTE",
      str_detect(model, "Up Sampling") ~ "Up Sampling",
      str_detect(model, "Down Sampling") ~ "Down Sampling",
      TRUE ~ NA_character_
    ),
    Metric = factor(.metric,
                    levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                    labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))
  ) %>%
  filter(!is.na(Resample), !is.na(.estimate))

# Identify top score per metric
highlight_xgb <- xgb_results_long %>%
  group_by(Metric) %>%
  filter(.estimate == max(.estimate, na.rm = TRUE)) %>%
  mutate(is_best = TRUE)

# Merge with main table
xgb_results_highlight <- xgb_results_long %>%
  left_join(highlight_xgb %>% select(Resample, Metric, is_best), by = c("Resample", "Metric")) %>%
  mutate(is_best = ifelse(is.na(is_best), FALSE, is_best))

# Plot
ggplot(xgb_results_highlight, aes(x = Resample, y = .estimate, fill = Metric)) +
  geom_col(width = 0.7, color = NA) +
  geom_col(
    data = subset(xgb_results_highlight, is_best),
    aes(x = Resample, y = .estimate),
    fill = NA,
    color = "black",
    linewidth = 1.2,
    width = 0.7
  ) +
  facet_wrap(~ Metric, nrow = 2, scales = "free_y") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_manual(values = rep("#C0392B", 6)) +
  labs(
    title = "XGB Performance by Metric and Resampling Method",
    subtitle = "Black outline shows the best result for each metric",
    x = "Resampling Strategy", y = "Score"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 30, hjust = 1),
    strip.text = element_text(face = "bold"),
    legend.position = "none"
  )
```

### XGBoost – Turnover Risk Modeling Summary

XGBoost delivered **high-performing classification results** with strong precision, recall, and F1 scores across most sampling techniques. It consistently captured at-risk employees well, though like other models, it struggled with **calibrated probability estimates**, as reflected in modest ROC AUC scores. Notably, **No Resampling achieved perfect recall**, while **Downsampling delivered the best ROC AUC and specificity**, highlighting varied tradeoffs by sampling strategy.

- **No Resampling** delivered **perfect recall (1.000)** and strong overall metrics including **F1 score (0.927)**, **accuracy (0.886)**, and **precision (0.863)**. However, **ROC AUC was low (0.197)**, indicating limited ability to rank risk probabilities effectively.

- **SMOTE** produced high **recall (0.945)** and **F1 score (0.904)** with strong **precision (0.867)**. **Specificity improved to 0.625**, but **ROC AUC declined slightly to 0.178**, continuing the trend of poor calibration despite high classification scores.

- **Upsampling** maintained excellent performance with **recall (0.910)** and **F1 score (0.889)**, while achieving the **highest specificity (0.643)** among oversampling methods. **ROC AUC was modest at 0.186**.

- **Downsampling** had slightly lower **recall (0.717)** and **F1 score (0.776)**, but delivered the **best ROC AUC (0.223)** and strong **specificity (0.661)**. This method reduced overfitting and improved the model’s probability ranking capability — a distinct advantage over other sampling methods.

XGBoost showed excellent classification metrics across all sampling methods, particularly with **No Resampling** and **SMOTE** for recall and F1. However, if **probability calibration or risk ranking is a priority**, **Downsampling emerged as the best option** due to its superior ROC AUC and specificity. All ROC AUC values remained below **0.25**, highlighting continued challenges in probability-based prediction.

## Monitoring Turnover Risk with Naive Bayes

### Naive Bayes - No Resampling

```{r 04-nb-model, message=FALSE}
# Define Naive Bayes Model
nb_model <- naive_Bayes() %>%
  set_engine("naivebayes") %>%
  set_mode("classification")

# Build Workflow
workflow_nb_none <- workflow() %>%
  add_model(nb_model) %>%
  add_recipe(base_recipe)

# 10-Fold Cross-Validation on Training Data
set.seed(42)
nb_cv_results <- fit_resamples(
  workflow_nb_none,
  resamples = mtr_folds,
  metrics = metric_set(accuracy, precision, recall, f_meas, yardstick::spec, roc_auc)
)

# Fit Final Model to Full Training Data
nb_fit_none <- fit(workflow_nb_none, data = mtr_train_data)

# Predict on Test Data
nb_preds_none <- predict(nb_fit_none, new_data = mtr_test_data, type = "prob") %>%
  bind_cols(predict(nb_fit_none, new_data = mtr_test_data)) %>%
  bind_cols(mtr_test_data %>% select(Monitoring_Turnover_Risk))
```

### Naive Bayes - No Resampling - Performance Metrics

```{r 04-nb-metrics, echo=TRUE}
# Metrics
nb_none_metrics <- all_metrics(nb_preds_none, truth = Monitoring_Turnover_Risk, estimate = .pred_class, .pred_Yes)

# Plot
nb_none_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#F4D03F", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)), hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Performance Metrics – NB (No Resampling)",
    x = "Score", y = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        axis.text = element_text(face = "bold"),
        panel.grid.major.y = element_blank())

```

```{r 04-nb-matrix-none, echo=TRUE}
# Confusion Matrix
cm_nb_none <- conf_mat(nb_preds_none, truth = Monitoring_Turnover_Risk, estimate = .pred_class)
cm_df_nb_none <- as.data.frame(cm_nb_none$table)
cm_df_nb_none$Truth <- factor(cm_df_nb_none$Truth, levels = c("Yes", "No"))
cm_df_nb_none$Prediction <- factor(cm_df_nb_none$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_nb_none, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#FFF9C4", high = "#F4D03F", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(title = "Confusion Matrix – NB (No Resampling)",
       x = "Predicted Class", y = "Actual Class") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(face = "bold"),
        legend.title = element_text(face = "bold"),
        panel.grid = element_blank())

```

```{r 04-nb-roc-curve-none, echo=TRUE}
# Roc
roc_curve(nb_preds_none, truth = Monitoring_Turnover_Risk, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(title = "ROC Curve – NB (No Resampling)",
       subtitle = "Model performance on predicting Monitoring Turnover Risk",
       x = "False Positive Rate (1 - Specificity)",
       y = "True Positive Rate (Sensitivity)") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        plot.subtitle = element_text(hjust = 0.5, size = 11),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(face = "bold"))

```

### Naive Bayes - Smote

```{r 04-nb-smote, message=FALSE}
# Create Upsampling Recipe
recipe_nb_smote <- base_recipe %>%
  step_smote(Monitoring_Turnover_Risk)

# Build Workflow with Upsampling
workflow_nb_smote <- workflow() %>%
  add_model(nb_model) %>%
  add_recipe(recipe_nb_smote)

# 10-Fold Cross-Validation
set.seed(42)
nb_cv_smote <- fit_resamples(
  workflow_nb_smote,
  resamples = mtr_folds,
  metrics = metric_set(accuracy, precision, recall, f_meas, yardstick::spec, roc_auc)
)

# Final Fit on Full Training Data
nb_fit_smote <- fit(workflow_nb_smote, data = mtr_train_data)

# Predict on Test Set
nb_preds_smote <- predict(nb_fit_smote, new_data = mtr_test_data, type = "prob") %>%
  bind_cols(predict(nb_fit_smote, new_data = mtr_test_data)) %>%
  bind_cols(mtr_test_data %>% select(Monitoring_Turnover_Risk))
```

### Naive Bayes - SMOTE - Performance Metrics

```{r 04-nb-smote-metrics, echo=TRUE}
# Metrics
nb_smote_metrics <- all_metrics(nb_preds_smote, truth = Monitoring_Turnover_Risk, estimate = .pred_class, .pred_Yes)

# Print
nb_smote_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#F4D03F", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)), hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Performance Metrics – NB (SMOTE)",
       x = "Score", y = NULL) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        axis.text = element_text(face = "bold"),
        panel.grid.major.y = element_blank())

```

```{r 04-nb-smote-matrix, echo=TRUE}
# Confusion Matrix
cm_nb_smote <- conf_mat(nb_preds_smote, truth = Monitoring_Turnover_Risk, estimate = .pred_class)
cm_df_nb_smote <- as.data.frame(cm_nb_smote$table)
cm_df_nb_smote$Truth <- factor(cm_df_nb_smote$Truth, levels = c("Yes", "No"))
cm_df_nb_smote$Prediction <- factor(cm_df_nb_smote$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_nb_smote, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#FFF9C4", high = "#F4D03F", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(title = "Confusion Matrix – NB (SMOTE)",
       x = "Predicted Class", y = "Actual Class") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(face = "bold"),
        legend.title = element_text(face = "bold"),
        panel.grid = element_blank())

```

```{r 04-nb-roc-curve-smote, echo=TRUE}
# Roc
roc_curve(nb_preds_smote, truth = Monitoring_Turnover_Risk, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(title = "ROC Curve – NB (SMOTE)",
       subtitle = "Model performance on predicting Monitoring Turnover Risk",
       x = "False Positive Rate (1 - Specificity)",
       y = "True Positive Rate (Sensitivity)") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        plot.subtitle = element_text(hjust = 0.5, size = 11),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(face = "bold"))

```

### Naive Bayes - Upsampling

```{r 04-nb-upsample}
# Create Upsampling Recipe
recipe_nb_upsample <- base_recipe %>%
  step_upsample(Monitoring_Turnover_Risk)

# Build Workflow with Upsampling
workflow_nb_upsample <- workflow() %>%
  add_model(nb_model) %>%
  add_recipe(recipe_nb_upsample)

# 10-Fold Cross-Validation
set.seed(42)
nb_cv_upsample <- fit_resamples(
  workflow_nb_upsample,
  resamples = mtr_folds,
  metrics = metric_set(accuracy, precision, recall, f_meas, yardstick::spec, roc_auc)
)

# Final Fit on Full Training Data
nb_fit_upsample <- fit(workflow_nb_upsample, data = mtr_train_data)

# Predict on Test Set
nb_preds_upsample <- predict(nb_fit_upsample, new_data = mtr_test_data, type = "prob") %>%
  bind_cols(predict(nb_fit_upsample, new_data = mtr_test_data)) %>%
  bind_cols(mtr_test_data %>% select(Monitoring_Turnover_Risk))
```

### Naive Bayes - Upsampling - Performance Metrics

```{r 04-nb-upsample-metrics, echo=TRUE}
# Metrics
nb_upsample_metrics <- all_metrics(nb_preds_upsample, truth = Monitoring_Turnover_Risk, estimate = .pred_class, .pred_Yes)

# Print
nb_upsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#F4D03F", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)), hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Performance Metrics – NB (Upsampling)",
       x = "Score", y = NULL) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        axis.text = element_text(face = "bold"),
        panel.grid.major.y = element_blank())

```

```{r 04-nb-matrix, echo=TRUE}
# Confusion Matrix
cm_nb_upsample <- conf_mat(nb_preds_upsample, truth = Monitoring_Turnover_Risk, estimate = .pred_class)
cm_df_nb_upsample <- as.data.frame(cm_nb_upsample$table)
cm_df_nb_upsample$Truth <- factor(cm_df_nb_upsample$Truth, levels = c("Yes", "No"))
cm_df_nb_upsample$Prediction <- factor(cm_df_nb_upsample$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_nb_upsample, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#FFF9C4", high = "#F4D03F", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(title = "Confusion Matrix – NB (Upsampling)",
       x = "Predicted Class", y = "Actual Class") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(face = "bold"),
        legend.title = element_text(face = "bold"),
        panel.grid = element_blank())

```

```{r 04-nb-roc-curve-upsample, echo=TRUE}
# Roc
roc_curve(nb_preds_upsample, truth = Monitoring_Turnover_Risk, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(title = "ROC Curve – NB (Upsampling)",
       subtitle = "Model performance on predicting Monitoring Turnover Risk",
       x = "False Positive Rate (1 - Specificity)",
       y = "True Positive Rate (Sensitivity)") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        plot.subtitle = element_text(hjust = 0.5, size = 11),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(face = "bold"))

```

### Naive Bayes - Downsampling

```{r 04-nb-down-sample, message=FALSE}
# Create Downsampling Recipe
recipe_nb_downsample <- base_recipe %>%
  step_downsample(Monitoring_Turnover_Risk)

# Build Workflow with Downsampling
workflow_nb_downsample <- workflow() %>%
  add_model(nb_model) %>%
  add_recipe(recipe_nb_downsample)

# 10-Fold Cross-Validation
set.seed(42)
nb_cv_downsample <- fit_resamples(
  workflow_nb_downsample,
  resamples = mtr_folds,
  metrics = metric_set(accuracy, precision, recall, f_meas, yardstick::spec, roc_auc)
)

# Fit Final Model on Full Training Data
nb_fit_downsample <- fit(workflow_nb_downsample, data = mtr_train_data)

# Predict on Test Set
nb_preds_downsample <- predict(nb_fit_downsample, new_data = mtr_test_data, type = "prob") %>%
  bind_cols(predict(nb_fit_downsample, new_data = mtr_test_data)) %>%
  bind_cols(mtr_test_data %>% select(Monitoring_Turnover_Risk))
```

### Naive Bayes - Downsampling - Performance Metrics

```{r 04-nb-down-sample-metrics, echo=TRUE}
# Metrics
nb_downsample_metrics <- all_metrics(nb_preds_downsample, truth = Monitoring_Turnover_Risk, estimate = .pred_class, .pred_Yes)

# Print
nb_downsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
                          levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#F4D03F", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)), hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Performance Metrics – NB (Downsampling)",
       x = "Score", y = NULL) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        axis.text = element_text(face = "bold"),
        panel.grid.major.y = element_blank())

```

```{r 04-nb-down-sample-matrix, echo=TRUE}
# Confusion Marix
cm_nb_downsample <- conf_mat(nb_preds_downsample, truth = Monitoring_Turnover_Risk, estimate = .pred_class)
cm_df_nb_downsample <- as.data.frame(cm_nb_downsample$table)
cm_df_nb_downsample$Truth <- factor(cm_df_nb_downsample$Truth, levels = c("Yes", "No"))
cm_df_nb_downsample$Prediction <- factor(cm_df_nb_downsample$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_nb_downsample, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#FFF9C4", high = "#F4D03F", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(title = "Confusion Matrix – NB (Downsampling)",
       x = "Predicted Class", y = "Actual Class") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(face = "bold"),
        legend.title = element_text(face = "bold"),
        panel.grid = element_blank())

```

```{r 04-nb-down-sample-roc-curve, echo=TRUE}
# Roc
roc_curve(nb_preds_downsample, truth = Monitoring_Turnover_Risk, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(title = "ROC Curve – NB (Downsampling)",
       subtitle = "Model performance on predicting Monitoring Turnover Risk",
       x = "False Positive Rate (1 - Specificity)",
       y = "True Positive Rate (Sensitivity)") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
        plot.subtitle = element_text(hjust = 0.5, size = 11),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(face = "bold"))

```

### Results Table for Naive Bayes (Monitoring Turnover Risk)

```{r 04-nb-results-table, results='asis', echo=TRUE}
nb_results_long <- bind_rows(
  nb_none_metrics %>% mutate(Resample = "No Sampling"),
  nb_smote_metrics %>% mutate(Resample = "SMOTE"),
  nb_upsample_metrics %>% mutate(Resample = "Up Sampling"),
  nb_downsample_metrics %>% mutate(Resample = "Down Sampling")
) %>%
  mutate(
    Metric = factor(.metric,
                    levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                    labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC")),
    Score = .estimate,
    Resample = factor(Resample, levels = c("No Sampling", "SMOTE", "Up Sampling", "Down Sampling"))
  ) %>%
  filter(!is.na(Score))

highlight_nb <- nb_results_long %>%
  group_by(Metric) %>%
  filter(Score == max(Score, na.rm = TRUE)) %>%
  mutate(is_best = TRUE)

nb_results_highlight <- left_join(nb_results_long, highlight_nb %>% select(Resample, Metric, is_best),
                                  by = c("Resample", "Metric")) %>%
  mutate(is_best = ifelse(is.na(is_best), FALSE, is_best))

ggplot(nb_results_highlight, aes(x = Resample, y = Score, fill = Metric)) +
  geom_col(width = 0.7) +
  geom_col(data = subset(nb_results_highlight, is_best),
           fill = NA, color = "black", linewidth = 1.2, width = 0.7) +
  facet_wrap(~ Metric, nrow = 2, scales = "free_y") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_manual(values = rep("#F4D03F", 6)) +
  labs(
    title = "NB Performance by Metric and Resampling Method",
    subtitle = "Black outline shows the best result for each metric",
    x = "Resampling Strategy", y = "Score"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 30, hjust = 1),
    strip.text = element_text(face = "bold"),
    legend.position = "none"
  )

```

### Executive Summary for Naive Bayes (Monitoring Turnover Risk)

Naive Bayes delivered **mixed classification performance** when predicting turnover risk. Its simple structure and strong assumptions (e.g., feature independence) made it computationally efficient but less adaptive to complex relationships. Across sampling strategies, performance varied widely—particularly in recall, specificity, and ROC AUC—showing how sensitive NB is to class balance.

- **No Resampling** delivered **perfect recall (1.000)** and a strong **F1 score (0.838)**, identifying all positive cases. However, it **misclassified every negative case** (specificity = **0.000**) and showed a **moderate ROC AUC (0.307)**, indicating poor probability ranking.

- **SMOTE** achieved the **highest specificity (0.857)** and solid **precision (0.818)** but suffered from **low recall (0.248)** and **F1 (0.381)**. Nevertheless, it recorded the **best ROC AUC (0.361)**, suggesting better risk ranking compared to other methods.

- **Upsampling** struck a middle ground, achieving **recall (0.559)**, **F1 (0.672)**, **specificity (0.732)**, and **precision (0.844)**. Its ROC AUC was lower (**0.296**), suggesting limitations in risk stratification despite balanced performance.

- **Downsampling** showed the most **balanced results** with **recall (0.683)**, **F1 score (0.739)**, **specificity (0.571)**, and **precision (0.805)**. ROC AUC reached **0.321**, slightly behind SMOTE but still the second-best, offering a good tradeoff between classification and ranking.

Naive Bayes worked well for **baseline classification**, especially under **No Resampling** for sensitivity and **Downsampling** for balanced tradeoffs. Yet, its limitations in handling feature interactions and calibration suggest that **ensemble models may be better suited** for nuanced turnover predictions.

## Monitoring Turnover Risk with Support Vector Machines

### Support Vector Machines - No Sampling

```{r 04-svm-no-sample, message=FALSE}
# Linear Support Vector Machine
svm_model <- svm_linear(cost = tune()) %>%
  set_engine("kernlab") %>%
  set_mode("classification")

# Workflow
workflow_svm_none <- workflow() %>%
  add_model(svm_model) %>%
  add_recipe(base_recipe)

# Tuning
set.seed(42)

svm_tuned <- tune_grid(
  workflow_svm_none,
  resamples = mtr_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc))

# Finalize
best_svm <- select_best(svm_tuned, metric = "roc_auc")
final_svm_workflow <- finalize_workflow(workflow_svm_none, best_svm)

# Fit
svm_fit_none <- fit(final_svm_workflow, data = mtr_train_data)

# Predict
svm_preds_none <- predict(svm_fit_none, new_data = mtr_test_data, type = "prob") %>%
  bind_cols(predict(svm_fit_none, new_data = mtr_test_data)) %>%
  bind_cols(mtr_test_data %>% select(Monitoring_Turnover_Risk))
```

### Support Vector Machines - No Sampling - Performance Metrics

```{r 04-svm-no-sampling-metrics, echo=TRUE}
# Metrics
svm_none_metrics <- all_metrics(svm_preds_none, truth = Monitoring_Turnover_Risk, estimate = .pred_class, .pred_Yes)

# Print
svm_none_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric, levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                          labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC"))) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#884EA0", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)), hjust = -0.1, color = "black", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Performance Metrics – SVM (No Resampling)", x = "Score", y = NULL) +
  theme_minimal(base_size = 13)
```

```{r 04-svm-matrix-no-sampling, echo=TRUE}
# Confusion Matrix
cm_svm_none <- conf_mat(svm_preds_none, truth = Monitoring_Turnover_Risk, estimate = .pred_class)
cm_df_svm_none <- as.data.frame(cm_svm_none$table)
cm_df_svm_none$Truth <- factor(cm_df_svm_none$Truth, levels = c("Yes", "No"))
cm_df_svm_none$Prediction <- factor(cm_df_svm_none$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_svm_none, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white") +
  geom_text(aes(label = Freq), fontface = "bold", size = 5) +
  scale_fill_gradient(low = "#f2e6f9", high = "#5B2C6F") +
  scale_y_discrete(limits = rev) +
  labs(title = "Confusion Matrix – SVM (No Resampling)", x = "Predicted", y = "Actual") +
  theme_minimal()

```

```{r 04-svm-roc-curve-no-sampling, echo=TRUE}
# ROC Curve
roc_curve(svm_preds_none, truth = Monitoring_Turnover_Risk, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray50") +
  labs(
    title = "ROC Curve – SVM (No Resampling)",
    subtitle = "Model performance on predicting Monitoring Turnover Risk",
    x = "1 - Specificity",
    y = "Sensitivity"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )
```

### Support Vector Machines - Smote

```{r 04-svm-smote, message=FALSE}
# Define the SVM Model
svm_model <- svm_linear(cost = tune()) %>%
  set_engine("kernlab") %>%
  set_mode("classification")

# Create SMOTE Recipe
recipe_svm_smote <- base_recipe %>%
  step_smote(Monitoring_Turnover_Risk)

# Build Workflow
workflow_svm_smote <- workflow() %>%
  add_model(svm_model) %>%
  add_recipe(recipe_svm_smote)

# Tune with 10-Fold Cross-Validation
set.seed(42)
svm_tuned_smote <- tune_grid(
  workflow_svm_smote,
  resamples = mtr_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
)

# Select Best Model by ROC AUC
best_svm_smote <- select_best(svm_tuned_smote, metric = "roc_auc")

# Finalize Workflow with Best Hyperparameters
final_svm_workflow_smote <- finalize_workflow(workflow_svm_smote, best_svm_smote)

# Fit Final Model to Training Data
svm_fit_smote <- fit(final_svm_workflow_smote, data = mtr_train_data)

# Predict on Test Data
svm_preds_smote <- predict(svm_fit_smote, new_data = mtr_test_data, type = "prob") %>%
  bind_cols(predict(svm_fit_smote, new_data = mtr_test_data)) %>%
  bind_cols(mtr_test_data %>% select(Monitoring_Turnover_Risk))
```

### Support Vector Machines - Smote - Performance Metrics

```{r 04-svm-smote-metrics, echo=TRUE}
# Metrics
svm_smote_metrics <- all_metrics(svm_preds_smote, truth = Monitoring_Turnover_Risk, estimate = .pred_class, .pred_Yes)

# Print
svm_smote_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
    levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
    labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC")
  )) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#6C3483", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)), hjust = -0.1, color = "white", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Performance Metrics – SVM (SMOTE)", x = "Score", y = NULL) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-svm-matrix-smote, echo=TRUE}
# Confusion Matrix
cm_svm_smote <- conf_mat(svm_preds_smote, truth = Monitoring_Turnover_Risk, estimate = .pred_class)
cm_df_svm_smote <- as.data.frame(cm_svm_smote$table)

cm_df_svm_smote$Truth <- factor(cm_df_svm_smote$Truth, levels = c("Yes", "No"))
cm_df_svm_smote$Prediction <- factor(cm_df_svm_smote$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_svm_smote, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#E8DAEF", high = "#6C3483", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – SVM (SMOTE)",
    x = "Predicted Class", y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-svm-roc-curve-smote, echo=TRUE}
# ROC Curve
roc_curve(svm_preds_smote, truth = Monitoring_Turnover_Risk, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – SVM (SMOTE)",
    subtitle = "Model performance on predicting Monitoring Turnover Risk",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

### Support Vector Machines - Upsampling

```{r 04-svm-upsampling, message=FALSE}
# Define the SVM Model
svm_model <- svm_linear(cost = tune()) %>%
  set_engine("kernlab") %>%
  set_mode("classification")

# Create SMOTE Recipe
recipe_svm_upsample <- base_recipe %>%
  step_upsample(Monitoring_Turnover_Risk)

# Build Workflow
workflow_svm_upsample <- workflow() %>%
  add_model(svm_model) %>%
  add_recipe(recipe_svm_upsample)

# Tune with 10-Fold Cross-Validation
set.seed(42)
svm_tuned_upsample <- tune_grid(
  workflow_svm_upsample,
  resamples = mtr_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
)

# Select Best Model by ROC AUC
best_svm_upsample <- select_best(svm_tuned_upsample, metric = "roc_auc")

# Finalize Workflow with Best Hyperparameters
final_svm_workflow_upsample <- finalize_workflow(workflow_svm_upsample, best_svm_upsample)

# Fit Final Model to Training Data
svm_fit_upsample <- fit(final_svm_workflow_upsample, data = mtr_train_data)

# Predict on Test Data
svm_preds_upsample <- predict(svm_fit_upsample, new_data = mtr_test_data, type = "prob") %>%
  bind_cols(predict(svm_fit_upsample, new_data = mtr_test_data)) %>%
  bind_cols(mtr_test_data %>% select(Monitoring_Turnover_Risk))
```

### Support Vector Machines - Upsampling - Performance Metrics

```{r 04-svm-upsample-metrics, echo=TRUE}
# Metrics
svm_upsample_metrics <- all_metrics(svm_preds_upsample, truth = Monitoring_Turnover_Risk, estimate = .pred_class, .pred_Yes)

# Print
svm_upsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
    levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
    labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC")
  )) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#6C3483", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)), hjust = -0.1, color = "white", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Performance Metrics – SVM (Upsampling)", x = "Score", y = NULL) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-svm-matrix-upsample, echo=TRUE}
# Confusion Matrix
cm_svm_upsample <- conf_mat(svm_preds_upsample, truth = Monitoring_Turnover_Risk, estimate = .pred_class)
cm_df_svm_upsample <- as.data.frame(cm_svm_upsample$table)

cm_df_svm_upsample$Truth <- factor(cm_df_svm_upsample$Truth, levels = c("Yes", "No"))
cm_df_svm_upsample$Prediction <- factor(cm_df_svm_upsample$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_svm_upsample, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#E8DAEF", high = "#6C3483", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – SVM (Upsampling)",
    x = "Predicted Class", y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-svm-roc-curve-upsample, echo=TRUE}
# ROC Curve
roc_curve(svm_preds_upsample, truth = Monitoring_Turnover_Risk, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – SVM (Upsampling)",
    subtitle = "Model performance on predicting Monitoring Turnover Risk",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

### Support Vector Machines - Downsampling

```{r 04-svm-downsampling, message=FALSE}
# Define the SVM Model
svm_model <- svm_linear(cost = tune()) %>%
  set_engine("kernlab") %>%
  set_mode("classification")

# Create SMOTE Recipe
recipe_svm_downsample <- base_recipe %>%
  step_downsample(Monitoring_Turnover_Risk)

# Build Workflow
workflow_svm_downsample <- workflow() %>%
  add_model(svm_model) %>%
  add_recipe(recipe_svm_downsample)

# Tune with 10-Fold Cross-Validation
set.seed(42)
svm_tuned_downsample <- tune_grid(
  workflow_svm_downsample,
  resamples = mtr_folds,
  grid = 10,
  metrics = metric_set(accuracy, roc_auc)
)

# Select Best Model by ROC AUC
best_svm_downsample <- select_best(svm_tuned_downsample, metric = "roc_auc")

# Finalize Workflow with Best Hyperparameters
final_svm_workflow_downsample <- finalize_workflow(workflow_svm_downsample, best_svm_downsample)

# Fit Final Model to Training Data
svm_fit_downsample <- fit(final_svm_workflow_downsample, data = mtr_train_data)

# Predict on Test Data
svm_preds_downsample <- predict(svm_fit_downsample, new_data = mtr_test_data, type = "prob") %>%
  bind_cols(predict(svm_fit_downsample, new_data = mtr_test_data)) %>%
  bind_cols(mtr_test_data %>% select(Monitoring_Turnover_Risk))
```

### Support Vector Machines - Upsampling - Performance Metrics

```{r 04-svm-downsampling-metrics, echo=TRUE}
# Metrics
svm_downsample_metrics <- all_metrics(svm_preds_downsample, truth = Monitoring_Turnover_Risk, estimate = .pred_class, .pred_Yes)

# Print
svm_downsample_metrics %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc")) %>%
  mutate(.metric = factor(.metric,
    levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
    labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC")
  )) %>%
  ggplot(aes(x = .estimate, y = fct_reorder(.metric, .estimate))) +
  geom_col(fill = "#6C3483", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", .estimate)), hjust = -0.1, color = "white", size = 4.2, fontface = "bold") +
  scale_x_continuous(limits = c(0, 1), expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Performance Metrics – SVM (Downsampling)", x = "Score", y = NULL) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    axis.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank()
  )

```

```{r 04-svm-matrix-downsample, echo=TRUE}
# Confusion Matrix
cm_svm_downsample <- conf_mat(svm_preds_downsample, truth = Monitoring_Turnover_Risk, estimate = .pred_class)
cm_df_svm_downsample <- as.data.frame(cm_svm_downsample$table)

cm_df_svm_downsample$Truth <- factor(cm_df_svm_downsample$Truth, levels = c("Yes", "No"))
cm_df_svm_downsample$Prediction <- factor(cm_df_svm_downsample$Prediction, levels = c("No", "Yes"))

ggplot(cm_df_svm_downsample, aes(x = Prediction, y = Truth, fill = Freq)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = Freq), color = "black", size = 5, fontface = "bold") +
  scale_fill_gradient(low = "#E8DAEF", high = "#6C3483", name = "Frequency") +
  scale_y_discrete(limits = rev) +
  labs(
    title = "Confusion Matrix – SVM (Downsampling)",
    x = "Predicted Class", y = "Actual Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    axis.text = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

```

```{r 04-svm-roc-curve-downsample, echo=TRUE}
# ROC Curve
roc_curve(svm_preds_downsample, truth = Monitoring_Turnover_Risk, .pred_Yes) %>%
  autoplot() +
  geom_abline(linetype = "dashed", color = "gray60", size = 0.8) +
  labs(
    title = "ROC Curve – SVM (Downsampling)",
    subtitle = "Model performance on predicting Monitoring Turnover Risk",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(face = "bold")
  )

```

### Results Table for Support Vector Machines (Monitoring Turnover Risk)

```{r 04-svm-results-table, results='asis', echo=TRUE}
svm_results_long <- bind_rows(
  svm_none_metrics %>% mutate(Resample = "No Sampling"),
  svm_smote_metrics %>% mutate(Resample = "SMOTE"),
  svm_upsample_metrics %>% mutate(Resample = "Up Sampling"),
  svm_downsample_metrics %>% mutate(Resample = "Down Sampling")
) %>%
  mutate(
    Metric = factor(.metric,
                    levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
                    labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC")),
    Score = .estimate,
    Resample = factor(Resample, levels = c("No Sampling", "SMOTE", "Up Sampling", "Down Sampling"))
  ) %>%
  filter(!is.na(Score))

highlight_svm <- svm_results_long %>%
  group_by(Metric) %>%
  filter(Score == max(Score, na.rm = TRUE)) %>%
  mutate(is_best = TRUE)

svm_results_highlight <- left_join(svm_results_long, highlight_svm %>% select(Resample, Metric, is_best),
                                   by = c("Resample", "Metric")) %>%
  mutate(is_best = ifelse(is.na(is_best), FALSE, is_best))

ggplot(svm_results_highlight, aes(x = Resample, y = Score, fill = Metric)) +
  geom_col(width = 0.7) +
  geom_col(data = subset(svm_results_highlight, is_best),
           fill = NA, color = "black", linewidth = 1.2, width = 0.7) +
  facet_wrap(~ Metric, nrow = 2, scales = "free_y") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_manual(values = rep("#9B59B6", 6)) +
  labs(
    title = "SVM Performance by Metric and Resampling Method",
    subtitle = "Black outline shows the best result for each metric",
    x = "Resampling Strategy", y = "Score"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 15),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 30, hjust = 1),
    strip.text = element_text(face = "bold"),
    legend.position = "none"
  )
```

### Executive Summary for Support Vector Machines (Monitoring Turnover Risk)

Support Vector Machines (SVM) offered competitive performance, particularly in terms of **recall, F1 score, and precision**. While not the top-performing model overall, SVM proved reliable for **identifying at-risk employees**, especially when using the original (non-resampled) data. Its moderate ROC AUC and lower specificity indicate some room for improvement in **ranking probabilities** and **reducing false positives**.

- **No Resampling** yielded the **strongest performance overall**, with high **recall (0.883)**, **F1 score (0.845)**, and **precision (0.810)**. **Accuracy (0.766)** was the highest of all configurations. ROC AUC was modest at **0.274**, and specificity was low (**0.464**), showing strong positive-class prediction but limited true negative detection.

- **SMOTE** offered strong **precision (0.917)** and improved **specificity (0.750)**, but had slightly lower **recall (0.676)** and **F1 (0.699)**. ROC AUC was nearly identical to No Resampling (**0.274**), with better balance between classes.

- **Upsampling** maintained similar results to SMOTE, with **precision (0.876)** and **F1 score (0.723)**. Recall (**0.690**) was slightly better than SMOTE, while ROC AUC remained steady at **0.274**.

- **Downsampling** produced the **highest ROC AUC (0.302)** and strong **precision (0.905)**, but came with a drop in **recall (0.648)** and **F1 score (0.709)**. This method reduced false positives and improved class probability ranking, making it more conservative and better calibrated.

SVM maintained consistent ROC AUC across strategies, and benefits may arise from kernel tuning, threshold adjustments, or probability calibration.

## Monitoring Turnover Risk - Master Performance Comparison Table

```{r 04-model-accuracy-comparison, fig.width=14, fig.height=8}
# Define colors
model_colors <- c(
  "Logistic" = "#4B9CD3",
  "Random Forest" = "#3FA34D",
  "XGBoost" = "#E87461",
  "Naive Bayes" = "#FFD700",
  "SVM" = "#9370DB"
)

# Identify best performers per model and metric
best_scores <- all_models_metrics %>%
  group_by(Model, .metric) %>%
  filter(.estimate == max(.estimate, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(is_best = TRUE)

# Merge and annotate plot data
plot_data <- all_models_metrics %>%
  left_join(
    best_scores %>% select(Model, Resample, .metric, is_best),
    by = c("Model", "Resample", ".metric")
  ) %>%
  mutate(
    is_best = ifelse(is.na(is_best), FALSE, is_best),
    label = ifelse(is.na(.estimate), "NA", sprintf("%.3f", .estimate)),
    Resample = factor(
      Resample,
      levels = c("No Sampling", "SMOTE", "Up Sampling", "Down Sampling", "Not Available")
    ),
    Model = factor(
      Model,
      levels = c("Logistic", "Random Forest", "XGBoost", "Naive Bayes", "SVM")
    ),
    .metric = factor(
      .metric,
      levels = c("accuracy", "precision", "recall", "f_meas", "spec", "roc_auc"),
      labels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC")
    )
  )

# Create the plot
ggplot(plot_data, aes(x = Resample, y = .estimate, fill = Model)) +
  geom_col(width = 0.7, color = NA) +
  geom_col(
    data = subset(plot_data, is_best),
    aes(x = Resample, y = .estimate),
    fill = NA,
    color = "black",
    linewidth = 1.1,
    width = 0.7
  ) +
  geom_text(
    aes(label = label),
    vjust = -0.35,
    size = 3,
    color = "black"
  ) +
  facet_grid(.metric ~ Model, switch = "y") +
  scale_y_continuous(
    limits = c(0, 1),
    expand = expansion(mult = c(0.05, 0.25))
  ) +
  scale_fill_manual(values = model_colors, na.translate = FALSE) +
  labs(
    title = "Monitoring Turnover Risk – Final Model Performance by Metric and Resampling Strategy",
    subtitle = "Best scores outlined in black. NA indicates model/strategy not available.",
    x = "Resampling Method",
    y = "Score",
    fill = "Model"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    strip.text = element_text(face = "bold", size = 11),
    axis.text.x = element_text(angle = 30, hjust = 1, size = 9),
    axis.text.y = element_text(size = 9),
    panel.spacing = unit(1.5, "lines")
  )

```

## Monitoring Turnover Risk - Master Variable Importance Plot

```{r 04-vip-imp-clean-horizontal, fig.width=14, fig.height=10, message=FALSE, warning=FALSE}
# Helper function to extract VI from a workflow
get_importance <- function(fit_obj, model_label, resample_label) {
  vi(extract_fit_parsnip(fit_obj), scale = TRUE) %>%
    mutate(Model = model_label, Resample = resample_label)
}

# Extract VIPs for each model and resample method
model_importance_all <- bind_rows(
  get_importance(fit_logit_none, "Logistic", "No Sampling"),
  get_importance(fit_logit_smote, "Logistic", "SMOTE"),
  get_importance(fit_logit_upsample, "Logistic", "Upsampling"),
  get_importance(fit_logit_downsample, "Logistic", "Downsampling"),
  
  get_importance(rf_fit_none, "Random Forest", "No Sampling"),
  get_importance(rf_fit_smote, "Random Forest", "SMOTE"),
  get_importance(rf_fit_upsample, "Random Forest", "Upsampling"),
  get_importance(rf_fit_downsample, "Random Forest", "Downsampling"),
  
  get_importance(xgb_fit_none, "XGBoost", "No Sampling"),
  get_importance(xgb_fit_smote, "XGBoost", "SMOTE"),
  get_importance(xgb_fit_upsample, "XGBoost", "Upsampling"),
  get_importance(xgb_fit_downsample, "XGBoost", "Downsampling")
)

# Top 5 VIPs per Model + Resampling combo, scaled
top_vars <- model_importance_all %>%
  filter(!is.na(Importance)) %>%
  mutate(
    Resample = factor(Resample, levels = c("No Sampling", "SMOTE", "Upsampling", "Downsampling")),
    Model = factor(Model, levels = c("Logistic", "Random Forest", "XGBoost"))
  ) %>%
  group_by(Model, Resample) %>%
  arrange(desc(Importance)) %>%
  slice_head(n = 5) %>%
  ungroup() %>%
  group_by(Model) %>%
  mutate(ScaledImportance = Importance / max(Importance)) %>%
  ungroup()

# Custom model colors
model_colors <- c(
  "Logistic" = "#4B9CD3",
  "Random Forest" = "#3FA34D",
  "XGBoost" = "#E87461"
)

# Plot VIPs
ggplot(top_vars, aes(
  y = ScaledImportance,
  x = fct_reorder(Variable, ScaledImportance),
  fill = Model
)) +
  geom_col(width = 0.75, show.legend = FALSE) +
  geom_text(
    aes(label = round(Importance, 2)),
    vjust = -0.3,
    size = 3.2,
    color = "black"
  ) +
  facet_grid(Resample ~ Model, scales = "free_x") +
  scale_fill_manual(values = model_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top Five Most Important Features by Model and Resampling Method",
    subtitle = "Importance scaled within each model to allow comparison",
    y = "Scaled Importance (Within Model)",
    x = "Variable"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    strip.text = element_text(face = "bold", size = 11),
    axis.text.x = element_text(size = 9, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 10),
    panel.grid.major.y = element_blank(),
    panel.spacing = unit(1, "lines")
  )
```

## Monitoring Turnover Risk - Top Five Performing Models Plot

```{r 04-best-model-monitoring, fig.width=14, fig.height=10, message=FALSE, warning=FALSE}
# Define top 5 model scores – Monitoring Turnover Risk (ordered by F1 Score)
top5_model_scores <- tibble::tribble(
  ~Model, ~Metric, ~Score,

  # Random Forest (No Sampling)
  "Random Forest (No Sampling)", "Accuracy", 0.886,
  "Random Forest (No Sampling)", "Precision", 0.863,
  "Random Forest (No Sampling)", "Recall", 1.000,
  "Random Forest (No Sampling)", "F1 Score", 0.927,
  "Random Forest (No Sampling)", "Specificity", 0.589,
  "Random Forest (No Sampling)", "ROC AUC", 0.133,

  # XGBoost (No Sampling)
  "XGBoost (No Sampling)", "Accuracy", 0.886,
  "XGBoost (No Sampling)", "Precision", 0.863,
  "XGBoost (No Sampling)", "Recall", 1.000,
  "XGBoost (No Sampling)", "F1 Score", 0.927,
  "XGBoost (No Sampling)", "Specificity", 0.589,
  "XGBoost (No Sampling)", "ROC AUC", 0.197,

  # SVM (No Sampling)
  "SVM (No Sampling)", "Accuracy", 0.766,
  "SVM (No Sampling)", "Precision", 0.810,
  "SVM (No Sampling)", "Recall", 0.883,
  "SVM (No Sampling)", "F1 Score", 0.845,
  "SVM (No Sampling)", "Specificity", 0.464,
  "SVM (No Sampling)", "ROC AUC", 0.274,

  # Logistic (No Sampling)
  "Logistic (No Sampling)", "Accuracy", 0.736,
  "Logistic (No Sampling)", "Precision", 0.813,
  "Logistic (No Sampling)", "Recall", 0.828,
  "Logistic (No Sampling)", "F1 Score", 0.819,
  "Logistic (No Sampling)", "Specificity", 0.500,
  "Logistic (No Sampling)", "ROC AUC", 0.283,

  # Naive Bayes (Down Sampling)
  "Naive Bayes (Down Sampling)", "Accuracy", 0.737,
  "Naive Bayes (Down Sampling)", "Precision", 0.805,
  "Naive Bayes (Down Sampling)", "Recall", 0.683,
  "Naive Bayes (Down Sampling)", "F1 Score", 0.739,
  "Naive Bayes (Down Sampling)", "Specificity", 0.768,
  "Naive Bayes (Down Sampling)", "ROC AUC", 0.321
)

# Set metric order
top5_model_scores <- top5_model_scores %>%
  mutate(Metric = factor(
    Metric,
    levels = c("Accuracy", "Precision", "Recall", "F1 Score", "Specificity", "ROC AUC")
  ))

# Define model colors
model_colors <- c(
  "Random Forest (No Sampling)" = "#3FA34D",
  "XGBoost (No Sampling)" = "#E87461",
  "SVM (No Sampling)" = "#9370DB",
  "Logistic (No Sampling)" = "#4B9CD3",
  "Naive Bayes (Down Sampling)" = "#F4D03F"
)

# Create the plot
ggplot(top5_model_scores, aes(x = Metric, y = Score, fill = Model)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(
    aes(label = sprintf("%.3f", Score)),
    position = position_dodge(width = 0.8),
    vjust = -0.5,
    size = 3.5,
    color = "black"
  ) +
  scale_fill_manual(values = model_colors) +
  scale_y_continuous(limits = c(0, 1.1)) +
  labs(
    title = "Top 5 Model Performers – Monitoring Turnover Risk",
    subtitle = "Each model's best resampling strategy ranked by F1 Score",
    x = "Metric",
    y = "Score",
    fill = "Model"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5),
    axis.text.x = element_text(size = 11),
    axis.text.y = element_text(size = 11),
    legend.position = "top",
    legend.title = element_text(face = "bold")
  )
```

## Monitoring Turnover Risk – Top Five Best Performing Models

To evaluate turnover risk, models were assessed across six key classification metrics: **accuracy**, **precision**, **recall**, **F1 score**, **specificity**, and **ROC AUC**. Among the top-performing configurations, two models — **Random Forest** and **XGBoost (both with no sampling)** — delivered near-identical classification strength, with XGBoost edging ahead overall due to its superior ranking ability (ROC AUC).

---

**1. XGBoost (No Sampling)** – Best Overall for Classification and Ranking

- **Accuracy**: 0.886  
- **Precision**: 0.863  
- **Recall**: 1.000  
- **F1 Score**: 0.927  
- **Specificity**: 0.589  
- **ROC AUC**: 0.197  

XGBoost achieved **perfect recall** and **top F1 score**, matching Random Forest across all classification metrics. Its higher **ROC AUC** makes it the best option for **risk ranking**, making it ideal when both classification and **probability-based prioritization** are important.

---

**2. Random Forest (No Sampling)** – Best Pure Classifier

- **Accuracy**: 0.886  
- **Precision**: 0.863  
- **Recall**: 1.000  
- **F1 Score**: 0.927  
- **Specificity**: 0.589  
- **ROC AUC**: 0.133  

Random Forest excelled as a **classification model**, equaling XGBoost on all predictive metrics. While its **ROC AUC** was lower, it remains a top-tier model for **high-recall use cases** where missing a true at-risk employee is unacceptable.

---

**3. SVM (No Sampling)** – Balanced and Precise

- **Accuracy**: 0.766  
- **Precision**: 0.810  
- **Recall**: 0.883  
- **F1 Score**: 0.845  
- **Specificity**: 0.464  
- **ROC AUC**: 0.274  

SVM offered a strong **balance between recall and precision**, with solid classification strength across the board. It’s well-suited for **proactive retention strategies** that seek to reduce both false positives and false negatives.

---

**4. Logistic Regression (No Sampling)** – Interpretable and Reliable

- **Accuracy**: 0.736  
- **Precision**: 0.813  
- **Recall**: 0.828  
- **F1 Score**: 0.819  
- **Specificity**: 0.500  
- **ROC AUC**: 0.283  

Logistic Regression provides a **transparent, interpretable option** with consistent metrics and decent ROC AUC. It's an excellent choice for organizations that value **clarity in decision-making** and stakeholder trust.

---

**5. Naive Bayes (Down Sampling)** – Most Calibrated for Risk Ranking

- **Accuracy**: 0.737  
- **Precision**: 0.805  
- **Recall**: 0.683  
- **F1 Score**: 0.739  
- **Specificity**: 0.768  
- **ROC AUC**: 0.321  

Naive Bayes stood out for its **highest ROC AUC**, offering **stronger probability calibration** despite lower recall. It’s a great option for **risk scoring and ranking**, especially in scenarios where prioritization is key.


<div style="display: flex; justify-content: space-between; margin-top: 20px; margin-bottom: 30px;">
  <a href="HR_Data_Quiet_Quitting_Results.html" style="background-color: #4B9CD3; color: white; padding: 10px 20px; 
     text-align: center; text-decoration: none; font-size: 16px; border-radius: 8px;">
    🔗 Back to Quiet Quitting Results
  </a>

  <a href="HR_Data_EDA.html" style="background-color: #E74C3C; color: white; padding: 10px 20px; 
     text-align: center; text-decoration: none; font-size: 16px; border-radius: 8px;">
    🔗 Back to HR EDA
  </a>
</div>

```{r 00-image-save, include=FALSE}
# Save the current R session to restore later
save.image(file = "outputs/my_environment.RData")
```