---
title: "Workforce Analytics Using Synthetic HR Data: Exploratory Data Analysis"
author: "Kevin Allen Cauley"
date: "`r format(Sys.Date(), '%B %d, %Y')`"

output:
  html_document:
    always_allow_html: true
    self_contained: false
    theme: flatly
    highlight: null
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    css: "assets/styles.css"
    output_file: HR_Data_EDA.html
    output_dir: ../docs

number_sections: true
---

```{r 00-setup, include=FALSE}
# Libraries
library(tidyverse)
library(reshape2)
library(readxl)
library(here)
library(janitor)
library(psych)
library(GGally)
library(car)
library(e1071)
library(fastDummies)
library(scales)
library(patchwork)
library(stringr)
library(knitr)
library(kableExtra)
library(corrplot)
library(ggcorrplot)
library(naniar)

# Function Helpers
source(here::here("analysis", "helper.R"))

# Reproducibility
set.seed(123)

# Root Directory
library(rprojroot)
knitr::opts_knit$set(
  root.dir = rprojroot::find_rstudio_root_file()
)

# Output Directories
plot_dir <- here::here("docs", "figures", "eda")
dir.create(plot_dir, recursive = TRUE, showWarnings = FALSE)

# Global Chunk Options
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  out.width = "100%",
  fig.width = 10,
  fig.height = 6,
  dpi = 300,
  cache = FALSE,
  fig.path = "figures/eda/"
)

# Image Display Helper
show_plot <- function(filename) {
  knitr::include_graphics(
    file.path("figures", "eda", filename)
  )
}

# Color System
hr_palette <- list(
  primary = "#5FA8D3",
  secondary = "#F4A261",
  danger = "#D7263D",
  success = "#2A9D8F",
  neutral = "#6C757D"
)
```

```{r 00-install-packages, include=FALSE}
required_packages <- c(
  "tidyverse",
  "reshape2",
  "readxl",
  "here",
  "broom",
  "DT",
  "knitr",
  "kableExtra",
  "janitor",
  "skimr",
  "corrplot",
  "ggcorrplot",
  "GGally",
  "psych",
  "randomcoloR",
  "RColorBrewer",
  "ggthemes",
  "patchwork",
  "cowplot",
  "gridExtra",
  "htmltools",
  "car",
  "e1071",
  "fastDummies",
  "naniar",
  "scales"
)

missing_packages <- required_packages[
  !required_packages %in% installed.packages()[, "Package"]
]

if (length(missing_packages) > 0) {
  stop("Missing packages: ", paste(missing_packages, collapse = ", "))
}
```

```{r 00-import-data, include=FALSE}
file_path <- here("data", "raw", "HR_Data.xlsx")

data_demographics <- read_excel(file_path, sheet = "Data & Demographics")
data_metrics <- read_excel(file_path, sheet = "Metrics")
data_satisfaction <- read_excel(file_path, sheet = "Satisfaction")
data_wellness <- read_excel(file_path, sheet = "Wellness")
data_monitoring <- read_excel(file_path, sheet = "Monitoring")
data_privacy <- read_excel(file_path, sheet = "Privacy")
data_attrition <- read_excel(file_path, sheet = "Attrition")
```

```{r 00-merge-datasets, include=FALSE}
hr_df <- data_demographics %>%
  left_join(data_metrics, by = "Employee_ID") %>%
  left_join(data_satisfaction, by = "Employee_ID") %>%
  left_join(data_wellness, by = "Employee_ID") %>%
  left_join(data_monitoring, by = "Employee_ID") %>%
  left_join(data_privacy, by = "Employee_ID") %>%
  left_join(data_attrition, by = "Employee_ID") %>%
  select(
    -Employee_ID,
    -First_Name,
    -Last_Name,
    -Voluntary_Resignation,
    -Involuntary_Termination
  ) %>%
  janitor::clean_names() %>%
  mutate(across(where(is.character), as.factor))

data <- hr_df

dir.create(here("data", "processed"), recursive = TRUE, showWarnings = FALSE)
saveRDS(hr_df, here("data", "processed", "hr_dataset_clean.rds"))
```

```{r 00-base-dataset, include=FALSE}
# Kills dependency on hr_df and creates self-contained chunks
get_hr_df <- function() {
  hr_df %>%
    mutate(across(where(is.character), as.factor))
}

# Safe Numeric Selector
safe_numeric <- function(df) {
  df %>%
    select(where(is.numeric)) %>%
    select(where(~ var(.x, na.rm = TRUE) > 0)) %>%
    drop_na()
}

# Safe Factor Dummy Builder
safe_dummies <- function(df) {
  df %>%
    select(where(is.factor)) %>%
    fastDummies::dummy_cols(
      remove_first_dummy = TRUE,
      remove_selected_columns = TRUE
    ) %>%
    select(where(~ var(.x, na.rm = TRUE) > 0)) %>%
    drop_na()
}
```

```{r 00-metadata-tables, fig.show="hold", include=FALSE}
employee_data_demo_meta <- tribble(
  ~Variable, ~Type, ~Measurement, ~Purpose,
  "Employee_ID", "Character", "Nominal", "Identifier/Drop",
  "First_Name", "Character", "Nominal", "Identifier/Drop",
  "Last_Name", "Character", "Nominal", "Identifier/Drop",
  "Occupation", "Categorical", "Nominal", "Predictor",
  "Department", "Categorical", "Nominal", "Predictor",
  "Average_Salary", "Numeric", "Ratio", "Predictor",
  "Tenure_Years", "Numeric", "Ratio", "Predictor",
  "Seniority", "Categorical", "Ordinal", "Predictor",
  "Work_Arrangement", "Categorical", "Nominal", "Predictor",
  "Distance_From_Home", "Numeric", "Ratio", "Predictor",
  "Age_Group", "Categorical", "Ordinal", "Predictor",
  "Gender", "Categorical", "Nominal", "Predictor",
  "Race", "Categorical", "Nominal", "Predictor",
  "Ethnicity", "Categorical", "Nominal", "Predictor")

employee_metrics_meta <- tribble(
  ~Variable, ~Type, ~Measurement, ~Purpose,
  "Actual_Training_Hours", "Numeric", "Ratio", "Predictor",
  "Overtime_Hours", "Numeric", "Ratio", "Predictor",
  "Performance_Score", "Numeric", "Ordinal", "Predictor",
  "Productivity_Score", "Numeric", "Ordinal", "Predictor",
  "Discretionary_Effort", "Numeric", "Ordinal", "Predictor",
  "Engagement_Level", "Numeric", "Ordinal", "Predictor",
  "Organizational_Commitment", "Numeric", "Ordinal", "Predictor",
  "Role_Clarity", "Numeric", "Ordinal", "Predictor")

employee_satisfaction_meta <- tribble(
  ~Variable, ~Type, ~Measurement, ~Purpose,
  "Job_Satisfaction", "Numeric", "Ordinal", "Predictor",
  "Workplace_Environment_Satisfaction", "Numeric", "Ordinal", "Predictor",
  "Work_Life_Balance_Satisfaction", "Numeric", "Ordinal", "Predictor",
  "Perceived_Organizational_Support", "Numeric", "Ordinal", "Predictor",
  "Relationship_With_Manager", "Numeric", "Ordinal", "Predictor",
  "Relationship_With_Coworkers", "Numeric", "Ordinal", "Predictor",
  "Team_Collaboration", "Numeric", "Ordinal", "Predictor",
  "Perceived_Career_Growth_Opportunities", "Numeric", "Ordinal", "Predictor",
  "Promotion_Opportunity", "Numeric", "Ordinal", "Predictor",
  "Learning_Opportunities", "Numeric", "Ordinal", "Predictor",
  "Manager_Support_Score", "Numeric", "Ordinal", "Predictor",
  "Perceived_Fairness_Score", "Numeric", "Ordinal", "Predictor",
  "Trust_In_Management", "Numeric", "Ordinal", "Predictor",
  "Job_Involvement", "Numeric", "Ordinal", "Predictor",
  "Recognition_Score", "Numeric", "Ordinal", "Predictor",
  "Voice_In_Decision_Making", "Numeric", "Ordinal", "Predictor",
  "Alignment_With_Mission", "Numeric", "Ordinal", "Predictor",
  "Manager_Communication_Quality", "Numeric", "Ordinal", "Predictor",
  "Feeling_Valued_At_Work", "Numeric", "Ordinal", "Predictor")

employee_wellness_meta <- tribble(
  ~Variable, ~Type, ~Measurement, ~Purpose,
  "Burnout_Score", "Numeric", "Ordinal", "Future Target",
  "Wellbeing_Score", "Numeric", "Ordinal", "Future Target",
  "Anxiety_Level", "Numeric", "Ordinal", "Predictor",
  "Depression_Level", "Numeric", "Ordinal", "Predictor",
  "Stress_Level", "Numeric", "Ordinal", "Predictor",
  "Mental_Health_Days_Taken", "Numeric", "Ratio", "Predictor",
  "Absenteeism_Due_To_Mental_Health", "Numeric", "Ratio", "Predictor",
  "Access_To_Counseling", "Categorical", "Nominal", "Predictor",
  "Workload_Pressure", "Numeric", "Ordinal", "Predictor",
  "Coping_Skills", "Numeric", "Ordinal", "Predictor",
  "Peer_Support", "Numeric", "Ordinal", "Predictor",
  "Manager_Empathy", "Numeric", "Ordinal", "Predictor",
  "Leadership_Style", "Categorical", "Nominal", "Predictor",
  "Stressors_Experienced", "Numeric", "Ratio", "Predictor")

employee_monitoring_meta <- tribble(
  ~Variable, ~Type, ~Measurement, ~Purpose,
  "Time_Tracking_Method_Used", "Categorical", "Nominal", "Predictor",
  "Surveillance_Method_Used", "Categorical", "Nominal", "Predictor",
  "Total_Monitoring_Tools", "Numeric", "Ratio", "Predictor",
  "Monitoring_Frequency", "Numeric", "Ordinal", "Predictor",
  "Monitoring_Intrusiveness", "Numeric", "Ordinal", "Predictor",
  "Monitoring_Productivity_Impact", "Numeric", "Ordinal", "Predictor",
  "Monitoring_Satisfaction_Impact", "Numeric", "Ordinal", "Predictor",
  "Violations_Logged_By_Monitoring_Software", "Numeric", "Ratio", "Predictor",
  "Perceived_Monitoring_Purpose", "Categorical", "Nominal", "Predictor",
  "AI_Supervision", "Categorical", "Nominal", "Predictor")

employee_privacy_meta <- tribble(
  ~Variable, ~Type, ~Measurement, ~Purpose,
  "Monitoring_Transparency", "Numeric", "Ordinal", "Predictor",
  "Privacy_Concern_Score", "Numeric", "Ordinal", "Target",
  "Perceived_Intrusiveness", "Numeric", "Ordinal", "Predictor",
  "Trust_In_Monitoring", "Numeric", "Ordinal", "Predictor",
  "Privacy_Complaints_Score", "Numeric", "Ratio", "Target")

employee_attrition_meta <- tribble(
  ~Variable, ~Type, ~Measurement, ~Purpose,
  "Turnover_Intention", "Categorical", "Nominal", "Target",
  "Voluntary_Resignation", "Categorical", "Nominal", "Drop",
  "Involuntary_Termination", "Categorical", "Nominal", "Drop",
  "Quiet_Quitting", "Categorical", "Nominal", "Target",
  "Employer_Promise_Broken", "Numeric", "Ordinal", "Predictor",
  "Internal_Job_Applications", "Numeric", "Ratio", "Predictor",
  "Promotion_Last_12_Months", "Categorical", "Nominal", "Predictor",
  "Monitoring_Turnover_Risk", "Categorical", "Nominal", "Target")
```

```{r 01-project-overview, echo=FALSE}
project_overview <- tibble(
  Attribute = c(
    "Project Type",
    "Dataset Source",
    "Observations",
    "Variables",
    "Primary Tools",
    "Analysis Scope"
  ),
  Description = c(
    "Workforce Analytics / HR Predictive Modeling",
    "Research-informed synthetic HR dataset",
    paste0(nrow(hr_df), " employees"),
    paste0(ncol(hr_df), " workforce variables"),
    "R, tidyverse, ggplot2, fastDummies",
    "Exploratory analysis, correlation analysis, multicollinearity diagnostics, and predictive modeling preparation"
  )
)

knitr::kable(
  project_overview,
  format = "html",
  caption = "Project Overview"
) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = TRUE,
    position = "center",
    font_size = 13
  )
```

```{=html}
<div class="main-container">

  <div class="hero-section">
    <h1>Exploratory Data Analysis</h1>
    <p class="hero-subtitle">
      This section explores workforce patterns, distributions, and relationships 
      to establish a foundational understanding of employee behavior and key 
      organizational variables.
    </p>
  </div>

  <div style="
    display:grid;
    grid-template-columns: 1fr auto 1fr;
    align-items:center;
    margin:24px 0 24px 0;
  ">

    <!-- LEFT -->
    <div style="text-align:left;">
      <a href="index.html" style="
        display:inline-block;
        padding:12px 28px;
        border-radius:999px;
        font-weight:600;
        color:white;
        text-decoration:none;
        background:linear-gradient(135deg, #74B9FF, #2E86C1);
      ">
        ← Back Home
      </a>
    </div>

    <!-- RIGHT -->
    <div style="text-align:right;">
      <a href="HR_Data_Turnover_Intention_Results.html" style="
        display:inline-block;
        padding:12px 28px;
        border-radius:999px;
        font-weight:600;
        color:white;
        text-decoration:none;
        background:linear-gradient(135deg, #FF7675, #922B21);
      ">
        Next →
      </a>
    </div>

  </div>
```

# Introduction

Organizations increasingly rely on workforce data to drive decisions related to retention, performance, and employee well-being. However, access to real-world HR datasets is often restricted due to privacy and compliance constraints.

This project demonstrates a **full workforce analytics pipeline** using a research-informed synthetic dataset designed to replicate real organizational dynamics.

The objective is not simply exploratory analysis — it is to show how data can move from **raw structure → validated insights → predictive readiness**.

This analysis focuses on three critical workforce outcomes:

- Employee turnover intention  
- Quiet quitting behavior  
- Monitoring-driven attrition risk  

By combining labor statistics, organizational research, and behavioral indicators, this project illustrates how modern analytics workflows can support **evidence-based HR decision-making at scale**.

# Why This Analysis Matters

Workforce analytics is no longer optional — it is a competitive advantage.

Organizations that fail to understand employee behavior risk:
- Increased turnover costs  
- Decreased productivity  
- Lower employee engagement  
- Reduced organizational trust  

This project demonstrates how structured data analysis can uncover:

- Hidden drivers of employee disengagement  
- Relationships between monitoring practices and trust  
- Early indicators of burnout and attrition risk  

The goal is to bridge the gap between **data exploration and real-world organizational impact**.

# Exploratory Analysis Objectives

The goal of the exploratory data analysis (EDA) phase is to better understand the structure, distribution, and relationships within the synthetic HR dataset before developing predictive models.

The analysis focuses on four key objectives:

-   **Understand variable distributions** across continuous, ordinal, and categorical employee attributes.

-   **Identify potential data quality issues**, including missing values and extreme outliers.

-   **Evaluate relationships among predictors**, particularly correlations and multicollinearity that could affect model performance.

-   **Assess class imbalance and skewness**, which may require resampling or robust modeling approaches during predictive analysis.

These steps ensure that the dataset is properly validated before model development begins and help guide feature selection and modeling strategies.

# Data Overview

The dataset used in this analysis represents a synthetic workforce dataset informed by real-world labor statistics, organizational research literature, and survey-based employee engagement measures. The dataset was designed to simulate realistic relationships between employee characteristics, workplace experiences, monitoring practices, and organizational outcomes such as turnover intention.

These variables collectively support predictive modeling of employee attitudes, behavior, and attrition risk.

## Data Dictionary

The following metadata tables document the structure of the synthetic HR dataset.

Each variable is described according to:

-   Variable name
-   Data type
-   Measurement level
-   Modeling role (predictor, target, or excluded variable)

Documenting these attributes improves transparency and ensures reproducibility in later modeling stages.

```{r 02-data-dictionary, fig.show='hold', results='asis'}
make_pretty_html_table <- function(data, title) {
  knitr::kable(
    data,
    format = "html",
    caption = paste0(
      "<strong style='color:#145A32;font-size:16px;'>",
      title,
      "</strong>"
    ),
    align = c("l", "c", "c", "c")
  ) %>%
    kableExtra::kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = TRUE,
      position = "center",
      font_size = 14
    ) %>%
    kableExtra::row_spec(0, bold = TRUE, color = "white", background = "#145A32")
}

make_pretty_html_table(employee_data_demo_meta,    "1. Data & Demographics")
make_pretty_html_table(employee_metrics_meta,      "2. Metrics")
make_pretty_html_table(employee_satisfaction_meta, "3. Satisfaction")
make_pretty_html_table(employee_wellness_meta,     "4. Wellness")
make_pretty_html_table(employee_monitoring_meta,   "5. Monitoring")
make_pretty_html_table(employee_privacy_meta,      "6. Privacy")
make_pretty_html_table(employee_attrition_meta,    "7. Attrition")
```

### Workforce Dataset Overview

This section provides a high-level summary of the workforce dataset used in the analysis. 
Understanding the scale and structure of the data is essential before exploring deeper relationships. 
The dataset was designed to reflect realistic organizational conditions, including a diverse set of employee attributes and behavioral indicators. 
These characteristics ensure the dataset is suitable for both exploratory analysis and predictive modeling.

```{r 02-dataset-preview, echo=TRUE}
DT::datatable(
  head(hr_df, 100),
  rownames = FALSE,
  class = "compact stripe hover",
  extensions = c("Scroller"),
  options = list(
    deferRender = TRUE,
    scrollX = TRUE,
    pageLength = 12,
    autoWidth = TRUE,
    dom = "tip",
    columnDefs = list(list(className = "dt-center", targets = "_all"))
  )
)
```

### Dataset Structure

Before performing statistical analysis, it is important to verify the structure and data types of each variable. 
This step ensures that variables are correctly formatted for modeling and helps identify potential data quality issues such as missing values. 
A clear understanding of variable types also informs appropriate visualization and transformation strategies.

```{r 02-dataset-structure, echo=TRUE}
structure_df <- tibble(
  Variable = names(hr_df),
  Class = purrr::map_chr(hr_df, ~ class(.x)[1]),
  Missing_Values = purrr::map_int(hr_df, ~ sum(is.na(.x)))
) %>%
  mutate(Variable = clean_label(Variable))

knitr::kable(
  structure_df,
  format = "html",
  caption = "Dataset Structure"
) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = TRUE,
    position = "left"
  ) %>%
  kableExtra::scroll_box(width = "100%", height = "380px")
```

### Dataset Summary Statistics

Summary statistics provide insight into the central tendency and variability of numeric variables within the dataset. 
These metrics help identify potential outliers, skewness, and unusual distributions that may influence modeling performance. 
Understanding these characteristics is critical for selecting appropriate modeling techniques and transformations.

```{r 02-numeric-summary, echo=TRUE}
numeric_summary <- hr_df %>%
  select(where(is.numeric)) %>%
  psych::describe() %>%
  as.data.frame() %>%
  rownames_to_column("Variable") %>%
  mutate(Variable = clean_label(Variable))

knitr::kable(
  numeric_summary,
  format = "html",
  caption = "Numeric Variable Summary Statistics"
) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = TRUE,
    position = "center",
    font_size = 13
  ) %>%
  kableExtra::scroll_box(width = "100%", height = "380px")
```

# Data Engineering & Preparation

## Variable Definitions

We define groups of numeric predictors, categorical predictors, and target variables to support later exploratory analysis and modeling.

```{r 03-define-variable-groups, echo=TRUE}
num_vars <- c(
  "average_salary", "tenure_years", "distance_from_home", "actual_training_hours",
  "overtime_hours", "mental_health_days_taken", "absenteeism_due_to_mental_health",
  "stressors_experienced", "violations_logged_by_monitoring_software",
  "privacy_complaints_score", "burnout_score", "wellbeing_score", "anxiety_level",
  "depression_level", "stress_level", "performance_score", "productivity_score",
  "discretionary_effort", "engagement_level", "organizational_commitment",
  "role_clarity", "job_satisfaction", "workplace_environment_satisfaction",
  "work_life_balance_satisfaction", "perceived_organizational_support",
  "relationship_with_manager", "relationship_with_coworkers", "team_collaboration",
  "perceived_career_growth_opportunities", "promotion_opportunity",
  "learning_opportunities", "manager_support_score", "perceived_fairness_score",
  "trust_in_management", "job_involvement", "recognition_score",
  "voice_in_decision_making", "alignment_with_mission",
  "manager_communication_quality", "feeling_valued_at_work",
  "monitoring_frequency", "monitoring_intrusiveness",
  "monitoring_productivity_impact", "monitoring_satisfaction_impact",
  "monitoring_transparency", "privacy_concern_score",
  "trust_in_monitoring", "workload_pressure", "coping_skills",
  "peer_support", "manager_empathy"
)

cat_vars <- c(
  "age_group", "gender", "race", "ethnicity", "occupation", "department", "seniority",
  "work_arrangement", "time_tracking_method_used", "surveillance_method_used",
  "access_to_counseling", "leadership_style", "perceived_monitoring_purpose",
  "ai_supervision", "promotion_last_12_months"
)

cat_target_vars <- c(
  "turnover_intention", "quiet_quitting", "monitoring_turnover_risk"
)

num_target_vars <- c(
  "burnout_score", "wellbeing_score", "privacy_concern_score", "privacy_complaints_score"
)

expected_vars <- unique(c(num_vars, cat_vars, cat_target_vars, num_target_vars))
missing_vars <- setdiff(expected_vars, names(hr_df))

if (length(missing_vars) > 0) {
  stop("Missing variables: ", paste(missing_vars, collapse = ", "))
}
```

## Variable Validation

This validation step confirms that all expected predictor variables exist in the merged dataset before downstream analysis begins. Detecting missing variables early prevents silent modeling errors later in the pipeline.

## Data Preprocessing

Prior to exploratory analysis, character variables are converted to factors to ensure appropriate handling during statistical modeling and visualization.

```{r 03-factor-conversion, echo=FALSE}
hr_df <- hr_df %>%
  mutate(across(where(is.character), as.factor))
```

#  Exploratory Workforce Analysis

This section explores trends, patterns, distributions, and possible anomalies across both numeric and categorical variables. The goal is to better understand the synthetic dataset before predictive modeling.

## Variable Type Classification

Before plotting, numeric variables are divided into likely continuous variables and Likert-style variables based on the number of distinct values.

```{r 04-variable-type-classification, echo=FALSE}
hr_df <- get_hr_df()

numeric_vars <- hr_df %>%
  select(where(is.numeric)) %>%
  select(where(~ var(.x, na.rm = TRUE) > 0))

continuous_vars <- numeric_vars %>%
  select(where(~ dplyr::n_distinct(.x, na.rm = TRUE) > 15)) %>%
  names()

likert_vars <- setdiff(names(numeric_vars), continuous_vars)
```

## Variable Classification

For exploratory analysis, variables were grouped into four categories based on their measurement scale and analytical purpose:

-   **Continuous variables** representing measurable workforce characteristics such as salary, tenure, and distance from home.

-   **Likert-scale variables** capturing employee attitudes, perceptions, and workplace satisfaction measures.

-   **Categorical variables** describing demographic attributes and organizational context such as department, occupation, and work arrangement.

-   **Target variables** representing key organizational outcomes including turnover intention, quiet quitting behavior, and monitoring-related turnover risk.

This classification structure guides the exploratory analysis and ensures that each variable type is visualized using the most appropriate statistical and graphical techniques.

```{r 04-variable-selection, include=FALSE}
hr_df <- get_hr_df()

numeric_df <- hr_df %>%
  select(where(is.numeric)) %>%
  select(where(~ var(.x, na.rm = TRUE) > 0))

continuous_vars <- numeric_df %>%
  select(where(~ dplyr::n_distinct(.x, na.rm = TRUE) > 15)) %>%
  names()

top_continuous <- numeric_df %>%
  summarise(across(all_of(continuous_vars), ~ var(.x, na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "name", values_to = "value") %>%
  arrange(desc(value)) %>%
  slice_head(n = 6) %>%
  pull(name)

likert_vars <- numeric_df %>%
  select(where(~ is.numeric(.x) && max(.x, na.rm = TRUE) <= 7)) %>%
  names()

top_likert <- numeric_df %>%
  summarise(across(all_of(likert_vars), ~ var(.x, na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "name", values_to = "value") %>%
  arrange(desc(value)) %>%
  slice_head(n = 6) %>%
  pull(name)

top_outliers <- top_likert
```

## Continuous Variable Distributions

We use histogram and density plots to inspect the spread and shape of continuous variables.

```{r 04-continuous-distributions, echo=FALSE}
hr_df <- get_hr_df()

numeric_df <- hr_df %>%
  select(where(is.numeric)) %>%
  select(where(~ var(.x, na.rm = TRUE) > 0))

continuous_vars <- numeric_df %>%
  select(where(~ dplyr::n_distinct(.x, na.rm = TRUE) > 15)) %>%
  names()

top_continuous <- numeric_df %>%
  summarise(across(all_of(continuous_vars), ~ var(.x, na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "name", values_to = "value") %>%
  arrange(desc(value)) %>%
  slice_head(n = 9) %>%
  pull(name)

plots <- lapply(top_continuous, function(var) {
  ggplot(hr_df, aes(x = .data[[var]])) +
    geom_histogram(fill = hr_palette$primary, bins = 30, alpha = 0.8) +
    geom_density(color = "black", linewidth = 1.1) +
    labs(title = clean_label(var), x = NULL, y = "Density") +
    theme_hr(10)
})

p <- wrap_plots(plots, ncol = 3)

print(p)
save_plot(p, "continuous_distributions.png", 12, 8, plot_dir)
```

## Likert Scale Variable Distributions

Likert-scale variables are plotted as bar charts to show the distribution of employee attitudes, perceptions, and satisfaction levels.

```{r 04-likert-distributions, echo=FALSE}
hr_df <- get_hr_df()

numeric_df <- hr_df %>%
  select(where(is.numeric)) %>%
  select(where(~ var(.x, na.rm = TRUE) > 0))

likert_vars <- numeric_df %>%
  select(where(~ max(.x, na.rm = TRUE) <= 7)) %>%
  names()

top_likert <- numeric_df %>%
  summarise(across(all_of(likert_vars), ~ var(.x, na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "name", values_to = "value") %>%
  arrange(desc(value)) %>%
  slice_head(n = 6) %>%
  pull(name)

plots <- lapply(top_likert, function(var) {
  df <- hr_df %>%
    filter(!is.na(.data[[var]])) %>%
    count(.data[[var]])

  ggplot(df, aes(x = factor(.data[[var]]), y = n, fill = n)) +
    geom_col() +
    scale_fill_gradient(low = "#A8DADC", high = "#1D3557") +
    labs(title = clean_label(var), x = "Response", y = "Count") +
    theme_hr(10) +
    theme(legend.position = "none")
})

p <- wrap_plots(plots, ncol = 3)

print(p)
save_plot(p, "likert_distributions.png", 12, 8, plot_dir)
```

### Interpretation

Several productivity and engagement metrics appear normally distributed, suggesting that they may perform well in regression-based models. However, variables such as **Overtime_Hours** and **Mental_Health_Days_Taken** show right-skewed distributions, indicating potential behavioral outliers. These patterns may justify transformation or robust modeling approaches in later stages.

## Outlier Detection

Outliers can influence statistical inference, model stability, and interpretation of workforce trends.\
Boxplots combined with jittered observations allow us to visually inspect potential extreme values while preserving the underlying distribution of employee observations.

This step helps identify whether unusually high or low values represent realistic workforce variation or potential data anomalies that may require transformation or robust modeling techniques.

### Likert Variables (Outliers)

```{r 04-outliers, echo=FALSE}
hr_df <- get_hr_df()

numeric_df <- hr_df %>%
  select(where(is.numeric)) %>%
  select(where(~ var(.x, na.rm = TRUE) > 0))

likert_vars <- numeric_df %>%
  select(where(~ is.numeric(.x) && max(.x, na.rm = TRUE) <= 7)) %>%
  names()

top_outliers <- numeric_df %>%
  summarise(across(all_of(likert_vars), ~ var(.x, na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "name", values_to = "value") %>%
  arrange(desc(value)) %>%
  slice_head(n = 6) %>%
  pull(name)

likert_long <- hr_df %>%
  select(all_of(top_outliers)) %>%
  pivot_longer(
    cols = everything(),
    names_to = "name",
    values_to = "value"
  ) %>%
  filter(!is.na(value))

p <- ggplot(likert_long, aes(x = value, y = clean_label(name))) +
  geom_boxplot(
    fill = "#CCCCCC",
    alpha = 0.4,
    outlier.shape = NA
  ) +
  geom_jitter(
    aes(color = value),
    width = 0.15,
    alpha = 0.5,
    size = 1.3
  ) +
  scale_color_viridis_c(option = "C", end = 0.9) +
  theme_hr() +
  labs(
    title = "Likert Variable Outliers",
    x = "Response Level",
    y = NULL
  )

print(p)
save_plot(p, "likert_outliers.png", 12, 8, plot_dir)
```

# Workforce Relationship Analysis

Correlation analysis helps identify relationships between workforce variables and highlights clusters of related predictors that may influence employee behavior. Strong correlations may indicate shared underlying factors such as organizational support, engagement, or workplace monitoring practices.

Understanding these relationships is critical before predictive modeling, as highly correlated predictors can introduce multicollinearity and reduce model interpretability.

Because hierarchical clustering is used to organize the correlation matrix, we first remove numeric variables with **zero variance**. Variables with no variation cannot produce valid correlations and may introduce missing values that interfere with clustering algorithms.

## Correlation Matrix for Numeric Variables

The following visualization displays pairwise correlations between numeric variables. 
Variables are automatically reordered using hierarchical clustering so that related features appear closer together in the matrix, improving interpretability.

To improve readability while preserving meaningful relationships, only correlations with an absolute value of |r| ≥ 0.10 are displayed. 
This threshold reduces visual noise while retaining subtle but important behavioral signals commonly observed in workforce and organizational data.
Stronger relationships are emphasized visually, while weaker correlations are intentionally de-emphasized.

```{r correlation-matrix-elite, fig.width=12, fig.height=10, dpi=300}
hr_df <- get_hr_df()

numeric_data <- safe_numeric(hr_df)

corr_matrix <- cor(numeric_data, use = "complete.obs")

hc <- hclust(dist(corr_matrix))
corr_matrix <- corr_matrix[hc$order, hc$order]

corr_df <- reshape2::melt(corr_matrix)

corr_df <- corr_df %>%
  mutate(abs_corr = abs(value)) %>%
  filter(abs_corr >= 0.10) %>%
  mutate(
    Var1 = clean_label(Var1),
    Var2 = clean_label(Var2)
  )

p <- ggplot(corr_df, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(aes(alpha = abs(value)), color = "white", linewidth = 0.2) +
  scale_alpha(range = c(0.35, 1), guide = "none") +
  scale_fill_gradient2(
    low = "#D7263D",
    mid = "white",
    high = "#2A9D8F",
    midpoint = 0,
    limits = c(-1, 1)
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
    axis.text.y = element_text(size = 7),
    panel.grid = element_blank()
  )

print(p)
save_plot(p, "correlation_matrix.png", 12, 10, plot_dir)
```

### Interpreting Correlation Strength

Correlation coefficients (r) measure the strength and direction of relationships between variables. 
In workforce analytics, even small correlations can represent meaningful behavioral patterns due to the complexity of human systems.

- |r| ≈ 0.10 → Weak but meaningful signal  
- |r| ≈ 0.15–0.20 → Small but consistent relationship  
- |r| ≈ 0.25–0.30 → Moderate relationship  
- |r| ≥ 0.35 → Moderately strong relationship  
- |r| ≥ 0.50 → Strong relationship (potential multicollinearity)

This interpretation framework ensures that the visualization highlights meaningful workforce relationships without overstating weaker associations.

```{=html}
<div class="takeaway-box">
<p><strong>Key Insight</strong></p>

<p>
Most variables exhibit relatively low pairwise correlations, indicating a low risk of widespread multicollinearity across the dataset. 
This suggests that the majority of features contribute unique information and can be included in downstream modeling without significant redundancy concerns.
</p>

<p>
However, distinct clustering patterns are visible among conceptually related variables. In particular, mental health indicators (e.g., stress level, anxiety level, burnout score) show moderate positive relationships, reflecting a shared underlying construct of employee well-being. Similar clustering is observed among engagement, satisfaction, and organizational support variables, as well as among monitoring-related features.
</p>

<p>
These localized correlation structures suggest that while the dataset is broadly well-conditioned, certain groups of variables capture overlapping dimensions of employee experience. This presents a strong opportunity for dimensionality reduction, composite index creation, or feature selection strategies in future modeling to improve interpretability and reduce redundancy.
</p>

</div>
```

## Pairwise Relationship Exploration

```{r 05-pairwise-correlation, echo=FALSE}
hr_df <- get_hr_df()

numeric_df <- safe_numeric(hr_df)

corr_matrix <- cor(numeric_df, use = "complete.obs")

top_corr_vars <- names(
  sort(apply(abs(corr_matrix), 1, mean), decreasing = TRUE)
)[1:4]

pairwise_df <- numeric_df %>%
  select(all_of(top_corr_vars))

p <- GGally::ggpairs(
  pairwise_df,
  upper = list(continuous = wrap("cor", size = 3)),
  lower = list(
    continuous = function(data, mapping, ...) {
      ggplot(data, mapping) +
        geom_point(alpha = 0.4, color = "#2A9D8F") +
        geom_smooth(method = "lm", color = "black")
    }
  ),
  diag = list(
    continuous = function(data, mapping, ...) {
      ggplot(data, mapping) +
        geom_density(fill = "#4B9CD3", alpha = 0.8)
    }
  )
)

print(p)
save_plot(p, "pairwise.png", 10, 8, plot_dir)
```

### Key Correlation Insights

The correlation threshold exploration allows clearer interpretation of relationships between variables in the dataset. At the strictest threshold (∣r∣≥0.50), only one strong relationship emerges: Average Salary and Tenure Years. This relationship is expected in many organizational contexts, as employees with longer tenure often receive salary increases through accumulated experience, promotions, or periodic compensation adjustments.

As the threshold is gradually relaxed, additional relationships begin to appear among variables related to employee mental health, workplace stress, and monitoring perceptions.

At moderate thresholds (∣r∣≈0.30–0.35), correlations begin to emerge among variables associated with employee well-being. In particular, Depression Level, Stress Level, and Anxiety Level begin to show consistent associations with each other and with related mental health indicators such as Absenteeism Due to Mental Health and Mental Health Days Taken. These relationships suggest that the psychological well-being measures in the dataset behave coherently and may represent a shared underlying construct related to employee mental health.

At slightly lower thresholds (∣r∣≈0.25–0.30), additional relationships appear among workplace pressure variables. For example, Workload Pressure and Stressors Experienced begin to show associations with mental health measures, suggesting that increased workload demands and workplace stressors may contribute to elevated psychological strain among employees.

Relationships also begin to appear between monitoring-related variables and privacy perception metrics. In particular, Perceived Intrusiveness and Privacy Complaints Score show emerging correlations, indicating that employees who perceive workplace monitoring systems as more intrusive may also be more likely to report concerns related to privacy.

Taken together, the correlation analysis suggests three broad clusters of related variables within the dataset:

-   Compensation and tenure dynamics (Average Salary, Tenure Years)
-   Employee mental health and wellbeing indicators (Stress, Anxiety, Depression, Absenteeism)
-   Workplace monitoring and privacy perception measures (Monitoring Intrusiveness, Privacy Complaints)

These patterns indicate that the synthetic dataset captures several structural relationships commonly observed in organizational behavior research, particularly the connections between workplace stressors, employee well-being, and perceptions of organizational monitoring practices.

# Target Variable Exploration

This section examines the distribution of the key outcome variables used in predictive modeling. Understanding the balance of these classes is important because class imbalance can influence model training and evaluation metrics.

The following plots visualize the distribution of:

-   Turnover Intention
-   Quiet Quitting
-   Monitoring Turnover Risk

These visualizations help assess whether the outcomes occur with roughly balanced frequencies or whether resampling techniques may be necessary during model development.

## Target Variable Distributions

Below we visualize the target categorical outcomes: Turnover Intention, Quiet Quitting, and Monitoring Turnover Risk.

```{r 06-target-variable-distributions, echo=FALSE}
hr_df <- get_hr_df()

df_long <- hr_df %>%
  select(turnover_intention, quiet_quitting, monitoring_turnover_risk) %>%
  pivot_longer(everything()) %>%
  mutate(
    name = clean_label(name),
    value = clean_label(value)
  )

p <- ggplot(df_long, aes(x = value, fill = value)) +
  geom_bar() +
  facet_wrap(~name) +
  scale_fill_manual(values = c(
    "No" = hr_palette$primary,
    "Yes" = hr_palette$danger
  )) +
  theme_hr() +
  labs(
    title = "Target Variable Distributions",
    x = NULL,
    y = "Count"
  )

print(p)
save_plot(p, "target_distribution.png", 10, 6, plot_dir)
```

### Turnover Intention by Department

```{r 06-turnover-by-department, echo=FALSE}
hr_df <- get_hr_df()

df <- hr_df %>%
  filter(!is.na(turnover_intention), !is.na(department)) %>%
  mutate(
    department = clean_label(department),
    turnover_yes = turnover_intention == "Yes"
  ) %>%
  group_by(department) %>%
  summarise(turnover_rate = mean(turnover_yes), .groups = "drop") %>%
  arrange(turnover_rate)

p <- ggplot(df, aes(x = turnover_rate, y = reorder(department, turnover_rate), fill = turnover_rate)) +
  geom_col() +
  scale_fill_gradient(
    low = hr_palette$primary,
    high = hr_palette$danger
  ) +
  theme_hr(10) +
  labs(
    title = "Turnover Intention by Department",
    x = "Turnover Rate",
    y = NULL
  )

print(p)
save_plot(p, "turnover_by_department.png", 10, 6, plot_dir)
```

# Model Stability Diagnostics

Multicollinearity directly impacts model stability and interpretability.  
Before building predictive models, it is critical to ensure that predictor variables provide **independent signal rather than redundant information**. Variance Inflation Factor (VIF) is used to quantify this relationship.

## VIF for Numeric Predictors

```{r vif-numeric, fig.width=12, fig.height=10}
library(car)

hr_df <- get_hr_df()

numeric_data <- safe_numeric(hr_df)

numeric_data$dummy_response <- rnorm(nrow(numeric_data))

vif_model <- lm(dummy_response ~ ., data = numeric_data)

aliased <- is.na(coef(vif_model))

if (any(aliased)) {
  numeric_data <- numeric_data[, !names(numeric_data) %in% names(coef(vif_model))[aliased]]
  vif_model <- lm(dummy_response ~ ., data = numeric_data)
}

vif_values <- car::vif(vif_model)

vif_df <- tibble(
  Variable = names(vif_values),
  VIF = as.numeric(vif_values)
) %>%
  mutate(Variable = gsub("`|'", "", Variable)) %>%
  arrange(desc(VIF))

p <- ggplot(vif_df, aes(x = VIF, y = reorder(Variable, VIF), fill = VIF)) +
  geom_col() +
  scale_fill_gradient(low = "#A8DADC", high = "#1D3557") +
  geom_vline(xintercept = 5, linetype = "dashed", color = "#E63946", linewidth = 0.8) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.y = element_text(size = 8),
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "none"
  ) +
  labs(
    title = "Variance Inflation Factor (Numeric Predictors)",
    x = "Variance Inflation Factor",
    y = NULL
  )

print(p)
save_plot(p, "vif_numeric.png", 12, 10, plot_dir)
```

```{=html}
<div class="takeaway-box">
<p><strong>Key Insight</strong></p>

<p>
Most numeric predictors fall within acceptable VIF thresholds, indicating limited multicollinearity concerns.
</p>

<p>
A small number of variables show moderate inflation, which may reflect overlapping constructs rather than true redundancy. These variables should be monitored but do not currently require removal.
</p>

</div>
```

## VIF for Categorical Predictors

To assess multicollinearity among categorical predictors, we convert factor variables into dummy variables using one-hot encoding.

```{r vif-categorical, fig.width=14, fig.height=12}
library(car)
library(fastDummies)

hr_df <- get_hr_df()

cat_df <- safe_dummies(hr_df)

cat_df$dummy_response <- rnorm(nrow(cat_df))

vif_model <- lm(dummy_response ~ ., data = cat_df)

aliased <- is.na(coef(vif_model))
if (any(aliased)) {
  cat_df <- cat_df[, !names(cat_df) %in% names(coef(vif_model))[aliased]]
  vif_model <- lm(dummy_response ~ ., data = cat_df)
}

vif_values <- car::vif(vif_model)

vif_df <- tibble(
  Variable = names(vif_values),
  VIF = as.numeric(vif_values)
) %>%
  mutate(Variable = gsub("`|'", "", Variable))

p <- ggplot(vif_df, aes(x = VIF, y = reorder(Variable, VIF), fill = VIF)) +
  geom_col() +
  scale_fill_gradient(low = "#A8DADC", high = "#1D3557") +
  geom_vline(xintercept = 5, linetype = "dashed", color = "#F4A261") +
  geom_vline(xintercept = 10, linetype = "dashed", color = "#D7263D") +
  theme_minimal(base_size = 10) +
  theme(axis.text.y = element_text(size = 6)) +
  labs(title = "VIF – Categorical Predictors", x = "VIF", y = NULL)

print(p)
save_plot(p, "vif_categorical.png", 14, 12, plot_dir)
```

```{=html}
<div class="takeaway-box">
<p><strong>Key Insight</strong></p>

<p>
Most numeric predictors fall within acceptable VIF thresholds, indicating limited multicollinearity concerns.
</p>

<p>
A small number of variables show moderate inflation, which may reflect overlapping constructs rather than true redundancy. These variables should be monitored but do not currently require removal.
</p>

</div>
```

## Addressing Age Group Multicollinearity

The Variance Inflation Factor (VIF) measures the extent to which predictor variables are linearly associated with one another. Using a threshold of 5, the numeric predictors show no meaningful multicollinearity concerns.

For categorical predictors, however, multiple age_group dummy variables produced elevated VIF values. This occurs because the original age categories represent mutually exclusive levels of the same underlying construct. When many highly related dummy variables are included simultaneously, strong linear dependencies can arise.

To improve model stability and reduce redundancy, the original age-group categories are collapsed into three broader career-stage groupings:

-   Early-Career
-   Mid-Career
-   Senior-Career

This transformation preserves the conceptual meaning of the variable while reducing unnecessary multicollinearity among age-related predictors.

```{r 07-update-age-groups, fig.show="hold", echo=TRUE}
hr_df <- get_hr_df()

hr_df <- hr_df %>%
  mutate(
    age_group = case_when(
      age_group %in% c("16-19", "20-24", "25-34") ~ "Early-Career",
      age_group %in% c("35-44", "45-54") ~ "Mid-Career",
      age_group %in% c("55-64", "65+") ~ "Senior-Career",
      TRUE ~ as.character(age_group)
    ),
    age_group = factor(
      age_group,
      levels = c("Early-Career", "Mid-Career", "Senior-Career")
    )
  )
```

## Updated Categorical VIF Check

After collapsing the age categories into broader career stages, the multicollinearity diagnostics are recalculated for the categorical predictors to confirm that the issue has been resolved.

```{r vif-categorical-updated, fig.width=10, fig.height=8}
library(car)
library(fastDummies)

hr_df <- get_hr_df()

hr_df <- hr_df %>%
  mutate(
    age_group = case_when(
      age_group %in% c("16-19", "20-24", "25-34") ~ "Early-Career",
      age_group %in% c("35-44", "45-54") ~ "Mid-Career",
      age_group %in% c("55-64", "65+") ~ "Senior-Career",
      TRUE ~ as.character(age_group)
    ),
    age_group = factor(
      age_group,
      levels = c("Early-Career", "Mid-Career", "Senior-Career")
    )
  )

cat_df <- safe_dummies(hr_df)

cat_df$dummy_response <- rnorm(nrow(cat_df))

vif_model <- lm(dummy_response ~ ., data = cat_df)

aliased <- is.na(coef(vif_model))
if (any(aliased)) {
  cat_df <- cat_df[, !names(cat_df) %in% names(coef(vif_model))[aliased]]
  vif_model <- lm(dummy_response ~ ., data = cat_df)
}

vif_values <- car::vif(vif_model)

vif_df <- tibble(
  Variable = names(vif_values),
  VIF = as.numeric(vif_values)
) %>%
  mutate(Variable = gsub("`|'", "", Variable)) %>%
  arrange(VIF)

p <- ggplot(vif_df, aes(x = VIF, y = reorder(Variable, VIF), fill = VIF)) +
  geom_col() +
  scale_fill_gradient(low = "#A8DADC", high = "#1D3557") +
  geom_vline(xintercept = 5, linetype = "dashed", color = "#F4A261") +
  geom_vline(xintercept = 10, linetype = "dashed", color = "#D7263D") +
  theme_minimal(base_size = 11) +
  labs(title = "Updated VIF – Categorical Predictors", x = "VIF", y = NULL)

print(p)
save_plot(p, "vif_categorical_updated.png", 10, 8, plot_dir)
```

# Distribution & Risk Imbalance Analysis

Understanding distribution imbalance is important for predictive modeling because skewed variables and dominant categories can influence model performance, stability, and interpretation.  

In workforce analytics, imbalance is expected due to real-world behavioral patterns—such as extreme overtime, absenteeism, or mental health indicators occurring in a smaller subset of employees.

## Numeric Variable Skewness

```{r skewness-plot, fig.width=10, fig.height=6}
library(e1071)

hr_df <- get_hr_df()

numeric_df <- safe_numeric(hr_df)

skew_df <- numeric_df %>%
  summarise(across(everything(), ~ e1071::skewness(.x, na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "Skewness")

p <- ggplot(skew_df, aes(x = reorder(Variable, Skewness), y = Skewness, fill = Skewness)) +
  geom_col() +
  coord_flip() +
  scale_fill_gradient2(low = "#D7263D", mid = "white", high = "#2A9D8F", midpoint = 0) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "none") +
  labs(title = "Skewness of Numeric Variables", x = NULL, y = "Skewness")

print(p)
save_plot(p, "skewness.png", 10, 6, plot_dir)
```

```{=html}
<div class="takeaway-box">
<p><strong>Key Insight</strong></p>

<p>
Several behavioral workforce variables (e.g., overtime hours, absenteeism, and mental health days taken) exhibit right-skewed distributions, indicating that extreme outcomes are concentrated among a smaller subset of employees.
</p>

<p>
This pattern reflects realistic organizational dynamics, where most employees operate within typical ranges while a smaller group experiences elevated workload, stress, or disengagement.
</p>

<p>
From a modeling perspective, these distributions may benefit from transformation or robust modeling approaches, particularly when using regression-based techniques.
</p>

</div>
```

## Categorical Class Imbalance

Categorical variables can exhibit imbalance when one category dominates the distribution. While this is common in workforce datasets, it can influence how predictive models learn patterns and evaluate performance.

The following visualization highlights the most dominant category within each categorical variable.

### Dominant Category Proportion

```{r categorical-distributions, fig.width=12, fig.height=8}
hr_df <- get_hr_df()

cat_vars <- hr_df %>%
  select(where(is.factor)) %>%
  names()

plots <- lapply(cat_vars, function(var) {
  df <- hr_df %>%
    filter(!is.na(.data[[var]])) %>%
    count(.data[[var]]) %>%
    mutate(prop = n / sum(n)) %>%
    arrange(desc(prop)) %>%
    slice_head(n = 6)

  ggplot(df, aes(x = reorder(.data[[var]], prop), y = prop, fill = prop)) +
    geom_col() +
    coord_flip() +
    scale_fill_gradient(low = "#5FA8D3", high = "#D7263D") +
    theme_minimal(base_size = 10) +
    theme(legend.position = "none") +
    labs(title = var, x = NULL, y = "Proportion")
})

plots <- plots[1:min(length(plots), 6)]

p <- patchwork::wrap_plots(plots, ncol = 2)

print(p)
save_plot(p, "categorical_distributions.png", 12, 8, plot_dir)
```

```{=html}
<div class="takeaway-box">
<p><strong>Key Insight</strong></p>

<p>
Several categorical variables exhibit moderate class imbalance, with dominant categories representing a large share of employees. This reflects realistic workforce composition rather than data issues.
</p>

<p>
From a modeling perspective, imbalance may affect classification performance, making metrics such as precision, recall, and F1-score more informative than accuracy alone. Techniques such as resampling or class weighting may be considered if imbalance impacts results.
</p>

</div>
```

# Modeling Implications

Several categorical predictors display moderate imbalance, with certain categories representing a large share of the observations. This pattern is expected, as the synthetic dataset is informed by real-world workforce distributions where organizational roles, departments, and demographic attributes are rarely uniform.

From a modeling perspective, imbalance influences how algorithms learn patterns from the data. Dominant categories may contribute disproportionately to model training, while minority categories may contain important but underrepresented signals.

For categorical target variables, class imbalance can also distort evaluation metrics. Models may achieve high overall accuracy while performing poorly on minority outcomes. As a result, later modeling stages will consider more robust evaluation approaches, including precision, recall, and F1-score, as well as techniques such as class weighting or resampling.

These distribution characteristics reflect realistic workforce dynamics and provide impor

## Key Observations from Exploratory Analysis

The exploratory analysis reveals several meaningful patterns within the synthetic workforce dataset:

- **Workforce engagement variables** (e.g., organizational commitment, role clarity, job satisfaction) show balanced distributions, indicating strong variability and potential predictive signal.

- **Behavioral indicators** (e.g., overtime hours, absenteeism, mental health days) are right-skewed, suggesting that extreme outcomes are concentrated within a smaller subset of employees.

- **Correlation analysis** indicates generally weak relationships among numeric predictors, reducing the risk of unstable model estimates and improving interpretability.

- **Multicollinearity diagnostics** identified redundancy within detailed age group categories. Collapsing these into broader career stages significantly improved model stability.

- **Categorical imbalance patterns** reflect realistic workforce composition but may influence model learning and evaluation.

Overall, the dataset demonstrates structural properties consistent with real organizational data, supporting its use for predictive modeling experiments.

# Summary of Exploratory Diagnostics

The exploratory analysis evaluated several diagnostic aspects of the dataset, including variable distributions, correlation structure, and multicollinearity. The table below summarizes the primary findings and their implications for predictive modeling.

```{r 09-eda-summary-table, echo=FALSE}
eda_summary <- tibble(
  Diagnostic = c(
    "Correlation Structure",
    "Numeric Multicollinearity",
    "Categorical Multicollinearity",
    "Numeric Distribution Shape",
    "Categorical Class Balance"
  ),
  Key_Findings = c(
    "Most predictors display weak pairwise correlations; strongest between salary and tenure.",
    "Numeric predictors show low multicollinearity (VIF below thresholds).",
    "Age-group encoding initially caused redundancy in categorical variables.",
    "Behavioral metrics such as overtime and absenteeism are right-skewed.",
    "Some categorical variables show dominant categories."
  ),
  Modeling_Implications = c(
    "Low correlation improves regression stability.",
    "Numeric predictors can be safely included in models.",
    "Feature engineering improves categorical stability.",
    "Tree-based models handle skewness effectively.",
    "Evaluation metrics should account for class imbalance."
  )
)

knitr::kable(
  eda_summary,
  format = "html",
  caption = "Summary of Exploratory Diagnostics"
) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = TRUE
  )
```

# Transition to Predictive Modeling

The exploratory analysis confirms that the synthetic workforce dataset reflects realistic patterns observed in organizational environments.

Distribution diagnostics show meaningful variability across employee behaviors, while several behavioral metrics exhibit expected right-skewness. Correlation analysis indicates generally weak relationships among predictors, supporting model stability. Multicollinearity diagnostics further confirm that both numeric and categorical predictors are suitable for modeling following feature engineering adjustments.

Class imbalance patterns observed across categorical variables are consistent with real-world workforce distributions and will be addressed during model evaluation.

Together, these findings indicate that the dataset is well-positioned for predictive modeling.

The next phase focuses on developing machine learning models to predict key workforce outcomes:

- **Turnover Intention**
- **Quiet Quitting**
- **Monitoring-Related Turnover Risk**

These models will be evaluated across multiple algorithms to identify the most informative predictors of employee retention risk.

# Final Takeaways

This analysis demonstrates that the synthetic workforce dataset successfully replicates key patterns observed in real organizational environments.

Several critical insights emerge:

- Workforce behavior is not uniformly distributed — extreme outcomes are concentrated among smaller employee segments  
- Mental health variables form a tightly connected cluster, reinforcing their importance in predictive modeling  
- Monitoring and privacy perceptions show measurable relationships with employee sentiment  
- Multicollinearity risks can be mitigated through thoughtful feature engineering  

From a modeling perspective, the dataset is:

- Statistically stable  
- Behaviorally realistic  
- Structurally suitable for predictive analysis  

This positions the dataset for the next stage: **machine learning model development and evaluation**.

<hr>

---

## Summary of Key Findings

```{=html}
<div class="takeaway-box">

<p><strong>Summary Insight</strong></p>

<p>
Exploratory analysis reveals a well-structured dataset with minimal multicollinearity and realistic workforce behavior patterns.
</p>

<ul>
  <li><strong>Behavioral variables are right-skewed</strong>, indicating that extreme outcomes (e.g., high stress or absenteeism) are concentrated within a smaller subset of employees.</li>
  
  <li><strong>Correlation analysis shows low overall redundancy</strong>, with moderate clustering among well-being variables such as stress, burnout, and anxiety.</li>
  
  <li><strong>Variance Inflation Factor (VIF) diagnostics confirm acceptable levels of multicollinearity</strong>, even after accounting for categorical expansion.</li>
  
  <li><strong>Categorical variables exhibit moderate class imbalance</strong>, reflecting realistic workforce composition and informing future model evaluation strategies.</li>
</ul>

<p>
Overall, the dataset is well-suited for predictive modeling, with opportunities for feature engineering and transformation to further enhance model performance.
</p>

</div>
```

```{=html}
<hr>

<div style="
  display:grid;
  grid-template-columns: 1fr auto 1fr;
  align-items:center;
  margin:30px 0;
">

  <!-- LEFT -->
  <div style="text-align:left;">
    <a href="index.html" style="
      display:inline-block;
      padding:12px 28px;
      border-radius:999px;
      font-weight:600;
      color:white;
      text-decoration:none;
      background:linear-gradient(135deg, #74B9FF, #2E86C1);
    ">
      ← Back Home
    </a>
  </div>

  <!-- CENTER (empty for alignment) -->
  <div></div>

  <!-- RIGHT -->
  <div style="text-align:right;">
    <a href="HR_Data_Turnover_Intention_Results.html" style="
      display:inline-block;
      padding:12px 28px;
      border-radius:999px;
      font-weight:600;
      color:white;
      text-decoration:none;
      background:linear-gradient(135deg, #FF7675, #922B21);
    ">
      Next →
    </a>
  </div>

</div>
```