---
title: ""

output:
  html_document:
    always_allow_html: true
    self_contained: false
    toc: false
    toc_float: false
    theme: null
    highlight: null
    df_print: paged
    fig_caption: true
    css: assets/styles.css
    includes:
      in_header: ../docs/assets/header.html

number_sections: true
---

```{=html}
<div class="report-container">

  <div class="page-layout">

    <!-- SIDEBAR -->
    <div class="sidebar">
      <div class="sidebar-card">
        <div class="sidebar-label">On this page</div>
        <nav class="sidebar-nav">
          <a href="#executive-summary">Executive Summary</a>
          <a href="#modeling-overview">Modeling Overview</a>
          <a href="#data-prep-framework">Data Preparation</a>
          <a href="#model-development">Model Development</a>
          <a href="#model-evaluation">Model Evaluation</a>
          <a href="#model-evaluation-results">Evaluation Results</a>
          <a href="#sampling-strategy-framework">Sampling Strategy Framework</a>
          <a href="#precision-recall">Precision–Recall</a>
          <a href="#model-selection">Model Selection</a>
          <a href="#model-summary">Model Summary</a>
          <a href="#driver-interpretation">Driver Interpretation</a>
          <a href="#feature-action">Feature → Action</a>
          <a href="#business-impact">Business Impact</a>
        </nav>
      </div>
    </div>

    <!-- MAIN CONTENT -->
    <div class="page-content">

      <!-- MATCH EDA STRUCTURE -->
      <div class="hero-section">
        <p class="hero-meta">
          Kevin Allen Cauley • `r format(Sys.Date(), '%B %d, %Y')`
        </p>

        <h1 class="hero-title">Quiet Quitting Modeling</h1>

        <p class="hero-subtitle">
          This section develops and evaluates predictive models to identify employees at risk of quiet quitting, interpret key disengagement risk patterns, and translate model outputs into executive-level engagement recommendations.
        </p>
      </div>

      <!-- REPLACED page-nav WITH cta-row -->
      <div class="cta-row">
        <a href="HR_Data_Turnover_Intention.html" class="cta-btn btn-blue">← Back</a>
        <a href="index.html" class="cta-btn btn-teal">Home</a>
        <a href="HR_Data_Monitoring_Privacy.html" class="cta-btn btn-red">Next →</a>
      </div>
```

```{r 00-setup, include=FALSE}
source(here::here("analysis", "helper.R"))

PLOT_SUBFOLDER <- "quiet_quitting"

cat("Using plot folder:", PLOT_SUBFOLDER, "\n")

# Required Packages
required_pkgs <- c(
  "conflicted", "tidyverse", "tidymodels", "themis", "here",
  "klaR", "discrim", "ranger", "kernlab", "xgboost", "gt", "scales"
)

for (pkg in required_pkgs) {
  if (!pkg %in% rownames(installed.packages())) {
    install.packages(pkg)
  }
}

library(conflicted)
library(tidyverse)
library(tidymodels)
library(themis)
library(here)
library(klaR)
library(discrim)
library(gt)
library(scales)

conflicted::conflicts_prefer(
  dplyr::filter,
  dplyr::select,
  dplyr::mutate,
  dplyr::summarise,
  dplyr::arrange,
  dplyr::slice,
  dplyr::slice_head,
  dplyr::slice_max,
  dplyr::slice_min,
  dplyr::recode,
  dplyr::rename,
  dplyr::lag
)

# Global Options
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  results = "hide",
  fig.align = "center",
  out.width = "100%",
  fig.retina = 2,
  fig.width = 9,
  fig.height = 5.75,
  dpi = 300,
  cache = FALSE,
  dev = "png"
)

# Reproducibility
set.seed(123)

# Load CLEAN dataset from EDA (DO NOT TOUCH THIS)
df <- readRDS(here::here("data", "processed", "hr_dataset_clean.rds"))

# Base dataset
base_df <- df %>%
  mutate(across(where(is.character), as.factor))
```

# Executive Summary {#executive-summary}

This analysis develops a privacy-preserving predictive analytics framework designed to identify workforce disengagement patterns associated with quiet quitting using a synthetic, research-informed HR dataset.

Rather than modeling resignation behavior directly, this framework focuses on identifying employees who may demonstrate reduced discretionary effort, lower engagement, or elevated disengagement risk while remaining employed.

Multiple machine learning workflows—including Logistic Regression, Random Forest, XGBoost, Support Vector Machines, and Naive Bayes—were evaluated under varying sampling strategies to assess how workforce-risk detection changes under different modeling conditions.

Results show that disengagement-risk modeling is influenced as much by workforce-data strategy as by algorithm selection. Resampling techniques materially changed detection behavior, precision, recall, and operational usability across model configurations.

The final model-selection framework prioritizes balanced workforce-risk detection while confirming that selected configurations remain practically usable when applied to unseen workforce patterns.

Importantly, this project demonstrates how synthetic workforce data can support advanced HR analytics experimentation without exposing sensitive employee information or requiring access to confidential organizational records.

From a business perspective, the framework supports earlier workforce-risk visibility, more targeted engagement review, improved intervention prioritization, and stronger organizational decision support.

# Modeling Overview {#modeling-overview}

This section builds on the exploratory analysis by applying predictive modeling techniques to identify employees at risk of quiet quitting.

Rather than relying on a single model, multiple algorithms are evaluated and compared to determine the most effective approach for workforce decision-making.

Each model is assessed using cross-validation and evaluated across multiple sampling strategies to address class imbalance.

The objective is not only to identify the most accurate model, but to determine which approach best aligns with organizational priorities.

# Data Preparation & Modeling Framework {#data-prep-framework}

## Target Definition

The target variable for this analysis is **quiet quitting**, representing elevated disengagement-risk behavior while employees remain organizationally active.

Within this framework, quiet quitting should not be interpreted as a formally observable employment event. Instead, it reflects a collection of behavioral, engagement, wellbeing, role clarity, discretionary-effort, and organizational-risk indicators associated with declining workforce connection.

This distinction is important because HR analytics often focuses on identifying emerging workforce-risk patterns before they become visible through turnover, burnout, performance decline, or employee-relations concerns.

The outcome is structured as a binary classification target with levels defined as **"No"** and **"Yes"**, where **"Yes"** represents employees flagged as higher risk for quiet-quitting behavior in the synthetic dataset.

```{r 01-target, echo=FALSE}
# Ensure correct factor levels
base_df <- base_df %>%
  mutate(
    quiet_quitting = factor(quiet_quitting, levels = c("No", "Yes"))
  )
```

## Data Partitioning

To evaluate model performance objectively, the dataset is split into training and testing subsets.

The training set is used to fit and tune models, while the test set is reserved for final evaluation. Stratified sampling is applied to preserve the original class distribution of quiet quitting intention, ensuring that model performance reflects realistic conditions.

```{r 02-split, echo=FALSE}
split <- initial_split(base_df, prop = 0.8, strata = quiet_quitting)

train_data <- training(split)
test_data  <- testing(split)
```

## Feature Engineering Pipeline

A standardized preprocessing workflow was applied to ensure consistent evaluation across all modeling approaches.

Workforce variables representing organizational context, employee wellbeing, engagement, managerial support, workload conditions, role clarity, and behavioral indicators were transformed into model-compatible formats while preserving interpretability for downstream HR analysis.

Categorical workforce attributes were converted into numeric representations using dummy encoding, while continuous variables were normalized where required for scale-sensitive algorithms such as Logistic Regression and Support Vector Machines.

Predictors with zero variance or limited informational value were removed to reduce noise and improve model stability across resampling conditions.

This preprocessing framework ensures that observed performance differences are driven primarily by modeling behavior and sampling strategy rather than inconsistent data preparation.

```{r 03-recipe, echo=FALSE}
rec <- recipe(quiet_quitting ~ ., data = train_data) %>%
  
  # Handle factors
  step_dummy(all_nominal_predictors()) %>%
  
  # Normalize numeric features (important for SVM, Logistic)
  step_normalize(all_numeric_predictors()) %>%
  
  # Remove zero variance
  step_zv(all_predictors())
```

## Cross-Validation Strategy

Cross-validation is used to evaluate model performance during training and reduce the risk of overfitting.

By partitioning the training data into multiple folds, each model is evaluated across different subsets of the data. This provides a more robust estimate of model performance compared to a single train-test split.

```{r 04-resampling, echo=FALSE}
cv_folds <- vfold_cv(train_data, v = 5, strata = quiet_quitting)
```

# Model Development {#model-development}

Multiple predictive modeling approaches were evaluated to determine how different algorithms interpret disengagement-risk patterns within a synthetic HR environment.

The goal was not simply to identify the highest-performing model, but to compare how various modeling strategies balance interpretability, detection capability, stability, and operational usefulness within a privacy-preserving analytics framework.

Each model was trained using a standardized preprocessing workflow to ensure that performance differences reflected modeling behavior rather than inconsistent data preparation.

- Logistic Regression (baseline, interpretable)
- Random Forest (ensemble tree-based)
- XGBoost (boosted trees for performance)
- Support Vector Machine (nonlinear classification)
- Naive Bayes (probabilistic baseline)

This allows for a fair comparison across algorithms with varying levels of complexity and interpretability.

---

```{r 05-logistic, echo=FALSE}
log_model <- logistic_reg() %>%
  set_engine("glm")
```

```{r 06-rf, echo=FALSE}
rf_model <- rand_forest(mtry = tune(), trees = 500, min_n = tune()) %>%
  set_engine("ranger") %>%
  set_mode("classification")
```

```{r 07-xgb, echo=FALSE}
xgb_model <- boost_tree(
  trees = 500,
  mtry = tune(),
  learn_rate = tune(),
  min_n = tune()
) %>%
  set_engine("xgboost") %>%
  set_mode("classification")
```

```{r 08-nb, echo=FALSE}
nb_model <- naive_Bayes() %>%
  set_engine("klaR")
```

```{r 09-svm, echo=FALSE}
svm_model <- svm_rbf(cost = tune(), rbf_sigma = tune()) %>%
  set_engine("kernlab") %>%
  set_mode("classification")
```

## Model Pipelines

Workflows are used to combine preprocessing steps and model specifications into a single pipeline.

This ensures that all transformations are applied consistently during both training and prediction, reducing the risk of data leakage and simplifying model management.

```{r 10-workflows, echo=FALSE}
log_wf <- workflow() %>% add_model(log_model) %>% add_recipe(rec)
rf_wf  <- workflow() %>% add_model(rf_model)  %>% add_recipe(rec)
xgb_wf <- workflow() %>% add_model(xgb_model) %>% add_recipe(rec)
svm_wf <- workflow() %>% add_model(svm_model) %>% add_recipe(rec)
nb_wf  <- workflow() %>% add_model(nb_model)  %>% add_recipe(rec)
```

## Evaluation Metrics

Model performance is evaluated using a combination of classification metrics.

Accuracy provides an overall performance measure, while precision and recall capture trade-offs between false positives and false negatives. The F1 Score balances these two metrics, and ROC AUC evaluates the model’s ability to distinguish between classes.

Using multiple metrics ensures a comprehensive evaluation aligned with business priorities.

```{r 11-metrics, echo=FALSE}
# Classification metrics
# IMPORTANT:
# quiet_quitting levels are c("No", "Yes")
# event_level = "second" ensures the positive class is "Yes" = at risk of quiet quitting

options(yardstick.event_first = FALSE)

class_metrics <- metric_set(
  accuracy,
  precision,
  recall,
  f_meas,
  roc_auc
)

target_event_level <- "second"

model_colors <- c(
  "Logistic" = "#4B9CD3",
  "Random Forest" = "#3FA34D",
  "XGBoost" = "#E87461",
  "Naive Bayes" = "#F4D03F",
  "SVM" = "#9370DB"
)

sampling_order <- c(
  "No Sampling",
  "Up Sampling",
  "Down Sampling",
  "SMOTE"
)

metric_labels <- c(
  accuracy = "Accuracy",
  precision = "Precision",
  recall = "Recall",
  f_meas = "F1 Score",
  roc_auc = "ROC AUC"
)
```

## Hyperparameter Tuning

For models with tunable parameters, grid search is used to identify optimal configurations.

This process evaluates multiple parameter combinations using cross-validation, selecting those that maximize predictive performance. Tuning helps improve model accuracy while maintaining generalizability.

## Model Training

Multiple supervised learning algorithms were evaluated, including Logistic Regression, Random Forest, XGBoost, Support Vector Machines, and Naive Bayes.

Each model was trained under multiple class-balancing conditions to evaluate how resampling strategies influence disengagement-risk detection behavior. This is particularly important in quiet quitting modeling because employees exhibiting elevated disengagement patterns often represent a smaller portion of the workforce population.

Rather than focusing on a single performance statistic, the modeling framework evaluates how each configuration balances detection capability, prediction reliability, and operational practicality.

This comparative approach reinforces the broader objective of the project: demonstrating how research-informed synthetic workforce data can support realistic HR analytics workflows while preserving employee privacy and organizational confidentiality.

```{r 12-modeling-framework, include=FALSE, results='hide', message=FALSE, warning=FALSE}
# This framework evaluates each model + sampling strategy combination together using cross-validation. Tunable models are tuned separately under each sampling condition.

# Final model selection is based on cross-validated F1 Score with a holdout viability check before reporting the selected model.

# Sampling recipes
sampling_recipes <- list(
  "No Sampling" = rec,
  "Up Sampling" = rec %>% step_upsample(quiet_quitting),
  "Down Sampling" = rec %>% step_downsample(quiet_quitting),
  "SMOTE" = rec %>% step_smote(quiet_quitting)
)

# Model specifications
model_specs <- list(
  "Logistic" = log_model,
  "Naive Bayes" = nb_model,
  "Random Forest" = rf_model,
  "XGBoost" = xgb_model,
  "SVM" = svm_model
)

# Tunable model names
tunable_models <- c("Random Forest", "XGBoost", "SVM")

# Unified prediction function
get_preds <- function(model, data) {
  predict(model, data, type = "prob") %>%
    bind_cols(predict(model, data)) %>%
    bind_cols(data %>% select(quiet_quitting))
}

# Unified scoring function
score_predictions <- function(preds, model_name, sampling_name) {
  class_metrics(
    preds,
    truth = quiet_quitting,
    estimate = .pred_class,
    .pred_Yes,
    event_level = target_event_level
  ) %>%
    mutate(
      Model = model_name,
      Sampling = sampling_name
    )
}

# Evaluate one model + sampling combination using CV
run_cv_model_sampling <- function(model_name, sampling_name) {
  
  model_spec <- model_specs[[model_name]]
  recipe_used <- sampling_recipes[[sampling_name]]
  
  wf <- workflow() %>%
    add_model(model_spec) %>%
    add_recipe(recipe_used)
  
  set.seed(123)
  
  if (model_name %in% tunable_models) {
    
    cv_fit <- tune_grid(
      wf,
      resamples = cv_folds,
      grid = 10,
      metrics = class_metrics,
      control = control_grid(save_pred = TRUE)
    )
    
    best_params <- select_best(cv_fit, metric = "f_meas")
    
    # Keep only metrics for the best tuned parameter combination.
    # This prevents duplicate model + sampling rows from appearing later.
    cv_metrics <- collect_metrics(cv_fit) %>%
      inner_join(best_params, by = names(best_params)) %>%
      mutate(
        Model = model_name,
        Sampling = sampling_name,
        Model_Config = paste(model_name, sampling_name, sep = " + ")
      )
    
    final_wf <- finalize_workflow(wf, best_params)
    
  } else {
    
    cv_fit <- fit_resamples(
      wf,
      resamples = cv_folds,
      metrics = class_metrics,
      control = control_resamples(save_pred = TRUE)
    )
    
    cv_metrics <- collect_metrics(cv_fit) %>%
      mutate(
        Model = model_name,
        Sampling = sampling_name,
        Model_Config = paste(model_name, sampling_name, sep = " + ")
      )
    
    final_wf <- wf
  }
  
  # Fit finalized workflow to full training data
  set.seed(123)
  final_fit <- fit(final_wf, data = train_data)
  
  # Evaluate finalized workflow on holdout test data
  test_preds <- get_preds(final_fit, test_data)
  
  test_metrics <- score_predictions(
    preds = test_preds,
    model_name = model_name,
    sampling_name = sampling_name
  ) %>%
    mutate(Model_Config = paste(model_name, sampling_name, sep = " + "))
  
  list(
    cv_fit = cv_fit,
    cv_metrics = cv_metrics,
    final_workflow = final_wf,
    final_fit = final_fit,
    test_preds = test_preds,
    test_metrics = test_metrics
  )
}

# Run all model + sampling combinations
model_grid <- tidyr::expand_grid(
  Model = names(model_specs),
  Sampling = names(sampling_recipes)
)

set.seed(123)

model_training_output <- capture.output({
  model_results <- purrr::pmap(
    list(model_grid$Model, model_grid$Sampling),
    run_cv_model_sampling
  )
})

names(model_results) <- paste(model_grid$Model, model_grid$Sampling, sep = " + ")

# Cross-validation results for all best model + sampling configurations
cv_results_all <- purrr::map_dfr(model_results, "cv_metrics") %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "roc_auc")) %>%
  mutate(
    Metric = dplyr::recode(.metric, !!!metric_labels),
    Sampling = factor(Sampling, levels = sampling_order),
    Model = factor(Model, levels = names(model_colors))
  )

# Holdout test results for all finalized configurations
results_all <- purrr::map_dfr(model_results, "test_metrics") %>%
  filter(.metric %in% c("accuracy", "precision", "recall", "f_meas", "roc_auc")) %>%
  mutate(
    Metric = dplyr::recode(.metric, !!!metric_labels),
    Sampling = factor(Sampling, levels = sampling_order),
    Model = factor(Model, levels = names(model_colors))
  )
```

---

# Model Evaluation {#model-evaluation}

This evaluation framework examines how different modeling strategies perform on unseen workforce data while emphasizing practical HR decision-support considerations rather than purely statistical optimization.

Because workforce intervention resources are finite, the analysis prioritizes models that balance early disengagement-risk detection with prediction reliability. The objective is to support proactive engagement review processes while minimizing unnecessary intervention burden.

Evaluation therefore focuses not only on predictive accuracy, but also on how effectively each configuration translates into operationally useful workforce prioritization signals.

Rather than focusing solely on accuracy, evaluation emphasizes the **precision–recall tradeoff**, balancing the ability to detect at-risk employees with the cost of unnecessary intervention.

This ensures model selection aligns with operational priorities, where both detection capability and decision efficiency matter.

Key metrics include:

- **Accuracy** → Overall correctness  
- **Precision** → Reliability of positive predictions  
- **Recall** → Ability to detect at-risk employees  
- **F1 Score** → Balance between precision and recall  
- **ROC AUC** → Overall classification strength  

This comparison highlights the trade-offs between detection, prediction reliability, and practical intervention burden.

## Test Set Performance

Final model assessment is conducted using the holdout test dataset to evaluate how each configuration performs on previously unseen workforce observations.

This stage acts as a practical stability check rather than the primary model-selection mechanism. Cross-validation identifies the strongest generalized configurations during training, while holdout evaluation confirms whether those patterns remain operationally useful under unseen conditions.

This distinction is especially important in workforce analytics because some configurations may perform well during resampling yet produce unstable or overly aggressive behavior when applied to new employee populations.

The evaluation framework therefore separates model discovery from final operational validation, creating a more defensible and realistic assessment process for workforce-risk modeling.

---

# Sampling Strategy Framework {#sampling-strategy-framework}

Sampling techniques are applied within the preprocessing pipeline to ensure consistency across models.

By integrating sampling directly into the workflow, transformations are applied only to the training data, preventing data leakage and preserving the integrity of model evaluation.

## Sampling Recipes and Evaluation Setup

Each sampling strategy is embedded directly into the modeling recipe. This ensures that class balancing occurs only within the training folds during cross-validation, preventing leakage into validation or test data.

# Model Evaluation Results & Insights {#model-evaluation-results}

This section synthesizes model performance across sampling strategies and evaluation metrics, providing a comparative view of how different approaches perform under realistic conditions.

The focus shifts from individual model behavior to cross-model insights, highlighting trade-offs that directly impact workforce decision-making.

The goal is to identify a model that balances detection capability with operational efficiency—minimizing both missed risk and unnecessary intervention.

```{r 13-validation-check, include=FALSE}
stopifnot(exists("cv_results_all"))
stopifnot(exists("results_all"))
stopifnot(nrow(cv_results_all) > 0)
stopifnot(nrow(results_all) > 0)
stopifnot(all(c("Model", "Sampling", ".metric", ".estimate") %in% names(results_all)))
stopifnot(all(c("Model", "Sampling", ".metric", "mean") %in% names(cv_results_all)))
```

```{r 14-final-model-selection, include=FALSE}
# -------------------------------------------------------------------
# Final model selection
# -------------------------------------------------------------------
# Gold-standard selection is based primarily on cross-validated F1 Score,
# but the selected model must also meet minimum holdout stability checks.
#
# This prevents a model from being selected if it performs well in CV
# but fails to produce usable positive-class predictions on the holdout set.
# -------------------------------------------------------------------

# Minimum holdout requirements for a model to be considered usable
min_holdout_f1 <- 0.40
min_holdout_recall <- 0.30
min_holdout_precision <- 0.20

# Create holdout metric screen
holdout_screen <- results_all %>%
  filter(.metric %in% c("precision", "recall", "f_meas", "roc_auc")) %>%
  select(Model, Sampling, .metric, .estimate) %>%
  pivot_wider(names_from = .metric, values_from = .estimate) %>%
  mutate(
    Model = as.character(Model),
    Sampling = as.character(Sampling),
    viable_holdout = !is.na(precision) &
      !is.na(recall) &
      !is.na(f_meas) &
      precision > 0 &
      recall > 0 &
      f_meas > 0,
    stable_holdout = viable_holdout &
      f_meas >= min_holdout_f1 &
      recall >= min_holdout_recall &
      precision >= min_holdout_precision
  )

# Prefer models that meet the stronger holdout stability screen.
eligible_configs <- holdout_screen %>%
  dplyr::filter(stable_holdout) %>%
  dplyr::select(Model, Sampling)

# Fallback: if no model meets the stronger screen, use basic viability.
if (nrow(eligible_configs) == 0) {
  eligible_configs <- holdout_screen %>%
    dplyr::filter(viable_holdout) %>%
    dplyr::select(Model, Sampling)
}

# Stop if no model is even viable on the holdout set.
if (nrow(eligible_configs) == 0) {
  stop(
    "No model configuration produced viable positive-class holdout predictions. ",
    "Review class balance, sampling methods, thresholding, and model specifications before selecting a final model."
  )
}

# Cross-validated F1 table
cv_f1_table <- cv_results_all %>%
  filter(.metric == "f_meas") %>%
  mutate(
    Model = as.character(Model),
    Sampling = as.character(Sampling)
  ) %>%
  select(Model, Sampling, cv_f1 = mean)

# Select the best cross-validated model among eligible holdout configurations
top_model_cv <- cv_f1_table %>%
  inner_join(eligible_configs, by = c("Model", "Sampling")) %>%
  left_join(
    holdout_screen %>%
      dplyr::select(
        Model,
        Sampling,
        holdout_precision = precision,
        holdout_recall = recall,
        holdout_f1 = f_meas,
        holdout_roc_auc = roc_auc,
        stable_holdout,
        viable_holdout
      ),
    by = c("Model", "Sampling")
  ) %>%
  dplyr::arrange(dplyr::desc(cv_f1), dplyr::desc(holdout_f1)) %>%
  utils::head(1)

top_model_name <- as.character(top_model_cv$Model)
top_sampling_method <- as.character(top_model_cv$Sampling)
top_model_config <- paste(top_model_name, top_sampling_method, sep = " + ")

top_f1 <- round(top_model_cv$cv_f1, 3)
top_holdout_f1 <- round(top_model_cv$holdout_f1, 3)
top_holdout_precision <- round(top_model_cv$holdout_precision, 3)
top_holdout_recall <- round(top_model_cv$holdout_recall, 3)
top_holdout_roc_auc <- round(top_model_cv$holdout_roc_auc, 3)

selection_strategy_label <- ifelse(
  isTRUE(top_model_cv$stable_holdout),
  "Holdout-Stable Risk Prioritization",
  "Holdout-Viable Risk Prioritization"
)

selection_rule_note <- ifelse(
  isTRUE(top_model_cv$stable_holdout),
  "The selected model met the holdout stability screen and ranked strongest by cross-validated F1 Score among stable candidates.",
  "No configuration met the full holdout stability screen, so the selected model is the strongest viable holdout performer among cross-validated candidates."
)

# Retrieve selected model object from saved model results
selected_result <- model_results[[top_model_config]]

final_wf <- selected_result$final_workflow
final_fit <- selected_result$final_fit
final_preds <- selected_result$test_preds

# Holdout test metrics for selected model
final_model_metrics <- results_all %>%
  filter(
    as.character(Model) == top_model_name,
    as.character(Sampling) == top_sampling_method,
    .metric %in% c("accuracy", "precision", "recall", "f_meas", "roc_auc")
  ) %>%
  mutate(
    Metric = dplyr::recode(.metric, !!!metric_labels),
    Score = round(.estimate, 3)
  )
```

```{r 14b-selected-model-consistency-check, include=FALSE}
# Final selected model consistency check
stopifnot(exists("top_model_name"))
stopifnot(exists("top_sampling_method"))
stopifnot(exists("top_model_config"))
stopifnot(exists("final_model_metrics"))
stopifnot(exists("final_preds"))

stopifnot(length(top_model_name) == 1)
stopifnot(length(top_sampling_method) == 1)
stopifnot(nrow(final_model_metrics) == 5)

stopifnot(
  all(c("Accuracy", "Precision", "Recall", "F1 Score", "ROC AUC") %in% final_model_metrics$Metric)
)
```

## Sampling Strategy Comparison {#sampling-strategy-comparison}

Class imbalance is a major challenge in quiet-quitting prediction because employees flagged for disengagement risk typically represent a smaller portion of the workforce population.

To better understand how class-balancing strategies influence disengagement-risk detection, multiple resampling approaches were evaluated across all algorithms. The goal is not to artificially inflate performance, but to examine how different balancing strategies affect the tradeoff between detection capability and prediction reliability.

The models include:

- No Sampling (baseline)
- Up Sampling (increase minority class)
- Down Sampling (reduce majority class)
- SMOTE (synthetic minority generation)

**The goal is to understand how balancing the dataset impacts model performance.**

```{r 15-plot-sampling-comparison, echo=FALSE}
plot_df <- results_all %>%
  filter(.metric %in% c("precision", "recall", "f_meas", "roc_auc")) %>%
  mutate(
    Metric = factor(
      Metric,
      levels = c("Precision", "Recall", "F1 Score", "ROC AUC")
    ),
    Model_Sampling = paste(Model, Sampling, sep = " + ")
  )

f1_order <- plot_df %>%
  filter(.metric == "f_meas") %>%
  arrange(desc(.estimate)) %>%
  pull(Model_Sampling)

plot_df <- plot_df %>%
  mutate(
    Model_Sampling = factor(Model_Sampling, levels = rev(f1_order)),
    Plot_Label = ifelse(
      is.na(.estimate),
      "",
      scales::number(.estimate, accuracy = 0.001)
    )
  )

p <- ggplot(plot_df, aes(x = Metric, y = Model_Sampling, fill = .estimate)) +
  geom_tile(color = "white", linewidth = 0.6) +
  geom_text(
    aes(label = Plot_Label),
    fontface = "bold",
    size = 3.5,
    color = "#243447"
  ) +
  scale_fill_gradient(
    low = "#EAF2F8",
    high = "#2E86C1",
    limits = c(0, 1),
    na.value = "#F2F4F4",
    name = "Score"
  ) +
  labs(
    title = "Holdout Performance Across Model Configurations",
    subtitle = "Rows are ranked by holdout F1 Score; higher values indicate stronger performance",
    x = NULL,
    y = NULL
  ) +
  theme_hr(10) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(
      size = 12,
      color = "#5a6b7c"
    ),
    axis.text.y = element_text(size = 10, face = "bold"),
    axis.text.x = element_text(size = 10, face = "bold"),
    legend.position = "right",
    panel.grid = element_blank()
  )

print(p)
save_plot(p, "sampling_strategy_performance_heatmap.png", 10.5, 8)
```

The heatmap shows holdout test performance across model and sampling combinations. It is used as a confirmation view rather than the sole model-selection mechanism.

The final model is selected using cross-validated F1 Score among configurations that also produce usable holdout performance. This distinction matters because a model can look strong on one metric during development but still weaken when applied to new workforce observations.

The results show that no single configuration dominates every metric. Some approaches improve recall but reduce precision, while others are more conservative. This reinforces that quiet-quitting modeling should be interpreted as a decision-support layer rather than as an automatic classification system.

## Precision–Recall Curve Analysis {#precision-recall}

```{r 16-precision-recall, echo=FALSE}
pr_df <- results_all %>%
  filter(.metric %in% c("precision", "recall", "f_meas")) %>%
  select(Model, Sampling, .metric, .estimate) %>%
  pivot_wider(names_from = .metric, values_from = .estimate) %>%
  filter(!is.na(precision), !is.na(recall), !is.na(f_meas)) %>%
  mutate(
    Model = factor(as.character(Model), levels = names(model_colors)),
    Sampling = factor(as.character(Sampling), levels = sampling_order),
    Model_Config = paste(Model, Sampling, sep = " + "),
    Is_Selected = Model_Config == top_model_config
  )

selected_point <- pr_df %>%
  dplyr::filter(Is_Selected) %>%
  utils::head(1)

p <- ggplot(
  pr_df,
  aes(
    x = recall,
    y = precision,
    color = Model,
    shape = Sampling
  )
) +
  geom_vline(
    xintercept = 0.50,
    linetype = "dashed",
    color = "#D5DBDB",
    linewidth = 0.5
  ) +
  geom_hline(
    yintercept = 0.50,
    linetype = "dashed",
    color = "#D5DBDB",
    linewidth = 0.5
  ) +
  annotate(
    "rect",
    xmin = 0.45,
    xmax = 0.65,
    ymin = 0.45,
    ymax = 0.65,
    alpha = 0.08,
    fill = "#2A9D8F"
  ) +
  annotate(
    "text",
    x = 0.55,
    y = 0.67,
    label = "Balanced Operational Zone",
    size = 4.2,
    fontface = "bold",
    color = "#1B4332"
  ) +
  geom_point(
    aes(size = f_meas),
    alpha = 0.78
  ) +
  geom_point(
    data = selected_point,
    aes(x = recall, y = precision),
    inherit.aes = FALSE,
    shape = 21,
    size = 7,
    stroke = 1.3,
    fill = "white",
    color = "#1F2D3D"
  ) +
  geom_label(
    data = selected_point,
    aes(
      x = pmin(recall + 0.08, 0.86),
      y = pmin(precision + 0.08, 0.92),
      label = paste0(
        "Selected Model\n",
        Model_Config,
        "\nHoldout F1 = ",
        scales::number(f_meas, accuracy = 0.001)
      )
    ),
    inherit.aes = FALSE,
    size = 3.1,
    fontface = "bold",
    color = "#243447",
    fill = "white",
    label.size = 0.25
  ) +
  scale_color_manual(values = model_colors) +
  scale_shape_manual(
    values = c(
      "No Sampling" = 16,
      "Up Sampling" = 17,
      "Down Sampling" = 15,
      "SMOTE" = 3
    )
  ) +
  scale_size_continuous(
    range = c(3.2, 6.8),
    guide = "none"
  ) +
  scale_x_continuous(
    limits = c(0, 1),
    labels = scales::percent_format(accuracy = 1)
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    labels = scales::percent_format(accuracy = 1)
  ) +
  labs(
    title = "Precision–Recall Tradeoff by Model Configuration",
    subtitle = "Holdout results compare detection capability against prediction reliability",
    x = "Recall: Share of quiet-quitting cases detected",
    y = "Precision: Reliability of quiet-quitting predictions",
    color = "Model",
    shape = "Sampling"
  ) +
  guides(
    color = guide_legend(order = 1, override.aes = list(size = 4, alpha = 1)),
    shape = guide_legend(order = 2, override.aes = list(size = 4, alpha = 1))
  ) +
  theme_hr(11) +
  theme(
    plot.title = element_text(face = "bold", size = 15),
    plot.subtitle = element_text(
  size = 12,
  color = "#5a6b7c",
  margin = margin(b = 12)
),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 9),
    legend.position = "right",
    legend.box = "vertical",
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 8.5),
    panel.grid.minor = element_blank(),
    plot.margin = margin(10, 20, 15, 10)
  )

print(p)
save_plot(p, "precision_recall_tradeoff.png", 10.5, 5.8)
```

The precision–recall view shows how each finalized configuration performs on the holdout test set after cross-validation and tuning. Configurations farther to the right detect more quiet-quitting cases, while configurations higher on the chart produce more reliable positive predictions.

For this target, the holdout results show a difficult tradeoff: models that detect more potential quiet-quitting cases may also produce less reliable positive predictions. This is why the selected model should be interpreted as a prioritization screen rather than a definitive classifier.

The selected configuration is highlighted on the plot. It was not chosen from holdout results alone; it was chosen through a two-step process that compares cross-validated performance and then confirms that the selected workflow remains practically usable on unseen data.

## Cross-Validated Model Comparison

```{r 17-model-decision-breakdown, echo=FALSE}
decision_table <- cv_results_all %>%
  group_by(Model, Sampling) %>%
  summarise(
    Accuracy = mean(mean[.metric == "accuracy"], na.rm = TRUE),
    Precision = mean(mean[.metric == "precision"], na.rm = TRUE),
    Recall = mean(mean[.metric == "recall"], na.rm = TRUE),
    F1 = mean(mean[.metric == "f_meas"], na.rm = TRUE),
    ROC_AUC = mean(mean[.metric == "roc_auc"], na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(F1))

decision_table %>%
  dplyr::slice_head(n = 12) %>%
  gt() %>%
  tab_header(
    title = "Cross-Validated Model Comparison",
    subtitle = "Top configurations ranked by cross-validated F1 Score"
  ) %>%
  fmt_number(
    columns = c(Accuracy, Precision, Recall, F1, ROC_AUC),
    decimals = 3
  ) %>%
  data_color(
    columns = c(Accuracy, Precision, Recall, F1, ROC_AUC),
    colors = scales::col_numeric(
      palette = c("#EAF2F8", "#A9CCE3", "#2E86C1"),
      domain = c(0, 1)
    )
  ) %>%
  cols_align("center") %>%
  tab_options(
    table.width = pct(92),
    table.font.size = px(12),
    data_row.padding = px(8)
  )
```

### Model Selection {#model-selection}

### Model Selection {#model-selection}

Model selection was guided by a two-step rule designed to balance statistical rigor with business usefulness.

First, candidate configurations were ranked by cross-validated F1 Score, which serves as the primary model-selection metric because it balances precision and recall under class imbalance. Second, the leading candidates were reviewed on the holdout test set to confirm that they still produced usable positive-class predictions on unseen data.

This second step matters. A configuration can rank highly during resampling yet behave too aggressively or too weakly on the holdout set. For this reason, the final model is selected as the strongest cross-validated configuration that also satisfies minimum holdout stability requirements.

In this run, the recommended configuration was **`r top_model_name` with `r top_sampling_method`**. The model should be interpreted as a **risk prioritization tool**, meaning it helps identify employees or workforce segments that may warrant earlier engagement review, while leaving final decisions to HR leaders, managers, and organizational context.

## Selected Model Holdout Performance

```{r 18-final-model-metrics-plot, echo=FALSE}
final_model_plot_df <- final_model_metrics %>%
  mutate(
    Metric = factor(
      Metric,
      levels = c("Accuracy", "Precision", "Recall", "F1 Score", "ROC AUC")
    ),
    Highlight = ifelse(Metric == "F1 Score", "Selected Metric", "Supporting Metric"),
    Plot_Label = ifelse(is.na(Score), "N/A", scales::number(Score, accuracy = 0.001))
  )

y_max <- min(
  1.08,
  max(final_model_plot_df$Score, na.rm = TRUE) + 0.12
)

p <- ggplot(final_model_plot_df, aes(x = Metric, y = Score, fill = Highlight)) +
  geom_col(width = 0.65, na.rm = TRUE) +
  geom_text(
    aes(label = Plot_Label),
    vjust = -0.45,
    fontface = "bold",
    size = 4
  ) +
  scale_fill_manual(values = c(
    "Selected Metric" = "#2E86C1",
    "Supporting Metric" = "#AAB7B8"
  )) +
  scale_y_continuous(
    limits = c(0, y_max),
    labels = scales::percent_format(accuracy = 1)
  ) +
  labs(
    title = paste("Final Holdout Performance:", top_model_name),
    subtitle = paste("Sampling Strategy:", top_sampling_method),
    x = NULL,
    y = "Score"
  ) +
  theme_hr(11) +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(color = "#5a6b7c"),
    axis.text.x = element_text(face = "bold"),
    plot.margin = margin(10, 20, 15, 10)
  )

print(p)
save_plot(p, "final_model_metrics.png", 8, 4.6)
```

## Risk Distribution

The selected model assigns relative disengagement-risk scores across the workforce population.

Rather than interpreting probabilities as deterministic outcomes, this framework treats prediction scores as workforce-prioritization signals that help HR teams identify where additional engagement review may be warranted.

Because quiet quitting reflects behavioral and perception-based disengagement patterns, risk interpretation should emphasize relative prioritization rather than absolute certainty.

```{r 19-risk-distribution, echo=FALSE}
risk_distribution <- final_preds %>%
  mutate(
    predicted_probability = .pred_Yes,
    risk_percentile = percent_rank(predicted_probability),
    review_priority = case_when(
      risk_percentile >= 0.90 ~ "High Review Priority",
      risk_percentile >= 0.60 ~ "Moderate Review Priority",
      TRUE ~ "Lower Review Priority"
    ),
    review_priority = factor(
      review_priority,
      levels = c(
        "Lower Review Priority",
        "Moderate Review Priority",
        "High Review Priority"
      )
    )
  )

p <- ggplot(
  risk_distribution,
  aes(x = predicted_probability, fill = review_priority)
) +
  geom_histogram(
    bins = 28,
    color = "white",
    alpha = 0.90
  ) +
  scale_fill_manual(
    values = c(
      "Lower Review Priority" = "#D5E8D4",
      "Moderate Review Priority" = "#F9E79F",
      "High Review Priority" = "#F5B7B1"
    )
  ) +
  scale_x_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, 1),
    expand = expansion(mult = c(0.01, 0.01))
  ) +
  labs(
    title = "Predicted Quiet Quitting Review-Priority Distribution",
    subtitle = "Relative disengagement-risk signals across the workforce population",
    x = "Predicted Probability of Quiet Quitting",
    y = "Employee Count",
    fill = "Review Priority"
  ) +
  theme_hr(11) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(color = "#5a6b7c", size = 11),
    legend.position = "top",
    panel.grid.minor = element_blank(),
    plot.margin = margin(10, 20, 15, 10)
  )

print(p)
save_plot(p, "risk_distribution.png", 8.2, 5.2)
```

## Review-Priority Segmentation

The segmentation view translates model probabilities into practical review-priority tiers. These tiers are not automatic decisions; they organize attention for HR, manager, or workforce planning review.

Because quiet quitting probabilities may cluster within a narrow range, relative ranking is more useful than fixed cutoffs. In practice, HR teams could adjust review queues based on available capacity, intervention goals, employee-relations sensitivity, and tolerance for missed disengagement risk.

```{r 20-risk-segmentation, echo=FALSE, message=FALSE, warning=FALSE}
risk_segment_summary <- risk_distribution %>%
  count(review_priority, .drop = FALSE) %>%
  mutate(
    percent = n / sum(n),
    label = paste0(n, "\n", scales::percent(percent, accuracy = 1))
  )

p <- ggplot(
  risk_segment_summary,
  aes(
    x = review_priority,
    y = n,
    fill = review_priority
  )
) +
  geom_col(width = 0.65) +
  geom_text(
    aes(label = label),
    vjust = -0.45,
    fontface = "bold",
    size = 4
  ) +
  scale_fill_manual(
    values = c(
      "Lower Review Priority" = "#7FBF7F",
      "Moderate Review Priority" = "#F39C12",
      "High Review Priority" = "#D35454"
    )
  ) +
  scale_y_continuous(
    limits = c(0, max(risk_segment_summary$n, na.rm = TRUE) * 1.18),
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(
    title = "Employee Engagement Review-Priority Segmentation",
    subtitle = "Risk tiers translate model scores into practical engagement review priorities",
    x = NULL,
    y = "Employee Count"
  ) +
  theme_hr(11) +
  theme(
    plot.title = element_text(face = "bold", size = 15),
    plot.subtitle = element_text(color = "#5a6b7c", size = 11),
    axis.text.x = element_text(size = 10, face = "bold"),
    legend.position = "none",
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    plot.margin = margin(10, 18, 18, 10)
  )

print(p)
save_plot(p, "risk_segments.png", 7.8, 5)
```

For quiet quitting, these tiers should be treated as engagement-prioritization signals. Lower-priority employees may require normal engagement maintenance, moderate-priority employees may warrant trend monitoring, and high-priority employees may warrant earlier manager or HR review.

# Final Model Summary & Key Takeaways {#model-summary}

The selected model represents the most practical configuration after comparing cross-validated performance with holdout test behavior. For quiet quitting, this distinction matters because some configurations performed strongly during resampling but weakened when applied to unseen data.

While the absolute holdout performance values remain moderate, the model is still useful as a **risk prioritization tool** because it helps rank employees by relative likelihood of quiet quitting intention. This distinction is important: the model should not be used to make automatic employment decisions. Instead, it supports earlier review, targeted engagement conversations, and more focused workforce planning.

## 📊 Key Takeaways

```{=html}
<div class="model-highlight-card">

  <div class="model-badge">Recommended Model Selected</div>

  <h3 class="model-title">
    `r top_model_name`
  </h3>

  <p class="model-subtitle">
    Optimized using <strong>`r top_sampling_method`</strong> with a focus on <strong>F1 Score</strong>
  </p>

  <div class="model-metrics">
    <div class="metric-box">
      <span class="metric-label">CV F1 Score</span>
      <span class="metric-value">
        `r top_f1`
      </span>
    </div>

    <div class="metric-box">
      <span class="metric-label">Strategy</span>
      <span class="metric-value">
        `r selection_strategy_label`
      </span>
    </div>

    <div class="metric-box">
      <span class="metric-label">Use Case</span>
      <span class="metric-value">
        Disengagement Risk Targeting
      </span>
    </div>
  </div>

</div>
```

## Why This Model Was Selected

```{=html}
<div class="decision-box">

<h3>Why This Model Was Selected</h3>

<p>
The final model was selected using a two-step process: cross-validation identified strong candidate configurations, and the holdout test set confirmed whether those candidates remained practically usable on unseen data.
</p>

<p>
In this run, the selected configuration was <strong>`r top_model_name` with `r top_sampling_method`</strong>. It achieved a cross-validated F1 Score of <strong>`r top_f1`</strong>, while its holdout test results produced precision of <strong>`r top_holdout_precision`</strong>, recall of <strong>`r top_holdout_recall`</strong>, and F1 Score of <strong>`r top_holdout_f1`</strong>.
</p>

<p>
`r selection_rule_note`
</p>

<p>
Because quiet quitting is a behavioral risk signal rather than a directly observable event in this synthetic dataset, predictions should be treated as prompts for further HR or manager review, not as standalone conclusions about employee behavior.
</p>

</div>
```

## Selected Model Performance Metrics

The table below reports holdout test performance for the selected configuration. These values should be interpreted as final confirmation metrics, not as standalone proof that the model is definitive. For quiet quitting, the model is most useful for prioritization and early review rather than individual-level certainty.

```{r 21-final-model-performance-table, echo=FALSE}
final_model_metrics %>%
  select(Metric, Score) %>%
  mutate(
    Score = ifelse(is.na(Score), NA_real_, Score)
  ) %>%
  gt() %>%
  tab_header(
    title = "Selected Model Performance Summary",
    subtitle = paste("Holdout test performance for", top_model_name, "with", top_sampling_method)
  ) %>%
  fmt_number(columns = Score, decimals = 3) %>%
  sub_missing(columns = Score, missing_text = "N/A") %>%
  cols_align("center") %>%
  tab_options(
    table.width = pct(55),
    table.font.size = px(13)
  )
```

## Top Model Configurations

```{r 22-top-models-table, echo=FALSE}
top_config_table_all <- cv_results_all %>%
  filter(.metric == "f_meas") %>%
  transmute(
    Model = as.character(Model),
    Sampling = as.character(Sampling),
    `CV F1 Score` = round(mean, 3)
  ) %>%
  left_join(
    holdout_screen %>%
      transmute(
        Model,
        Sampling,
        `Holdout Precision` = round(precision, 3),
        `Holdout Recall` = round(recall, 3),
        `Holdout F1` = round(f_meas, 3),
        Status = case_when(
          Model == top_model_name & Sampling == top_sampling_method ~ "Selected Final Model",
          stable_holdout ~ "Holdout-Stable Alternative",
          viable_holdout ~ "Viable but Below Stability Screen",
          TRUE ~ "Not Selected"
        )
      ),
    by = c("Model", "Sampling")
  ) %>%
  mutate(
    Selected_Row = Model == top_model_name & Sampling == top_sampling_method
  )

selected_row <- top_config_table_all %>%
  filter(Selected_Row)

comparison_rows <- top_config_table_all %>%
  dplyr::filter(!Selected_Row) %>%
  dplyr::arrange(dplyr::desc(`CV F1 Score`), dplyr::desc(`Holdout F1`)) %>%
  dplyr::slice_head(n = 9)

top_config_table <- bind_rows(selected_row, comparison_rows) %>%
  mutate(
    Status = factor(
      Status,
      levels = c(
        "Selected Final Model",
        "Holdout-Stable Alternative",
        "Viable but Below Stability Screen",
        "Not Selected"
      )
    )
  )

top_config_table %>%
  select(-Selected_Row) %>%
  gt() %>%
  tab_header(
    title = "Top Model Configurations",
    subtitle = "Selected model shown with leading cross-validated alternatives"
  ) %>%
  fmt_number(
    columns = c(`CV F1 Score`, `Holdout Precision`, `Holdout Recall`, `Holdout F1`),
    decimals = 3
  ) %>%
  sub_missing(
    columns = c(`Holdout Precision`, `Holdout Recall`, `Holdout F1`),
    missing_text = "N/A"
  ) %>%
  cols_align("center") %>%
  data_color(
    columns = c(`CV F1 Score`, `Holdout Precision`, `Holdout Recall`, `Holdout F1`),
    colors = scales::col_numeric(
      palette = c("#EAF2F8", "#A9CCE3", "#2E86C1"),
      domain = c(0, 1),
      na.color = "#F2F4F4"
    )
  ) %>%
  cols_width(
    Model ~ px(130),
    Sampling ~ px(150),
    `CV F1 Score` ~ px(120),
    `Holdout Precision` ~ px(130),
    `Holdout Recall` ~ px(120),
    `Holdout F1` ~ px(110),
    Status ~ px(220)
  ) %>%
  tab_style(
    style = list(
      cell_fill(color = "#EBF5FB"),
      cell_text(weight = "bold")
    ),
    locations = cells_body(
      rows = Status == "Selected Final Model"
    )
  ) %>%
  tab_options(
    table.width = pct(92),
    table.font.size = px(12)
  )
```

Performance metrics show that several configurations performed strongly during cross-validation but did not transfer cleanly to the holdout test set. This is especially important for quiet quitting, where the outcome reflects behavioral and perception-based disengagement rather than a directly observed employment event.

For this reason, the selected model should be interpreted as a holdout-viable prioritization screen. It detects quiet-quitting cases aggressively, but its low precision means results should be used to prioritize review, not to make definitive employee-level conclusions.

# Driver Interpretation {#driver-interpretation}

Because the selected predictive workflow does not provide a straightforward, model-agnostic feature-importance ranking, driver interpretation is based on the broader analytical pattern observed across the modeling and exploratory analysis.

Quiet quitting intention is not driven by a single variable. It reflects a network of behavioral, organizational, wellbeing, and perception-based signals that collectively indicate disengagement risk.

The modeling framework demonstrates how predictive analytics can move beyond static reporting and support proactive workforce engagement planning. Even when prediction is imperfect, the model provides a structured way to identify potential disengagement risk earlier, prioritize attention, and guide more targeted engagement strategies.

Used responsibly, this type of model becomes a decision-support tool: not a replacement for HR expertise, but a way to help organizations focus attention sooner, allocate support more effectively, and improve long-term employee experience outcomes.

## Feature → Action Mapping {#feature-action}

To translate model insights into practical action, key quiet-quitting risk areas are mapped to targeted workforce interventions.

This bridges the gap between predictive analytics and real-world decision-making by connecting model interpretation to practical HR response strategies focused on engagement, discretionary effort, role clarity, wellbeing, and manager support.

```{r 23-driver-action-table, echo=FALSE}
driver_actions <- tibble::tribble(
  ~Risk_Area, ~What_It_Suggests, ~Recommended_Action,
  "Engagement and discretionary effort",
  "Employees showing lower engagement or reduced discretionary effort may be more likely to withdraw effort while remaining employed.",
  "Strengthen recognition, role connection, team engagement practices, and meaningful work conversations.",

  "Burnout, stress, and wellbeing",
  "Elevated strain may increase emotional withdrawal and reduce motivation.",
  "Review workload balance, staffing pressure, wellbeing resources, and manager support expectations.",

  "Role clarity and career growth",
  "Employees with unclear responsibilities or limited growth perception may disengage over time.",
  "Improve career pathing, role expectations, development planning, and internal mobility conversations.",

  "Manager and leadership support",
  "Weak support signals may reduce trust, commitment, and willingness to go beyond minimum expectations.",
  "Improve manager coaching, feedback quality, communication cadence, and follow-up on employee concerns.",

  "Work environment and fairness",
  "Perceived unfairness or poor work environment conditions may contribute to quiet-quitting behavior.",
  "Evaluate team climate, fairness perceptions, workload equity, and decision-making transparency."
)

driver_actions %>%
  gt() %>%
  tab_header(
    title = "Quiet Quitting Risk Driver-to-Action Map",
    subtitle = "Translating model insights into workforce intervention strategy"
  ) %>%
  cols_label(
    Risk_Area = "Risk Area",
    What_It_Suggests = "What It Suggests",
    Recommended_Action = "Recommended Action"
  ) %>%
  cols_width(
    Risk_Area ~ px(190),
    What_It_Suggests ~ px(360),
    Recommended_Action ~ px(430)
  ) %>%
  tab_options(
    table.width = pct(90),
    table.font.size = px(13)
  )
```

# Business Impact & Decision Guidance {#business-impact}

The modeling framework is designed to support workforce decision-making at a strategic level while preserving employee privacy through synthetic data simulation.

Rather than functioning as an automated employment-decision system, the selected model should be interpreted as a workforce intelligence layer that helps HR teams prioritize where additional engagement review, manager follow-up, or workforce support analysis may be beneficial.

The practical value of the framework lies in its ability to:

- Identify workforce segments that may warrant earlier engagement review
- Support proactive disengagement-risk monitoring
- Improve allocation of HR and manager follow-up resources
- Enhance organizational visibility into workforce experience patterns
- Demonstrate privacy-preserving HR analytics workflows

Importantly, the model is intended to support human decision-making, not replace it. Quiet-quitting risk remains dependent on organizational context, leadership interpretation, employee communication, and ethical review processes.

This project demonstrates how predictive analytics can be integrated into HR environments responsibly using synthetic, research-informed data that protects employee confidentiality while still enabling advanced workforce analysis.

## Strategic Decision Guidance {#decision-guidance}

```{=html}
<div class="decision-box">

<p>
This model is best deployed as a <strong>risk prioritization tool</strong>, not as a standalone decision system.
</p>

<ul>
  <li><strong>High-risk employees</strong> → Prioritize for HR, manager, or engagement review</li>
  <li><strong>Moderate-risk employees</strong> → Monitor engagement, workload, role clarity, and support trends</li>
  <li><strong>Low-risk employees</strong> → Maintain baseline engagement practices and regular manager check-ins</li>
</ul>

<p>
The value of the model is not in replacing human judgment, but in helping decision-makers focus attention earlier, identify possible disengagement patterns, and support employees before withdrawal becomes more severe.
</p>

</div>
```

## Model Limitations {#model-limitations}

```{=html}
<div class="decision-box">

<h3>Model Limitations</h3>

<p>
The model should be interpreted within the context of synthetic data, simulated workforce patterns, and organizational variability.
</p>

<ul>
  <li>Predictions reflect patterns in synthetic data, not observed outcomes from a real organization</li>
  <li>Performance may shift when applied to different employee populations, industries, or engagement cultures</li>
  <li>Risk scores should support review and prioritization, not automatic employment decisions</li>
  <li>Human context remains essential before any intervention is taken</li>
</ul>

</div>
```

## Final Takeaway

Quiet quitting is not driven by a single variable. It reflects a network of behavioral, organizational, wellbeing, engagement, role clarity, and perception-based signals that collectively indicate disengagement risk.

The modeling framework demonstrates how predictive analytics can move beyond static reporting and support proactive workforce engagement planning. Even with imperfect prediction, the model provides a structured way to identify potential disengagement risk earlier, prioritize attention, and guide more targeted engagement strategies.

Used responsibly, this type of model becomes a decision-support tool: not a replacement for HR expertise, but a way to help organizations act earlier, allocate support more effectively, strengthen manager follow-up, and improve long-term employee experience outcomes.

```{r 24-export-final-model-summary, include=FALSE}
summary_dir <- here::here("data", "processed", "model_summaries")
dir.create(summary_dir, recursive = TRUE, showWarnings = FALSE)

top_holdout_accuracy <- final_model_metrics %>%
  filter(Metric == "Accuracy") %>%
  pull(Score) %>%
  round(3)

final_model_summary <- tibble::tibble(
  Outcome = "Quiet Quitting",
  Model = top_model_name,
  Sampling = top_sampling_method,
  Model_Config = top_model_config,
  Selection_Strategy = selection_strategy_label,
  Selection_Rule = selection_rule_note,
  CV_F1 = top_f1,
  Holdout_Accuracy = top_holdout_accuracy,
  Holdout_Precision = top_holdout_precision,
  Holdout_Recall = top_holdout_recall,
  Holdout_F1 = top_holdout_f1,
  Holdout_ROC_AUC = top_holdout_roc_auc
)

saveRDS(
  final_model_summary,
  here::here("data", "processed", "model_summaries", "quiet_quitting_model_summary.rds")
)
```

```{=html}
<hr class="section-divider">

<div class="page-nav">

  <div class="nav-left">
    <a href="HR_Data_Turnover_Intention.html" class="cta-btn btn-blue">← Back</a>
  </div>

  <div class="nav-center">
    <a href="index.html" class="cta-btn btn-teal">Home</a>
  </div>

  <div class="nav-right">
    <a href="HR_Data_Monitoring_Privacy.html" class="cta-btn btn-red">Next →</a>
  </div>

</div>

</div> <!-- page-content -->
</div> <!-- page-layout -->
</div> <!-- main-container -->
```